### 2018-07-18 Introduction/Introduction.tex
First commit

\documentclass[letterpaper]{article}
\usepackage{graphicx}
\usepackage{subfigure}
\graphicspath{ {images/} }
\usepackage{enumerate}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{minted}
\usepackage{tcolorbox}
\usepackage[margin=1in]{geometry}
\usepackage{float}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{indentfirst}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
\renewcommand{\arraystretch}{1.2}

\begin{document}

\title{Master's Thesis}
\author{Sun Myung Park}
\date{}

\maketitle

\section*{Introduction}

\end{document}

### 2018-08-08 Introduction/Introduction.tex
Introduction work

@@
 
+With the world population slated to continue growing for the rest of the 21st century, our energy consumption is likely to follow suit.
+The bulk of energy production today comes from burning fossil fuels.
+However, the resulting release of carbon dioxide, a greenhouse gas, has contributed to rising temperatures and more severe weather phenomena around the globe.
+It is imperative that we eliminate the major source of greenhouse gases in fossil fuel burning power plants as quickly as possible to prevent worsening climate change.
+
 \end{document}

### 2019-07-24 introduction.tex
Intro first draft

Rapidly rising greenhouse gas emissions since the Industrial Revolution in the
1800s have contributed significantly to global warming, the consequences of
which have been acutely felt in recent years through increasingly frequent and
record breaking extreme weather events. Urgent measures are necessary to limit
CO$_2$ emissions and the negative impacts of global warming.

Since electricity generation from burning fossil fuels represents the
largest source of CO$_2$ emissions (38\% in 2018 \cite{noauthor_global_2018}), replacing them with
low-carbon
alternatives would effectively curb a large fraction of these emissions. Out
of all low-carbon power sources, nuclear power in general is best suited to
replace fossil fuel burning; it provides consistent base load
power independent of weather and local geographical conditions. Solar and wind
power are dependent on favorable weather conditions, while hydropower requires
relatively more area and is limited
to geologically appropriate locations. This argument
provides a strong case for nuclear power being the most appropriate and
realistic alternative to burning fossil fuels for our current and future
electricity needs.

Large scale reactor deployments would be necessary to displace the
presently large share of energy production from fossil fuel power plants.
However, several obstacles stand in the way of mass reactor deployments. These
include perceived safety risks, sustainability concerns, nuclear proliferation
risks and the ability to compete economically with other sources of energy.
The Molten Salt Reactor (MSR) concept, one of six advanced reactor designs
selected by the Generation IV International Forum
(GIF)* for continued research and development, is a potential solution to the
aforementioned issues.

The primary coolant in MSRs is a molten salt mixture,
with fissile and/or fertile material directly dissolved in the coolant.
MSRs possess an inherently robust safety feature in the
strongly negative fuel temperature reactivity coefficient from Doppler
broadening and thermal fuel expansion that greatly reduces
the risk of a reactor power excursion. Many designs can also incorporate the
thorium fuel cycle for improved sustainability arising from the use of abundant
natural thorium resources and reduced transuranic waste. The resultant reduced
radiotoxicity from transuranic waste also reduces costs associated with
long-term nuclear waste storage.
In addition, the ability to operate at near atmospheric pressures eliminates
the need for a thick pressure vessel and drives down construction
costs, while online fuel reprocessing reduces reactor downtime during reactor
operation.

### 2019-10-21 acros.tex
chp 1, 2, 3 update

%\newacronym{<++>}{<++>}{<++>}
\newacronym[longplural={metric tons of heavy metal}]{MTHM}{MTHM}{metric ton of heavy metal}
\newacronym{ABM}{ABM}{agent-based modeling}
\newacronym{ACDIS}{ACDIS}{Program in Arms Control \& Domestic and International Security}
\newacronym{AHTR}{AHTR}{Advanced High Temperature Reactor}
\newacronym{ANDRA}{ANDRA}{Agence Nationale pour la gestion des D\'echets RAdioactifs, the French National Agency for Radioactive Waste Management}
\newacronym{ANL}{ANL}{Argonne National Laboratory}
\newacronym{API}{API}{application programming interface}
\newacronym{ARE}{ARE}{Aircraft Reactor Experiment}
\newacronym{ASME}{ASME}{American Society of Mechanical Engineers}
\newacronym{ATWS}{ATWS}{Anticipated Transient Without Scram}
\newacronym{BDBE}{BDBE}{Beyond Design Basis Event}
\newacronym{BIDS}{BIDS}{Berkeley Institute for Data Science}
\newacronym{CAFCA}{CAFCA}{ Code for Advanced Fuel Cycles Assessment }
\newacronym{CDTN}{CDTN}{Centro de Desenvolvimento da Tecnologia Nuclear}
\newacronym{CEA}{CEA}{Commissariat \`a l'\'Energie Atomique et aux \'Energies Alternatives}
\newacronym{CFD}{CFD}{Computational Fluid Dynamics}
\newacronym{CI}{CI}{continuous integration}
\newacronym{CNEN}{CNEN}{Comiss\~{a}o Nacional de Energia Nuclear}
\newacronym{CNERG}{CNERG}{Computational Nuclear Engineering Research Group}
\newacronym{COMSOL}{COMSOL}{COMmon SOLution}
\newacronym{COSI}{COSI}{Commelini-Sicard}
\newacronym{COTS}{COTS}{commercial, off-the-shelf}
\newacronym{CSNF}{CSNF}{commercial spent nuclear fuel}
\newacronym{CTAH}{CTAHs}{Coiled Tube Air Heaters}
\newacronym{CUBIT}{CUBIT}{CUBIT Geometry and Mesh Generation Toolkit}
\newacronym{CURIE}{CURIE}{Centralized Used Fuel Resource for Information Exchange}
\newacronym{DAG}{DAG}{directed acyclic graph}
\newacronym{DANESS}{DANESS}{Dynamic Analysis of Nuclear Energy System Strategies}
\newacronym{DBE}{DBE}{Design Basis Event}
\newacronym{DESAE}{DESAE}{Dynamic Analysis of Nuclear Energy Systems Strategies}
\newacronym{DHS}{DHS}{Department of Homeland Security}
\newacronym{DNP}{DNP}{delayed neutron precursor}
\newacronym{DOE}{DOE}{Department of Energy}
\newacronym{DRACS}{DRACS}{Direct Reactor Auxiliary Cooling System}
\newacronym{DRE}{DRE}{dynamic resource exchange}
\newacronym{DSNF}{DSNF}{DOE spent nuclear fuel}
\newacronym{DYMOND}{DYMOND}{Dynamic Model of Nuclear Development }
\newacronym{EBS}{EBS}{Engineered Barrier System}
\newacronym{EDZ}{EDZ}{Excavation Disturbed Zone}
\newacronym{EPA}{EPA}{Environmental Protection Agency}
\newacronym{EP}{EP}{Engineering Physics}
\newacronym{EVOL}{EVOL}{Evaluation and Viability of Liquid Fuel Fast Reactor System}
\newacronym{FCO}{FCO}{Fuel Cycle Options}
\newacronym{FCT}{FCT}{Fuel Cycle Technology}
\newacronym{FEHM}{FEHM}{Finite Element Heat and Mass Transfer}
\newacronym{FEPs}{FEPs}{Features, Events, and Processes}
\newacronym{FHR}{FHR}{Fluoride-Salt-Cooled High-Temperature Reactor}
\newacronym{FLiBe}{FLiBe}{Fluoride-Lithium-Beryllium}
\newacronym{GDSE}{GDSE}{Generic Disposal System Environment}
\newacronym{GDSM}{GDSM}{Generic Disposal System Model}
\newacronym{GENIUSv1}{GENIUSv1}{Global Evaluation of Nuclear Infrastructure Utilization Scenarios, Version 1}
\newacronym{GENIUSv2}{GENIUSv2}{Global Evaluation of Nuclear Infrastructure Utilization Scenarios, Version 2}
\newacronym{GENIUS}{GENIUS}{Global Evaluation of Nuclear Infrastructure Utilization Scenarios}
\newacronym{GIF}{GIF}{Generation IV International Forum}
\newacronym{GPAM}{GPAM}{Generic Performance Assessment Model}
\newacronym{GRSAC}{GRSAC}{Graphite Reactor Severe Accident Code}
\newacronym{GUI}{GUI}{graphical user interface}
\newacronym{HLW}{HLW}{high level waste}
\newacronym{HPC}{HPC}{high-performance computing}
\newacronym{HTC}{HTC}{high-throughput computing}
\newacronym{HTGR}{HTGR}{High Temperature Gas-Cooled Reactor}
\newacronym{IAEA}{IAEA}{International Atomic Energy Agency}
\newacronym{IEMA}{IEMA}{Illinois Emergency Mangament Agency}
\newacronym{IHLRWM}{IHLRWM}{International High Level Radioactive Waste Management}
\newacronym{INL}{INL}{Idaho National Laboratory}
\newacronym{IPRR1}{IRP-R1}{Instituto de Pesquisas Radioativas Reator 1}
\newacronym{IRP}{IRP}{Integrated Research Project}
\newacronym{ISFSI}{ISFSI}{Independent Spent Fuel Storage Installation}
\newacronym{ISRG}{ISRG}{Independent Student Research Group}
\newacronym{JFNK}{JFNK}{Jacobian-Free Newton Krylov}
\newacronym{LANL}{LANL}{Los Alamos National Laboratory}
\newacronym{LBNL}{LBNL}{Lawrence Berkeley National Laboratory}
\newacronym{LCOE}{LCOE}{levelized cost of electricity}
\newacronym{LDRD}{LDRD}{laboratory directed research and development}
\newacronym{LEU}{LEU}{low-enriched uranium}
\newacronym{LFR}{LFR}{Lead-Cooled Fast Reactor}
\newacronym{LLNL}{LLNL}{Lawrence Livermore National Laboratory}
\newacronym{LMFBR}{LMFBR}{Liquid Metal Fast Breeder Reactor}
\newacronym{LOFC}{LOFC}{Loss of Forced Cooling}
\newacronym{LOHS}{LOHS}{Loss of Heat Sink}
\newacronym{LOLA}{LOLA}{Loss of Large Area}
\newacronym{LP}{LP}{linear program}
\newacronym{LWR}{LWR}{Light Water Reactor}
\newacronym{MAGNOX}{MAGNOX}{Magnesium Alloy Graphie Moderated Gas Cooled Uranium Oxide Reactor}
\newacronym{MA}{MA}{minor actinide}
\newacronym{MCNP}{MCNP}{Monte Carlo N-Particle code}
\newacronym{MILP}{MILP}{mixed-integer linear program}
\newacronym{MIT}{MIT}{the Massachusetts Institute of Technology}
\newacronym{MOAB}{MOAB}{Mesh-Oriented datABase}
\newacronym{MOOSE}{MOOSE}{Multiphysics Object-Oriented Simulation Environment}
\newacronym{MOSART}{MOSART}{Molten Salt Actinide Recycler and Transmuter}
\newacronym{MOX}{MOX}{mixed oxide}
\newacronym{MPI}{MPI}{Message Passing Interface}
\newacronym{MRPP}{MRPP}{Multiregion Processing Plant}
\newacronym{MSBR}{MSBR}{Molten Salt Breeder Reactor}
\newacronym{MSFR}{MSFR}{Molten Salt Fast Reactor}
\newacronym{MSRE}{MSRE}{Molten Salt Reactor Experiment}
\newacronym{MSR}{MSR}{Molten Salt Reactor}
\newacronym{NAGRA}{NAGRA}{National Cooperative for the Disposal of Radioactive Waste}
\newacronym{NEAMS}{NEAMS}{Nuclear Energy Advanced Modeling and Simulation}
\newacronym{NEUP}{NEUP}{Nuclear Energy University Programs}
\newacronym{NFCSim}{NFCSim}{Nuclear Fuel Cycle Simulator}
\newacronym{NGNP}{NGNP}{Next Generation Nuclear Plant}
\newacronym{NMWPC}{NMWPC}{Nuclear MW Per Capita}
\newacronym{NNSA}{NNSA}{National Nuclear Security Administration}
\newacronym{NPP}{NPP}{Nuclear Power Plant}
\newacronym{NPRE}{NPRE}{Department of Nuclear, Plasma, and Radiological Engineering}
\newacronym{NQA1}{NQA-1}{Nuclear Quality Assurance - 1}
\newacronym{NRC}{NRC}{Nuclear Regulatory Commission}
\newacronym{NSF}{NSF}{National Science Foundation}
\newacronym{NSSC}{NSSC}{Nuclear Science and Security Consortium}
\newacronym{NUWASTE}{NUWASTE}{Nuclear Waste Assessment System for Technical Evaluation}
\newacronym{NWF}{NWF}{Nuclear Waste Fund}
\newacronym{NWTRB}{NWTRB}{Nuclear Waste Technical Review Board}
\newacronym{OCRWM}{OCRWM}{Office of Civilian Radioactive Waste Management}
\newacronym{ORION}{ORION}{ORION}
\newacronym{ORNL}{ORNL}{Oak Ridge National Laboratory}
\newacronym{PARCS}{PARCS}{Purdue Advanced Reactor Core Simulator}
\newacronym{PBAHTR}{PB-AHTR}{Pebble Bed Advanced High Temperature Reactor}
\newacronym{PBFHR}{PB-FHR}{Pebble-Bed Fluoride-Salt-Cooled High-Temperature Reactor}
\newacronym{PDE}{PDE}{partial differential equation}
\newacronym{PEI}{PEI}{Peak Environmental Impact}
\newacronym{PH}{PRONGHORN}{PRONGHORN}
\newacronym{PRIS}{PRIS}{Power Reactor Information System}
\newacronym{PRKE}{PRKE}{Point Reactor Kinetics Equations}
\newacronym{PSPG}{PSPG}{Pressure-Stabilizing/Petrov-Galerkin}
\newacronym{PWAR}{PWAR}{Pratt and Whitney Aircraft Reactor}
\newacronym{PWR}{PWR}{Pressurized Water Reactor}
\newacronym{PyNE}{PyNE}{Python toolkit for Nuclear Engineering}
\newacronym{PyRK}{PyRK}{Python for Reactor Kinetics}
\newacronym{QA}{QA}{quality assurance}
\newacronym{RDD}{RD\&D}{Research Development and Demonstration}
\newacronym{RD}{R\&D}{Research and Development}
\newacronym{REE}{REE}{rare earth element}
\newacronym{RELAP}{RELAP}{Reactor Excursion and Leak Analysis Program}
\newacronym{RIA}{RIA}{Reactivity Insertion Accident}
\newacronym{RIF}{RIF}{Region-Institution-Facility}
\newacronym{ROD}{ROD}{Reactor Optimum Design}
\newacronym{SAMOFAR}{SAMOFAR}{Safety Assessment of the Molten Salt Fast Reactor}
\newacronym{SAM}{SAM}{System Analysis Module}
\newacronym{SFR}{SFR}{Sodium-Cooled Fast Reactor}
\newacronym{SINDAG}{SINDA{\textbackslash}G}{Systems Improved Numerical Differencing Analyzer $\backslash$ Gaski}
\newacronym{SKB}{SKB}{Svensk K\"{a}rnbr\"{a}nslehantering AB}
\newacronym{SNF}{SNF}{spent nuclear fuel}
\newacronym{SNL}{SNL}{Sandia National Laboratory}
\newacronym{STC}{STC}{specific temperature change}
\newacronym{SUPG}{SUPG}{Streamline-Upwind/Petrov-Galerkin}
\newacronym{SWF}{SWF}{Separations and Waste Forms}
\newacronym{SWU}{SWU}{Separative Work Unit}
\newacronym{TRIGA}{TRIGA}{Training Research Isotope General Atomic}
\newacronym{TRISO}{TRISO}{Tristructural Isotropic}
\newacronym{TRU}{TRU}{transuranic}
\newacronym{TSM}{TSM}{Total System Model}
\newacronym{TSPA}{TSPA}{Total System Performance Assessment for the Yucca Mountain License Application}
\newacronym{ThOX}{ThOX}{thorium oxide}
\newacronym{UFD}{UFD}{Used Fuel Disposition}
\newacronym{UML}{UML}{Unified Modeling Language}
\newacronym{UOX}{UOX}{uranium oxide}
\newacronym{UQ}{UQ}{uncertainty quantification}
\newacronym{US}{US}{United States}
\newacronym{UW}{UW}{University of Wisconsin}
\newacronym{VISION}{VISION}{the Verifiable Fuel Cycle Simulation Model}
\newacronym{VVER}{VVER}{Voda-Vodyanoi Energetichesky Reaktor (Russian Pressurized Water Reactor)}
\newacronym{VV}{V\&V}{verification and validation}
\newacronym{WIPP}{WIPP}{Waste Isolation Pilot Plant}
\newacronym{YMR}{YMR}{Yucca Mountain Repository Site}
\newacronym{BOL}{BOL}{Beginning-of-Life}
\newacronym{ULOF}{ULOF}{Unprotected Loss of Flow}
\newacronym{LOSCA}{LOSCA}{Loss of Secondary Cooling Accident}
\newacronym{ULOHS}{ULOHS}{Unprotected Loss of Heat Sink}

### 2019-10-21 introduction.tex
chp 1, 2, 3 update

@@
+\section{Background and Motivation}
+
 Rapidly rising greenhouse gas emissions since the Industrial Revolution in the
@@
 Since electricity generation from burning fossil fuels represents the
-largest source of CO$_2$ emissions (38\% in 2018 \cite{noauthor_global_2018}), replacing them with
+largest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019}), replacing them with
 low-carbon
@@
 operation.
+
+The development of MSR designs requires efficient and accurate simulation tools
+for the complex neutronics and thermal-hydraulics behaviors expected in the
+reactor core. Among other uses, these tools are essential in identifying and
+characterizing potential accident scenarios to improve the safety of the design
+and inform efforts towards creating MSR licensing frameworks.
+
+Over the past decade, there has been a significant increase in the number of
+simulation tools made available for transient MSR analyses. They generally fall
+into two categories: loosely coupled, separate, stand-alone codes for
+neutronics and thermal-hydraulic, and multiphysics software for concurrently
+solving neutron transport and thermal-hydraulics. Fiorina et al.
+\cite{fiorina_modelling_2014} presented simulation results from two
+models developed independently at \textit{Politechnico di Milano} and
+\textit{Technical University of Delft} in the context of the Molten Salt Fast
+Reactor (MSFR). The MSFR is a circulating fuel reactor concept under active
+development in the H2020 SAMOFAR project. The fuel consists of a molten salt
+mixture of thorium and uranium fluorides dissolved in lithium fluoride (LiF).
+The former employed the commercial
+multiphysics modeling software COMSOL \cite{comsol_comsol_nodate}, while the
+latter coupled two in-house codes DALTON and HEAT for neutron transport and
+CFD calculations respectively. Both models showed generally good agreement with
+each other with a simplified, 2-D axisymmetric model for several transient
+scenarios.
+
+Another paper by Aufiero et al. showed simulation results for a full 3-D model
+of the MSFR using the multiphysics toolkit OpenFOAM. While running a 3-D model
+is more computationally intensive, it allows for proper modeling of 3-D
+turbulent effects, and the analysis of asymmetric transient scenarios such as
+the single pump failure scenario in the MSFR (out of 16
+pumps). The authors also demonstrated that there exists a clear discrepancy in
+the results between implicit and explicit coupling schemes for the neutronics
+and thermal-hydraulics. This is especially true for MSRs due to the strong
+coupling arising from large temperature feedback coefficients.
+
+The preceding discussion highlights the challenges faced in MSR simulation, and
+the need for a highly efficient simulation tool that incorporates implicitly
+coupled multiphysics with good computational scalability over multiple
+processing units. This paper presents the open source MSR simulation tool,
+Moltres, as a strong contender to overcoming these challenges for simulating
+the MSFR. Moltres is an application code
+built in the Multiphysics Object-Oriented Simulation Environment (MOOSE)
+parallel finite element framework. It solves the deterministic neutronics and
+thermal-hydraulics governing equations with implicit
+time-stepping on a single mesh for up to three dimensions. The MOOSE framework
+is also highly parallelizable with its use of MPI for parallel computing.
+
+\section{Objectives}
+
+In this thesis, we demonstrate the coupled multiphysics modeling capabilities
+of Moltres through the reference configuration of the MSFR design. Moltres
+differs from many other codes used for simulating MSRs due to the fully
+implicit coupling between the neutronics and thermal-hydraulics governing
+equations, to fully account for the tight coupling expected in MSRs due to the
+movement of fuel in the salt. The Reynolds-Averaged Navier-Stokes (RANS)-based
+k-epsilon ($k-\epsilon$)
+turbulence model was implemented for better simulating turbulence effects in
+the high Reynolds number flows expected in MSFRs. The model simulates the mean
+flow in turbulent flow with the introduction of two additional variables, the
+turbulence kinetic energy ($k$) and the rate of dissipation of turbulence
+energy ($\epsilon$). Moltres models
+neutronics and thermal-hydraulics behaviors through the deterministic
+multi-group neutron diffusion and convection-diffusion partial
+differential equations respectively.
+
+The objectives of this thesis are: 1) to demonstrate and verify the new
+$k-\epsilon$ turbulence modeling feature in Moltres, 2) and to study
+steady-state and transient cases with the reference MSFR model and identify
+general characteristic behaviors in MSRs and fast spectrum reactors. The
+results can be verified against common test cases for
+turbulence models, and the MSFR simulation results by Rouch et al. and Aufiero
+et al., both of which couple neutronics with the $k-\epsilon$ turbulence model.
+
+** insert thesis outline here **

### 2019-10-21 model.tex
chp 1, 2, 3 update

The main simulation tool used in this work for studying the \gls{MSFR} is
Moltres. While we used Moltres for the steady state and transient analysis, it
requires a dedicated neutron transport solver such as Serpent 2 to generate
neutron energy group constant data for the deterministic, multi-group
neutronics calculations in Moltres. This chapter provides brief descriptions
of the Serpent 2 code and Moltres.

\section{Serpent 2}

Serpent 2 is a continuous-energy Monte Carlo reactor physics code under
active development at the VTT Technical Research Centre of Finland. It was
born in 2004 out of the need for better general-purpose neutron transport
codes for group constant generation in lattice geometries, and has since grown
to support more general capabilities, for an active user base of more than 500
people worldwide. Serpent is highly parallelizable due to its support for
both MPI and OpenMP parallel programming APIs, and highly validated and
verified against experimental data and well-established codes.

In Serpent, each neutron is tracked through a combination of ray-tracing-based
surface tracking and rejection sampling-based delta-tracking. Users may define
the number of neutrons and the number of active and inactive cycles for each
simulation. Inactive cycles are required for fission source distribution
convergence, before interactions are tallied in the active cycles.
Interaction types and locations are
determined stochastically based on neutron interaction data from established
nuclear data libraries (e.g. ENDF, JEFF). These nuclear data libraries provide
continuous-energy cross section data at discrete temperatures. For other
temperatures, Serpent has a built-in Doppler-broadening preprocessor that
extrapolates the relevant cross section data from a lower temperature.

In the context of this project, we used Serpent with the JEFF-3.1.2 nuclear
data library* to generate group constants
for Moltres. The reactor geometry is based on the axisymmetric \gls{MSFR}
neutronics benchmark geometry published in several previous papers involving
the \gls{MSFR}. The relevant group constant data are collapsed into six
neutron energy groups, provided in table *, and calculated for temperature
values of 900 K to 1200 K at 50 K intervals.

The group constants relevant for neutronics calculations in Moltres are the
macroscopic fission, absorption, and scattering neutron cross sections,
neutron diffusion coefficients, average fission energies, average neutron
yields, inverse neutron speeds, flux spectra, \gls{DNP} decay constants, and
effective delayed neutron fractions. These group constants are extracted from
the Serpent output files using a Python script available from the Github
repository that holds the Moltres source code. The script rewrites the group
constants into a format that Moltres can read.

\section{Moltres}

Moltres is a \gls{MOOSE}-based application for coupled
neutronics/thermal-hydraulics simulations of \glspl{MSR}. \gls{MOOSE} is a
highly parallelizable,
finite-element framework developed at \gls{INL} for simplifying the process of
creating fully-coupled, non-linear, multiphysics solvers. The framework
provides a user-friendly interface for this task through object-oriented
programming in C++. All aspects of a typical multiphysics problem, such as the
terms in the \glspl{PDE}, the initial and boundary conditions, the material
properties, etc., are represented in \gls{MOOSE} as C++ objects. This approach
is helpful for many researchers, a significant fraction of whom do not possess
high-level programming backgrounds, as they are unencumbered by the
technical details and complexicities involved in programming mesh handling
and \gls{PDE} solving in finite element analysis.

\gls{MOOSE} itself relies heavily on LibMesh and PETSc for
its mesh handling and \gls{PDE} solver functionalities. As a result,
\gls{MOOSE} supports adaptive meshing schemes and automatic variable scaling
amongst other features for improved accuracy and performance times. Full
coupling is maintained by the execution of Newton-based solves on the
weak formulations of the multiple \glspl{PDE} to minimize the residual values.
Fully-coupled solves are essential for accurately resolving systems with
strongly interacting physics. The \gls{MSR} concept is one such example, where
the neutronics and thermal-hydraulics are tightly coupled through the Doppler
effect and the temperature dependence of liquid fuel salt density.

\gls{MOOSE}, and Moltres by extension, are capable of up to 3D geometry
modelling. They support a wide range of input mesh file formats listed in a
\gls{MOOSE} webpage, including the commonly used Exodus II file format.
Specifically for the 2D case, axial symmetry is easily imposed by changing
one line of code in the input file, without any changes in the Cartesian
representations of the \glspl{PDE} and boundary conditions in their original
C++ implementations. This feature provides significant computational time
savings for 3D systems that exhibit high axial symmetry. Another important
feature for reducing computational time is the use of MPI for parallel
computing. All \gls{MOOSE}-based codes can be easily compiled and run on high
performance computing clusters.

Both \gls{MOOSE} and Moltres are open source codes licensed under Lesser GNU
Public License (LPGL) terms. In the case of Moltres, this was done in support
of open and transparent research practices as well as to leverage on the ease
of access for peer review and encourage continued development of the code.
Moltres is hosted on a Github repository of the same name. Github provides
version control and continuous integration testing, which ensure existing
physics capabilities are protected from potential bugs inadvertently
introduced when developing new physics.

\subsection{Neutronics}

Focussing our attention on Moltres, the neutronics calculations are
implemented through the standard formulation of the time-dependent multi-group
neutron diffusion equations and \gls{DNP} concentration equations as shown in
equations * and *:
%
\begin{align}
	\frac{1}{v_g} \frac{\partial \phi_g}{\partial t} - \nabla \cdot D_g \nabla
	\phi_g + \Sigma^r_g \phi_g
	= \sum^G_{g \neq g'} \Sigma^s_{g' \rightarrow g} \phi_{g'} + \chi^p_g
	\sum^G_{g'=1} (1-\beta) \nu \Sigma^f_{g'} \phi_{g'} + \chi^d_g \sum^I_i
	\lambda_i C_i, \label{eq1}
\end{align}
%
\begin{align}
	\frac{\partial C_i}{\partial t} = \sum^G_{g'=1} \beta_i \nu \Sigma^f_{g'}
	\phi_{g'} - \lambda_i C_i - \nabla \cdot u C_i. \label{eq2}
\end{align}
%
where
{\small
%
\begin{align*}
	v_g &= \text{average speed of neutrons in group }g \\
	\phi_g &= \text{flux of neutrons in group }g \\
	t &= \text{time} \\
	D_g &= \text{diffusion coefficient of neutrons in group }g \\
	\Sigma^r_g &= \text{macroscopic cross-section for removal of neutrons from
	group }g \\
	\Sigma^s_{g' \rightarrow g} &= \text{macroscopic cross-section of
	scattering from }g' \text{ to }g \\
	\chi^p_g &= \text{prompt fission spectrum neutrons in group }g \\
	G &= \text{total number of discrete neutron groups} \\
	\nu &= \text{average number of neutrons produced per fission} \\
	\Sigma^f_{g} &= \text{macroscopic fission cross-section for neutron in group }g \\
	\chi^d_g &= \text{delayed fission spectrum neutrons in group }g \\
	I &= \text{total number of delayed neutron precursor groups} \\
	\beta &= \text{total delayed neutron fraction} \\
	\beta_i &= \text{delayed neutron fraction of precursor group i} \\
	\lambda_i &= \text{average decay constant of delayed neutron precursors in precursor group }i \\
	C_i &= \text{concentration of delayed neutron precursors in precursor group }i
\end{align*}
}
%

Moltres users may call for an arbitrary number of neutron energy groups as
long as they provide Moltres with the appropriate group constant data. The
number of precursor groups is also arbitrary, though usually pre-determined by
the choice of nuclear data library in the group constant generation step.
Moltres automatically interpolates the group constant data for required
temperatures using one of the many pre-defined interpolation methods available
in \gls{MOOSE}. Once again, users have the freedom to select their
interpolation method of choice.

\subsection{Thermal-Hydraulics}

The governing equation for temperature in the molten salt is given as:
%
\begin{align}
	\rho c_{p} \frac{\partial T}{\partial t} + \nabla \cdot \big( \rho
	c_{p} \overrightarrow{u} \cdot T - k \nabla T \big) = Q_s - Q_{hx},
	\label{eq3}
\end{align}
%
	where
{\small
\begin{align*}
	\rho &= \text{density of molten salt} \\
	c_{p} &= \text{specific heat capacity of molten salt} \\
	T &= \text{temperature of molten salt} \\
	\overrightarrow{u} &= \text{velocity of molten salt} \\
	k &= \text{thermal conductivity of molten salt} \\
	Q_{hx} &= \text{heat sink},
\end{align*}
}
	and the source term $Q_s$ is given as:
%
\begin{align}
Q_s = \sum^G_{g=1} \epsilon_g \Sigma_g^f \phi_g. \label{eq4}
\end{align}

For steady-state solves, there are two options: the inverse power
method (\texttt{InversePowerMethod}) and the nonlinear Newton method
(\texttt{NonlinearEigen}).

### 2019-10-21 msr.tex
chp 1, 2, 3 update

\glspl{MSR} are one of six advanced reactor designs shortlisted by
the \gls{GIF} in 2001 for promising significant advances in safety,
sustainability, efficiency, and cost over existing designs in operation
today. This has attracted significant attention and resources towards
\gls{MSR} research, most noticeable by the number of start-up companies that
have emerged in recent years touting various \gls{MSR} designs. This chapter
provides a brief history of \glspl{MSR}, followed by the distinctive features
that earned the concept the label of being a Generation IV reactor. Lastly,
we present the reference specifications of the \gls{MSFR} concept studied in
this work.

\section{History}

The first \gls{MSR}, named the \gls{ARE}, dates back to the 1940s,
as part of the US Aircraft Nuclear Propulsion program; the molten salt
concept was considered due to the stability of molten salts at high
temperatures and neutron radiation. The 2.5
MW$_{\text{th}}$ reactor was built at \gls{ORNL}, where it achieved
criticality on November 1954 and generated 100 MWh over nine days. The fuel
consisted of enriched uranium in a molten salt mixture of NaF, ZrF$_4$, and
UF$_4$, and moderated by blocks of beryllium oxide. The project ultimately
never came to fruition as the development of intercontinental ballistic
missiles effectively eliminated the need for long-range nuclear-powered
bomber aircraft.

However, the successful demonstration of the \gls{ARE} spurred further
research into adapting \glspl{MSR} for civilian power generation. One of the
key findings from the
research was that better economy could be achieved from breeding $^{233}$U
from $^{232}$Th in thermal-spectrum reactors than $^{239}$Pu from $^{238}$U.
Ultimately, these efforts culminated in the \gls{MSRE}, a graphite-moderated
thermal \gls{MSR}.
Although no breeding was attempted with the \gls{MSRE}, scientists at
\gls{ORNL} obtained a wealth of experimental data and new insights from the
study of this reactor. The \gls{MSRE} had a graphite-moderated design with
LiF-BeF$_2$-ZrF$_4$-UF$_4$ fuel salt mixture, initially rated at 10
MW$_{\text{th}}$ but later restricted to 8 MW$_{\text{th}}$ due to a
miscalculation of heat transfer capabilities.

Design of the \gls{MSRE} commenced in the summer of 1960, with construction
starting in early 1962. The reactor achieved zero-power criticality in June
1965, and 30 days of continuous operation at full power in December 1966.
The reactor operated at full power for the most of the following 15 months,
during which the researchers carried out various experiments. Soon after
shutdown, the $^{235}$U fuel was replaced with $^{233}$U and in January
1969, the \gls{MSRE} became the first reactor to run on $^{233}$U fuel.

Although the \gls{MSRE} was a huge success, \gls{ORNL} failed to secure new
funding for the construction and operation of the \gls{MSBR}. The \gls{MSR}
development program lost out to the competing \gls{LMFBR} program which had
a head start and garnered wider political and technical support.
Nevertheless, from a technical perspective, two independent technology
evaluation and design studies of the \gls{MSR} had ``reported favorably on
the promise of the system".

\section{Features}

As mentioned in the introduction section, the most significant difference
between \glspl{MSR} and other reactor
concepts is the liquid fuel in \glspl{MSR}; fissile and/or fertile material
is dissolved in
high temperature, commonly eutectic mixtures of molten salts. Most \gls{MSR}
designs are circulating-fuel reactors. The primary coolant loop containing
the fuel salt transfers heat through a heat exchanger to the clean,
secondary/intermediate loop.

The flexibility of \glspl{MSR} is best illustrated by the various designs
under development today. Graphite-moderated thermal-spectrum \glspl{MSR} are
typically straight-forward \gls{LEU} burners, or $^{232}$Th/$^{233}$U
breeders, while epithermal- and fast-spectrum \glspl{MSR} have the additional
options of operating as \gls{TRU} fuel burners or $^{238}$U/$^{239}$Pu
breeders. Breeder designs can be further categorized into one- or two-fluid
designs; two-fluid designs feature blanket molten salt mixtures that contain
higher proportions of fertile material than the fuel salt mixture.

### 2020-01-27 model.tex
Edit chp 3 and equations

@@
 repository that holds the Moltres source code. The script rewrites the group
-constants into a format that Moltres can read.
+constants into a Moltres-compatible format.
 
@@
 
-Focussing our attention on Moltres, the neutronics calculations are
-implemented through the standard formulation of the time-dependent multi-group
+Focusing our attention on Moltres, the neutron flux calculations are
+performed using the standard formulations for the time-dependent multi-group
 neutron diffusion equations and \gls{DNP} concentration equations as shown in
-equations * and *:
+equations \ref{eq:neut} and \ref{eq:dnp}:
 %
 \begin{align}
-	\frac{1}{v_g} \frac{\partial \phi_g}{\partial t} - \nabla \cdot D_g \nabla
-	\phi_g + \Sigma^r_g \phi_g
-	= \sum^G_{g \neq g'} \Sigma^s_{g' \rightarrow g} \phi_{g'} + \chi^p_g
+	\frac{1}{v_g} \frac{\partial \phi_g}{\partial t} &= \nabla \cdot D_g
+	\nabla \phi_g - \Sigma^r_g \phi_g
+	\sum^G_{g \neq g'} \Sigma^s_{g' \rightarrow g} \phi_{g'} + \chi^p_g
 	\sum^G_{g'=1} (1-\beta) \nu \Sigma^f_{g'} \phi_{g'} + \chi^d_g \sum^I_i
-	\lambda_i C_i, \label{eq1}
+	\lambda_i C_i, \label{eq:neut} \\
+	\frac{\partial C_i}{\partial t} &= \beta_i \sum^G_{g'=1} \nu \Sigma^f_{g'}
+	\phi_{g'} - \lambda_i C_i - \boldsymbol{u} \cdot \nabla C_i + \nabla \cdot
+	\frac{\mu_t}{\rho Sc_t} \nabla C_i, \label{eq:dnp} \\
+	\intertext{where} v_g &= \text{average speed of neutrons in group }g 
+	\nonumber \\
+	\phi_g &= \text{flux of neutrons in group }g \nonumber \\
+	t &= \text{time} \nonumber \\
+	D_g &= \text{diffusion coefficient of neutrons in group }g \nonumber \\
+	\Sigma^r_g &= \text{macroscopic cross-section for removal of neutrons from
+	group }g \nonumber \\
+	\Sigma^s_{g' \rightarrow g} &= \text{macroscopic cross-section of
+	scattering from }g' \text{ to }g \nonumber \\
+	\chi^p_g &= \text{prompt fission spectrum neutrons in group }g \nonumber
+	\\
+	G &= \text{total number of discrete neutron groups} \nonumber \\
+	\nu &= \text{average number of neutrons produced per fission} \nonumber \\
+	\Sigma^f_{g} &= \text{macroscopic fission cross-section for neutron in
+	group }g \nonumber \\
+	\chi^d_g &= \text{delayed fission spectrum neutrons in group }g \nonumber
+	\\
+	I &= \text{total number of delayed neutron precursor groups} \nonumber \\
+	\beta &= \text{total delayed neutron fraction} \nonumber \\
+	\beta_i &= \text{delayed neutron fraction of precursor group i} \nonumber
+	\\
+	\lambda_i &= \text{average decay constant of delayed neutron precursors in
+	precursor group }i \nonumber \\
+	C_i &= \text{concentration of delayed neutron precursors in precursor
+	group }i \nonumber
 \end{align}
 %
-\begin{align}
-	\frac{\partial C_i}{\partial t} = \sum^G_{g'=1} \beta_i \nu \Sigma^f_{g'}
-	\phi_{g'} - \lambda_i C_i - \nabla \cdot u C_i. \label{eq2}
-\end{align}
-%
-where
-{\small
-%
-\begin{align*}
-	v_g &= \text{average speed of neutrons in group }g \\
-	\phi_g &= \text{flux of neutrons in group }g \\
-	t &= \text{time} \\
-	D_g &= \text{diffusion coefficient of neutrons in group }g \\
-	\Sigma^r_g &= \text{macroscopic cross-section for removal of neutrons from
-	group }g \\
-	\Sigma^s_{g' \rightarrow g} &= \text{macroscopic cross-section of
-	scattering from }g' \text{ to }g \\
-	\chi^p_g &= \text{prompt fission spectrum neutrons in group }g \\
-	G &= \text{total number of discrete neutron groups} \\
-	\nu &= \text{average number of neutrons produced per fission} \\
-	\Sigma^f_{g} &= \text{macroscopic fission cross-section for neutron in group }g \\
-	\chi^d_g &= \text{delayed fission spectrum neutrons in group }g \\
-	I &= \text{total number of delayed neutron precursor groups} \\
-	\beta &= \text{total delayed neutron fraction} \\
-	\beta_i &= \text{delayed neutron fraction of precursor group i} \\
-	\lambda_i &= \text{average decay constant of delayed neutron precursors in precursor group }i \\
-	C_i &= \text{concentration of delayed neutron precursors in precursor group }i
-\end{align*}
-}
-%
+
+The limitations of the multi-group neutron diffusion compared to
+S$_{\text{N}}$ and Monte Carlo methods are well-documented but they provide
+decent accuracy at lower computational costs. A comparison of the
+multiplication factor and reactivity coefficients in the context of the
+\gls{MSFR} is provided in the results section.
+
+The \gls{DNP} concentration equation has an additional advection term to
+account for the movement of \glspl{DNP} in the primary coolant loop. Moltres
+also accounts for the decay of \glspl{DNP} outside the active core
+region by simulating its flow in a separate 1D geometry. The length and
+velocity defined in this geometry may differ according to reactor
+specifications.
 
@@
 in \gls{MOOSE}. Once again, users have the freedom to select their
-interpolation method of choice.
+interpolation method of choice. For this work, we have selected six neutron
+energy groups according to the energy boundaries in table *, eight \gls{DNP}
+groups as defined by the JEFF-3.1.2 library, and the spline interpolation
+method.
+
+For the decay heat model, previous studies on the MSFR * have shown that
+using three decay heat precursor groups with different half-lives can model
+decay heat in the MSFR for up to 300 seconds after shutdown with a relative
+error of less than 2 \%. Thus, we implemented the decay heat model using
+equations similar to the \gls{DNP} equations, as shown in eq
+\ref{eq:decayheat}:
+%
+\begin{align}
+	\frac{\partial D^k}{\partial t} &= f_k \sum^G_{g'=1} \epsilon_{g'}
+	\Sigma^f_{g'} \phi_{g'} - \lambda^k D^k - \boldsymbol{u} \cdot \nabla D^k
+	+ \nabla \cdot \frac{\mu_t}{\rho Sc_t} \nabla D^k, \label{eq:decayheat} \\
+	\intertext{where}
+	D^k &= \text{concentration of decay heat precursors in group }k \nonumber
+	\\
+	f_k &= \text{fraction of decay heat to total power at steady state}
+	\nonumber \\
+	\epsilon_g &= \text{average fission energy} \nonumber \\
+	\lambda^k &= \text{average decay constant of decay heat precursors in
+	group }k. \nonumber
+\end{align}
 
@@
 \begin{align}
-	\rho c_{p} \frac{\partial T}{\partial t} + \nabla \cdot \big( \rho
-	c_{p} \overrightarrow{u} \cdot T - k \nabla T \big) = Q_s - Q_{hx},
+	\rho c_{p} \frac{\partial T}{\partial t} &= - \nabla \cdot \big( \rho
+	c_{p} \overrightarrow{u} \cdot T - k \nabla T \big) + Q_s - Q_{hx},
 	\label{eq3}
+	\intertext{where}
+	\rho &= \text{density of molten salt} \nonumber \\
+	c_{p} &= \text{specific heat capacity of molten salt} \nonumber \\
+	T &= \text{temperature of molten salt} \nonumber \\
+	\overrightarrow{u} &= \text{velocity of molten salt} \nonumber \\
+	k &= \text{thermal conductivity of molten salt} \nonumber \\
+	Q_{hx} &= \text{heat sink},
 \end{align}
-%
-	where
-{\small
-\begin{align*}
-	\rho &= \text{density of molten salt} \\
-	c_{p} &= \text{specific heat capacity of molten salt} \\
-	T &= \text{temperature of molten salt} \\
-	\overrightarrow{u} &= \text{velocity of molten salt} \\
-	k &= \text{thermal conductivity of molten salt} \\
-	Q_{hx} &= \text{heat sink},
-\end{align*}
-}
 	and the source term $Q_s$ is given as:
@@
 
+In the \gls{MSFR}, salt flow in the core has a Reynolds number on the order of
+10$^6$*. Thus the Navier-Stokes module in \gls{MOOSE} cannot sufficiently
+capture the turbulent flow behavior. Furthermore, the \gls{MSFR} benchmark
+geometry encourages the formation of turbulent recirculation regions in the
+fuel salt near the blanket tank regions. Due to this present limitation, only
+the flow in the active core region is simulated and it is assumed to be
+vertical flow. This eliminates the presence of recirculation zones reported by
+*, which will likely be a source of discrepancy between our results and
+theirs. Another consequence is the absence of turbulent diffusion, resulting
+in higher-than-expected temperatures along the central axis of the reactor
+where the neutron flux and heat generation is highest. This is further
+exacebated by the high Prandtl number of the molten salt; heat transfer is
+dominated by convective flow. This means that there would be very low heat
+transfer observed by diffusion due to the negligible contributions from
+heat conduction.
+
 For steady-state solves, there are two options: the inverse power

### 2020-01-27 msr.tex
Edit chp 3 and equations

@@
 higher proportions of fertile material than the fuel salt mixture.
+
+\section{Molten Salt Fast Reactor}
+
+The \gls{MSFR} is a reference fast-spectrum \gls{MSR} concept developed
+under the \gls{EVOL} project. Inspired by the \gls{MSBR}, the \gls{MSFR} is
+intended to run primarily on a closed thorium fuel cycle with
+continuous online fuel reprocessing. Several reasons motivated the omission of
+graphite moderators from the \gls{MSFR} design. Graphite is susceptible to
+long-term radiation damage and replacement is likely to be necessary during
+the operating lifetime of the reactor. The fast neutron spectrum improves
+breeding ratios due to higher fission neutron yield in $^{233}$U and lower
+parasitic absorption in non-fuel materials. Lastly, graphite has a positive
+temperature coefficient of reactivity; eliminating graphite from the design
+ensures a greater safety margin.
+
+Power output of the \gls{MSFR} is rated at 3000 MW$_{\text{th}}$ and 1500
+MW$_{\text{e}}$. The high thermal efficiency is achieved

### 2020-01-29 msr.tex
msr chapter edits

@@
 The \gls{MSFR} is a reference fast-spectrum \gls{MSR} concept developed
-under the \gls{EVOL} project. Inspired by the \gls{MSBR}, the \gls{MSFR} is
+under the \gls{EVOL} and \gls{SAMOFAR} projects. The main reactor
+specifications and schematic view are shown in Table * and Fig. *
+respectively. Inspired by the \gls{MSBR}, the \gls{MSFR} is
 intended to run primarily on a closed thorium fuel cycle with
@@
 long-term radiation damage and replacement is likely to be necessary during
-the operating lifetime of the reactor. The fast neutron spectrum improves
-breeding ratios due to higher fission neutron yield in $^{233}$U and lower
-parasitic absorption in non-fuel materials. Lastly, graphite has a positive
-temperature coefficient of reactivity; eliminating graphite from the design
-ensures a greater safety margin.
+the operating lifetime of the reactor. The relatively fast neutron spectrum
+improves breeding ratios due to higher fission neutron yield in $^{233}$U and
+lower parasitic absorption in non-fuel materials. Lastly, graphite has a
+positive temperature coefficient of reactivity; eliminating graphite from the
+design ensures a greater safety margin.
+
+In the \gls{MSFR} design, fuel salt flows upwards through a 9 m$^3$ central
+core region. At the top of the core, the flow separates into 16 smaller
+external loops, each of which passes through a heat exchanger before being
+pumped back into the bottom of the core. Other instrumentation are situated
+along the external loop for online salt reprocessing and gas sparging. The
+core is surrounded axially by nickel alloy reflectors, and radially by a
+toroidal blanket tank containing fertile salt for breeding. There is a layer
+of boron carbide behind the blanket tanks to protect the peripheral equipment
+from excessive neutron damage. In case of severe accidents, there is an
+actively cooled freeze plug at the bottom of the core that melts when
+temperatures exceed a certain threshold. The fuel salt would drain into a
+containment vessel designed to keep it subcritical. Reactivity control under
+normal operating conditions is performed by varying pump speeds to advantage
+of strong thermal feedback. Coupled with the fact that there is no excess
+reactivity reserve due to online fuel reprocessing, control rods are not
+included in the \gls{MSFR}.
+
+Although the \gls{MSFR} is primarily designed to operate on the thorium fuel
+cycle, it can support a range of start-up fuel and feed compositions. This
+versatility is particularly important for the first few \glspl{MSFR} to be
+deployed due to the lack of $^{233}$U reserves required for the initial core
+loading. In general, the fuel and blanket salts are approximately composed of
+eutectic mixtures of 77.5\% LiF - 22.5\% AcF$_4$, where AcF$_4$ represents
+actinide fluorides such as uranium, thorium, plutonium, and other \gls{TRU}
+fluorides. For an initial composition consisting of $^{232}$Th and $^{233}$U,
+the benchmark value for the amount of uranium for criticality under
+normal operating conditions is 2.515 mol\%. However, most code verification
+studies adjust the ratio of $^{232}$Th to $^{233}$U to achieve exact
+criticality at a uniform temperature of 973 K; this ensures that subsequent
+neutronics and safety analyses are not affected by the difference in
+k$_{\textt{eff}}$ values. We performed the same exercise in this paper.
 
 Power output of the \gls{MSFR} is rated at 3000 MW$_{\text{th}}$ and 1500
-MW$_{\text{e}}$. The high thermal efficiency is achieved
+MW$_{\text{e}}$. It has a high thermal efficiency due to its high operating
+temperature. \glspl{MSR} in general are not restricted by the same pressure
+constraints seen in \glspl{LWR}. The inlet and outlet temperature
+specifications of the fuel salt are 923 K and 1023 K respectively. This was
+motivated by the need for a minimum 50 K temperature buffer between the
+operating temperatures and the melting point of the salt. The \gls{MSFR} has
+heat exchangers and an intermediate coolant loop to isolate the power
+conversion system from the highly radioactive fuel salt. This also serves as
+a layer of containment between the radioactive material and the outside
+environment. The exact composition of the intermediate coolant is still under
+active study and not finalized yet.

### 2020-02-04 msr.tex
msr chapter edits, add bib entries

@@
 The first \gls{MSR}, named the \gls{ARE}, dates back to the 1940s,
-as part of the US Aircraft Nuclear Propulsion program; the molten salt
-concept was considered due to the stability of molten salts at high
-temperatures and neutron radiation. The 2.5
-MW$_{\text{th}}$ reactor was built at \gls{ORNL}, where it achieved
-criticality on November 1954 and generated 100 MWh over nine days. The fuel
+as part of the US Aircraft Nuclear Propulsion program
+\cite{rosenthal_molten-salt_1970}; the molten salt concept was considered due
+to the stability of molten salts at high temperatures and neutron radiation.
+The 2.5 MW$_{\text{th}}$ reactor was built at \gls{ORNL}, where it achieved
+criticality on November 1954 and generated 100 MWh over nine days
+\cite{rosenthal_molten-salt_1970}. The fuel
 consisted of enriched uranium in a molten salt mixture of NaF, ZrF$_4$, and
@@
 However, the successful demonstration of the \gls{ARE} spurred further
-research into adapting \glspl{MSR} for civilian power generation. One of the
-key findings from the
+research into adapting \glspl{MSR} for civilian power generation
+\cite{rosenthal_molten-salt_1970}. One of the key findings from the
 research was that better economy could be achieved from breeding $^{233}$U
-from $^{232}$Th in thermal-spectrum reactors than $^{239}$Pu from $^{238}$U.
-Ultimately, these efforts culminated in the \gls{MSRE}, a graphite-moderated
-thermal \gls{MSR}.
-Although no breeding was attempted with the \gls{MSRE}, scientists at
-\gls{ORNL} obtained a wealth of experimental data and new insights from the
-study of this reactor. The \gls{MSRE} had a graphite-moderated design with
-LiF-BeF$_2$-ZrF$_4$-UF$_4$ fuel salt mixture, initially rated at 10
-MW$_{\text{th}}$ but later restricted to 8 MW$_{\text{th}}$ due to a
-miscalculation of heat transfer capabilities.
+from $^{232}$Th in thermal-spectrum reactors than $^{239}$Pu from $^{238}$U
+\cite{macpherson_molten_1985}. Ultimately, these efforts culminated in the
+design, construction, and successful operation of the \gls{MSRE}, a graphite
+moderated thermal \gls{MSR}. Although no breeding was attempted with the
+\gls{MSRE}, scientists at \gls{ORNL} obtained a wealth of experimental data
+and new insights from the study of this reactor. The \gls{MSRE} had a
+graphite-moderated design with LiF-BeF$_2$-ZrF$_4$-UF$_4$ fuel salt mixture,
+initially rated at 10 MW$_{\text{th}}$ but later restricted to 8 M
+$_{\text{th}}$ due to a miscalculation of heat transfer capabilities
+\cite{haubenreich_experience_1970}.
 
 Design of the \gls{MSRE} commenced in the summer of 1960, with construction
-starting in early 1962. The reactor achieved zero-power criticality in June
-1965, and 30 days of continuous operation at full power in December 1966.
-The reactor operated at full power for the most of the following 15 months,
-during which the researchers carried out various experiments. Soon after
-shutdown, the $^{235}$U fuel was replaced with $^{233}$U and in January
-1969, the \gls{MSRE} became the first reactor to run on $^{233}$U fuel.
+starting in early 1962 \cite{haubenreich_experience_1970}. The reactor
+achieved zero-power criticality in June 1965, and 30 days of continuous
+operation at full power in December 1966. The reactor operated at full power
+for the most of the following 15 months, during which the researchers carried
+out various experiments. Soon after shutdown, the $^{235}$U fuel was replaced
+with $^{233}$U and in January 1969, the \gls{MSRE} became the first reactor to
+run on $^{233}$U fuel. Further experiments were run, including xenon
+stripping, fission product deposition, tritium behavior, and plutonium
+addition studies, before the \gls{MSRE} was permanently shut down to conserve
+remaining funding for other related activities \cite{macpherson_molten_1985}.
 
-Although the \gls{MSRE} was a huge success, \gls{ORNL} failed to secure new
-funding for the construction and operation of the \gls{MSBR}. The \gls{MSR}
-development program lost out to the competing \gls{LMFBR} program which had
-a head start and garnered wider political and technical support.
-Nevertheless, from a technical perspective, two independent technology
-evaluation and design studies of the \gls{MSR} had ``reported favorably on
-the promise of the system".
+Building on the \gls{MSRE}'s hugely successful run, \gls{ORNL} proposed a
+major development project for the construction and operation of the
+\gls{MSBR} \cite{macpherson_molten_1985}. However, they failed to secure
+funding on two separate occasions in 1972 and 1974; they lost out to the
+competing \gls{LMFBR} program which had a head start and wider
+political and technical support. Nevertheless, from a technical perspective,
+two independent technology evaluation and design studies of the \gls{MSR} had
+``reported favorably on the promise of the system''
+\cite{macpherson_molten_1985}.
 
@@
 under the \gls{EVOL} and \gls{SAMOFAR} projects. The main reactor
-specifications and schematic view are shown in Table * and Fig. *
-respectively. Inspired by the \gls{MSBR}, the \gls{MSFR} is
+specifications and schematic view are shown in Table \ref{table:msfr} and Fig.
+\ref{fig:msfr} respectively. Inspired by the \gls{MSBR}, the \gls{MSFR} is
 intended to run primarily on a closed thorium fuel cycle with
@@
 design ensures a greater safety margin.
+%
+\begin{figure}[htb!] 
+	\centering
+	\includegraphics[width=0.6\textwidth]{MSFR}
+	\caption{Schematic view of the MSFR concept \cite{serp_molten_2014}.}
+	\label{fig:msfr}
+\end{figure}
+%
+\begin{table}[htb!]
+	\caption{Main specifications of the \gls{MSFR} concept
+				\cite{serp_molten_2014}.}
+	\centering
+	\begin{tabular}{ l r }
+		\hline
+		Parameter & Value \\
+		\hline
+		Thermal/Electric output [MW$_{\text{th}}$/MW$_{\text{e}}$] & 3000 /
+		1500 
+		\\
+		Salt volume [m$^3$] & 18 \\
+		Salt fraction in core & 0.5 \\
+		Number of circulation loops & 16 \\
+		Nominal flow rate [kg s$^{-1}$] & 18500  \\
+		Nominal circulation time [s] & 4.0 \\
+		Inlet/outlet temperature [K] & 923 / 1023 \\
+		Blanket volume [m$^3$] & 7.3\\
+		\hline
+	\end{tabular}
+	\label{table:msfr}
+\end{table}
 
@@
 reactivity reserve due to online fuel reprocessing, control rods are not
-included in the \gls{MSFR}.
+included in the \gls{MSFR} design.
 
@@
 neutronics and safety analyses are not affected by the difference in
-k$_{\textt{eff}}$ values. We performed the same exercise in this paper.
+k$_{\text{eff}}$ values. We performed the same exercise in this paper.
 
 Power output of the \gls{MSFR} is rated at 3000 MW$_{\text{th}}$ and 1500
-MW$_{\text{e}}$. It has a high thermal efficiency due to its high operating
+MW$_{\text{e}}$. It has a high thermal efficiency due to the high operating
 temperature. \glspl{MSR} in general are not restricted by the same pressure
@@
 active study and not finalized yet.
+
+\subsection{Model Geometry}
+
+In this thesis, we study the 2-D axisymmetric representation of the
+\gls{MSFR} as shown in Fig. \ref{fig:msfrgeom}. This is the same model
+geometry used in a neutronic benchmark
+study of the \gls{MSFR} and the coupled analyses by Fiorina et al., Aufiero
+et al., and Zhang et al. Even though this is an outdated design that had been
+later optimized for better flow distribution *, the analysis results are still
+valid for code verification purposes. Future efforts will adopt the optimized
+hourglass-shaped design for a more realistic assessment of the \gls{MSFR}.
+%
+\begin{figure}[htb!] 
+	\centering
+	\includegraphics[width=0.5\textwidth]{msfr-geom}
+	\caption{2-D axisymmetric model of the \gls{MSFR} core used for the
+	simulations in Serpent and Moltres. All dimensions are in meters.
+	\cite{brovchenko_neutronic_2019}}
+	\label{fig:msfrgeom}
+\end{figure}
+
+For the 2-D axisymmetric model, the sixteen individual loops are homogenized
+into a single loop while keeping the nominal volumetric flow rate constant.
+Keeping with the model used in the previous studies, we assumed that the
+region for heat exchangers and pumps on the loop consists of 100\% fuel salt.
+However, this assumption has a negligible impact on the multiplication factor
+due to its location behind the strong boron carbide absorber layer;

### 2020-02-09 msr.tex
fixes and expand msr chapter

@@
 graphite-moderated design with LiF-BeF$_2$-ZrF$_4$-UF$_4$ fuel salt mixture,
-initially rated at 10 MW$_{\text{th}}$ but later restricted to 8 M
-$_{\text{th}}$ due to a miscalculation of heat transfer capabilities
+initially rated at 10 MW$_{\text{th}}$ but later restricted to 8
+MW$_{\text{th}}$ due to a miscalculation of heat transfer capabilities
 \cite{haubenreich_experience_1970}.
@@
 Building on the \gls{MSRE}'s hugely successful run, \gls{ORNL} proposed a
-major development project for the construction and operation of the
-\gls{MSBR} \cite{macpherson_molten_1985}. However, they failed to secure
-funding on two separate occasions in 1972 and 1974; they lost out to the
-competing \gls{LMFBR} program which had a head start and wider
-political and technical support. Nevertheless, from a technical perspective,
-two independent technology evaluation and design studies of the \gls{MSR} had
-``reported favorably on the promise of the system''
-\cite{macpherson_molten_1985}.
+new program for the construction and operation of a demonstration reactor
+based on the \gls{MSBR} concept that they had developed
+\cite{macpherson_molten_1985}. The \gls{MSBR} is a thermal-spectrum, single
+fluid reactor with fertile $^{232}$Th isotopes mixed directly into the FLiBe
+molten salt for $^{233}$U breeding \cite{gehin_liquid_2016}. Like the
+\gls{MSRE}, the \gls{MSBR} relies on continuous online reprocessing to add
+fertile material and remove fission product neutron poisons. As a breeder
+reactor, the doubling time (the minimum amount of time required to produce
+enough fissile material to start up another \gls{MSBR}) was estimated to be
+approximately 22 years. However, \gls{ORNL} failed to secure funding on two
+separate occasions in 1972 and 1974; they lost out to the competing
+\gls{LMFBR} program which had a head start and wider political and technical
+support. Nevertheless, from a technical perspective, two independent
+technology evaluation and design studies of the \gls{MSR} had ``reported
+favorably on the promise of the system'' \cite{macpherson_molten_1985}.
+
+In spite of this setback, research into \glspl{MSR} continued through the late
+1970s. In 1980, \gls{ORNL} published a report describing a new \gls{MSR}
+concept, called the \gls{DMSR} \cite{gehin_liquid_2016}. This design was
+developed in response to the fuel reprocessing restrictions introduced by
+President Ford in 1976; the \gls{DMSR} is designed to operate as a
+once-through converter system without fuel reprocessing. While it is largely
+fueled by 19.75 \% \gls{HALEU}, the initial core loading includes thorium to
+boost its conversion ratio throughout its lifetime. It has a continuous
+online feed consists of \gls{HALEU} to maintain criticality, and denatured
+$^{235}$U to keep uranium enrichement levels below nuclear non-proliferation
+policy thresholds. The design also includes a gas sparging system for removing
+gaseous \glspl{FP}, while noble metals naturally plate out onto the walls of
+the coolant loop. A significant drawback in the \gls{MSBR} design is the
+extensive neutron damage in the graphite moderator that necessitated frequent
+replacement (every four years) throughout its operational lifetime. The
+\gls{DMSR} avoids this issue by having a lower power density while maintaining
+the overall power output of 2250 MW$_{\text{th}}$. As a result, the graphite
+moderator was projected to last for the entirety of the \gls{DMSR}'s design
+lifetime.
+
+There was a concurrent program at the UK Atomic Energy Authority for the
+development of a 2500MW$_{\text{e}}$ lead-cooled
+\gls{MCFR} concept \cite{smith_assessment_1974}. It is a dual fluid system,
+with separate loops for the fuel salt and the blanket salt. The blanket is a
+1 m-wide tank surrounding the core. The relatively harder neutron spectrum
+arising from the absence of moderators and the choice of chloride over
+fluoride salt favors $^{239}$Pu breeding over the thorium cycle. Some
+experiments were performed to study molten salt chemistry but no reactor
+prototypes were built. The UK program was eventually shut down just like its
+US counterpart partly due to the successful demonstration of the Prototype
+Fast Reactor which had achieved criticality in 1974.
+
+Following a lull lasting through the late 20th century, \gls{MSR} research
+picked up pace due to renewed interest initiated by the \gls{GIF} in 2001.
+Today, there are numerous \gls{MSR} concepts under active development led by
+various national and commercial bodies. Many differences exist between
+different concepts and the only common denominator is the use of nuclear fuel
+dissolved in molten salt. Table * is a non-exhaustive list of \glspl{MSR}
+under development.
 
@@
 As mentioned in the introduction section, the most significant difference
-between \glspl{MSR} and other reactor
-concepts is the liquid fuel in \glspl{MSR}; fissile and/or fertile material
-is dissolved in
-high temperature, commonly eutectic mixtures of molten salts. Most \gls{MSR}
-designs are circulating-fuel reactors. The primary coolant loop containing
-the fuel salt transfers heat through a heat exchanger to the clean,
-secondary/intermediate loop.
+between \glspl{MSR} and other reactor concepts is the liquid fuel in
+\glspl{MSR}; fissile and/or fertile material is dissolved in high temperature,
+commonly eutectic mixtures of molten salts. Most \gls{MSR} designs are
+circulating-fuel reactors. The primary coolant loop containing the fuel salt
+transfers heat through a heat exchanger to the clean, secondary/intermediate
+loop.
 

### 2020-02-12 msr.tex
msr features section edit

@@
 transfers heat through a heat exchanger to the clean, secondary/intermediate
-loop.
+loop. The liquid fuel form allows for continuous online fuel reprocessing,
+and the removal of gaseous \glspl{FP} via a gas sparging system.
 
@@
 typically straight-forward \gls{LEU} burners, or $^{232}$Th/$^{233}$U
-breeders, while epithermal- and fast-spectrum \glspl{MSR} have the additional
-options of operating as \gls{TRU} fuel burners or $^{238}$U/$^{239}$Pu
-breeders. Breeder designs can be further categorized into one- or two-fluid
-designs; two-fluid designs feature blanket molten salt mixtures that contain
-higher proportions of fertile material than the fuel salt mixture.
+iso-breeders/breeders, while epithermal- and fast-spectrum \glspl{MSR} have
+the additional options of operating as \gls{TRU} fuel burners or
+$^{238}$U/$^{239}$Pu breeders. Breeder designs can be further categorized into
+one- or two-fluid designs; two-fluid designs feature separate blanket molten
+salt mixtures that contain higher proportions of fertile material than the
+fuel salt mixture.
+
+\subsection{Safety}
+
+\glspl{MSR} hold a number of safety advantages over current \glspl{LWR}. The
+most immediate and significant advantage is the strong negative temperature
+reactivity feedback of the fuel salt. This is due to greater
+temperature-induced density changes in liquid fuel than solid fuel. Combined
+with the Doppler broadening of resonance capture cross sections present in
+both fuel forms, we expect to observe a smaller temperature increase following
+an unprotected reactivity insertion. The overall temperature reactivity
+coefficient varies widely between different \gls{MSR} designs due to other
+structures, such as moderators and reflectors, present in the core. In
+particular, graphite moderators tend to have slightly positive temperature
+reactivity coefficients. This was notably observed in the \gls{MSBR} concept,
+but the total temperature reactivity coefficient was still relatively large
+and negative.
+
+Continuous online fuel reprocessing allows operators to maintain low excess
+reactivity inventories in the core as additional fuel can be added on an ad
+hoc basis. This factor, in addition to the strong negative temperature
+reactivity coefficient, diminishes the likelihood and severity of unprotected
+criticality accidents in \glspl{MSR}.
 

### 2020-02-13 introduction.tex
Features: safety

@@
 The former employed the commercial
-multiphysics modeling software COMSOL \cite{comsol_comsol_nodate}, while the
+multiphysics modeling software COMSOL \cite{comsol_ab_comsol_2018}, while the
 latter coupled two in-house codes DALTON and HEAT for neutron transport and

### 2020-02-13 msr.tex
Features: safety

@@
 
-\glspl{MSR} hold a number of safety advantages over current \glspl{LWR}. The
-most immediate and significant advantage is the strong negative temperature
-reactivity feedback of the fuel salt. This is due to greater
-temperature-induced density changes in liquid fuel than solid fuel. Combined
-with the Doppler broadening of resonance capture cross sections present in
-both fuel forms, we expect to observe a smaller temperature increase following
-an unprotected reactivity insertion. The overall temperature reactivity
-coefficient varies widely between different \gls{MSR} designs due to other
-structures, such as moderators and reflectors, present in the core. In
-particular, graphite moderators tend to have slightly positive temperature
-reactivity coefficients. This was notably observed in the \gls{MSBR} concept,
-but the total temperature reactivity coefficient was still relatively large
-and negative.
+\glspl{MSR} are generally deemed to be safer than \glspl{LWR} due to their
+reliance on natural physical phenomena for passive safety.
+The most immediate and significant safety characteristic is the strong
+negative temperature reactivity feedback of the fuel salt. This is due to
+greater temperature-induced volumetric expansion in liquid fuel than solid
+fuel. Combined with the Doppler broadening of resonance capture cross sections
+present in both fuel forms, we expect to observe a smaller temperature
+increase following an unprotected reactivity insertion. The overall
+temperature reactivity coefficient varies widely between different \gls{MSR}
+designs due to other structures, such as moderators and reflectors, present in
+the core. In particular, graphite moderators tend to have slightly positive
+temperature reactivity coefficients. This was notably observed in the
+\gls{MSBR} concept, but the total temperature reactivity coefficient was still
+relatively large and negative \cite{rykhlevskii_modeling_2019}. This
+phenomenon provides a large degree of control and stability as it is always
+present in an \gls{MSR} regardless whether the operating conditions.
 
@@
 reactivity inventories in the core as additional fuel can be added on an ad
-hoc basis. This factor, in addition to the strong negative temperature
+hoc basis. Reprocessing and gas sparging systems help reduce fissile
+requirements by continuously removing neutron poisons. These factors, in
+addition to the strong negative temperature
 reactivity coefficient, diminishes the likelihood and severity of unprotected
-criticality accidents in \glspl{MSR}.
+criticality accidents in \glspl{MSR} \cite{elsheikh_safety_2013}. In the
+unlikely situation where an \gls{MSR} encounters a severe runaway reaction,
+we can rely on another passive safety feature, freeze plugs. A freeze plug is
+a plug filled with solidified salt at the bottom of the core
+\cite{elsheikh_safety_2013}. It is actively
+cooled by fans or other cooling systems to keep it just below the freezing
+point of the salt. When temperatures in the core exceed a certain threshold
+during a dangerous transient, the freeze plug melts and the molten salt in the
+core drains into a containment tank designed to keep the salt in a
+sub-critical configuration. This is especially easy to achieve for
+thermal-spectrum \glspl{MSR} as the absence of moderators in the containment
+tank would automatically bring the multiplication factor down below unity
+\cite{elsheikh_safety_2013}. \glspl{MSR} also typically have large
+margin-to-boiling under nominal operating conditions so that fuel salt boiling
+does not occur. Furthermore, the reactor vessel is consequently subject to
+much lower stresses as \glspl{MSR} operate at near-atmospheric pressure
+levels. Thus, the probability of pipe ruptures due to high pressure is low.
+
+\glspl{MSR} may be more resilient to pump failure accidents as natural
+circulation could passively sustain enough heat transfer through the heat
+exchanger to remove any remaining fission and decay heat after the failure. If
+natural circulation proves insufficient, the aforementioned freeze plug can
+drain the salt out of the core. Decay heat in \glspl{MSR} with online
+reprocessing isexpected to be lower than that in \glspl{LWR} due to the
+continuous removal of \glspl{FP}. For example, the decay heat in an \gls{MSFR}
+after reaching equilibrium salt composition is approximately 3.5 \% compared
+to 6 \% in \glspl{LWR} \cite{brovchenko_design-related_2013}.
+
+However, there are also some safety disadvantages associated with \glspl{MSR}.
+Firstly, \glspl{MSR} have smaller fractions of \gls{DNP} in the active core
+region as some of them decay in the external loop regions. This complicates
+reactor control and may result in faster transients due to the decrease in
+average neutron lifetime. Secondly, structural materials in the core must be
+resistant to salt corrosion under high temperatures and neutron irradiation
+over the lifetime of the reactor. To a lesser extent, pipes should also be
+corrosion resistant to prevent pipe ruptures. Lastly, fuel salt overcooling is
+a potentially dangerous accident scenario as the strong temperature
+coefficient can effectively cause a large reactivity insertion. Overcooling
+may also cause salt to freeze and restrict flow, thereby causing a complete
+loss of flow accident.
+
+\subsection{Sustainability}
 

### 2020-02-21 acros.tex
Edit chp 3 and equations
fixes and expand msr chapter
edits to msr, msfr, model chapters

@@
 \newacronym{CNERG}{CNERG}{Computational Nuclear Engineering Research Group}
+\newacronym{CNRS}{CNRS}{Centre National de la Recherche Scientifique, the French National Centre for Scientific Research}
 \newacronym{COMSOL}{COMSOL}{COMmon SOLution}
@@
 \newacronym{DOE}{DOE}{Department of Energy}
+\newacronym{DMSR}{DMSR}{Denatured Molten Salt Reactor}
 \newacronym{DRACS}{DRACS}{Direct Reactor Auxiliary Cooling System}
@@
 \newacronym{FLiBe}{FLiBe}{Fluoride-Lithium-Beryllium}
+\newacronym{FP}{FP}{fission product}
 \newacronym{GDSE}{GDSE}{Generic Disposal System Environment}
@@
 \newacronym{GUI}{GUI}{graphical user interface}
+\newacronym{HALEU}{HALEU}{high-assay low-enriched uranium}
 \newacronym{HLW}{HLW}{high level waste}
@@
 \newacronym{MA}{MA}{minor actinide}
+\newacronym{MCFR}{MCFR}{Molten Chloride Fast Reactor}
 \newacronym{MCNP}{MCNP}{Monte Carlo N-Particle code}
@@
 \newacronym{MPI}{MPI}{Message Passing Interface}
+\newacronym{MPM}{MPM}{Multi-Physics Modelling}
 \newacronym{MRPP}{MRPP}{Multiregion Processing Plant}
@@
 \newacronym{QA}{QA}{quality assurance}
+\newacronym{RANS}{RANS}{Reynolds-averaged Navier-Stokes}
 \newacronym{RDD}{RD\&D}{Research Development and Demonstration}

### 2020-02-21 introduction.tex
edits to msr, msfr, model chapters

--- 2019-07-24 introduction.tex
@@
 Since electricity generation from burning fossil fuels represents the
-largest source of CO$_2$ emissions (38\% in 2018 \cite{noauthor_global_2018}), replacing them with
-low-carbon
+largest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019}),
+replacing them with low-carbon
 alternatives would effectively curb a large fraction of these emissions. Out
@@
 the risk of a reactor power excursion. Many designs can also incorporate the
-thorium fuel cycle for improved sustainability arising from the use of abundant
-natural thorium resources and reduced transuranic waste. The resultant reduced
-radiotoxicity from transuranic waste also reduces costs associated with
-long-term nuclear waste storage.
-In addition, the ability to operate at near atmospheric pressures eliminates
-the need for a thick pressure vessel and drives down construction
-costs, while online fuel reprocessing reduces reactor downtime during reactor
-operation.
+thorium fuel cycle for improved sustainability arising from the use of
+abundant natural thorium resources and reduced transuranic waste. The
+resultant reduced radiotoxicity from transuranic waste also reduces costs
+associated with long-term nuclear waste storage. In addition, the ability to
+operate at near atmospheric pressures eliminates the need for a thick pressure
+vessel and drives down construction costs, while online fuel reprocessing
+reduces reactor downtime during reactor operation.
+
+However, the liquid fuel form also brought about novel computational
+challenges in simulating the transient behavior of \glspl{MSR}; the
+neutronics and thermal-hydraulics are more tightly coupled due to the strong
+temperature reactivity coefficient, and the additional advection coupling
+terms. Furthermore, we have to account for the movement of \glspl{DNP} as
+they are now generated directly within the primary coolant loop. Therefore,
+the choice of coupling methods for each set of physics requires careful
+consideration. Existing reactor system-level codes and modeling approaches
+for conventional \glspl{LWR} analysis contain simplifying assumptions in
+multiphysics coupling and other areas that render the techniques unsuitable
+for simulating \gls{MSR}. Thus, making minor modifications to these codes
+without changing the underlying algorithms is not the best approach for
+\gls{MSR} safety analysis.
+
+In the past two decades, researchers have developed several
+new tools for simulating steady-state and transient behavior in \glspl{MSR}.
+Many of the earlier efforts featured simplifications in simulating
+thermal-hydraulics by solving 1-D Navier-Stokes equations or using
+predetermined uniform velocity fields \cite{krepel_dyn3d-msr_2007}
+\cite{kophazi_development_2009}. In more recent years, there has been
+significant progress towards fully coupled, spatial codes that feature
+2-D axisymmetric or full 3-D models. In 2011, Cammi et al.
+\cite{cammi_multi-physics_2011} performed an ``\gls{MPM}'' analysis
+of a simplified 2-D axisymmetric model of a single MSBR fuel channel using
+the commercial finite-element analysis software COMSOL Multiphysics. The
+physics were implemented through the two-group neutron diffusion
+equations, and the \gls{RANS} standard $k-\epsilon$ turbulence model, for the
+neutronics and thermal-hydraulics respectively. The
+authors emphasized the need for proper full coupling of the multiphysics, and
+presented both steady-state and transient results in various
+scenarios such as reactivity insertions, changes in pumping rate, and the
+presence of periodic perturbations. This approach featured again in a later
+paper by Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D
+axisymmetric model of the
+\gls{MSFR}. The authors presented results from the Politecnico di Milano
+COMSOL-based approach, and another approach by researchers from Delft
+University of Technology, where they coupled their in-house neutronics and
+thermal-hydraulics codes, DALTON and HEAT respectively. With multi-group
+neutron diffusion and \gls{RANS} formulations on ultra-fine meshes, both
+models showed good agreement in the steady-state neutron flux, temperature,
+and \gls{DNP} distributions, and in the power responses following various
+accident transient initiations. Aufiero et al. \cite{aufiero_development_2014}
+concurrently developed
+a full-core 3-D model of the \gls{MSFR} on OpenFOAM, albeit with one-group
+neutron diffusion to reduce computational load. With the 3-D model, the
+authors could simulate the asymmetric reactor response to the failure of a
+single pump in the sixteen-pump \gls{MSFR} configuration. The authors also
+provided some quantitative data supporting the use of implicit coupling over
+explicit coupling to obtain accurate solutions of the transient cases.
+Recognizing the huge computational burden required for full 3-D simulations, 
+later authors came up with innovative ways to alleviate this issue such as
+selective geometrical reduced order modeling for various components of a
+reactor based on the importance of the physical phenomena being simulated
+\cite{zanetti_geometric_2015}, or using a novel, efficient method for
+neutronics calculations \cite{laureau_transient_2017}.
+
+The preceding discussion highlights the challenges faced in MSR simulation,
+and the need for a highly efficient simulation tool that incorporates
+implicitly coupled multiphysics with good computational scalability over
+multiple processing units. This paper presents the open source MSR simulation
+tool, Moltres, as a strong contender to overcoming these challenges for
+simulating the MSFR. Moltres is an application code
+built in the Multiphysics Object-Oriented Simulation Environment (MOOSE)
+parallel finite element framework. Similar to COMSOL and OpenFOAM, it solves
+the deterministic multi-group neutron diffusion and thermal-hydraulics
+\glspl{PDE} simultaneously on the same mesh. It supports up to 3-D meshes and
+scales well for a large number of processors.
+
+\section{Objectives}
+
+In this thesis, we demonstrate the coupled multiphysics modeling capabilities
+of Moltres through the reference configuration of the MSFR design. Moltres
+employs fully implicit coupling between the neutronics and thermal-hydraulics
+governing equations, to fully account for the tight coupling expected in MSRs
+due to the movement of fuel in the salt. We implemented flow using the
+Navier-Stokes equations for an accurate representation of the flow profile as
+opposed to imposing a predetermined velocity field.
+
+The objectives of this thesis are: 1) study the steady-state operating
+conditions with the reference \gls{MSFR} model, 2) study the transient
+behavior of the \gls{MSFR} under several unprotected accident scenarios, and
+identify general characteristic behaviors in MSRs and fast spectrum reactors.
+The simulation results are verified against the results by Fiorina et al.
+\cite{fiorina_modelling_2014} and Aufiero et al.
+\cite{aufiero_development_2014}.
+
+** insert thesis outline here **

### 2020-02-21 model.tex
edits to msr, msfr, model chapters

@@
 
-Serpent 2 is a continuous-energy Monte Carlo reactor physics code under
+Serpent 2 \cite{leppanen_serpent_2014} is a continuous-energy Monte Carlo
+reactor physics code under
 active development at the VTT Technical Research Centre of Finland. It was
@@
 both MPI and OpenMP parallel programming APIs, and highly validated and
-verified against experimental data and well-established codes.
+verified against experimental data and other well-established codes.
 
@@
 In the context of this project, we used Serpent with the JEFF-3.1.2 nuclear
-data library* to generate group constants
+data library \cite{oecd/nea_jeff-3.1.2_2014} to generate group constants
 for Moltres. The reactor geometry is based on the axisymmetric \gls{MSFR}
@@
 the \gls{MSFR}. The relevant group constant data are collapsed into six
-neutron energy groups, provided in table *, and calculated for temperature
-values of 900 K to 1200 K at 50 K intervals.
+neutron energy groups, and calculated for
+temperature values of 900 K to 1200 K at 50 K intervals.
 
@@
 performance computing clusters.
-
-Both \gls{MOOSE} and Moltres are open source codes licensed under Lesser GNU
-Public License (LPGL) terms. In the case of Moltres, this was done in support
-of open and transparent research practices as well as to leverage on the ease
-of access for peer review and encourage continued development of the code.
-Moltres is hosted on a Github repository of the same name. Github provides
-version control and continuous integration testing, which ensure existing
-physics capabilities are protected from potential bugs inadvertently
-introduced when developing new physics.
 
@@
 interpolation method of choice. For this work, we have selected six neutron
-energy groups according to the energy boundaries in table *, eight \gls{DNP}
-groups as defined by the JEFF-3.1.2 library, and the spline interpolation
-method.
+energy groups according to the energy boundaries in table \ref{table:bound},
+eight \gls{DNP} groups as defined by the JEFF-3.1.2 library, and the spline
+interpolation method.
+%
+\begin{table}[htb!]
+	\centering
+	\caption{Neutron energy group upper bounds used in Serpent.}
+	\begin{tabular}{c S}
+		\hline
+		{Group number} & {Upper bound [MeV]}\\
+		\hline
+		1 & 20\\
+		2 & 2.2313\\
+		3 & 0.4979\\
+		4 & 2.47875$\times 10^{-2}$\\
+		5 & 5.5308$\times 10^{-3}$\\
+		6 & 7.485$\times 10^{-4}$\\
+		\hline
+	\end{tabular}
+	\label{table:bound}
+\end{table}
 

### 2020-02-21 msfr.tex
edits to msr, msfr, model chapters

The \gls{MSFR} is a reference fast-spectrum \gls{MSR} concept developed
under the \gls{EVOL} and \gls{SAMOFAR} projects. The main reactor
specifications and schematic view are shown in Table \ref{table:msfr} and Fig.
\ref{fig:msfr} respectively. Developed from the \gls{MSBR}, the \gls{MSFR} is
intended to run primarily on a closed thorium fuel cycle with
continuous online fuel reprocessing. Several reasons motivated the omission of
graphite moderators from the \gls{MSFR} design. Graphite is susceptible to
long-term radiation damage and replacement is likely to be necessary during
the operating lifetime of the reactor. Graphite also has a positive
temperature coefficient of reactivity; eliminating graphite from the design
ensures a greater safety margin \cite{mathieu_thorium_2006}. While negative
temperature coefficients are attainable with very thermalized spectra,
breeding ratios deteriorated significantly due to parasitic absorption in the
large volume of graphite needed for thermalization
\cite{mathieu_thorium_2006}.
%
\begin{figure}[htb!] 
	\centering
	\includegraphics[width=0.6\textwidth]{MSFR}
	\caption{Schematic view of the MSFR concept \cite{serp_molten_2014}.}
	\label{fig:msfr}
\end{figure}
%
\begin{table}[htb!]
	\caption{Main specifications of the \gls{MSFR} concept
				\cite{serp_molten_2014}.}
	\centering
	\begin{tabular}{ l r }
		\hline
		Parameter & Value \\
		\hline
		Thermal/Electric output [MW$_{\text{th}}$/MW$_{\text{e}}$] & 3000 /
		1500 
		\\
		Salt volume [m$^3$] & 18 \\
		Salt fraction in core & 0.5 \\
		Number of circulation loops & 16 \\
		Nominal flow rate [kg s$^{-1}$] & 18500  \\
		Nominal circulation time [s] & 4.0 \\
		Inlet/outlet temperature [K] & 923 / 1023 \\
		Blanket volume [m$^3$] & 7.3\\
		\hline
	\end{tabular}
	\label{table:msfr}
\end{table}

In the \gls{MSFR} design, fuel salt flows upwards through a 9 m$^3$ central
core region. At the top of the core, the flow separates into 16 smaller
external loops, each of which passes through a heat exchanger before being
pumped back into the bottom of the core. Other instrumentation are situated
along the external loop for online salt reprocessing and gas sparging. The
core is surrounded axially by nickel alloy reflectors, and radially by a
toroidal blanket tank containing fertile salt for breeding. There is a layer
of boron carbide behind the blanket tanks to protect the peripheral equipment
from excessive neutron damage. In case of severe accidents, there is an
actively cooled freeze plug at the bottom of the core that melts when
temperatures exceed a certain threshold. The fuel salt would drain into a
containment vessel designed to keep it subcritical. Reactivity control under
normal operating conditions is performed by varying pump speeds to advantage
of strong thermal feedback. Coupled with the fact that there is no excess
reactivity reserve due to online fuel reprocessing, there are no control rods
in the \gls{MSFR} design.

Although the \gls{MSFR} is primarily designed to operate on the thorium fuel
cycle, it can support a range of start-up fuel and feed compositions. This
versatility is particularly important for the first few \glspl{MSFR} to be
deployed due to the lack of $^{233}$U reserves required for the initial core
loading. In general, the fuel and blanket salts are approximately composed of
eutectic mixtures of 77.5\% LiF - 22.5\% AcF$_4$, where AcF$_4$ represents
actinide fluorides such as uranium, thorium, plutonium, and other \gls{TRU}
fluorides. For an initial composition consisting of $^{232}$Th and $^{233}$U,
the benchmark value for the amount of uranium for criticality under
normal operating conditions is 2.515 mol\%. However, most code verification
studies adjust the ratio of $^{232}$Th to $^{233}$U to achieve exact
criticality at a uniform temperature of 973 K; this ensures that subsequent
neutronics and safety analyses are not affected by the difference in
k$_{\text{eff}}$ values. We performed the same exercise in this thesis.

Power output of the \gls{MSFR} is rated at 3000 MW$_{\text{th}}$ and 1500
MW$_{\text{e}}$. It has a high thermal efficiency due to the high operating
temperature. \glspl{MSR} in general are not restricted by the same pressure
constraints seen in \glspl{LWR}. The inlet and outlet temperature
specifications of the fuel salt are 923 K and 1023 K respectively. This was
motivated by the need for a minimum 50 K temperature buffer between the
operating temperatures and the melting point of the salt. The \gls{MSFR} has
heat exchangers and an intermediate coolant loop to isolate the power
conversion system from the highly radioactive fuel salt. This also serves as
a layer of containment between the radioactive material and the outside
environment. The exact composition of the intermediate coolant is still under
active study and not finalized yet.

\section{Model Geometry}

For this work, we used a model similar to the reference square-cylindrical
\gls{MSFR} design to benchmark our results against results published by
Fiorina et al. and Aufiero et al. The reference design is a 2-D axisymmetric
model with the sixteen individual external loops homogenized into a single
outer loop as shown in Fig. \ref{fig:msfrgeom}. For the multi-group
cross sections and group constants calculations in Serpent, we extended this
2-D axisymmetric model into a 3-D model by a simple full rotation about the
central axis. The material definitions are the same as those specified in the
reference \gls{MSFR} model. Accordingly, the pump and heat exchanger regions
are assumed to be composed of 100\% fuel salt. While this may not be entirely
accurate, the exact details of the pump and heat exchanger systems are still
be researched, and this external loop region is presumed to be of little
neutronic importance due to its position behind the strong boron carbide
neutron absorber layer.
%
\begin{figure}[htb!] 
	\centering
	\includegraphics[width=0.5\textwidth]{msfr-geom}
	\caption{2-D axisymmetric model of the \gls{MSFR} core used for the
	simulations in Serpent and Moltres. All dimensions are in meters.
	\cite{brovchenko_neutronic_2019}}
	\label{fig:msfrgeom}
\end{figure}

There are two key differences between our \gls{MSFR} model geometry in
Moltres and the Polimi/TUDelft models. The first difference is a relative
minor change to the mesh by the exclusion of the 2 cm thick structural
material around the blanket tank that separates the fuel and blanket salts.
We removed this feature in our finite element mesh as we had difficulty
meshing this layer that is relatively much thinner than the rest of the model.
Any impact on the
neutron flux is expected to be minimal. Furthermore, we solved for the
temperature distribution only in the primary loop and applied homogeneous
Neumann boundary conditions for temperature on the core walls, as was done in
the Polimi/TUDelft models. Therefore, we believe the overall impact on the
results is negligible.

The second difference pertains to the modeling of the external loop. In its
current implementation, Moltres lacks pump- and heat exchanger-equivalents in
the code. Thus, the external loop, beyond the central region of the primary
loop where most of the fissions take place, is modeled as a 1-D pipe with a
point heat sink to represent the heat exchanger. Instead of pumps, the flow is
driven by Dirichlet boundary conditions for velocity on the inlet and outlet
boundaries of the central regions of the primary loop, and the 1-D external
pipe geometry. All flow-dependent variables such as temperature, \glspl{DNP},
and decay heat precursors are fully conserved as they loop around between the
two regions. As a result, this approach shares some similarities with the
geometric multiscale modeling approach by Zanetti et al.
\cite{zanetti_geometric_2015}. Future models could
create a better representation of the primary loop by implementing a whole
continuous loop with pressure increases and drops corresponding to the pumps
and heat exchangers.

\section{Material Specifications}

This section discusses the material specifications of the various reactor
components in the \gls{MSFR}.

\subsection{Molten Salt}

As mentioned before, the fuel and blanket salts are comprised of a 77.5\% LiF
- 22.5\% AcF$_4$ mixture. This is the reference salt composition of the
\gls{MSFR} at start-up. Typically, researchers working with the \gls{MSFR}
model calculate the exact actinide composition by varying the $^{232}$Th to
$^{233}$U ratio to obtain a $k_{\text{eff}}$ value of 1 at a uniform
temperature of 973 K. Thus, the exact actinide compositions vary depending on
the nuclear data library and neutron transport code. The exact composition
used for this work can be found in the Results chapter. We assume that the
effects of these composition variations on the reference physical properties
of the fuel and blanket salt are negligible. Table \ref{table:prip} shows
relevant physical properties of the fuel and blanket salts.
%
\begin{table}[htb!]
\small
\centering
\caption{Properties of the fuel and blanket salts LiF-AcF$_4$.}
\begin{tabular}{l l r r}
\toprule
Property & Expression & {Value at 973 K} & Validity Range\\
\midrule
Melting temperature [K] & 841 & {N/A} & 1 bar \\
Boiling temperature [K] & 1874 & {N/A} & 1 bar \\
Density, $\rho$ [kg m$^{-3}$] & $4094-0.882 \cdot (T-1008)$ & $4125$ & 893-1123 K \\
Dynamic viscosity, $\mu$ [Pa s] & $\rho \cdot 5.55 \times 10^{-8} \cdot e^{3689/T}$ & $1.015$ & 898-1121 K \\
Thermal conductivity, $k$ [W m$^{-1}$ K$^{-1}$] & $0.928+8.397 \times 10^{-5} \cdot T$ & $0.01010$ & 891-1020 K \\
Specific heat, $c_p$ [J kg$^{-1}$ K$^{-1}$] & $-1111+2.78\cdot T$ & $1555$* & 867-907 K \\
\bottomrule
\end{tabular}
\label{table:prip}
\end{table}

\subsection{Structural Materials}

The reflectors on the periphery of the reactor core and the blanket tank are
made of a nickel-based alloy provided from reference specifications
\cite{delpech_reactor_2009}. Table \ref{table:refl} details the compositions
of the nickel-based alloy and boron carbide absorber. The alloy has a density
of 10 g cm$^{-3}$. The material composition of the reflectors may be subject
to minor changes but it is not a major concern as the reflectors are not
situated in regions of high neutron flux. The \gls{MSFR} also includes a 20 cm
layer of boron carbide (B$_4$C) to protect the heat exchangers and pumps from
neutron irradiation. The reference specifications indicate that natural boron
is used, which is composed of 19.8 \% B$^{10}$ and 80.2 \% B$^{11}$, with an
overall density of 2.52 g cm$^{-3}$. 
%
\begin{table}[htb!]
\small
\centering
\caption{Composition (mol \%) of the nickel-based alloy used in the simulation
of the \gls{MSFR} in this work.}
\begin{tabular}{l l l l l l l l l l l l l}
\toprule
Ni & W & Cr & Mo & Fe & Ti & C & Mn & Si & Al & B & P & S \\
\midrule
79.432 & 9.976 & 8.014 & 0.736 & 0.632 & 0.295 & 0.294 & 0.257 & 0.252 & 0.052 & 0.033 & 0.023 & 0.004 \\
\bottomrule
\end{tabular}
\label{table:refl}
\end{table}

### 2020-02-21 msr.tex
edits to msr, msfr, model chapters

@@
 a plug filled with solidified salt at the bottom of the core
-\cite{elsheikh_safety_2013}. It is actively
+\cite{aji_experimental_2020}. It is actively
 cooled by fans or other cooling systems to keep it just below the freezing
@@
 
-\subsection{Sustainability}
+\subsection{Other factors}
 
-\section{Molten Salt Fast Reactor}
+Other factors for assessing a reactor design include sustainability,
+economics, waste management, and non-proliferation.
 
-The \gls{MSFR} is a reference fast-spectrum \gls{MSR} concept developed
-under the \gls{EVOL} and \gls{SAMOFAR} projects. The main reactor
-specifications and schematic view are shown in Table \ref{table:msfr} and Fig.
-\ref{fig:msfr} respectively. Inspired by the \gls{MSBR}, the \gls{MSFR} is
-intended to run primarily on a closed thorium fuel cycle with
-continuous online fuel reprocessing. Several reasons motivated the omission of
-graphite moderators from the \gls{MSFR} design. Graphite is susceptible to
-long-term radiation damage and replacement is likely to be necessary during
-the operating lifetime of the reactor. The relatively fast neutron spectrum
-improves breeding ratios due to higher fission neutron yield in $^{233}$U and
-lower parasitic absorption in non-fuel materials. Lastly, graphite has a
-positive temperature coefficient of reactivity; eliminating graphite from the
-design ensures a greater safety margin.
-%
-\begin{figure}[htb!] 
-	\centering
-	\includegraphics[width=0.6\textwidth]{MSFR}
-	\caption{Schematic view of the MSFR concept \cite{serp_molten_2014}.}
-	\label{fig:msfr}
-\end{figure}
-%
-\begin{table}[htb!]
-	\caption{Main specifications of the \gls{MSFR} concept
-				\cite{serp_molten_2014}.}
-	\centering
-	\begin{tabular}{ l r }
-		\hline
-		Parameter & Value \\
-		\hline
-		Thermal/Electric output [MW$_{\text{th}}$/MW$_{\text{e}}$] & 3000 /
-		1500 
-		\\
-		Salt volume [m$^3$] & 18 \\
-		Salt fraction in core & 0.5 \\
-		Number of circulation loops & 16 \\
-		Nominal flow rate [kg s$^{-1}$] & 18500  \\
-		Nominal circulation time [s] & 4.0 \\
-		Inlet/outlet temperature [K] & 923 / 1023 \\
-		Blanket volume [m$^3$] & 7.3\\
-		\hline
-	\end{tabular}
-	\label{table:msfr}
-\end{table}
+Breeder \gls{MSR} designs easily score well in the sustainability category.
+\glspl{MSR} generally have good neutron economy due to little structural
+material in the core and continuous online removal of neutron poisons. Noble
+metal neutron poisons also naturally plate out onto the inner walls of the
+loop which are generally regions of lower neutronic importance. Both
+$^{232}$Th/$^{233}$U and $^{238}$U/$^{239}$Pu fuel cycles are viable
+candidates for breeding in \gls{MSR}, with the former being more suited for
+thermal reactors and the latter being more suited for fast reactors. Another
+benefit of this is that some \glspl{MSR} can potentially start with an initial
+core loading of $^{239}$Pu with a fast spectrum before transitioning towards
+$^{233}$U fuel with a $^{232}$Th feed. This measure effectively circumvents
+the issue of having insufficient $^{233}$U inventory to start up the reactor.
 
-In the \gls{MSFR} design, fuel salt flows upwards through a 9 m$^3$ central
-core region. At the top of the core, the flow separates into 16 smaller
-external loops, each of which passes through a heat exchanger before being
-pumped back into the bottom of the core. Other instrumentation are situated
-along the external loop for online salt reprocessing and gas sparging. The
-core is surrounded axially by nickel alloy reflectors, and radially by a
-toroidal blanket tank containing fertile salt for breeding. There is a layer
-of boron carbide behind the blanket tanks to protect the peripheral equipment
-from excessive neutron damage. In case of severe accidents, there is an
-actively cooled freeze plug at the bottom of the core that melts when
-temperatures exceed a certain threshold. The fuel salt would drain into a
-containment vessel designed to keep it subcritical. Reactivity control under
-normal operating conditions is performed by varying pump speeds to advantage
-of strong thermal feedback. Coupled with the fact that there is no excess
-reactivity reserve due to online fuel reprocessing, control rods are not
-included in the \gls{MSFR} design.
+The $^{232}$Th/$^{233}$U fuel cycle produces significantly less \gls{TRU}
+waste than the other cycles due to the smaller atomic masses of $^{232}$Th and
+$^{233}$U. This reduces the overall radiotoxicity and long-term decay heat
+associated with long-lived plutonium and \gls{MA} isotopes. However, there are
+other radionuclides such as $^{231}$Pa, $^{229}$Th, and $^{230}$U that may
+pose long-term radiological concerns. Nevertheless, this feature complements
+\gls{TRU} burning in fast spectrum \glspl{MSR} to reduce overall levels
+of \gls{TRU} waste going into long-term storage in nuclear waste repositories.
 
-Although the \gls{MSFR} is primarily designed to operate on the thorium fuel
-cycle, it can support a range of start-up fuel and feed compositions. This
-versatility is particularly important for the first few \glspl{MSFR} to be
-deployed due to the lack of $^{233}$U reserves required for the initial core
-loading. In general, the fuel and blanket salts are approximately composed of
-eutectic mixtures of 77.5\% LiF - 22.5\% AcF$_4$, where AcF$_4$ represents
-actinide fluorides such as uranium, thorium, plutonium, and other \gls{TRU}
-fluorides. For an initial composition consisting of $^{232}$Th and $^{233}$U,
-the benchmark value for the amount of uranium for criticality under
-normal operating conditions is 2.515 mol\%. However, most code verification
-studies adjust the ratio of $^{232}$Th to $^{233}$U to achieve exact
-criticality at a uniform temperature of 973 K; this ensures that subsequent
-neutronics and safety analyses are not affected by the difference in
-k$_{\text{eff}}$ values. We performed the same exercise in this paper.
+There are some nuclear non-proliferation concerns with the thorium fuel cycle
+in \glspl{MSR}.
+The main concern involves the separation of the intermediate $^{233}$Pa
+isotope from the fuel salt. $^{233}$Pa decays into $^{233}$U with a half-life
+of approximately 27 days, and $^{233}$U is at least as potent as $^{235}$U for
+nuclear weapons production. The highly radioactive $^{232}$U provides some
+level of proliferation resistance but it can be sidestepped by
+separating away its $^{232}$Pa precursor, which has a half-life of only 1.31
+days. Strong regulatory framework and close monitoring of \glspl{MSR} would be
+essential to ensure non-proliferation.
 
-Power output of the \gls{MSFR} is rated at 3000 MW$_{\text{th}}$ and 1500
-MW$_{\text{e}}$. It has a high thermal efficiency due to the high operating
-temperature. \glspl{MSR} in general are not restricted by the same pressure
-constraints seen in \glspl{LWR}. The inlet and outlet temperature
-specifications of the fuel salt are 923 K and 1023 K respectively. This was
-motivated by the need for a minimum 50 K temperature buffer between the
-operating temperatures and the melting point of the salt. The \gls{MSFR} has
-heat exchangers and an intermediate coolant loop to isolate the power
-conversion system from the highly radioactive fuel salt. This also serves as
-a layer of containment between the radioactive material and the outside
-environment. The exact composition of the intermediate coolant is still under
-active study and not finalized yet.
+The economic prospects are uncertain at this stage of \gls{MSR} development
+but we can qualitatively list the potential impacts. Capital costs are
+expected to be greater due to the need for increased corrosion resistance and
+the supporting reprocessing facilities. This is in spite of the reduced
+pressure constraints on the reactor pressure vessel for \glspl{MSR} resulting
+from the significantly lower operating pressure. There are potential cost
+savings from eliminating fuel pellet and fuel assembly fabrication. Higher
+fuel utilization is achieveable as neutron poisons are continuously removed
+from the core via the continuous online reprocessing and gas sparging systems.
+Online fuel reprocessing also minimizes reactor downtime because refuelling
+can be performed at any time. The higher operating temperature results in
+greater thermal efficiencies and presents new opportunities beyond
+conventional electricity generation, namely industrial process heat
+applications and hydrogen production.
 
-\subsection{Model Geometry}
-
-In this thesis, we study the 2-D axisymmetric representation of the
-\gls{MSFR} as shown in Fig. \ref{fig:msfrgeom}. This is the same model
-geometry used in a neutronic benchmark
-study of the \gls{MSFR} and the coupled analyses by Fiorina et al., Aufiero
-et al., and Zhang et al. Even though this is an outdated design that had been
-later optimized for better flow distribution *, the analysis results are still
-valid for code verification purposes. Future efforts will adopt the optimized
-hourglass-shaped design for a more realistic assessment of the \gls{MSFR}.
-%
-\begin{figure}[htb!] 
-	\centering
-	\includegraphics[width=0.5\textwidth]{msfr-geom}
-	\caption{2-D axisymmetric model of the \gls{MSFR} core used for the
-	simulations in Serpent and Moltres. All dimensions are in meters.
-	\cite{brovchenko_neutronic_2019}}
-	\label{fig:msfrgeom}
-\end{figure}
-
-For the 2-D axisymmetric model, the sixteen individual loops are homogenized
-into a single loop while keeping the nominal volumetric flow rate constant.
-Keeping with the model used in the previous studies, we assumed that the
-region for heat exchangers and pumps on the loop consists of 100\% fuel salt.
-However, this assumption has a negligible impact on the multiplication factor
-due to its location behind the strong boron carbide absorber layer;
+Beyond these factors, \glspl{MSR} still require significant R\&D efforts and
+engineering demonstrations for experimental validation of various components
+before a full commercial model can be commissioned. Work towards creating
+safety and licensing framework for \glspl{MSR} have picked up pace only in
+recent years due to the growing interest from commercial \gls{MSR} developers.

### 2020-03-04 model.tex
expanding modeling approach chapter

@@
 active development at the VTT Technical Research Centre of Finland. It was
-born in 2004 out of the need for better general-purpose neutron transport
+created in 2004 out of the need for better general-purpose neutron transport
 codes for group constant generation in lattice geometries, and has since grown
@@
 neutronics benchmark geometry published in several previous papers involving
-the \gls{MSFR}. The relevant group constant data are collapsed into six
-neutron energy groups, and calculated for
-temperature values of 900 K to 1200 K at 50 K intervals.
+the \gls{MSFR}, as shown in Fig. \ref{fig:msfrgeom}. The relevant group
+constant data are collapsed into six neutron energy groups, and calculated for
+temperature values of 800 K to 1300 K at 50 K intervals.
 
@@
 Moltres is a \gls{MOOSE}-based application for coupled
-neutronics/thermal-hydraulics simulations of \glspl{MSR}. \gls{MOOSE} is a
-highly parallelizable,
+neutronics/thermal-hydraulics simulations of \glspl{MSR}. \gls{MOOSE}
+\cite{gaston_physics-based_2015} is a highly parallelizable,
 finite-element framework developed at \gls{INL} for simplifying the process of
@@
 terms in the \glspl{PDE}, the initial and boundary conditions, the material
-properties, etc., are represented in \gls{MOOSE} as C++ objects. This approach
+properties, etc., are represented in \gls{MOOSE} as C++ objects. New objects
+can inherit properties from related old objects to simplify implementation and
+reduce code duplication. Overall, this approach
 is helpful for many researchers, a significant fraction of whom do not possess
 high-level programming backgrounds, as they are unencumbered by the
-technical details and complexicities involved in programming mesh handling
+technical details and complexicities involved in programing mesh handling
 and \gls{PDE} solving in finite element analysis.
 
-\gls{MOOSE} itself relies heavily on LibMesh and PETSc for
-its mesh handling and \gls{PDE} solver functionalities. As a result,
-\gls{MOOSE} supports adaptive meshing schemes and automatic variable scaling
-amongst other features for improved accuracy and performance times. Full
+\gls{MOOSE} itself relies heavily on libMesh \cite{kirk_libmesh:_2006} and
+PETSc \cite{satish_petsc_2019} for its mesh handling and \gls{PDE} solver
+functionalities. As a result, \gls{MOOSE} supports adaptive meshing schemes
+and automatic variable scaling amongst other advanced features for improved
+accuracy and performance times. Full
 coupling is maintained by the execution of Newton-based solves on the
@@
 \gls{MOOSE} webpage, including the commonly used Exodus II file format.
-Specifically for the 2D case, axial symmetry is easily imposed by changing
-one line of code in the input file, without any changes in the Cartesian
-representations of the \glspl{PDE} and boundary conditions in their original
-C++ implementations. This feature provides significant computational time
-savings for 3D systems that exhibit high axial symmetry. Another important
-feature for reducing computational time is the use of MPI for parallel
-computing. All \gls{MOOSE}-based codes can be easily compiled and run on high
-performance computing clusters.
+Specifically for the 2D case, axial symmetry is easily and automatically
+imposed by changing one line of code in the input file, without any changes in
+the Cartesian representations of the \glspl{PDE} and boundary conditions in
+their original C++ implementations. This feature provides significant
+computational time savings for 3D systems that exhibit high axial symmetry.
+Another important feature for reducing computational time is the use of MPI
+for parallel computing. All \gls{MOOSE}-based codes can be compiled and
+run on high performance computing clusters.
 
@@
     \frac{1}{v_g} \frac{\partial \phi_g}{\partial t} &= \nabla \cdot D_g
-	\nabla \phi_g - \Sigma^r_g \phi_g
-	\sum^G_{g \neq g'} \Sigma^s_{g' \rightarrow g} \phi_{g'} + \chi^p_g
+    \nabla \phi_g - \Sigma^r_g \phi_g +
+    \sum^G_{g' \neq g} \Sigma^s_{g' \rightarrow g} \phi_{g'} + \chi^p_g
     \sum^G_{g'=1} (1-\beta) \nu \Sigma^f_{g'} \phi_{g'} + \chi^d_g \sum^I_i
@@
     \phi_{g'} - \lambda_i C_i - \boldsymbol{u} \cdot \nabla C_i + \nabla \cdot
-	\frac{\mu_t}{\rho Sc_t} \nabla C_i, \label{eq:dnp} \\
-	\intertext{where} v_g &= \text{average speed of neutrons in group }g 
+    D_t \nabla C_i, \label{eq:dnp} \\
+    \intertext{where}
+    v_g &= \text{average speed of neutrons in group $g$ [cm s$^{-1}$]} 
     \nonumber \\
-	\phi_g &= \text{flux of neutrons in group }g \nonumber \\
-	t &= \text{time} \nonumber \\
-	D_g &= \text{diffusion coefficient of neutrons in group }g \nonumber \\
+    \phi_g &= \text{neutron flux in group $g$ [cm$^{-2}$ s$^{-1}$]} \nonumber
+    \\
+    t &= \text{time [s]} \nonumber \\
+    D_g &= \text{diffusion coefficient of neutrons in group $g$ [cm$^2$
+    s$^{-1}$]} \nonumber \\
     \Sigma^r_g &= \text{macroscopic cross-section for removal of neutrons from
-	group }g \nonumber \\
+    group $g$ [cm$^{-1}$]} \nonumber \\
     \Sigma^s_{g' \rightarrow g} &= \text{macroscopic cross-section of
-	scattering from }g' \text{ to }g \nonumber \\
-	\chi^p_g &= \text{prompt fission spectrum neutrons in group }g \nonumber
-	\\
+    scattering from $g'$ to $g$ [cm$^{-1}$]} \nonumber \\
+    \chi^p_g &= \text{prompt fission spectrum for neutrons in group $g$}
+    \nonumber \\
     G &= \text{total number of discrete neutron groups} \nonumber \\
@@
     \Sigma^f_{g} &= \text{macroscopic fission cross-section for neutron in
-	group }g \nonumber \\
-	\chi^d_g &= \text{delayed fission spectrum neutrons in group }g \nonumber
-	\\
+    group $g$ [cm$^{-1}$]} \nonumber \\
+    \chi^d_g &= \text{delayed fission spectrum for neutrons in group }g
+    \nonumber \\
     I &= \text{total number of delayed neutron precursor groups} \nonumber \\
     \beta &= \text{total delayed neutron fraction} \nonumber \\
-	\beta_i &= \text{delayed neutron fraction of precursor group i} \nonumber
-	\\
+    \beta_i &= \text{delayed neutron fraction of precursor group $i$}
+    \nonumber \\
     \lambda_i &= \text{average decay constant of delayed neutron precursors in
-	precursor group }i \nonumber \\
+    precursor group $i$} \nonumber \\
     C_i &= \text{concentration of delayed neutron precursors in precursor
-	group }i \nonumber
+    group $i$} \nonumber \\
+    D_t &= \text{turbulent diffusion of the delayed neutron precursors [cm$^2$
+    s$^{-1}$]} \nonumber
 \end{align}
@@
 
-The limitations of the multi-group neutron diffusion compared to
-S$_{\text{N}}$ and Monte Carlo methods are well-documented but they provide
-decent accuracy at lower computational costs. A comparison of the
-multiplication factor and reactivity coefficients in the context of the
-\gls{MSFR} is provided in the results section.
+While the limitations of the multi-group neutron diffusion compared to other
+deterministic and Monte Carlo methods, particularly for flux values near
+boundaries, are well-documented, it provides acceptable
+accuracy at lower computational costs. Moreover, the
+\gls{MSFR} contains no material interfaces in the middle of the core where
+most of the fission interactions take place. The results section provides a
+comparison of the \gls{MSFR} multiplication factor values and reactivity
+coefficients between Moltres and Serpent.
 
-The \gls{DNP} concentration equation has an additional advection term to
-account for the movement of \glspl{DNP} in the primary coolant loop. Moltres
-also accounts for the decay of \glspl{DNP} outside the active core
-region by simulating its flow in a separate 1D geometry. The length and
+The \gls{DNP} concentration equation has additional advection and turbulent
+diffusion terms to account for the movement of \glspl{DNP} in the primary
+coolant loop. The turbulent diffusion value is governed by the following
+equation:
+%
+\begin{align}
+    D_t &= \frac{\mu_t}{\rho Sc_t}
+\end{align}
+%
+where $\mu_t$ is the eddy viscosity [kg cm$^{-1}$ s$^{-1}$] , $\rho$ is the
+density of the fuel salt [kg cm$^{-3}$], and $Sc_t$ is the turbulent Schmidt
+number. We assumed $Sc_t = 0.85$ for a fair comparison with the Polimi/TUDelft
+models which used the same value. Moltres also accounts for the decay of
+\glspl{DNP} outside the active core region by simulating its flow in a
+separate 1D pipe geometry. This outer loop pipe simulation is loosely coupled
+to the active core simulation through MOOSE's MultiApp functionality and
+inlet/outlet boundary values. The loose coupling is justified as the evolution
+of \gls{DNP} concentrations due to inflow is relatively slow. The length and
 velocity defined in this geometry may differ according to reactor
-specifications.
+specifications. For this work with the \gls{MSFR} model, we assumed a pipe
+length of 2.255 m with salt flowing at 1.1275 m s$^{-1}$ for an out-of-core
+residence time of 2 seconds to follow the design specifications.
 
-Moltres users may call for an arbitrary number of neutron energy groups as
+Moltres users may vary the total number of neutron energy groups as
 long as they provide Moltres with the appropriate group constant data. The
-number of precursor groups is also arbitrary, though usually pre-determined by
+number of precursor groups is also variable, though usually pre-determined by
 the choice of nuclear data library in the group constant generation step.
@@
 in \gls{MOOSE}. Once again, users have the freedom to select their
-interpolation method of choice. For this work, we have selected six neutron
-energy groups according to the energy boundaries in table \ref{table:bound},
-eight \gls{DNP} groups as defined by the JEFF-3.1.2 library, and the spline
-interpolation method.
-%
+interpolation method of choice.
+
+For this work, we have six neutron energy groups according to the energy
+boundaries in table \ref{table:bound}, eight \gls{DNP} groups as defined by
+the JEFF-3.1.2 library, and the spline interpolation method. The neutron flux
+and \gls{DNP} concentration values were approximated by first-order Lagrange
+and constant monomial shape functions respectively on the finite element mesh.
+We assumed vacuum boundary conditions for all six neutron group fluxes along
+the external boundaries of the geometry, and homogeneous Neumann boundary
+conditions along the axial symmetry boundary. For the \gls{DNP}
+concentrations, we imposed homogeneous Neumann boundary conditions on the
+walls, and inflow and outflow boundary conditions on the inlet and outlet
+boundaries respectively. The inlet \gls{DNP} concentration values were
+imported from the outlet values of the 1-D outer loop pipe at the same
+timestep.
+
 \begin{table}[htb!]
@@
 	\begin{tabular}{c S}
-		\hline
+		\toprule
 		{Group number} & {Upper bound [MeV]}\\
-		\hline
+		\midrule
 		1 & 20\\
@@
 		3 & 0.4979\\
-		4 & 2.47875$\times 10^{-2}$\\
-		5 & 5.5308$\times 10^{-3}$\\
-		6 & 7.485$\times 10^{-4}$\\
-		\hline
+		4 & 0.0247875\\
+		5 & 0.0055308\\
+		6 & 0.0007485\\
+		\bottomrule
 	\end{tabular}
@@
 
-For the decay heat model, previous studies on the MSFR * have shown that
-using three decay heat precursor groups with different half-lives can model
-decay heat in the MSFR for up to 300 seconds after shutdown with a relative
-error of less than 2 \%. Thus, we implemented the decay heat model using
-equations similar to the \gls{DNP} equations, as shown in eq
-\ref{eq:decayheat}:
+For the decay heat model, a previous study on the MSFR by Aufiero et al.
+\cite{aufiero_extended_2013} has shown that using three decay heat precursor
+groups with different half-lives in the form of exponential equations, can
+accurately model decay heat in the MSFR for up to 300 seconds after shutdown
+with a relative error of less than 2 \%. Thus, we implemented a decay heat
+model using the following set of equations:
 %
 \begin{align}
-	\frac{\partial D^k}{\partial t} &= f_k \sum^G_{g'=1} \epsilon_{g'}
-	\Sigma^f_{g'} \phi_{g'} - \lambda^k D^k - \boldsymbol{u} \cdot \nabla D^k
-	+ \nabla \cdot \frac{\mu_t}{\rho Sc_t} \nabla D^k, \label{eq:decayheat} \\
+	\frac{\partial \omega_k}{\partial t} &= f_k \sum^G_{g=1} \epsilon_{g}
+	\Sigma^f_{g} \phi_{g} - \lambda^k \omega_k - \boldsymbol{u} \cdot \nabla
+	\omega_k + \nabla \cdot D_t \nabla \omega_k, \label{eq:decayheat} \\
 	\intertext{where}
-	D^k &= \text{concentration of decay heat precursors in group }k \nonumber
-	\\
+    \omega_k &= \text{total decay heat power density from decay heat
+    precursors in group $k$ [W cm$^{-3}$]} \nonumber \\
 	f_k &= \text{fraction of decay heat to total power at steady state}
 	\nonumber \\
-	\epsilon_g &= \text{average fission energy} \nonumber \\
+	\epsilon_g &= \text{average fission energy per fission [W]} \nonumber \\
 	\lambda^k &= \text{average decay constant of decay heat precursors in
-	group }k. \nonumber
+	group $k$ [s$^{-1}$].} \nonumber
 \end{align}
+
+Like the neutron and \gls{DNP} groups, Moltres can take an arbitrary number of
+decay heat groups. In this work, as with the other parameters, we used the
+same decay heat fractions and decay constants used in the Polimi/TUDelft
+models for three decay heat groups.
+
+\begin{table}[htb!]
+	\centering
+	\caption{Decay heat group parameters \cite{fiorina_modelling_2014}.
+	$\lambda_i$ and $f_i$ are the decay constants and decay heat fractions
+	associated to group $i$.}
+	\begin{tabular}{S S S}
+		\toprule
+		{Decay heat group} & {$\lambda_i$ [s$^{-1}$]} & {$f_i$} \\
+		\midrule
+		1 & 0.1974 & 0.0117 \\
+		2 & 0.0168 & 0.0129 \\
+		3 & 3.58 $\times 10^{-4}$ & 0.0186 \\
+		\bottomrule
+	\end{tabular}
+	\label{table:decayheat}
+\end{table}
 
@@
 
-The governing equation for temperature in the molten salt is given as:
+The governing equation for temperature in the molten salt is:
 %
 \begin{align}
-	\rho c_{p} \frac{\partial T}{\partial t} &= - \nabla \cdot \big( \rho
-	c_{p} \overrightarrow{u} \cdot T - k \nabla T \big) + Q_s - Q_{hx},
-	\label{eq3}
+    \rho c_{p} \frac{\partial T}{\partial t} &= - \rho c_p \boldsymbol{u}
+    \cdot \nabla T + \nabla \cdot k \nabla T + Q_s - Q_{hx},
+    \label{eq:th}
     \intertext{where}
@@
 \end{align}
-	and the source term $Q_s$ is given as:
+	and the source term $Q_s$ is:
 %
 \begin{align}
-Q_s = \sum^G_{g=1} \epsilon_g \Sigma_g^f \phi_g. \label{eq4}
+    Q_s = \Big( 1 - \sum^K_{k=1} f_k \Big) \sum^G_{g=1} \epsilon_g \Sigma_g^f
+    \phi_g + \sum^K_{k=1} \omega_k, \label{eq:source}
+\end{align}
+%
+where $K$ is the total number of decay heat groups.
+
+Salt flow in the core is governed by the standard incompressible Navier-Stokes
+equations:
+%
+\begin{align}
+    \rho \frac{\partial \boldsymbol{u}}{\partial t} &= -\rho (\boldsymbol{u}
+    \cdot \nabla) \boldsymbol{u} + \nabla \cdot (-p \boldsymbol{I} + \mu [
+    \nabla \boldsymbol{u} + (\nabla \boldsymbol{u})^T] + \boldsymbol{f}
+    \label{eq:momemtum} \\
+    \nabla \cdot \boldsymbol{u} &= 0 \label{eq:divergence}
+    \intertext{where}
+    p &= \text{ pressure} \nonumber \\
+    \mu &= \text{ dynamic viscosity} \nonumber \\
+    \boldsymbol{f} &= \text{ body force per unit volume} \nonumber
 \end{align}
@@
 In the \gls{MSFR}, salt flow in the core has a Reynolds number on the order of
-10$^6$*. Thus the Navier-Stokes module in \gls{MOOSE} cannot sufficiently
+10$^6$. Thus the Navier-Stokes module in \gls{MOOSE} cannot sufficiently
 capture the turbulent flow behavior. Furthermore, the \gls{MSFR} benchmark
 geometry encourages the formation of turbulent recirculation regions in the
-fuel salt near the blanket tank regions. Due to this present limitation, only
-the flow in the active core region is simulated and it is assumed to be
-vertical flow. This eliminates the presence of recirculation zones reported by
-*, which will likely be a source of discrepancy between our results and
-theirs. Another consequence is the absence of turbulent diffusion, resulting
+fuel salt near the blanket tank regions as seen in the Polimi/TUDelft models
+\cite{fiorina_modelling_2014}. Moltres does not have a turbulence model in its
+current implementation. Therefore, instead of solving for the eddy viscosity
+with a two-equation \gls{RANS} model, we took a zeroth-order approximation of
+the eddy viscosity based on results reported by Fiorina et al.
+\cite{fiorina_modelling_2014}. We assumed $\mu_t = 40$ Pa s throughout the
+core.
+
+Another consequence is the absence of turbulent diffusion, resulting
 in higher-than-expected temperatures along the central axis of the reactor

### 2020-03-04 msfr.tex
expanding modeling approach chapter

@@
 	\caption{2-D axisymmetric model of the \gls{MSFR} core used for the
-	simulations in Serpent and Moltres. All dimensions are in meters.
+	simulations in Serpent. All dimensions are in meters.
 	\cite{brovchenko_neutronic_2019}}
@@
 \toprule
-Property & Expression & {Value at 973 K} & Validity Range\\
+Property & Formula & {Value at 973 K} & Validity Range\\
 \midrule

### 2020-03-09 introduction.tex
Matt's chp 1,2,3 review edits. add results chp.

@@
 
-Rapidly rising greenhouse gas emissions since the Industrial Revolution in the
-1800s have contributed significantly to global warming, the consequences of
-which have been acutely felt in recent years through increasingly frequent and
-record breaking extreme weather events. Urgent measures are necessary to limit
-CO$_2$ emissions and the negative impacts of global warming.
+Since the Industrial Revolution in the 1800s, rapidly rising greenhouse gas
+emissions have contributed significantly to global warming, the consequences
+of which have been acutely felt in recent years through increasingly frequent
+and record breaking extreme weather events. Urgent measures are necessary to
+limit CO$_2$ emissions and the negative impacts of global warming.
 
@@
 of all low-carbon power sources, nuclear power in general is best suited to
-replace fossil fuel burning; it provides consistent base load
+replace fossil fuel burning; it provides consistent base-load
 power independent of weather and local geographical conditions. Solar and wind
-power are dependent on favorable weather conditions, while hydropower requires
-relatively more area and is limited
-to geologically appropriate locations. This argument
-provides a strong case for nuclear power being the most appropriate and
-realistic alternative to burning fossil fuels for our current and future
-electricity needs.
+power are dependent on favorable weather conditions while hydropower requires
+relatively more area and is limited to geologically appropriate locations.
+This argument provides a strong case for nuclear power being the most
+appropriate and realistic alternative to burning fossil fuels for our current
+and future electricity needs.
 
 Large scale reactor deployments would be necessary to displace the
-presently large share of energy production from fossil fuel power plants.
-However, several obstacles stand in the way of mass reactor deployments. These
+presently large share of energy production from fossil fuel power plants;
+however, several obstacles stand in the way of mass reactor deployments. These
 include perceived safety risks, sustainability concerns, nuclear proliferation
-risks and the ability to compete economically with other sources of energy.
-The Molten Salt Reactor (MSR) concept, one of six advanced reactor designs
-selected by the Generation IV International Forum
-(GIF)* for continued research and development, is a potential solution to the
-aforementioned issues.
+risks, and the ability to compete economically with other sources of energy.
+The \gls{MSR} concept, one of six advanced reactor designs
+selected by the \gls{GIF} \cite{gif_technology_2002} for continued research
+and development, is a potential solution to the aforementioned issues.
 
@@
 consideration. Existing reactor system-level codes and modeling approaches
-for conventional \glspl{LWR} analysis contain simplifying assumptions in
+for conventional \gls{LWR} analysis contain simplifying assumptions in
 multiphysics coupling and other areas that render the techniques unsuitable
-for simulating \gls{MSR}. Thus, making minor modifications to these codes
+for simulating \glspl{MSR}. Thus, making minor modifications to these codes
 without changing the underlying algorithms is not the best approach for
@@
 2-D axisymmetric or full 3-D models. In 2011, Cammi et al.
-\cite{cammi_multi-physics_2011} performed an ``\gls{MPM}'' analysis
-of a simplified 2-D axisymmetric model of a single MSBR fuel channel using
-the commercial finite-element analysis software COMSOL Multiphysics. The
+\cite{cammi_multi-physics_2011} performed a ``\gls{MPM}'' analysis
+of a simplified 2-D axisymmetric model of a single \gls{MSBR} fuel channel
+using the commercial finite element analysis software COMSOL Multiphysics. The
 physics were implemented through the two-group neutron diffusion
@@
 neutronics and thermal-hydraulics respectively. The
-authors emphasized the need for proper full coupling of the multiphysics, and
+authors emphasized the need for proper full coupling of the multiphysics and
 presented both steady-state and transient results in various
 scenarios such as reactivity insertions, changes in pumping rate, and the
-presence of periodic perturbations. This approach featured again in a later
-paper by Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D
+presence of periodic perturbations. This approach was featured again in a
+later paper by Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D
 axisymmetric model of the
@@
 University of Technology, where they coupled their in-house neutronics and
-thermal-hydraulics codes, DALTON and HEAT respectively. With multi-group
-neutron diffusion and \gls{RANS} formulations on ultra-fine meshes, both
+thermal-hydraulics codes, DALTON and HEAT respectively. With multigroup
+neutron diffusion and \gls{RANS} formulations on ultra fine meshes, both
 models showed good agreement in the steady-state neutron flux, temperature,
@@
 parallel finite element framework. Similar to COMSOL and OpenFOAM, it solves
-the deterministic multi-group neutron diffusion and thermal-hydraulics
+the deterministic multigroup neutron diffusion and thermal-hydraulics
 \glspl{PDE} simultaneously on the same mesh. It supports up to 3-D meshes and

### 2020-03-09 model.tex
Matt's chp 1,2,3 review edits. add results chp.

@@
 The main simulation tool used in this work for studying the \gls{MSFR} is
-Moltres. While we used Moltres for the steady state and transient analysis, it
+Moltres. While we used Moltres for the steady-state and transient analysis, it
 requires a dedicated neutron transport solver such as Serpent 2 to generate
-neutron energy group constant data for the deterministic, multi-group
+neutron energy group constant data for the deterministic, multigroup
 neutronics calculations in Moltres. This chapter provides brief descriptions
@@
 \cite{gaston_physics-based_2015} is a highly parallelizable,
-finite-element framework developed at \gls{INL} for simplifying the process of
+finite element framework developed at \gls{INL} for simplifying the process of
 creating fully-coupled, non-linear, multiphysics solvers. The framework
@@
 high-level programming backgrounds, as they are unencumbered by the
-technical details and complexicities involved in programing mesh handling
+technical details and complexities involved in programing mesh handling
 and \gls{PDE} solving in finite element analysis.
@@
 
-\gls{MOOSE}, and Moltres by extension, are capable of up to 3D geometry
+\gls{MOOSE}, and Moltres by extension, are capable of up to 3-D geometry
 modelling. They support a wide range of input mesh file formats listed in a
 \gls{MOOSE} webpage, including the commonly used Exodus II file format.
-Specifically for the 2D case, axial symmetry is easily and automatically
+Specifically for the 2-D case, axial symmetry is easily and automatically
 imposed by changing one line of code in the input file, without any changes in
@@
 their original C++ implementations. This feature provides significant
-computational time savings for 3D systems that exhibit high axial symmetry.
+computational time savings for 3-D systems that exhibit high axial symmetry.
 Another important feature for reducing computational time is the use of MPI
@@
 Focusing our attention on Moltres, the neutron flux calculations are
-performed using the standard formulations for the time-dependent multi-group
+performed using the standard formulations for the time-dependent multigroup
 neutron diffusion equations and \gls{DNP} concentration equations as shown in
@@
     \intertext{where}
-    v_g &= \text{average speed of neutrons in group $g$ [cm s$^{-1}$]} 
+    v_g &= \text{average speed of neutrons in group $g$ [cm s$^{-1}$],} 
     \nonumber \\
-    \phi_g &= \text{neutron flux in group $g$ [cm$^{-2}$ s$^{-1}$]} \nonumber
+    \phi_g &= \text{neutron flux in group $g$ [cm$^{-2}$ s$^{-1}$],} \nonumber
     \\
-    t &= \text{time [s]} \nonumber \\
+    t &= \text{time [s],} \nonumber \\
     D_g &= \text{diffusion coefficient of neutrons in group $g$ [cm$^2$
-    s$^{-1}$]} \nonumber \\
-    \Sigma^r_g &= \text{macroscopic cross-section for removal of neutrons from
-    group $g$ [cm$^{-1}$]} \nonumber \\
-    \Sigma^s_{g' \rightarrow g} &= \text{macroscopic cross-section of
-    scattering from $g'$ to $g$ [cm$^{-1}$]} \nonumber \\
-    \chi^p_g &= \text{prompt fission spectrum for neutrons in group $g$}
+    s$^{-1}$],} \nonumber \\
+    \Sigma^r_g &= \text{macroscopic cross section for removal of neutrons from
+    group $g$ [cm$^{-1}$],} \nonumber \\
+    \Sigma^s_{g' \rightarrow g} &= \text{macroscopic cross section of
+    scattering from $g'$ to $g$ [cm$^{-1}$],} \nonumber \\
+    \chi^p_g &= \text{prompt fission spectrum for neutrons in group $g$,}
     \nonumber \\
-    G &= \text{total number of discrete neutron groups} \nonumber \\
-    \nu &= \text{average number of neutrons produced per fission} \nonumber \\
-    \Sigma^f_{g} &= \text{macroscopic fission cross-section for neutron in
-    group $g$ [cm$^{-1}$]} \nonumber \\
-    \chi^d_g &= \text{delayed fission spectrum for neutrons in group }g
+    G &= \text{total number of discrete neutron groups,} \nonumber \\
+    \nu &= \text{average number of neutrons produced per fission,} \nonumber
+    \\
+    \Sigma^f_{g} &= \text{macroscopic fission cross section for neutron in
+    group $g$ [cm$^{-1}$],} \nonumber \\
+    \chi^d_g &= \text{delayed fission spectrum for neutrons in group $g$,}
     \nonumber \\
-    I &= \text{total number of delayed neutron precursor groups} \nonumber \\
-    \beta &= \text{total delayed neutron fraction} \nonumber \\
-    \beta_i &= \text{delayed neutron fraction of precursor group $i$}
+    I &= \text{total number of delayed neutron precursor groups,} \nonumber \\
+    \beta &= \text{total delayed neutron fraction,} \nonumber \\
+    \beta_i &= \text{delayed neutron fraction of precursor group $i$,}
     \nonumber \\
     \lambda_i &= \text{average decay constant of delayed neutron precursors in
-    precursor group $i$} \nonumber \\
+    precursor group $i$,} \nonumber \\
     C_i &= \text{concentration of delayed neutron precursors in precursor
-    group $i$} \nonumber \\
+    group $i$,} \nonumber \\
     D_t &= \text{turbulent diffusion of the delayed neutron precursors [cm$^2$
-    s$^{-1}$]} \nonumber
+    s$^{-1}$].} \nonumber
 \end{align}
@@
 
-While the limitations of the multi-group neutron diffusion compared to other
+While the limitations of the multigroup neutron diffusion compared to other
 deterministic and Monte Carlo methods, particularly for flux values near
-boundaries, are well-documented, it provides acceptable
+boundaries, are well-documented, the diffusion model provides acceptable
 accuracy at lower computational costs. Moreover, the
 \gls{MSFR} contains no material interfaces in the middle of the core where
-most of the fission interactions take place. The results section provides a
+most of the fission interactions take place. The Results section provides a
 comparison of the \gls{MSFR} multiplication factor values and reactivity
@@
 \glspl{DNP} outside the active core region by simulating its flow in a
-separate 1D pipe geometry. This outer loop pipe simulation is loosely coupled
+separate 1-D pipe geometry. This outer loop pipe simulation is loosely coupled
 to the active core simulation through MOOSE's MultiApp functionality and
@@
 long as they provide Moltres with the appropriate group constant data. The
-number of precursor groups is also variable, though usually pre-determined by
+number of precursor groups is also variable, though usually predetermined by
 the choice of nuclear data library in the group constant generation step.
 Moltres automatically interpolates the group constant data for required
-temperatures using one of the many pre-defined interpolation methods available
+temperatures using one of the many predefined interpolation methods available
 in \gls{MOOSE}. Once again, users have the freedom to select their
@@
 and constant monomial shape functions respectively on the finite element mesh.
+Figure * shows the mesh adopted for the \gls{MSFR} model.
 We assumed vacuum boundary conditions for all six neutron group fluxes along
@@
 accurately model decay heat in the MSFR for up to 300 seconds after shutdown
-with a relative error of less than 2 \%. Thus, we implemented a decay heat
-model using the following set of equations:
+with a relative error of less than 2\%. Thus, we implemented a decay heat
+model using the following equation:
 %
@@
 		2 & 0.0168 & 0.0129 \\
-		3 & 3.58 $\times 10^{-4}$ & 0.0186 \\
+		3 & 0.000358 & 0.0186 \\
 		\bottomrule
@@
 
-The governing equation for temperature in the molten salt is:
+Fluid dynamics in Moltres can be modeled using the \gls{INS} equations with
+the Boussinesq hypothesis for eddy viscosity. Most of the Navier-Stokes
+capabilities in Moltres is derived from the MOOSE Navier-Stokes module
+\cite{peterson_overview_2017}. The standard \gls{INS} equations are:
+%
+\begin{align}
+    \text{Momentum eq.:} && \rho \frac{\partial \boldsymbol{u}}{\partial t} &=
+    -\rho (\boldsymbol{u}
+    \cdot \nabla) \boldsymbol{u} + \nabla \cdot (-p \boldsymbol{I} + \mu [
+    \nabla \boldsymbol{u} + (\nabla \boldsymbol{u})^T] + \boldsymbol{f} &&
+    \label{eq:momemtum} \\
+    \text{Divergence-free:} && \nabla \cdot \boldsymbol{u} &= 0 &&
+    \label{eq:divergence}
+    \intertext{where}
+    && p &= \text{ pressure} && \nonumber \\
+    && \mu &= \text{ dynamic viscosity} && \nonumber \\
+    && \boldsymbol{f} &= \text{ body force per unit volume} && \nonumber
+\end{align}
+
+In addition to the intrinsic molecular viscosity, we introduced an eddy
+viscosity term to approximate turbulent flow effects. The current
+implementation of Moltres does not have a turbulence model such as the
+\gls{RANS} models used in the Polimi/TUDelft models. Thus, we made a
+zeroth-order approximation of the eddy viscosity based on the results reported
+in the Polimi/TUDelft models. The eddy viscosity is assumed to be 40 Pa s.
+Despite the simplicity of this assumption, we were able to reproduce much of
+the flow profile observed in the Polimi/TUDelft models at steady state.
+
+The flow rate is dictated by an inflow boundary condition at the core inlet
+for a volumetric flow rate of 4.5 m$^3$ s$^{-1}$. We imposed no-slip boundary
+conditions on the walls of the core, and homogeneous Neumann boundary
+conditions on the core outlet. The energy balance equation for temperature is
+given in Eq. \ref{eq:temp}. The diffusion term includes turbulent heat
+diffusivity based on the eddy viscosity $\mu_t$ and the turbulent Prandtl
+number $Pr_t$, which we assume to be equal to 0.85. 
 %
@@
     \rho c_{p} \frac{\partial T}{\partial t} &= - \rho c_p \boldsymbol{u}
-    \cdot \nabla T + \nabla \cdot k \nabla T + Q_s - Q_{hx},
-    \label{eq:th}
+    \cdot \nabla T + \nabla \cdot [(k + k_t) \nabla T] + Q_s - Q_{hx}
+    \label{eq:temp} \\
+    k_t &= \frac{\mu_t}{\rho Pr_t} \\
+    Q_s &= \Big( 1 - \sum^K_{k=1} f_k \Big) \sum^G_{g=1} \epsilon_g \Sigma_g^f
+    \phi_g + \sum^K_{k=1} \omega_k, \label{eq:source}
     \intertext{where}
@@
     T &= \text{temperature of molten salt} \nonumber \\
-    \overrightarrow{u} &= \text{velocity of molten salt} \nonumber \\
+    \boldsymbol{u} &= \text{velocity of molten salt} \nonumber \\
     k &= \text{thermal conductivity of molten salt} \nonumber \\
-    Q_{hx} &= \text{heat sink},
-\end{align}
-	and the source term $Q_s$ is:
-%
-\begin{align}
-    Q_s = \Big( 1 - \sum^K_{k=1} f_k \Big) \sum^G_{g=1} \epsilon_g \Sigma_g^f
-    \phi_g + \sum^K_{k=1} \omega_k, \label{eq:source}
+    Q_{hx} &= \text{heat sink} \nonumber \\
 \end{align}
 %
-where $K$ is the total number of decay heat groups.
+where $K$ is the total number of decay heat groups. The first term in the heat
+source expression represents prompt fission heat, and the second term
+represents decay heat from the $K$ decay heat groups.
 
-Salt flow in the core is governed by the standard incompressible Navier-Stokes
-equations:
-%
-\begin{align}
-    \rho \frac{\partial \boldsymbol{u}}{\partial t} &= -\rho (\boldsymbol{u}
-    \cdot \nabla) \boldsymbol{u} + \nabla \cdot (-p \boldsymbol{I} + \mu [
-    \nabla \boldsymbol{u} + (\nabla \boldsymbol{u})^T] + \boldsymbol{f}
-    \label{eq:momemtum} \\
-    \nabla \cdot \boldsymbol{u} &= 0 \label{eq:divergence}
-    \intertext{where}
-    p &= \text{ pressure} \nonumber \\
-    \mu &= \text{ dynamic viscosity} \nonumber \\
-    \boldsymbol{f} &= \text{ body force per unit volume} \nonumber
-\end{align}
+We modeled the heat exchanger using a point source/sink at the center of the
+1-D outer loop pipe region. The rate of heat extraction is equivalent to the
+nominal 3 GW$_{\text{th}}$ power output with appropriate scaling to account
+for the 1-D geometry. *** include more detail when finalized ***
 
-In the \gls{MSFR}, salt flow in the core has a Reynolds number on the order of
-10$^6$. Thus the Navier-Stokes module in \gls{MOOSE} cannot sufficiently
-capture the turbulent flow behavior. Furthermore, the \gls{MSFR} benchmark
-geometry encourages the formation of turbulent recirculation regions in the
-fuel salt near the blanket tank regions as seen in the Polimi/TUDelft models
-\cite{fiorina_modelling_2014}. Moltres does not have a turbulence model in its
-current implementation. Therefore, instead of solving for the eddy viscosity
-with a two-equation \gls{RANS} model, we took a zeroth-order approximation of
-the eddy viscosity based on results reported by Fiorina et al.
-\cite{fiorina_modelling_2014}. We assumed $\mu_t = 40$ Pa s throughout the
-core.
-
-Another consequence is the absence of turbulent diffusion, resulting
-in higher-than-expected temperatures along the central axis of the reactor
-where the neutron flux and heat generation is highest. This is further
-exacebated by the high Prandtl number of the molten salt; heat transfer is
-dominated by convective flow. This means that there would be very low heat
-transfer observed by diffusion due to the negligible contributions from
-heat conduction.
-
-For steady-state solves, there are two options: the inverse power
-method (\texttt{InversePowerMethod}) and the nonlinear Newton method
-(\texttt{NonlinearEigen}).
+We expect good qualitative agreement with the Polimi/TUDelft models,
+including the large recirculation region near the blanket tank walls and the
+resulting high temperatures in that region. There would be some minor
+discrepancies where the viscosity values are under- or over-predicted, leading
+to slightly inaccurate temperature and precursor concentration values from
+turbulent diffusion.

### 2020-03-09 msfr.tex
Matt's chp 1,2,3 review edits. add results chp.

@@
 under the \gls{EVOL} and \gls{SAMOFAR} projects. The main reactor
-specifications and schematic view are shown in Table \ref{table:msfr} and Fig.
-\ref{fig:msfr} respectively. Developed from the \gls{MSBR}, the \gls{MSFR} is
-intended to run primarily on a closed thorium fuel cycle with
+specifications and schematic view are shown in Table \ref{table:msfr} and
+Figure \ref{fig:msfr}, respectively. Developed from the \gls{MSBR}, the
+\gls{MSFR} is intended to run primarily on a closed thorium fuel cycle with
 continuous online fuel reprocessing. Several reasons motivated the omission of
@@
 In the \gls{MSFR} design, fuel salt flows upwards through a 9 m$^3$ central
-core region. At the top of the core, the flow separates into 16 smaller
+core region. At the top of the core, the flow separates into sixteen smaller
 external loops, each of which passes through a heat exchanger before being
@@
 containment vessel designed to keep it subcritical. Reactivity control under
-normal operating conditions is performed by varying pump speeds to advantage
-of strong thermal feedback. Coupled with the fact that there is no excess
-reactivity reserve due to online fuel reprocessing, there are no control rods
-in the \gls{MSFR} design.
+normal operating conditions is performed by varying pump speeds to take
+advantage of the strong thermal feedback. Coupled with the fact that there is
+no excess reactivity reserve due to online fuel reprocessing, there are no
+control rods in the \gls{MSFR} design.
 
@@
 constraints seen in \glspl{LWR}. The inlet and outlet temperature
-specifications of the fuel salt are 923 K and 1023 K respectively. This was
+specifications of the fuel salt are 923 K and 1023 K, respectively. This was
 motivated by the need for a minimum 50 K temperature buffer between the
@@
 model with the sixteen individual external loops homogenized into a single
-outer loop as shown in Fig. \ref{fig:msfrgeom}. For the multi-group
+outer loop as shown in Figure \ref{fig:msfrgeom}. For the multigroup
 cross sections and group constants calculations in Serpent, we extended this
@@
 accurate, the exact details of the pump and heat exchanger systems are still
-be researched, and this external loop region is presumed to be of little
+under active study, and this external loop region is presumed to be of little
 neutronic importance due to its position behind the strong boron carbide
@@
 
-There are two key differences between our \gls{MSFR} model geometry in
-Moltres and the Polimi/TUDelft models. The first difference is a relative
+Although we used the exact reference model for generating group constant data
+from Serpent, there are two minor differences between the \gls{MSFR} model
+geometry we used in Moltres, and the Polimi/TUDelft models. The
+first difference is a relative
 minor change to the mesh by the exclusion of the 2 cm thick structural
 material around the blanket tank that separates the fuel and blanket salts.
-We removed this feature in our finite element mesh as we had difficulty
-meshing this layer that is relatively much thinner than the rest of the model.
-Any impact on the
+We removed this feature in our finite element mesh as we encountered
+difficulty meshing this layer that is relatively much thinner than the rest of
+the model. Any impact on the
 neutron flux is expected to be minimal. Furthermore, we solved for the
@@
 Thermal conductivity, $k$ [W m$^{-1}$ K$^{-1}$] & $0.928+8.397 \times 10^{-5} \cdot T$ & $0.01010$ & 891-1020 K \\
-Specific heat, $c_p$ [J kg$^{-1}$ K$^{-1}$] & $-1111+2.78\cdot T$ & $1555$* & 867-907 K \\
+Specific heat, $c_p$ [J kg$^{-1}$ K$^{-1}$] & $-1111+2.78\cdot T$ & $1594$ & 867-907 K \\
 \bottomrule
@@
 of 10 g cm$^{-3}$. The material composition of the reflectors may be subject
-to minor changes but it is not a major concern as the reflectors are not
+to minor changes, but it is not a major concern as the reflectors are not
 situated in regions of high neutron flux. The \gls{MSFR} also includes a 20 cm
@@
 neutron irradiation. The reference specifications indicate that natural boron
-is used, which is composed of 19.8 \% B$^{10}$ and 80.2 \% B$^{11}$, with an
+is used, which is composed of 19.8 \% $^{10}$B and 80.2 \% $^{11}$B, with an
 overall density of 2.52 g cm$^{-3}$. 

### 2020-03-09 msr.tex
Matt's chp 1,2,3 review edits. add results chp.

@@
 consisted of enriched uranium in a molten salt mixture of NaF, ZrF$_4$, and
-UF$_4$, and moderated by blocks of beryllium oxide. The project ultimately
+UF$_4$, and was moderated by blocks of beryllium oxide. The project ultimately
 never came to fruition as the development of intercontinental ballistic
@@
 \cite{rosenthal_molten-salt_1970}. One of the key findings from the
-research was that better economy could be achieved from breeding $^{233}$U
-from $^{232}$Th in thermal-spectrum reactors than $^{239}$Pu from $^{238}$U
-\cite{macpherson_molten_1985}. Ultimately, these efforts culminated in the
-design, construction, and successful operation of the \gls{MSRE}, a graphite
-moderated thermal \gls{MSR}. Although no breeding was attempted with the
-\gls{MSRE}, scientists at \gls{ORNL} obtained a wealth of experimental data
-and new insights from the study of this reactor. The \gls{MSRE} had a
-graphite-moderated design with LiF-BeF$_2$-ZrF$_4$-UF$_4$ fuel salt mixture,
+research was that breeding $^{233}$U from $^{232}$Th gave better performance
+than breeding $^{239}$Pu from $^{238}$U in thermal-spectrum reactors.
+Ultimately, these efforts culminated in the design, construction, and
+successful operation of the \gls{MSRE}, a graphite moderated thermal
+\gls{MSR}. Although the experiment did not include breeding, scientists at
+\gls{ORNL} obtained a wealth of other experimental data and new insights from
+the study of this reactor. The \gls{MSRE} had a
+graphite-moderated design with a LiF-BeF$_2$-ZrF$_4$-UF$_4$ fuel salt mixture,
 initially rated at 10 MW$_{\text{th}}$ but later restricted to 8
@@
 \gls{MSRE}, the \gls{MSBR} relies on continuous online reprocessing to add
-fertile material and remove fission product neutron poisons. As a breeder
+fertile material and remove \gls{FP} neutron poisons. As a breeder
 reactor, the doubling time (the minimum amount of time required to produce
@@
 boost its conversion ratio throughout its lifetime. It has a continuous
-online feed consists of \gls{HALEU} to maintain criticality, and denatured
-$^{235}$U to keep uranium enrichement levels below nuclear non-proliferation
+online feed consisting of \gls{HALEU} to maintain criticality, and denatured
+$^{235}$U to keep uranium enrichment levels below nuclear non-proliferation
 policy thresholds. The design also includes a gas sparging system for removing
@@
 There was a concurrent program at the UK Atomic Energy Authority for the
-development of a 2500MW$_{\text{e}}$ lead-cooled
+development of a 2500 MW$_{\text{e}}$ lead-cooled
 \gls{MCFR} concept \cite{smith_assessment_1974}. It is a dual fluid system,
@@
 under development today. Graphite-moderated thermal-spectrum \glspl{MSR} are
-typically straight-forward \gls{LEU} burners, or $^{232}$Th/$^{233}$U
+typically straightforward \gls{LEU} burners, or $^{232}$Th/$^{233}$U
 iso-breeders/breeders, while epithermal- and fast-spectrum \glspl{MSR} have
@@
 phenomenon provides a large degree of control and stability as it is always
-present in an \gls{MSR} regardless whether the operating conditions.
+present in an \gls{MSR} regardless of the operating conditions.
 
@@
 unlikely situation where an \gls{MSR} encounters a severe runaway reaction,
-we can rely on another passive safety feature, freeze plugs. A freeze plug is
-a plug filled with solidified salt at the bottom of the core
-\cite{aji_experimental_2020}. It is actively
-cooled by fans or other cooling systems to keep it just below the freezing
-point of the salt. When temperatures in the core exceed a certain threshold
-during a dangerous transient, the freeze plug melts and the molten salt in the
-core drains into a containment tank designed to keep the salt in a
-sub-critical configuration. This is especially easy to achieve for
+we can rely on another passive safety feature called freeze plugs. A freeze
+plug is a plug of solidified salt at the bottom of the core actively cooled by
+fans or other cooling systems to keep its temperature just below the freezing
+point of the salt \cite{aji_experimental_2020}. When temperatures in the core
+exceed a certain threshold during a dangerous transient, the freeze plug melts
+and the molten salt in the core drains into a containment tank designed to
+keep the salt in a subcritical configuration. This is especially easy to
+achieve for
 thermal-spectrum \glspl{MSR} as the absence of moderators in the containment
@@
 drain the salt out of the core. Decay heat in \glspl{MSR} with online
-reprocessing isexpected to be lower than that in \glspl{LWR} due to the
+reprocessing is typically lower than that in \glspl{LWR} due to the
 continuous removal of \glspl{FP}. For example, the decay heat in an \gls{MSFR}
-after reaching equilibrium salt composition is approximately 3.5 \% compared
-to 6 \% in \glspl{LWR} \cite{brovchenko_design-related_2013}.
+after reaching equilibrium salt composition is approximately 3.5\% compared
+to 6\% in \glspl{LWR} \cite{brovchenko_design-related_2013}.
 
 However, there are also some safety disadvantages associated with \glspl{MSR}.
-Firstly, \glspl{MSR} have smaller fractions of \gls{DNP} in the active core
-region as some of them decay in the external loop regions. This complicates
+Firstly, \glspl{MSR} have smaller fractions of \glspl{DNP} in the active core
+region as some of them decay in the external loop regions; this complicates
 reactor control and may result in faster transients due to the decrease in
@@
 
-\subsection{Other factors}
+\subsection{Other Factors}
 
@@
 $^{232}$Th/$^{233}$U and $^{238}$U/$^{239}$Pu fuel cycles are viable
-candidates for breeding in \gls{MSR}, with the former being more suited for
+candidates for breeding in \glspl{MSR}, with the former being more suited for
 thermal reactors and the latter being more suited for fast reactors. Another
@@
 core loading of $^{239}$Pu with a fast spectrum before transitioning towards
-$^{233}$U fuel with a $^{232}$Th feed. This measure effectively circumvents
+$^{233}$U fuel with a $^{232}$Th feed; this measure effectively circumvents
 the issue of having insufficient $^{233}$U inventory to start up the reactor.
@@
 isotope from the fuel salt. $^{233}$Pa decays into $^{233}$U with a half-life
-of approximately 27 days, and $^{233}$U is at least as potent as $^{235}$U for
+of approximately 27 days and $^{233}$U is at least as potent as $^{235}$U for
 nuclear weapons production. The highly radioactive $^{232}$U provides some
@@
 
-The economic prospects are uncertain at this stage of \gls{MSR} development
-but we can qualitatively list the potential impacts. Capital costs are
-expected to be greater due to the need for increased corrosion resistance and
-the supporting reprocessing facilities. This is in spite of the reduced
+The economic prospects are uncertain at the current stage of \gls{MSR}
+development but we can qualitatively list the potential impacts. Capital costs
+are expected to be greater due to the need for increased corrosion resistance
+and the supporting reprocessing facilities. This is in spite of the reduced
 pressure constraints on the reactor pressure vessel for \glspl{MSR} resulting
@@
 savings from eliminating fuel pellet and fuel assembly fabrication. Higher
-fuel utilization is achieveable as neutron poisons are continuously removed
+fuel utilization is achievable as neutron poisons are continuously removed
 from the core via the continuous online reprocessing and gas sparging systems.
@@
 engineering demonstrations for experimental validation of various components
-before a full commercial model can be commissioned. Work towards creating
-safety and licensing framework for \glspl{MSR} have picked up pace only in
+before a full commercial model can be commissioned. Work towards creating a
+safety and licensing framework for \glspl{MSR} has picked up pace only in
 recent years due to the growing interest from commercial \gls{MSR} developers.

### 2020-03-09 results.tex
Matt's chp 1,2,3 review edits. add results chp.

This chapter discusses three main sets of simulation results. First, we
compare the key neutronics results between Serpent and Moltres for a static
model of the \gls{MSFR}, i.e. no salt flow, to assess the accuracy of the
six-group neutron diffusion model in Moltres on a fast-spectrum reactor. This
exercise also serves as a checking exercise for the group constant data
generated for Moltres from Serpent. Second, we present steady-state results
from Moltres with salt flow and verify it against published results, namely
from the Polimi/TUDelft and OpenFOAM models developed by Fiorina et al.
\cite{fiorina_modelling_2014} and Aufiero et al.
\cite{aufiero_development_2014}, respectively. Lastly, we present and compare
transient simulation results from Moltres with the aforementioned models.
Transient accident scenarios include an unprotected reactivity insertion, an
unprotected loss of flow, an unprotected loss of heat sink, a chilled inlet,
and a pump overspeed scenario.

\section{Static Model}

We calculated estimates of $k_{\text{eff}}$ values in Serpent and Moltres for
the \gls{MSFR} model with static salt (no flow), a uniform temperature
distribution of 973 K, and a fuel salt density of $4.1249 \times 10^3$ kg
m$^{-3}$. Moltres performs eigenvalue calculations for finding the 
$k_{\text{eff}}$ as it does not currently have the capability to calculate
the adjoint flux. Table \ref{table:keff} shows the $k_{\text{eff}}$ values
from Serpent, and from Moltres with and without \glspl{DNP}. We observe a
small 127.9 pcm discrepancy between the two codes, which we attribute to two
main factors: the neutron diffusion model, and the omission of the blanket
tank structural material. It is well-known that the neutron diffusion model is
not as accurate as the other S$_{\text{N}}$ or SP$_{\text{N}}$ deterministic
methods nor the Monte Carlo approach in Serpent. Regarding the omission of the
blanket tank material, we replaced the 2 cm-thick structural material with
blanket salt. This decision may be partly responsible for the higher
$k_{\text{eff}}$ value calculated by Moltres. Nevertheless, the discrepancy is
smaller than the 228.5 pcm and 256.7 pcm discrepancy reported by Cervi et al.
\cite{cervi_development_2019} for their six-group $SP_3$ and neutron diffusion
methods, respectively, in OpenFOAM. The neutron diffusion model in OpenFOAM is
the same approach Aufiero et al. \cite{aufiero_development_2014} used for
their transient analysis of the \gls{MSFR}, albeit with one neutron energy
group.

\begin{table}[htb!]
	\centering
	\caption{$k_{\text{eff}}$ values from Serpent and Moltres.}
	\begin{tabular}{l S S}
		\toprule
		{Code} & {$k_{\text{eff}}$} & {Diff. wrt Serpent [pcm]}\\
		\midrule
		{Serpent} & 1.00667 \pm 0.000049 & 0\\
		{Moltres with \glspl{DNP}} & 1.007949 \pm 0.000001 & 127.9\\
		{Moltres without \glspl{DNP}} & 1.004937 \pm 0.000001 & {-}\\
		\bottomrule
	\end{tabular}
	\label{table:keff}
\end{table}
%
\begin{table}[htb!]
	\centering
	\caption{$\beta_{\text{eff}}$ values from Serpent and Moltres.}
	\begin{tabular}{l S S}
		\toprule
		{Code} & {$\beta_{\text{eff}}$ [pcm]} & {Diff. wrt Serpent [pcm]}\\
		\midrule
		{Serpent} & 303.2 \pm .8 & 0\\
		{Moltres} & 301.2 \pm .2 & 2.0\\
		\bottomrule
	\end{tabular}
	\label{table:betaeff}
\end{table}

The

\section{Steady-State Behavior}

\subsection{Neutronics}

\subsection{Thermal-Hydraulics}

\section{Transient Behavior}

\subsection{Unprotected Reactivity Insertion}

\subsection{Unprotected Loss of Flow}

\subsection{Unprotected Loss of Heat Sink}

\subsection{Chilled Inlet}

\subsection{Pump Overspeed}

### 2020-03-30 model.tex
Add static model and steady-state results

@@
     \lambda_i &= \text{average decay constant of delayed neutron precursors in
-    precursor group $i$,} \nonumber \\
+    precursor group $i$ [s$^{-1}$],} \nonumber \\
     C_i &= \text{concentration of delayed neutron precursors in precursor
-    group $i$,} \nonumber \\
+    group $i$ [cm$^{-3}$],} \nonumber \\
     D_t &= \text{turbulent diffusion of the delayed neutron precursors [cm$^2$
@@
     D_t &= \frac{\mu_t}{\rho Sc_t}
+    \intertext{where}
+    \mu_t &= \text{ eddy viscosity [Pa s],} \nonumber \\ 
+    \rho &= \text{ density of the fuel salt [kg m$^{-3}$],} \nonumber \\
+    Sc_t &= \text{ turbulent Schmidt number.} \nonumber
 \end{align}
 %
-where $\mu_t$ is the eddy viscosity [kg cm$^{-1}$ s$^{-1}$] , $\rho$ is the
-density of the fuel salt [kg cm$^{-3}$], and $Sc_t$ is the turbulent Schmidt
-number. We assumed $Sc_t = 0.85$ for a fair comparison with the Polimi/TUDelft
+
+We assumed $Sc_t = 0.85$ for a fair comparison with the Polimi/TUDelft
 models which used the same value. Moltres also accounts for the decay of
@@
 decay heat groups. In this work, as with the other parameters, we used the
-same decay heat fractions and decay constants used in the Polimi/TUDelft
-models for three decay heat groups.
+same decay heat fractions and decay constants, shown in Table
+\ref{eq:decayheat}, used in the Polimi/TUDelft models for three decay heat
+groups.
 
@@
     \intertext{where}
-    && p &= \text{ pressure} && \nonumber \\
-    && \mu &= \text{ dynamic viscosity} && \nonumber \\
-    && \boldsymbol{f} &= \text{ body force per unit volume} && \nonumber
+    && p &= \text{ pressure [Pa],} && \nonumber \\
+    && \mu &= \text{ dynamic viscosity [Pa s],} && \nonumber \\
+    && \boldsymbol{f} &= \text{ body force per unit volume [N m$^{-3}$].} &&
+    \nonumber
 \end{align}
@@
     \intertext{where}
-    \rho &= \text{density of molten salt} \nonumber \\
-    c_{p} &= \text{specific heat capacity of molten salt} \nonumber \\
-    T &= \text{temperature of molten salt} \nonumber \\
-    \boldsymbol{u} &= \text{velocity of molten salt} \nonumber \\
-    k &= \text{thermal conductivity of molten salt} \nonumber \\
-    Q_{hx} &= \text{heat sink} \nonumber \\
+    c_p &= \text{specific heat capacity of molten salt [J kg$^{-1}$ K$^{-1}$],
+    } \nonumber \\
+    T &= \text{temperature of molten salt [K]} \nonumber \\
+    \boldsymbol{u} &= \text{velocity of molten salt [m s$^{-1}$],} \nonumber
+    \\
+    k &= \text{thermal conductivity of molten salt [W m$^{-1}$ K$^{-1}$,}
+    \nonumber \\
+    Q_{hx} &= \text{heat loss rate through the heat exchanger [W].} \nonumber
 \end{align}

### 2020-03-30 results.tex
Add static model and steady-state results

@@
 We calculated estimates of $k_{\text{eff}}$ values in Serpent and Moltres for
-the \gls{MSFR} model with static salt (no flow), a uniform temperature
-distribution of 973 K, and a fuel salt density of $4.1249 \times 10^3$ kg
-m$^{-3}$. Moltres performs eigenvalue calculations for finding the 
-$k_{\text{eff}}$ as it does not currently have the capability to calculate
-the adjoint flux. Table \ref{table:keff} shows the $k_{\text{eff}}$ values
-from Serpent, and from Moltres with and without \glspl{DNP}. We observe a
-small 127.9 pcm discrepancy between the two codes, which we attribute to two
-main factors: the neutron diffusion model, and the omission of the blanket
-tank structural material. It is well-known that the neutron diffusion model is
-not as accurate as the other S$_{\text{N}}$ or SP$_{\text{N}}$ deterministic
-methods nor the Monte Carlo approach in Serpent. Regarding the omission of the
-blanket tank material, we replaced the 2 cm-thick structural material with
-blanket salt. This decision may be partly responsible for the higher
-$k_{\text{eff}}$ value calculated by Moltres. Nevertheless, the discrepancy is
-smaller than the 228.5 pcm and 256.7 pcm discrepancy reported by Cervi et al.
-\cite{cervi_development_2019} for their six-group $SP_3$ and neutron diffusion
-methods, respectively, in OpenFOAM. The neutron diffusion model in OpenFOAM is
-the same approach Aufiero et al. \cite{aufiero_development_2014} used for
-their transient analysis of the \gls{MSFR}, albeit with one neutron energy
-group.
+the \gls{MSFR} model with static salt (no flow), and uniform temperature
+distributions. Moltres solves the six-group neutron diffusion equations as a
+steady-state eigenvalue problem to find the $k_{\text{eff}}$ as it does not
+currently have the capability to calculate the adjoint flux. Table
+\ref{table:keff} shows the $k_{\text{eff}}$ values from Serpent and Moltres at
+973 K and the corresponding salt density, while Table \ref{table:keff} shows
+the $k_{\text{eff}}$ values for other temperatures at 100 K intervals. We
+observe small discrepancies on the order of 100 pcm between the two codes
+which we attribute to two main factors: the accuracy of the neutron diffusion
+model, and the omission of the blanket tank structural material. The neutron
+diffusion model is not as accurate as the other S$_{\text{N}}$ or
+SP$_{\text{N}}$ deterministic methods nor the Monte Carlo approach in Serpent.
+Regarding the omission of the blanket tank material, we replaced the
+2 cm-thick structural material with blanket salt. This replacement may be
+partly responsible for the higher $k_{\text{eff}}$ value calculated by Moltres
+as the macroscopic fission cross sections of the blanket salt are non-zero for
+the higher neutron energy groups. Nevertheless, the discrepancy is smaller
+than the 228.5 pcm and 256.7 pcm discrepancies reported by Cervi et al.
+\cite{cervi_development_2019} for their six-group $SP_3$ and neutron
+diffusion methods, respectively, in OpenFOAM. The neutron diffusion model in
+OpenFOAM is the same approach Aufiero et al. \cite{aufiero_development_2014}
+used for their transient analysis of the \gls{MSFR}, albeit with one neutron
+energy group.
 
 \begin{table}[htb!]
+    \small
 	\centering
-	\caption{$k_{\text{eff}}$ values from Serpent and Moltres.}
-	\begin{tabular}{l S S}
+	\caption{$k_{\text{eff}}$ values from Serpent and Moltres at 973 K.}
+	\begin{tabular}{l S c}
 		\toprule
-		{Code} & {$k_{\text{eff}}$} & {Diff. wrt Serpent [pcm]}\\
+		{Code} & {$k_{\text{eff}}$} & {Difference wrt Serpent [pcm]}\\
 		\midrule
-		{Serpent} & 1.00667 \pm 0.000049 & 0\\
-		{Moltres with \glspl{DNP}} & 1.007949 \pm 0.000001 & 127.9\\
-		{Moltres without \glspl{DNP}} & 1.004937 \pm 0.000001 & {-}\\
+		{Serpent} & 1.00662 \pm 0.00005 & {-}\\
+		{Moltres with \glspl{DNP}} & 1.0079488 \pm 0.0000010 & 133\\
+		{Moltres without \glspl{DNP}} & 1.0049369 \pm 0.0000010 & {-}\\
 		\bottomrule
@@
 \begin{table}[htb!]
+    \small
 	\centering
-	\caption{$\beta_{\text{eff}}$ values from Serpent and Moltres.}
+	\caption{$k_{\text{eff}}$ values from Serpent and Moltres at various
+	temperatures from 800 K to 1400 K.}
+	\begin{tabular}{S S S S}
+		\toprule
+		{Temperature [K]} & {$k_{\text{eff}}$ $\pm$ 1-$\sigma$ uncertainty
+		(Serpent)} & {$k_{\text{eff}}$ (Moltres)} & {Difference wrt Serpent
+		[pcm]}
+		\\
+		\midrule
+		800  & 1.01996 \pm 0.00005 & 1.02117 & 121 \\
+		900  & 1.01172 \pm 0.00005 & 1.01322 & 150 \\
+		1000 & 1.00428 \pm 0.00005 & 1.00544 & 116 \\
+		1100 & 0.99735 \pm 0.00005 & 0.99859 & 124 \\
+		1200 & 0.99006 \pm 0.00005 & 0.99119 & 113 \\
+		1300 & 0.98356 \pm 0.00005 & 0.98439 &  83 \\
+		1400 & 0.97702 \pm 0.00005 & 0.97820 & 118 \\
+		\bottomrule
+	\end{tabular}
+	\label{table:ktemp}
+\end{table}
+
+The absolute value of $k_{\text{eff}}$ largely impacts the steady-state
+temperature of the reactor. We may adjust $k_{\text{eff}}$ by controlling
+fissile inventory to raise or lower the operating temperatures and meet the
+design specifications for inlet and outlet temperatures. For transient
+analysis, the $\beta_{\text{eff}}$ and reactivity coefficients are arguably
+more important as they dictate the duration and magnitude of the reactor
+response to a transient initiator. The $\beta_{\text{eff}}$ values at 973 K,
+shown in Table \ref{table:betaeff}, are in excellent agreement with a 2.89
+pcm discrepancy.
+
+\begin{table}[htb!]
+	\centering
+	\caption{$\beta_{\text{eff}}$ values from Serpent and Moltres at 973 K.}
 	\begin{tabular}{l S S}
 		\toprule
-		{Code} & {$\beta_{\text{eff}}$ [pcm]} & {Diff. wrt Serpent [pcm]}\\
+		{Code} & {$\beta_{\text{eff}}$ [pcm]} & {Difference wrt Serpent [pcm]}
+		\\
 		\midrule
-		{Serpent} & 303.2 \pm .8 & 0\\
-		{Moltres} & 301.2 \pm .2 & 2.0\\
+		{Serpent} & 304.08 \pm .81 & {-}\\
+		{Moltres} & 301.19 \pm .20 & 2.89\\
 		\bottomrule
@@
 \end{table}
+%
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=.8\textwidth]{reactivity}
+    \caption{Reactivity values from Serpent and Moltres. The Doppler
+    reactivity values were calculated at a fixed density of 4.1249 g
+    cm$^{-3}$. The density reactivity values were calculated at a fixed
+    temperature of 973 K.}
+    \label{fig:reactivity}
+\end{figure}
+%
+\begin{table}[htb!]
+	\centering
+	\caption{Doppler, density, and total temperature coefficients
+	for the temperature range of 800 K to 1400 K.}
+	\begin{tabular}{l S S S S}
+		\toprule
+		{Code} & {$\alpha_D$ (log) [pcm]} & {$\alpha_D$ (linear) [pcm
+		K$^{-1}$]} & {$\alpha_\rho$ [pcm K$^{-1}$]} & {$\alpha_T$ [pcm
+		K$^{-1}$]} \\
+		\midrule
+		{Serpent} & -4034 \pm 14 & 3.737 \pm 0.013 & 3.424 \pm 0.013 & 7.165
+		\pm 0.013 \\
+		{Moltres} & {-} & {-} & {-} & 7.184\\
+		\bottomrule
+	\end{tabular}
+	\label{table:alpha}
+\end{table}
 
-The
+The temperature reactivity feedback arises mainly from Doppler broadening of
+resonance absorption peaks and temperature-induced density changes. Although
+Doppler coefficients typically show logarithmic dependence to temperature, we
+report them, along with the other coefficients, as linear gradient
+values of the reactivity given the relatively linear trend within the relevant
+temperature range (Figure \ref{fig:reactivity}). Table \ref{table:alpha} shows
+the temperature coefficients as described prior. The total temperature
+coefficients from Serpent and Moltres show excellent agreement with a
+discrepancy of 0.019 pcm K$^{-1}$.
 
-\section{Steady-State Behavior}
+Moltres also reproduced the six-group neutron spectrum very well from the
+Serpent group constants. Figure \ref{fig:ntspec} compares the energy spectra
+from Serpent and Moltres in the central fuel salt region. The six-group
+neutron spectra overlap exactly over each other. More generally, the plot
+shows the distinctive fast spectrum observed in the \gls{MSFR} with dips in
+the spectrum corresponding to elastic scattering resonances from lithium and
+fluorine. From this plot, we observe that the discrepancies in
+$k_{\text{eff}}$ arise mainly from discretizing neutron energy into groups
+rather than the neutron diffusion model itself. We could obtain a more
+accurate representation of the neutronics in the \gls{MSFR} by using more
+neutron energy groups but this would adversely impact simulation times in the
+subsequent multiphysics finite element analyses.
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=.8\textwidth]{nt-spec}
+    \caption{Reactivity values from Serpent and Moltres. The Doppler
+    reactivity values were calculated at a fixed density of 4.1249 g
+    cm$^{-3}$. The density reactivity values were calculated at a fixed
+    temperature of 973 K.}
+    \label{fig:ntspec}
+\end{figure}
+
+In summary, Moltres can replicate most of the relevant neutronics parameters
+accurately with the group constant data from Serpent. While the
+$k_{\text{eff}}$ values have discrepancies on the order of 100 pcm, they are
+relatively small compared to results from other codes. Furthermore, the
+$\beta_{\text{eff}}$ and temperature coefficients, which are important
+parameters for modeling transient reactor behavior, are in excellent
+agreement.
+
+\section{Steady-State Results}
+
+With the verification of Moltres' neutronics modeling capabilities in the
+context of the \gls{MSFR}, we now move on to the multiphysics simulation
+results. This section covers the steady-state multiphysics results from
+Moltres.
+
+The procedure for obtaining the steady-state results involved several steps
+due to the highly coupled \glspl{PDE}. First, we ran a preliminary transient
+simulation of fluid flow in the gls\MSFR} core, starting from zero inlet
+velocity and gradually ramping it up to match the nominal flow rate (4.5 m$^3$
+s$^{-1}$); otherwise Moltres had difficulty converging to the desired fully
+developed flow profile. We imposed a parabolic flow profile at the inlet.
+Next, we imported these fully developed flow values as initial values for
+velocity in the actual transient simulation modeling the full coupled
+neutronics and thermal-hydraulics. The initial values for the temperature and
+neutron group flux distributions are 953 K and $1 \times 10^{14}$ cm$^{-2}$
+s$^{-1}$ uniformly throughout the geometry. Finally, we assume that steady
+state is reached when the volume integral values of every variable remain
+constant (up to 6 sig. fig.) for at least four seconds in the simulation; this
+time period corresponds to the nominal circulation time of the \gls{MSFR}.
+
+For a direct comparison with the steady-state results from the Polimi/TUDelft
+models, the bulk of our steady-state results do not have decay heat modeling.
+Instead, we separately discuss the minor differences borne from decay heat
+modeling in the last subsection.
+
+\subsection{Thermal-Hydraulics}
+
+Figure \ref{fig:flow-temp} shows the temperature and velocity fields
+of the fuel salt in the core at steady state from Moltres and the
+Polimi/TUDelft models. The results from Moltres show good qualitative
+agreement with the Polimi/TUDelft models; we observe similar flow and hotspot
+features in all three models. There is a large recirculation region near the
+blanket tank walls arising from turbulent flow. Inertial forces dominate over
+viscous forces to form this large eddy. The figure also shows two relatively
+stagnant regions along the central axis at the top and bottom of the core.
+Temperature hotspots form in these regions of recirculation as convection is
+the dominant heat transfer mechanism. The maximum temperature from Moltres,
+1279 K near the bottom of the recirculation zone, is closer to the maximum
+temperature in the Polimi model ($\approx$ 1300 K) than the TUDelft model
+($\approx$ 1200K). The minimum temperature is 922 K at the inlet. Figure
+\ref{fig:flow} provides an alternate view of the flow profile through flow
+streamlines superimposed on the velocity magnitude distribution. The fastest
+salt flow occurs at the inlet, outlet, and at core half-height approximately
+0.40 m away from the central axis.
+
+\begin{figure}[htb!]
+    \centering
+    \begin{subfigure}[t]{.37\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{flow-temp}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.625\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{flow-temp-fiorina}
+    \end{subfigure}
+    \caption{Temperature and velocity fields in the core from Moltres
+    (left), Polimi (center), and TUDelft (right) models.}
+    \label{fig:flow-temp}
+\end{figure}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=.6\textwidth]{flow}
+    \caption{Fuel salt flow streamlines and velocity magnitude in the core.}
+    \label{fig:flow}
+\end{figure}
+
+While the temperatures at the hotspots are well below the melting point of the
+Ni-alloy structure (1500 K), they may cause undue thermal stress on the
+blanket tank structure and induce relatively faster salt corrosion rates. A
+sudden, large reactivity insertion could push fuel salt temperatures above the
+melting point of the Ni-alloy and cause irreversible damage. Furthermore,
+the reservoir of hot fuel salt may cause unpredictable behavior during
+transient scenarios when the flow profile undergoes a drastic change.
+Thus, Rouch et al. \cite{rouch_preliminary_2014} developed an improved
+hourglass-shaped design to optimize flow distribution and prevent these
+recirculation zones and hotspots from forming. A study of this new design
+using Moltres is a potential subject for future work when a proper turbulence
+model is in place.
 
@@
 
-\subsection{Thermal-Hydraulics}
+\subsection{Decay Heat}
 
-\section{Transient Behavior}
+A decay heat model effectively redistributes a fraction of the heat source
+from the center of the core to the entire loop. Thus, we expect to observe
+a slight flattening of the temperature distribution across the entire primary
+loop.
+
+\section{Transient Results}
+
+As noted by Fiorina et al. \cite{fiorina_modelling_2014}, explicit decay heat
+modeling has a negligible effect on reactivity-, pump-, and enchanced
+cooling-driven transients. The Polimi model presented results with the decay
+heat model for the loss of heat sink transient scenario only. For a fair
+comparison, we also
 

### 2020-04-11 model.tex
combine msr and msfr chapters, expand modeling chapter

@@
 requires a dedicated neutron transport solver such as Serpent 2 to generate
-neutron energy group constant data for the deterministic, multigroup
-neutronics calculations in Moltres. This chapter provides brief descriptions
-of the Serpent 2 code and Moltres.
+neutron energy group constant data for the multigroup neutron diffusion
+calculations in Moltres. This chapter provides brief introductions of the
+Serpent and Moltres codes, and the modeling approach for the multiphysics
+simulations.
 
@@
 
-\subsection{Neutronics}
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=.5\textwidth]{central-core}
+    \caption{2-D axisymmetric model of the MSFR. The red box indicates the
+    central core region relevant to the modeling approach in Moltres.}
+    \label{fig:core}
+\end{figure}
 
-Focusing our attention on Moltres, the neutron flux calculations are
+As mentioned in the previous chapter, we divided the fuel salt loop into two
+regions, the central core region where most of the fissions take place, and
+the outer loop region where the heat exchanger is located. The central core
+region specifically refers to the central region indicated by
+the red box in Figure \ref{fig:core}. We simplified the outer loop
+into a 1-D pipe as it is a subcritical region and its main purposes are to
+introduce a out-of-core residence time for the \glspl{DNP} and to cool the
+fuel salt before redirecting it back into the central core region.
+Accordingly, this section provides separate descriptions for the governing
+equations in the central core region and the outer loop region. There is also
+a third region comprising of the blanket salt, the Ni-alloy reflectors, 
+and other material . This region is important for capturing an accurate
+estimate of neutron leakage as opposed to simpler approximations such as
+imposing fixed albedo boundary conditions on the neutron fluxes.
+
+\subsection{Central Core Region}
+
+The central core region is of greatest interest to us during steady state and
+transient scenarios; the center of the reactor is naturally where most of the
+fissions and heat generation occur. Consequently, the \glspl{PDE} solved for
+in this region are in their most complete form relatively to simplifying
+assumptions found in the other regions.
+
+\subsubsection{Neutronics}
+
+The neutron flux calculations in the central core region are
 performed using the standard formulations for the time-dependent multigroup
@@
 \gls{MSFR} contains no material interfaces in the middle of the core where
-most of the fission interactions take place. The Results section provides a
+most of the fission interactions take place. The Results* section provides a
 comparison of the \gls{MSFR} multiplication factor values and reactivity
@@
 
-\subsection{Thermal-Hydraulics}
+\subsubsection{Thermal-Hydraulics}
 
@@
 turbulent diffusion.
+
+\subsection{Outer Loop Region}
+
+\subsubsection{Neutronics}
+
+\subsubsection{Thermal-Hydraulics}

### 2020-04-11 msr.tex
combine msr and msfr chapters, expand modeling chapter

@@
 recent years due to the growing interest from commercial \gls{MSR} developers.
+
+\section{Molten Salt Fast Reactor}
+
+The \gls{MSFR} is a reference fast-spectrum \gls{MSR} concept developed
+under the \gls{EVOL} and \gls{SAMOFAR} projects. The main reactor
+specifications and schematic view are shown in Table \ref{table:msfr} and
+Figure \ref{fig:msfr}, respectively. Developed from the \gls{MSBR}, the
+\gls{MSFR} is intended to run primarily on a closed thorium fuel cycle with
+continuous online fuel reprocessing. Several reasons motivated the omission of
+graphite moderators from the \gls{MSFR} design. Graphite is susceptible to
+long-term radiation damage and replacement is likely to be necessary during
+the operating lifetime of the reactor. Graphite also has a positive
+temperature coefficient of reactivity; eliminating graphite from the design
+ensures a greater safety margin \cite{mathieu_thorium_2006}. While negative
+temperature coefficients are attainable with very thermalized spectra,
+breeding ratios deteriorated significantly due to parasitic absorption in the
+large volume of graphite needed for thermalization
+\cite{mathieu_thorium_2006}.
+%
+\begin{figure}[htb!] 
+	\centering
+	\includegraphics[width=0.6\textwidth]{MSFR}
+	\caption{Schematic view of the MSFR concept \cite{serp_molten_2014}.}
+	\label{fig:msfr}
+\end{figure}
+%
+\begin{table}[htb!]
+    \small
+	\caption{Main specifications of the \gls{MSFR} concept
+				\cite{serp_molten_2014}.}
+	\centering
+	\begin{tabular}{ l r }
+		\hline
+		Parameter & Value \\
+		\hline
+		Thermal/Electric output [MW$_{\text{th}}$/MW$_{\text{e}}$] & 3000 /
+		1500 
+		\\
+		Salt volume [m$^3$] & 18 \\
+		Salt fraction in core & 0.5 \\
+		Number of circulation loops & 16 \\
+		Nominal flow rate [kg s$^{-1}$] & 18500  \\
+		Nominal circulation time [s] & 4.0 \\
+		Inlet/outlet temperature [K] & 923 / 1023 \\
+		Blanket volume [m$^3$] & 7.3\\
+		\hline
+	\end{tabular}
+	\label{table:msfr}
+\end{table}
+
+In the \gls{MSFR} design, fuel salt flows upwards through a 9 m$^3$ central
+core region. At the top of the core, the flow separates into sixteen smaller
+external loops, each of which passes through a heat exchanger before being
+pumped back into the bottom of the core. Other instrumentation are situated
+along the external loop for online salt reprocessing and gas sparging. The
+core is surrounded axially by nickel alloy reflectors, and radially by a
+toroidal blanket tank containing fertile salt for breeding. There is a layer
+of boron carbide behind the blanket tanks to protect the peripheral equipment
+from excessive neutron damage. In case of severe accidents, there is an
+actively cooled freeze plug at the bottom of the core that melts when
+temperatures exceed a certain threshold. The fuel salt would drain into a
+containment vessel designed to keep it subcritical. Reactivity control under
+normal operating conditions is performed by varying pump speeds to take
+advantage of the strong thermal feedback. Coupled with the fact that there is
+no excess reactivity reserve due to online fuel reprocessing, there are no
+control rods in the \gls{MSFR} design.
+
+Although the \gls{MSFR} is primarily designed to operate on the thorium fuel
+cycle, it can support a range of start-up fuel and feed compositions. This
+versatility is particularly important for the first few \glspl{MSFR} to be
+deployed due to the lack of $^{233}$U reserves required for the initial core
+loading. In general, the fuel and blanket salts are approximately composed of
+eutectic mixtures of 77.5\% LiF - 22.5\% AcF$_4$, where AcF$_4$ represents
+actinide fluorides such as uranium, thorium, plutonium, and other \gls{TRU}
+fluorides. For an initial composition consisting of $^{232}$Th and $^{233}$U,
+the benchmark value for the amount of uranium for criticality under
+normal operating conditions is 2.515 mol\%. However, most code verification
+studies adjust the ratio of $^{232}$Th to $^{233}$U to achieve exact
+criticality at a uniform temperature of 973 K; this ensures that subsequent
+neutronics and safety analyses are not affected by the difference in
+k$_{\text{eff}}$ values. We performed the same exercise in this thesis.
+
+Power output of the \gls{MSFR} is rated at 3000 MW$_{\text{th}}$ and 1500
+MW$_{\text{e}}$. It has a high thermal efficiency due to the high operating
+temperature. \glspl{MSR} in general are not restricted by the same pressure
+constraints seen in \glspl{LWR}. The inlet and outlet temperature
+specifications of the fuel salt are 923 K and 1023 K, respectively. This was
+motivated by the need for a minimum 50 K temperature buffer between the
+operating temperatures and the melting point of the salt. The \gls{MSFR} has
+heat exchangers and an intermediate coolant loop to isolate the power
+conversion system from the highly radioactive fuel salt. This also serves as
+a layer of containment between the radioactive material and the outside
+environment. The exact composition of the intermediate coolant is still under
+active study and not finalized yet.
+
+\subsection{Model Geometry}
+
+For this work, we used a model similar to the reference square-cylindrical
+\gls{MSFR} design to benchmark our results against results published by
+Fiorina et al. and Aufiero et al. The reference design is a 2-D axisymmetric
+model with the sixteen individual external loops homogenized into a single
+outer loop as shown in Figure \ref{fig:msfrgeom}. For the multigroup
+cross sections and group constants calculations in Serpent, we extended this
+2-D axisymmetric model into a 3-D model by a simple full rotation about the
+central axis. The material definitions are the same as those specified in the
+reference \gls{MSFR} model. Accordingly, the pump and heat exchanger regions
+are assumed to be composed of 100\% fuel salt. While this may not be entirely
+accurate, the exact details of the pump and heat exchanger systems are still
+under active study, and this external loop region is presumed to be of little
+neutronic importance due to its position behind the strong boron carbide
+neutron absorber layer.
+%
+\begin{figure}[t!] 
+	\centering
+	\includegraphics[width=0.5\textwidth]{msfr-geom}
+	\caption{2-D axisymmetric model of the \gls{MSFR} core used for the
+	simulations in Serpent. All dimensions are in meters.
+	\cite{brovchenko_neutronic_2019}}
+	\label{fig:msfrgeom}
+\end{figure}
+
+Although we used the exact reference model for generating group constant data
+from Serpent, there are two minor differences between the \gls{MSFR} model
+geometry we used in Moltres, and the Polimi/TUDelft models. The
+first difference is a relative
+minor change to the mesh by the exclusion of the 2 cm thick structural
+material around the blanket tank that separates the fuel and blanket salts.
+We removed this feature in our finite element mesh as we encountered
+difficulty meshing this layer that is relatively much thinner than the rest of
+the model. Any impact on the
+neutron flux is expected to be minimal. Furthermore, we solved for the
+temperature distribution only in the primary loop and applied homogeneous
+Neumann boundary conditions for temperature on the core walls, as was done in
+the Polimi/TUDelft models. Therefore, we believe the overall impact on the
+results is negligible.
+
+The second difference pertains to the modeling of the external loop. In its
+current implementation, Moltres lacks sophisticated pump- and heat exchanger-
+equivalents in the code. Thus, the external loop, beyond the central region of
+the primary loop where most of the fissions take place, is modeled as a 1-D
+pipe with a simple, point heat sink to represent the heat exchanger. Instead
+of pumps, a Dirichlet boundary condition for velocity at the inlet boundaries
+drives the flow in the central core region of the primary loop. All
+flow-dependent variables such as temperature, \glspl{DNP},
+and decay heat precursors are fully conserved as they loop around between the
+two regions. As a result, this approach shares some similarities with the
+geometric multiscale modeling approach by Zanetti et al.
+\cite{zanetti_geometric_2015}. Future models could create a better
+representation of the primary loop by implementing a whole continuous loop
+with pressure increases and drops corresponding to the pumps and heat
+exchangers.
+
+\subsection{Material Specifications}
+
+This section details the material specifications of the various reactor
+components in the \gls{MSFR}.
+
+\subsubsection{Molten Salt}
+
+As mentioned before, the fuel and blanket salts are comprised of a 77.5\% LiF
+- 22.5\% AcF$_4$ mixture. This is the reference salt composition of the
+\gls{MSFR} at start-up. Typically, researchers working with the \gls{MSFR}
+model calculate the exact actinide composition by varying the $^{232}$Th to
+$^{233}$U ratio to obtain a $k_{\text{eff}}$ value of 1 at a uniform
+temperature of 973 K. Thus, the exact actinide compositions vary depending on
+the nuclear data library and neutron transport code. The exact composition
+used for this work can be found in the Results chapter. We assume that the
+effects of these composition variations on the reference physical properties
+of the fuel and blanket salt are negligible. Table \ref{table:prip} shows
+relevant physical properties of the fuel and blanket salts.
+%
+\begin{table}[htb!]
+\small
+\centering
+\caption{Properties of the fuel and blanket salts LiF-AcF$_4$.}
+\begin{tabular}{l l r r}
+\toprule
+Property & Formula & {Value at 973 K} & Validity Range\\
+\midrule
+Melting temperature [K] & 841 & {N/A} & 1 bar \\
+Boiling temperature [K] & 1874 & {N/A} & 1 bar \\
+Density, $\rho$ [kg m$^{-3}$] & $4094-0.882 \cdot (T-1008)$ & $4125$ & 893-1123 K \\
+Dynamic viscosity, $\mu$ [Pa s] & $\rho \cdot 5.55 \times 10^{-8} \cdot e^{3689/T}$ & $1.015$ & 898-1121 K \\
+Thermal conductivity, $k$ [W m$^{-1}$ K$^{-1}$] & $0.928+8.397 \times 10^{-5} \cdot T$ & $0.01010$ & 891-1020 K \\
+Specific heat, $c_p$ [J kg$^{-1}$ K$^{-1}$] & $-1111+2.78\cdot T$ & $1594$ & 867-907 K \\
+\bottomrule
+\end{tabular}
+\label{table:prip}
+\end{table}
+
+\subsubsection{Structural Materials}
+
+The reflectors on the periphery of the reactor core and the blanket tank are
+made of a nickel-based alloy provided from reference specifications
+\cite{delpech_reactor_2009}. Table \ref{table:refl} details the compositions
+of the nickel-based alloy and boron carbide absorber. The alloy has a density
+of 10 g cm$^{-3}$. The material composition of the reflectors may be subject
+to minor changes, but it is not a major concern as the reflectors are not
+situated in regions of high neutron flux. The \gls{MSFR} also includes a 20 cm
+layer of boron carbide (B$_4$C) to protect the heat exchangers and pumps from
+neutron irradiation. The reference specifications indicate that natural boron
+is used, which is composed of 19.8 \% $^{10}$B and 80.2 \% $^{11}$B, with an
+overall density of 2.52 g cm$^{-3}$. 
+%
+\begin{table}[htb!]
+\footnotesize
+\centering
+\caption{Composition (mol \%) of the nickel-based alloy used in the simulation
+of the \gls{MSFR} in this work.}
+\begin{tabular}{l l l l l l l l l l l l l}
+\toprule
+Ni & W & Cr & Mo & Fe & Ti & C & Mn & Si & Al & B & P & S \\
+\midrule
+79.432 & 9.976 & 8.014 & 0.736 & 0.632 & 0.295 & 0.294 & 0.257 & 0.252 & 0.052 & 0.033 & 0.023 & 0.004 \\
+\bottomrule
+\end{tabular}
+\label{table:refl}
+\end{table}

### 2020-04-11 results.tex
combine msr and msfr chapters, expand modeling chapter

@@
 due to the highly coupled \glspl{PDE}. First, we ran a preliminary transient
-simulation of fluid flow in the gls\MSFR} core, starting from zero inlet
+simulation of fluid flow in the \gls{MSFR} core, starting from zero inlet
 velocity and gradually ramping it up to match the nominal flow rate (4.5 m$^3$
@@
 
+\subsection{Unprotected Pump Overspeed}
+
 \subsection{Unprotected Loss of Heat Sink}
 
-\subsection{Chilled Inlet}
-
-\subsection{Pump Overspeed}
+\subsection{Unprotected Chilled Inlet}

### 2020-04-13 introduction.tex
Completed first draft of Chapters 1-4 for review

@@
 
-In this thesis, we demonstrate the coupled multiphysics modeling capabilities
-of Moltres through the reference configuration of the MSFR design. Moltres
-employs fully implicit coupling between the neutronics and thermal-hydraulics
-governing equations, to fully account for the tight coupling expected in MSRs
-due to the movement of fuel in the salt. We implemented flow using the
-Navier-Stokes equations for an accurate representation of the flow profile as
-opposed to imposing a predetermined velocity field.
+Moltres employs fully implicit coupling between the neutronics and
+thermal-hydraulics governing equations, to fully account for the tightly
+coupled physics expected in MSRs due to the movement of fuel in the salt. We
+implemented flow using the Navier-Stokes equations and zeroth-order
+approximation of eddy viscosity for a more accurate representation of the flow
+profile compared to imposing a predetermined velocity field.
 
-The objectives of this thesis are: 1) study the steady-state operating
-conditions with the reference \gls{MSFR} model, 2) study the transient
-behavior of the \gls{MSFR} under several unprotected accident scenarios, and
-identify general characteristic behaviors in MSRs and fast spectrum reactors.
-The simulation results are verified against the results by Fiorina et al.
+The main objective of this thesis is to demonstrate Moltres' capabilities in
+modeling multiphysics, steady-state and transient behavior of fast-spectrum
+\glspl{MSR} through the study of the \gls{MSFR} concept. This is achieved by
+first verifying Moltres' neutronics results against Serpent in the context of
+the \gls{MSFR}, and then comparing the coupled neutronics/thermal-hydraulics
+steady-state and transient accident results of the \gls{MSFR} concept. The
+multiphysics results are verified against the results by Fiorina et al.
 \cite{fiorina_modelling_2014} and Aufiero et al.
@@
 
-** insert thesis outline here **
+\section{Thesis Outline}
+
+The outline of this thesis is as follows: Chapter 2 discusses the history and
+features of \glspl{MSR}. The \gls{MSFR} concept is covered in greater detail.
+Chapter 3 details the simulation codes and the general modeling approach for
+generating the results in this thesis. Chapter 4 provides a neutronics
+assessment by comparing key neutronics parameters from Moltres' eigenvalue
+calculations to Serpent's Monte Carlo calculations. Chapter 5 presents
+steady-state results of coupled neutronics/thermal-hydraulics \gls{MSFR}
+simulations in Moltres. Chapters 6 to 9 present transient accident simulation
+results for unprotected reactivity insertions, unprotected loss of flow,
+unprotected pump overspeed, and unprotected loss of heat sink, respectively.
+Lastly, Chapter 10 summarizes the key findings of this thesis and posits some
+potential avenues for future work.

### 2020-04-13 model.tex
Completed first draft of Chapters 1-4 for review

@@
 calculations in Moltres. This chapter provides brief introductions of the
-Serpent and Moltres codes, and the modeling approach for the multiphysics
-simulations.
+Serpent and Moltres codes, and the general modeling approach for the
+multiphysics simulations in Moltres.
 
@@
 the red box in Figure \ref{fig:core}. We simplified the outer loop
-into a 1-D pipe as it is a subcritical region and its main purposes are to
-introduce a out-of-core residence time for the \glspl{DNP} and to cool the
-fuel salt before redirecting it back into the central core region.
+into a 1-D pipe as it is a subcritical region and its main purposes are: to
+introduce a out-of-core residence time for the \glspl{DNP}, and to site the
+heat exchanger-equivalent object in Moltres.
 Accordingly, this section provides separate descriptions for the governing
@@
 
-\subsubsection{Neutronics}
+\subsubsection{Neutronics Model}
 
@@
 boundaries, are well-documented, the diffusion model provides acceptable
-accuracy at lower computational costs. Moreover, the
-\gls{MSFR} contains no material interfaces in the middle of the core where
-most of the fission interactions take place. The Results* section provides a
-comparison of the \gls{MSFR} multiplication factor values and reactivity
-coefficients between Moltres and Serpent.
+accuracy at lower computational costs. Moreover, the central core region
+contains no material interfaces except at its boundaries. The Neutronics
+results section provides a comparison of the \gls{MSFR} multiplication factor
+values and reactivity coefficients between Moltres and Serpent.
 
@@
 We assumed $Sc_t = 0.85$ for a fair comparison with the Polimi/TUDelft
-models which used the same value. Moltres also accounts for the decay of
-\glspl{DNP} outside the active core region by simulating its flow in a
-separate 1-D pipe geometry. This outer loop pipe simulation is loosely coupled
-to the active core simulation through MOOSE's MultiApp functionality and
-inlet/outlet boundary values. The loose coupling is justified as the evolution
-of \gls{DNP} concentrations due to inflow is relatively slow. The length and
-velocity defined in this geometry may differ according to reactor
-specifications. For this work with the \gls{MSFR} model, we assumed a pipe
-length of 2.255 m with salt flowing at 1.1275 m s$^{-1}$ for an out-of-core
-residence time of 2 seconds to follow the design specifications.
+models which used the same value.
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=.8\textwidth]{mesh}
+    \caption{Mesh adopted in Moltres and a close-up view of the mesh around
+    the boron carbide absorber.}
+    \label{fig:mesh}
+\end{figure}
 
@@
 interpolation method of choice.
-
-For this work, we have six neutron energy groups according to the energy
-boundaries in table \ref{table:bound}, eight \gls{DNP} groups as defined by
-the JEFF-3.1.2 library, and the spline interpolation method. The neutron flux
-and \gls{DNP} concentration values were approximated by first-order Lagrange
-and constant monomial shape functions respectively on the finite element mesh.
-Figure * shows the mesh adopted for the \gls{MSFR} model.
-We assumed vacuum boundary conditions for all six neutron group fluxes along
-the external boundaries of the geometry, and homogeneous Neumann boundary
-conditions along the axial symmetry boundary. For the \gls{DNP}
-concentrations, we imposed homogeneous Neumann boundary conditions on the
-walls, and inflow and outflow boundary conditions on the inlet and outlet
-boundaries respectively. The inlet \gls{DNP} concentration values were
-imported from the outlet values of the 1-D outer loop pipe at the same
-timestep.
 
@@
 
+For this work, we have six neutron energy groups according to the energy
+boundaries in table \ref{table:bound}, eight \gls{DNP} groups as defined by
+the JEFF-3.1.2 library, and the spline interpolation method. The neutron flux
+and \gls{DNP} concentration values were approximated by first-order Lagrange
+and constant monomial shape functions respectively on the finite element mesh.
+Figure \ref{fig:mesh} shows the mesh adopted for the \gls{MSFR} model.
+We assumed vacuum boundary conditions for all six neutron group fluxes along
+the external boundaries of the geometry, and homogeneous Neumann boundary
+conditions along the axial symmetry boundary. For the \gls{DNP}
+concentrations, we imposed homogeneous Neumann boundary conditions on the
+walls, and inflow and outflow boundary conditions on the inlet and outlet
+boundaries respectively. The inlet \gls{DNP} concentration values were
+imported from the outlet values of the 1-D outer loop pipe at the same
+timestep.
+
 For the decay heat model, a previous study on the MSFR by Aufiero et al.
-\cite{aufiero_extended_2013} has shown that using three decay heat precursor
+\cite{aufiero_extended_2013} had shown that using three decay heat precursor
 groups with different half-lives in the form of exponential equations, can
@@
 
-\subsubsection{Thermal-Hydraulics}
+\subsubsection{Thermal-Hydraulics Model}
 
@@
 
-The flow rate is dictated by an inflow boundary condition at the core inlet
-for a volumetric flow rate of 4.5 m$^3$ s$^{-1}$. We imposed no-slip boundary
-conditions on the walls of the core, and homogeneous Neumann boundary
-conditions on the core outlet. The energy balance equation for temperature is
+The energy balance equation for temperature is
 given in Eq. \ref{eq:temp}. The diffusion term includes turbulent heat
@@
     \rho c_{p} \frac{\partial T}{\partial t} &= - \rho c_p \boldsymbol{u}
-    \cdot \nabla T + \nabla \cdot [(k + k_t) \nabla T] + Q_s - Q_{hx}
+    \cdot \nabla T + \nabla \cdot [(k + k_t) \nabla T] + Q_s
     \label{eq:temp} \\
@@
     \\
-    k &= \text{thermal conductivity of molten salt [W m$^{-1}$ K$^{-1}$,}
+    k &= \text{thermal conductivity of molten salt [W m$^{-1}$ K$^{-1}$],}
     \nonumber \\
-    Q_{hx} &= \text{heat loss rate through the heat exchanger [W].} \nonumber
+    K &= \text{total number of decay heat groups.} \nonumber
 \end{align}
 %
-where $K$ is the total number of decay heat groups. The first term in the heat
-source expression represents prompt fission heat, and the second term
-represents decay heat from the $K$ decay heat groups.
-
-We modeled the heat exchanger using a point source/sink at the center of the
-1-D outer loop pipe region. The rate of heat extraction is equivalent to the
-nominal 3 GW$_{\text{th}}$ power output with appropriate scaling to account
-for the 1-D geometry. *** include more detail when finalized ***
+The first term in the heat source $Q_s$ represents prompt fission heat, and
+the second term represents decay heat from the $K$ decay heat groups.
 
@@
 
-\subsubsection{Neutronics}
+Moltres also accounts for the decay of
+\glspl{DNP} outside the central core region by simulating its flow in a
+separate 1-D pipe geometry. This outer loop pipe simulation is implicitly
+coupled to the active core simulation through Picard iterations in MOOSE's
+MultiApp functionality and inlet/outlet boundary values. For this work with
+the \gls{MSFR} model, we assumed a pipe
+length of 2.255 m with salt flowing at 1.1275 m s$^{-1}$ for an average
+out-of-core residence time of 2 seconds to follow the design specifications.
 
-\subsubsection{Thermal-Hydraulics}
+\subsubsection{Neutronics Model}
+
+Since this region is largely subcritical, the only significant
+neutronics-related phenomena are the drift, and decay of \glspl{DNP}.
+The governing equation for the \glspl{DNP} is:
+%
+\begin{align}
+    \frac{\partial C_i}{\partial t} &= - \lambda_i C_i - u
+    \frac{\partial C_i}{\partial x} + D_t \frac{\partial^2 C_i}{\partial x^2}.
+    \label{eq:dnploop}
+\end{align}
+%
+Equation \ref{eq:dnploop} is derived from equation \ref{eq:dnp} by removing
+the fission \gls{DNP} source term, and the conversion of the advection and
+diffusion terms to their 1-D forms. The decay constants and diffusion
+coefficient are the same values used in the central core region.
+
+\subsubsection{Thermal-Hydraulics Model}
+
+Instead of the Navier-Stokes equations, we have a predetermined velocity of
+1.1275 m s$^{-1}$ in the outer loop region. The governing equation for
+temperature, derived from equation \ref{eq:temp}, is:
+%
+\begin{align}
+    \rho c_{p} \frac{\partial T}{\partial t} &= - \rho c_p u
+    \frac{\partial T}{\partial x} - Q_{hx} \label{eq:temploop} \\
+    Q_{hx} &= \alpha (T - T_i) \delta (x_0) \label{eq:hx} \\
+    \intertext{where}
+    \alpha &= \text{heat transfer coefficient [W K$^{-1}$],} \nonumber \\
+    x_0 &= \text{position of the point heat exchanger [m].} \nonumber
+\end{align}
+
+The fission heat source term is replaced with a heat exchanger sink term
+which depends on the temperature difference between the fuel salt $T$ and the
+secondary loop salt $T_s$. For simplicity, we assumed a constant temperature
+of 823 K in the secondary loop. The heat transfer coefficient was determined
+by assuming that the fuel outlet temperature is 1023 K and calculating the
+heat removal rate to induce a 100 K drop at the given volumetric flow rate
+and heat capacity of the fuel salt. We opted to ignore the diffusion term due
+to the discontinuity of the temperature distribution across the point heat
+exchanger.
+
+\subsection{Boundary Conditions and Flow Transfers}
+
+This subsection details the various boundary conditions for all of the
+tracked variables, and the \gls{DNP} and temperature flow transfers between
+the central core and outer loop regions.
+
+Starting with the central core region, for the neutron group fluxes, we
+imposed vacuum boundary conditions on the
+outermost boundaries of the geometry in Figure \ref{fig:core} excluding the
+axial boundary. The \gls{DNP} variables have homogeneous Neumann boundary
+conditions along the axis and the walls in the central core region, and inflow
+and outflow boundary conditions on the inlet and outlet boundaries,
+respectively. The temperature variable shares the same type of boundary
+conditions as the \gls{DNP} variables.
+
+The flow rate is dictated by an inflow boundary condition at the core inlet
+for a volumetric flow rate of 4.5 m$^3$ s$^{-1}$. We imposed no-slip boundary
+conditions on the walls of the central core, and homogeneous Neumann boundary
+conditions on the outlet. Along the axial boundary, we have homogeneous
+Dirichlet boundary condition for the radial velocity component, and
+homogeneous Neumann boundary condition for the axial velocity component.
+
+At every timestep, Moltres also calculates weighted averages of the
+temperature and \gls{DNP} at the outlet. These values are weighted by the
+outflow velocity values at the outlet according to the following equation:
+%
+\begin{align}
+    \overline{\phi} &= \frac{\int_\mathcal{C} \phi(y) u(y) dy}{
+    \int_\mathcal{C} u(y) dy} \\
+    \intertext{where}
+    \phi &= \text{variable to be weighted} \nonumber \\
+    \mathcal{C} &= \text{outlet boundary curve} \nonumber \\
+    u &= \text{outflow velocity perpendicular to the outlet boundary [m
+    s$^{-1}$].} \nonumber
+\end{align}
+
+This outflow value from the central core region is transferred to the 1-D
+outer loop region as input for the inhomogeneous Dirichlet boundary
+condition at the inlet boundary. Likewise, the outflow value from the outer
+loop region is used for the inflow value in the central core region. No
+averaging is required for this step as the outer loop region is a 1-D system.
+We assume that the inflow temperature and \gls{DNP} are uniform at the inlet.
+The Picard iterations within every timestep ensure that the two systems are
+implicitly coupled even though they're solved separately.

### 2020-04-13 neutronics.tex
Completed first draft of Chapters 1-4 for review

--- 2020-03-30 results.tex
@@
-This chapter discusses three main sets of simulation results. First, we
-compare the key neutronics results between Serpent and Moltres for a static
-model of the \gls{MSFR}, i.e. no salt flow, to assess the accuracy of the
+In this chapter, we compare key neutronics results between Serpent and Moltres
+for a static model of the \gls{MSFR}, i.e. no salt flow, and uniform
+temperature distribution to assess the accuracy of the
 six-group neutron diffusion model in Moltres on a fast-spectrum reactor. This
-exercise also serves as a checking exercise for the group constant data
-generated for Moltres from Serpent. Second, we present steady-state results
-from Moltres with salt flow and verify it against published results, namely
-from the Polimi/TUDelft and OpenFOAM models developed by Fiorina et al.
-\cite{fiorina_modelling_2014} and Aufiero et al.
-\cite{aufiero_development_2014}, respectively. Lastly, we present and compare
-transient simulation results from Moltres with the aforementioned models.
-Transient accident scenarios include an unprotected reactivity insertion, an
-unprotected loss of flow, an unprotected loss of heat sink, a chilled inlet,
-and a pump overspeed scenario.
+exercise also serves as a verification for the group constant data
+generated for Moltres from Serpent.
 
-\section{Static Model}
+\section{Effective Multiplication Factor and Delayed Neutron Fraction}
 
@@
 \end{table}
-%
+
+\section{Reactivity Feedback Coefficients}
+
+The temperature reactivity feedback arises mainly from Doppler broadening of
+resonance absorption peaks and temperature-induced density changes. Although
+Doppler coefficients typically show logarithmic dependence to temperature, we
+report them, along with the other coefficients, as linear gradient
+values of the reactivity given the relatively linear trend within the relevant
+temperature range (Figure \ref{fig:reactivity}). Table \ref{table:alpha} shows
+the temperature coefficients as described prior. The total temperature
+coefficients from Serpent and Moltres show excellent agreement with a
+discrepancy of 0.019 pcm K$^{-1}$.
+
 \begin{figure}[htb!]
@@
 
-The temperature reactivity feedback arises mainly from Doppler broadening of
-resonance absorption peaks and temperature-induced density changes. Although
-Doppler coefficients typically show logarithmic dependence to temperature, we
-report them, along with the other coefficients, as linear gradient
-values of the reactivity given the relatively linear trend within the relevant
-temperature range (Figure \ref{fig:reactivity}). Table \ref{table:alpha} shows
-the temperature coefficients as described prior. The total temperature
-coefficients from Serpent and Moltres show excellent agreement with a
-discrepancy of 0.019 pcm K$^{-1}$.
+\section{Neutron Energy Spectrum}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=.8\textwidth]{nt-spec}
+    \caption{Reactivity values from Serpent and Moltres. The Doppler
+    reactivity values were calculated at a fixed density of 4.1249 g
+    cm$^{-3}$. The density reactivity values were calculated at a fixed
+    temperature of 973 K.}
+    \label{fig:ntspec}
+\end{figure}
 
@@
 
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=.8\textwidth]{nt-spec}
-    \caption{Reactivity values from Serpent and Moltres. The Doppler
-    reactivity values were calculated at a fixed density of 4.1249 g
-    cm$^{-3}$. The density reactivity values were calculated at a fixed
-    temperature of 973 K.}
-    \label{fig:ntspec}
-\end{figure}
-
 In summary, Moltres can replicate most of the relevant neutronics parameters
@@
 agreement.
-
-\section{Steady-State Results}
-
-With the verification of Moltres' neutronics modeling capabilities in the
-context of the \gls{MSFR}, we now move on to the multiphysics simulation
-results. This section covers the steady-state multiphysics results from
-Moltres.
-
-The procedure for obtaining the steady-state results involved several steps
-due to the highly coupled \glspl{PDE}. First, we ran a preliminary transient
-simulation of fluid flow in the gls\MSFR} core, starting from zero inlet
-velocity and gradually ramping it up to match the nominal flow rate (4.5 m$^3$
-s$^{-1}$); otherwise Moltres had difficulty converging to the desired fully
-developed flow profile. We imposed a parabolic flow profile at the inlet.
-Next, we imported these fully developed flow values as initial values for
-velocity in the actual transient simulation modeling the full coupled
-neutronics and thermal-hydraulics. The initial values for the temperature and
-neutron group flux distributions are 953 K and $1 \times 10^{14}$ cm$^{-2}$
-s$^{-1}$ uniformly throughout the geometry. Finally, we assume that steady
-state is reached when the volume integral values of every variable remain
-constant (up to 6 sig. fig.) for at least four seconds in the simulation; this
-time period corresponds to the nominal circulation time of the \gls{MSFR}.
-
-For a direct comparison with the steady-state results from the Polimi/TUDelft
-models, the bulk of our steady-state results do not have decay heat modeling.
-Instead, we separately discuss the minor differences borne from decay heat
-modeling in the last subsection.
-
-\subsection{Thermal-Hydraulics}
-
-Figure \ref{fig:flow-temp} shows the temperature and velocity fields
-of the fuel salt in the core at steady state from Moltres and the
-Polimi/TUDelft models. The results from Moltres show good qualitative
-agreement with the Polimi/TUDelft models; we observe similar flow and hotspot
-features in all three models. There is a large recirculation region near the
-blanket tank walls arising from turbulent flow. Inertial forces dominate over
-viscous forces to form this large eddy. The figure also shows two relatively
-stagnant regions along the central axis at the top and bottom of the core.
-Temperature hotspots form in these regions of recirculation as convection is
-the dominant heat transfer mechanism. The maximum temperature from Moltres,
-1279 K near the bottom of the recirculation zone, is closer to the maximum
-temperature in the Polimi model ($\approx$ 1300 K) than the TUDelft model
-($\approx$ 1200K). The minimum temperature is 922 K at the inlet. Figure
-\ref{fig:flow} provides an alternate view of the flow profile through flow
-streamlines superimposed on the velocity magnitude distribution. The fastest
-salt flow occurs at the inlet, outlet, and at core half-height approximately
-0.40 m away from the central axis.
-
-\begin{figure}[htb!]
-    \centering
-    \begin{subfigure}[t]{.37\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{flow-temp}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.625\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{flow-temp-fiorina}
-    \end{subfigure}
-    \caption{Temperature and velocity fields in the core from Moltres
-    (left), Polimi (center), and TUDelft (right) models.}
-    \label{fig:flow-temp}
-\end{figure}
-
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=.6\textwidth]{flow}
-    \caption{Fuel salt flow streamlines and velocity magnitude in the core.}
-    \label{fig:flow}
-\end{figure}
-
-While the temperatures at the hotspots are well below the melting point of the
-Ni-alloy structure (1500 K), they may cause undue thermal stress on the
-blanket tank structure and induce relatively faster salt corrosion rates. A
-sudden, large reactivity insertion could push fuel salt temperatures above the
-melting point of the Ni-alloy and cause irreversible damage. Furthermore,
-the reservoir of hot fuel salt may cause unpredictable behavior during
-transient scenarios when the flow profile undergoes a drastic change.
-Thus, Rouch et al. \cite{rouch_preliminary_2014} developed an improved
-hourglass-shaped design to optimize flow distribution and prevent these
-recirculation zones and hotspots from forming. A study of this new design
-using Moltres is a potential subject for future work when a proper turbulence
-model is in place.
-
-\subsection{Neutronics}
-
-\subsection{Decay Heat}
-
-A decay heat model effectively redistributes a fraction of the heat source
-from the center of the core to the entire loop. Thus, we expect to observe
-a slight flattening of the temperature distribution across the entire primary
-loop.
-
-\section{Transient Results}
-
-As noted by Fiorina et al. \cite{fiorina_modelling_2014}, explicit decay heat
-modeling has a negligible effect on reactivity-, pump-, and enchanced
-cooling-driven transients. The Polimi model presented results with the decay
-heat model for the loss of heat sink transient scenario only. For a fair
-comparison, we also
-
-\subsection{Unprotected Reactivity Insertion}
-
-\subsection{Unprotected Loss of Flow}
-
-\subsection{Unprotected Loss of Heat Sink}
-
-\subsection{Chilled Inlet}
-
-\subsection{Pump Overspeed}

### 2020-04-13 introduction.tex
minor grammatical edits

@@
 Since the Industrial Revolution in the 1800s, rapidly rising greenhouse gas
-emissions have contributed significantly to global warming, the consequences
-of which have been acutely felt in recent years through increasingly frequent
+emissions have contributed significantly to global warming. The consequences
+have been acutely felt in recent years through increasingly frequent
 and record breaking extreme weather events. Urgent measures are necessary to
@@
 replacing them with low-carbon
-alternatives would effectively curb a large fraction of these emissions. Out
-of all low-carbon power sources, nuclear power in general is best suited to
+alternatives would effectively curb a large fraction of emissions. Out
+of all low-carbon power sources, nuclear power is arguably best suited to
 replace fossil fuel burning; it provides consistent base-load
@@
 
-Large scale reactor deployments would be necessary to displace the
-presently large share of energy production from fossil fuel power plants;
-however, several obstacles stand in the way of mass reactor deployments. These
+Large-scale reactor deployments are necessary to displace the
+presently large share of energy production from fossil fuel power plants.
+However, several obstacles stand in the way of mass reactor deployments. These
 include perceived safety risks, sustainability concerns, nuclear proliferation
@@
 
-The primary coolant in MSRs is a molten salt mixture,
+The primary coolant in MSRs is a molten salt mixture
 with fissile and/or fertile material directly dissolved in the coolant.
@@
 
-However, the liquid fuel form also brought about novel computational
+However, the liquid fuel form also brings about novel computational
 challenges in simulating the transient behavior of \glspl{MSR}; the
@@
 consideration. Existing reactor system-level codes and modeling approaches
-for conventional \gls{LWR} analysis contain simplifying assumptions in
-multiphysics coupling and other areas that render the techniques unsuitable
+for conventional \gls{LWR} analysis contain reactor-specific assumptions in
+multiphysics coupling and other areas that render the techniques less suitable
 for simulating \glspl{MSR}. Thus, making minor modifications to these codes
@@
 COMSOL-based approach, and another approach by researchers from Delft
-University of Technology, where they coupled their in-house neutronics and
+University of Technology, in which they coupled their in-house neutronics and
 thermal-hydraulics codes, DALTON and HEAT respectively. With multigroup
@@
 
-\section{Objectives}
-
 Moltres employs fully implicit coupling between the neutronics and
@@
 profile compared to imposing a predetermined velocity field.
+
+\section{Objectives}
 

### 2020-04-13 msr.tex
Completed first draft of Chapters 1-4 for review
minor grammatical edits

@@
 
-The first \gls{MSR}, named the \gls{ARE}, dates back to the 1940s,
+The first \gls{MSR}, named the \gls{ARE}, dates back to the 1940s
 as part of the US Aircraft Nuclear Propulsion program
@@
 different concepts and the only common denominator is the use of nuclear fuel
-dissolved in molten salt. Table * is a non-exhaustive list of \glspl{MSR}
-under development.
+dissolved in molten salt.
+%Table * is a non-exhaustive list of \glspl{MSR} under development.
 
@@
 under development today. Graphite-moderated thermal-spectrum \glspl{MSR} are
-typically straightforward \gls{LEU} burners, or $^{232}$Th/$^{233}$U
+typically straightforward \gls{LEU} burners or $^{232}$Th/$^{233}$U
 iso-breeders/breeders, while epithermal- and fast-spectrum \glspl{MSR} have
@@
 continuous removal of \glspl{FP}. For example, the decay heat in an \gls{MSFR}
-after reaching equilibrium salt composition is approximately 3.5\% compared
-to 6\% in \glspl{LWR} \cite{brovchenko_design-related_2013}.
+after reaching equilibrium salt composition is expected to be approximately
+3.5\% of full reactor power compared to 6\% in \glspl{LWR}
+\cite{brovchenko_design-related_2013}.
 

### 2020-04-29 steadystate.tex
Steady-state and transient chapters text without figures

--- 2020-03-09 results.tex
@@
-This chapter discusses three main sets of simulation results. First, we
-compare the key neutronics results between Serpent and Moltres for a static
-model of the \gls{MSFR}, i.e. no salt flow, to assess the accuracy of the
-six-group neutron diffusion model in Moltres on a fast-spectrum reactor. This
-exercise also serves as a checking exercise for the group constant data
-generated for Moltres from Serpent. Second, we present steady-state results
-from Moltres with salt flow and verify it against published results, namely
-from the Polimi/TUDelft and OpenFOAM models developed by Fiorina et al.
-\cite{fiorina_modelling_2014} and Aufiero et al.
-\cite{aufiero_development_2014}, respectively. Lastly, we present and compare
-transient simulation results from Moltres with the aforementioned models.
-Transient accident scenarios include an unprotected reactivity insertion, an
-unprotected loss of flow, an unprotected loss of heat sink, a chilled inlet,
-and a pump overspeed scenario.
+With the verification of Moltres' neutronics modeling capabilities in the
+context of the \gls{MSFR}, we now move on to the multiphysics simulation
+results. This chapter covers the steady-state multiphysics simulation results
+from Moltres.
 
-\section{Static Model}
+The procedure for obtaining the steady-state results involved several steps
+due to the highly coupled \glspl{PDE}. First, we ran a preliminary transient
+simulation of fluid flow in the \gls{MSFR} core, starting from zero inlet
+velocity and gradually ramping it up to match the nominal flow rate (4.5 m$^3$
+s$^{-1}$); otherwise Moltres had difficulty converging to the desired fully
+developed flow profile. We imposed a parabolic flow profile at the inlet.
+Next, we imported these fully developed flow values as initial values for
+velocity in the actual transient simulation modeling the full coupled
+neutronics and thermal-hydraulics. The initial values for the temperature and
+neutron group flux distributions are 953 K and $1 \times 10^{14}$ cm$^{-2}$
+s$^{-1}$ uniformly throughout the geometry. Finally, we assume that steady
+state is reached when the volume integral values of every variable remain
+constant (up to 6 sig. fig.) for at least four seconds in the simulation; this
+time period corresponds to the nominal circulation time of the \gls{MSFR}.
 
-We calculated estimates of $k_{\text{eff}}$ values in Serpent and Moltres for
-the \gls{MSFR} model with static salt (no flow), a uniform temperature
-distribution of 973 K, and a fuel salt density of $4.1249 \times 10^3$ kg
-m$^{-3}$. Moltres performs eigenvalue calculations for finding the 
-$k_{\text{eff}}$ as it does not currently have the capability to calculate
-the adjoint flux. Table \ref{table:keff} shows the $k_{\text{eff}}$ values
-from Serpent, and from Moltres with and without \glspl{DNP}. We observe a
-small 127.9 pcm discrepancy between the two codes, which we attribute to two
-main factors: the neutron diffusion model, and the omission of the blanket
-tank structural material. It is well-known that the neutron diffusion model is
-not as accurate as the other S$_{\text{N}}$ or SP$_{\text{N}}$ deterministic
-methods nor the Monte Carlo approach in Serpent. Regarding the omission of the
-blanket tank material, we replaced the 2 cm-thick structural material with
-blanket salt. This decision may be partly responsible for the higher
-$k_{\text{eff}}$ value calculated by Moltres. Nevertheless, the discrepancy is
-smaller than the 228.5 pcm and 256.7 pcm discrepancy reported by Cervi et al.
-\cite{cervi_development_2019} for their six-group $SP_3$ and neutron diffusion
-methods, respectively, in OpenFOAM. The neutron diffusion model in OpenFOAM is
-the same approach Aufiero et al. \cite{aufiero_development_2014} used for
-their transient analysis of the \gls{MSFR}, albeit with one neutron energy
-group.
+For a direct comparison with the steady-state results from the Polimi/TUDelft
+models, the bulk of our steady-state results do not have decay heat modeling.
+Instead, we separately discuss the minor differences borne from decay heat
+modeling in the last subsection.
 
-\begin{table}[htb!]
+\section{Thermal-Hydraulics}
+
+Figure \ref{fig:flow-temp} shows the temperature and velocity fields
+of the fuel salt in the core at steady state from Moltres and the
+Polimi/TUDelft models. The results from Moltres show good qualitative
+agreement with the Polimi/TUDelft models; we observe similar flow and hotspot
+features in all three models. There is a large recirculation region near the
+blanket tank walls arising from turbulent flow. Inertial forces dominate over
+viscous forces to form this large eddy. The figure also shows two relatively
+stagnant regions along the central axis at the top and bottom of the core.
+Temperature hotspots form in these regions of recirculation as convection is
+the dominant heat transfer mechanism. The maximum temperature from Moltres,
+1279 K near the bottom of the recirculation zone, is closer to the maximum
+temperature in the Polimi model ($\approx$ 1300 K) than the TUDelft model
+($\approx$ 1200K). The minimum temperature is 922 K at the inlet. Figure
+\ref{fig:flow} provides an alternate view of the flow profile through flow
+streamlines superimposed on the velocity magnitude distribution. The fastest
+salt flow occurs at the inlet, outlet, and at core half-height approximately
+0.40 m away from the central axis.
+
+\begin{figure}[htb!]
     \centering
-	\caption{$k_{\text{eff}}$ values from Serpent and Moltres.}
-	\begin{tabular}{l S S}
-		\toprule
-		{Code} & {$k_{\text{eff}}$} & {Diff. wrt Serpent [pcm]}\\
-		\midrule
-		{Serpent} & 1.00667 \pm 0.000049 & 0\\
-		{Moltres with \glspl{DNP}} & 1.007949 \pm 0.000001 & 127.9\\
-		{Moltres without \glspl{DNP}} & 1.004937 \pm 0.000001 & {-}\\
-		\bottomrule
-	\end{tabular}
-	\label{table:keff}
-\end{table}
-%
-\begin{table}[htb!]
+    \begin{subfigure}[t]{.37\textwidth}
         \centering
-	\caption{$\beta_{\text{eff}}$ values from Serpent and Moltres.}
-	\begin{tabular}{l S S}
-		\toprule
-		{Code} & {$\beta_{\text{eff}}$ [pcm]} & {Diff. wrt Serpent [pcm]}\\
-		\midrule
-		{Serpent} & 303.2 \pm .8 & 0\\
-		{Moltres} & 301.2 \pm .2 & 2.0\\
-		\bottomrule
-	\end{tabular}
-	\label{table:betaeff}
-\end{table}
+        \includegraphics[width=\textwidth]{flow-temp}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.625\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{flow-temp-fiorina}
+    \end{subfigure}
+    \caption{Temperature and velocity fields in the core from Moltres
+    (left), Polimi (center), and TUDelft (right) models. The colors represent
+    temperature according to the respective colorbars while the arrows
+    represent velocity fields.}
+    \label{fig:flow-temp}
+\end{figure}
 
-The
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=.6\textwidth]{flow}
+    \caption{Fuel salt flow streamlines and velocity magnitude in the core.
+    The colors represent velocity magnitude according to the colorbar on the
+    right.}
+    \label{fig:flow}
+\end{figure}
 
-\section{Steady-State Behavior}
+While the temperatures at the hotspots are well below the melting point of the
+Ni-alloy structure (1500 K), they may cause undue thermal stress on the
+blanket tank structure and induce relatively faster salt corrosion rates. A
+sudden, large reactivity insertion could push fuel salt temperatures above the
+melting point of the Ni-alloy and cause irreversible damage. Furthermore,
+the reservoir of hot fuel salt may cause unpredictable behavior during
+transient scenarios when the flow profile undergoes a drastic change.
+Thus, Rouch et al. \cite{rouch_preliminary_2014} developed an improved
+hourglass-shaped design to optimize flow distribution and prevent these
+recirculation zones and hotspots from forming. A study of this new design
+using Moltres is a potential subject for future work when a proper turbulence
+model is in place.
 
-\subsection{Neutronics}
+\section{Neutronics}
 
-\subsection{Thermal-Hydraulics}
+The neutron flux distribution represents the heat source distribution in a
+nuclear reactor. Figure * shows the neutron flux distributions for all six
+neutron energy groups. The distribution forms the symmetrical shape we expect
+to see in a nuclear reactor with the peak at the center of the core. The
+relatively lower temperatures near the center of the core enhances the neutron
+flux peak but it is not a major concern as there are no structural parts
+vulnerable to neutron damage in that region. The total flux at the center is
+*, which is very close to values reported by Fiorina et al. 
+\cite{fiorina_molten_2013} and Aufiero \cite{aufiero_development_2014}, albeit
+their results are from their neutronics models at a fixed uniform temperature.
 
-\section{Transient Behavior}
+\begin{figure}[htb!]
+    \centering
+%    \includegraphics[width=.6\textwidth]{}
+    \caption{Neutron flux distributions in the core for neutron energy groups
+    1 to 6.}
+    \label{fig:neutronflux}
+\end{figure}
 
-\subsection{Unprotected Reactivity Insertion}
+As mentioned before, the \glspl{DNP} are mobile in \glspl{MSR} and their
+distributions do not directly correspond to the neutron flux distributions.
+The location where delayed neutrons are emitted impacts their neutron
+importance according to their proximity to fissile and parasitic isotopes.
+Figure * shows the \gls{DNP} distributions for all eight \gls{DNP} groups. The
+precursors from the shortest-lived group (Group 8) predominantly decay within
+the core as their half-lives are shorter than the time it takes to reach the
+outlet while the precursors from the longest-lived group (Group 1) are
+effectively uniformly distributed due to their long half-lives relative to the
+average circulation time. For the longer-lived groups, the mesh elements with
+relatively high and low \gls{DNP} concentrations near the outlet and inlet,
+respectively, suggest that the mesh is too coarse. Thus, we would recommend
+further refinement for future work involving similar geometries.
 
-\subsection{Unprotected Loss of Flow}
+\begin{figure}[htb!]
+    \centering
+%    \includegraphics[width=.6\textwidth]{}
+    \caption{\gls{DNP} distributions in the core for \gls{DNP} groups
+    1 to 8.}
+    \label{fig:dnp}
+\end{figure}
 
-\subsection{Unprotected Loss of Heat Sink}
+An important parameter for \gls{MSR} codes is the overall in-core
+$\beta_{\text{eff}}$ value. This value represents the true effective delayed
+neutron fraction in \glspl{MSR} that accounts for the loss of delayed neutrons
+due to \gls{DNP} decay outside the core where fissions cannot be sustained. 
+** DNP values and comparison **
 
-\subsection{Chilled Inlet}
+\section{Decay Heat}
 
-\subsection{Pump Overspeed}
+The inclusion of a decay heat model effectively redistributes a fraction of
+the volumetric heat source from the center of the core to the entire loop.
+Thus, we expect to observe a slight flattening of the temperature distribution
+across the entire primary loop.

### 2020-04-29 transient.tex
Steady-state and transient chapters text without figures

This chapter discusses the transient multiphysics simulation results of the
\gls{MSFR} from Moltres for four accident scenarios. These scenarios include
unprotected instances of reactivity insertion, loss of flow, pump overspeed,
and loss of heat sink. The term ``unprotected'' means that there is no
external intervention in these scenarios. As such, these simulations
give an insight on the \gls{MSFR}'s passive safety capabilities in the absence
of any active safety system. We used the steady-state configuration presented
in the previous chapter as the initial conditions for the transient
simulations discussed in this chapter. In essence, we imported all
steady-state spatial values for neutron flux, delayed neutron precursor
concentration, temperature, velocity, and pressure as the initial state of the
transient scenarios.

As noted by Fiorina et al. \cite{fiorina_modelling_2014}, explicit decay heat
modeling has a negligible effect on reactivity-, pump-, and enhanced
cooling-initiated transients. Furthermore, only their Polimi model had decay
heat modeling capabilities. Therefore, they presented results from their
Polimi and TUDelft model for the four accident scenarios (listed above)
without decay heat modeling. They only enabled decay heat modeling for the
loss of heat sink transient scenario only. For a fair comparison, we also ran
all transient scenarios without the decay heat model with the exception of the
loss of heat sink scenario in which we ran the simulations with and without
the decay heat model.

We detailed the setup for each transient simulation in their respective
sections.

\section{Unprotected Reactivity Insertion}

\glspl{RIA} are a type of nuclear accident caused by unintended positive
reactivity insertions. The excess reactivity causes the power output and
temperatures in nuclear reactors to rise to potentially dangerous levels. In
\glspl{MSR}, a positive reactivity insertion could occur when the online
refueling system injects excess fissile material into the core. Excessively
high neutron fluences and temperatures negatively impact reactor structural
integrity and increase the risk of containment breach.

We modeled two unprotected reactivity insertion scenarios in Moltres by
swapping out the original set of group constant data with two new, separate
sets of data from Serpent corresponding to 50 pcm and 200 pcm reactivity
insertions, respectively. We increased the reactivity of the Serpent
\gls{MSFR} models by increasing the $^{233}$U-to-$^{232}$Th ratio in the fuel
salt.

\section{Unprotected Loss of Flow}

A loss of forced flow accident can occur in the event of a station
blackout; the pumps would cease operating due to the loss of AC electrical
power. Natural circulation resulting from temperature-dependent density
changes becomes the dominant driving force for salt flow in the primary loop.
While Fiorina et al. \cite{fiorina_modelling_2014} and Aufiero et al.
\cite{aufiero_development_2014} applied the Boussinesq approximation for
bouyancy-driven flow in their models, it is not directly applicable to the
\gls{MSFR} model in this thesis because we partitioned the primary loop into
two separate geometries and used Dirichlet boundary conditions to drive flow.
Fiorina et al.'s Polimi and TUDelft models featured complete exponential
coast-downs of the pumps with a time constant of 5 s. The resulting flow rate
from natural circulation was approximately 18 times smaller than the initial
flow rate. Thus, for the \gls{MSFR} model in Moltres, we imposed a similar
exponential decay term with the same time constant on the inflow Dirichlet
boundary condition to reduce the flow rate to the same final flow rate
observed by Fiorina et al. Figure * shows the change in flow rate over the
transient simulation.

We also expect the reduced flow rate to decrease the heat transfer rate
between the primary and secondary loop through the heat exchanger as the heat
transfer coefficient is dependent on the flow rate. Given the limited
information available on the heat exchanger specifications, we assumed that
the heat transfer coefficient is directly proportional to the flow rate, and
the overall heat transfer rate retains its linear dependence to the
temperature difference between the primary and secondary loop at the location
of the heat exchanger as shown previously in \ref{eq:hx}.

\section{Unprotected Pump Overspeed}

\section{Unprotected Loss of Heat Sink}

\subsection{Without Decay Heat Model}

\subsection{With Decay Heat Model}

### 2020-06-19 introduction.tex
Some intro chapter edits

@@
 
-Since the Industrial Revolution in the 1800s, rapidly rising greenhouse gas
-emissions have contributed significantly to global warming. The consequences
-have been acutely felt in recent years through increasingly frequent
-and record breaking extreme weather events. Urgent measures are necessary to
-limit CO$_2$ emissions and the negative impacts of global warming.
+Greenhouse gas emission from human activities is the main cause of climate
+change, a global phenomenon characterized by shifting weather patterns and
+rising temperatures. Climate change has dire consequences on human health and
+safety due to extreme weather events and the overall impact on food
+production. As a result, urgent measures are necessary to limit greenhouse gas
+emissions, namely CO$_2$ emissions from electricity generation.
 
-Since electricity generation from burning fossil fuels represents the
-largest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019}),
-replacing them with low-carbon
-alternatives would effectively curb a large fraction of emissions. Out
-of all low-carbon power sources, nuclear power is arguably best suited to
-replace fossil fuel burning; it provides consistent base-load
-power independent of weather and local geographical conditions. Solar and wind
-power are dependent on favorable weather conditions while hydropower requires
-relatively more area and is limited to geologically appropriate locations.
-This argument provides a strong case for nuclear power being the most
-appropriate and realistic alternative to burning fossil fuels for our current
-and future electricity needs.
+Electricity generation from burning fossil fuels represents the
+largest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019});
+replacing it with low-carbon
+alternatives would curb a large fraction of emissions. Nuclear power is
+a viable low-carbon replacement for burning fossil fuels.
+Other low-carbon energy sources include solar-, wind-, and hydro-power. Each
+energy source has its own set of advantages and disadvantages; nuclear power's
+advantage over renewable energy sources is that it provides consistent
+base-load power independent of weather and geographical location (cite mit).
+Employing a diverse mix of these energy sources ensures energy security and
+reliability during and after our transition towards a low-carbon future (cite mit).
 
-Large-scale reactor deployments are necessary to displace the
+The world would have to ramp up the current rate of reactor deployments to
+displace some of the
 presently large share of energy production from fossil fuel power plants.
@@
 risks, and the ability to compete economically with other sources of energy.
-The \gls{MSR} concept, one of six advanced reactor designs
+A potential solution to the aforementioned issues is
+the \gls{MSR} concept, one of six advanced reactor designs
 selected by the \gls{GIF} \cite{gif_technology_2002} for continued research
-and development, is a potential solution to the aforementioned issues.
+and development.
 
-The primary coolant in MSRs is a molten salt mixture
+The primary coolants in MSRs consist of a molten salt mixture
 with fissile and/or fertile material directly dissolved in the coolant.
-MSRs possess an inherently robust safety feature in the
-strongly negative fuel temperature reactivity coefficient from Doppler
-broadening and thermal fuel expansion that greatly reduces
-the risk of a reactor power excursion. Many designs can also incorporate the
+MSRs possess an inherently robust safety feature in the strongly negative fuel
+temperature coefficient of reactivity. Some designs can also incorporate the
 thorium fuel cycle for improved sustainability arising from the use of
 abundant natural thorium resources and reduced transuranic waste. The
-resultant reduced radiotoxicity from transuranic waste also reduces costs
+latter also reduces economical costs
 associated with long-term nuclear waste storage. In addition, the ability to
@@
 challenges in simulating the transient behavior of \glspl{MSR}; the
-neutronics and thermal-hydraulics are more tightly coupled due to the strong
-temperature reactivity coefficient, and the additional advection coupling
-terms. Furthermore, we have to account for the movement of \glspl{DNP} as
+neutronics and thermal-hydraulics are tightly coupled due to the strong
+temperature reactivity coefficient. Furthermore, we have to account for the
+movement of \glspl{DNP} as
 they are now generated directly within the primary coolant loop. Therefore,
 the choice of coupling methods for each set of physics requires careful
-consideration. Existing reactor system-level codes and modeling approaches
+consideration. Existing reactor system-level software and modeling approaches
 for conventional \gls{LWR} analysis contain reactor-specific assumptions in

### 2020-06-19 neutronics.tex
Edit neutronics, add steady-state figures, edit transient

@@
 exercise also serves as a verification for the group constant data
-generated for Moltres from Serpent.
+generated for Moltres from Serpent. For visibility, we list the important
+details of the static model below.
+
+\textbf{Details of the static \gls{MSFR} model on Moltres}
+\begin{itemize}
+    \item No salt flow (static salt)
+    \item Uniform temperature of 973 K throughout the 2D core model
+    \item Six neutron energy groups
+    \item Eight delayed neutron precursor groups
+    \item Vacuum boundary conditions on the outer boundaries of the 2D core
+    mesh
+\end{itemize}
 
@@
 We calculated estimates of $k_{\text{eff}}$ values in Serpent and Moltres for
-the \gls{MSFR} model with static salt (no flow), and uniform temperature
-distributions. Moltres solves the six-group neutron diffusion equations as a
+the static \gls{MSFR} model. Moltres solves the six-group neutron diffusion equations as a
 steady-state eigenvalue problem to find the $k_{\text{eff}}$ as it does not
@@
 \ref{table:keff} shows the $k_{\text{eff}}$ values from Serpent and Moltres at
-973 K and the corresponding salt density, while Table \ref{table:keff} shows
+973 K and the corresponding salt density, while Table \ref{table:ktemp} shows
 the $k_{\text{eff}}$ values for other temperatures at 100 K intervals. We
@@
 \cite{cervi_development_2019} for their six-group $SP_3$ and neutron
-diffusion methods, respectively, in OpenFOAM. The neutron diffusion model in
-OpenFOAM is the same approach Aufiero et al. \cite{aufiero_development_2014}
-used for their transient analysis of the \gls{MSFR}, albeit with one neutron
-energy group.
+diffusion methods, respectively, in OpenFOAM. Aufiero et al.
+\cite{aufiero_development_2014} used the same neutron diffusion model in
+OpenFOAM for their transient analysis of the \gls{MSFR}, albeit with one
+neutron energy group to reduce computational load.
 
@@
 		{Serpent} & 1.00662 \pm 0.00005 & {-}\\
-		{Moltres with \glspl{DNP}} & 1.0079488 \pm 0.0000010 & 133\\
-		{Moltres without \glspl{DNP}} & 1.0049369 \pm 0.0000010 & {-}\\
+		{Moltres with \glspl{DNP}} & 1.0079400 \pm 0.0000010 & 132\\
+		{Moltres without \glspl{DNP}} & 1.0049197 \pm 0.0000010 & {-}\\
 		\bottomrule
@@
 The absolute value of $k_{\text{eff}}$ largely impacts the steady-state
-temperature of the reactor. We may adjust $k_{\text{eff}}$ by controlling
-fissile inventory to raise or lower the operating temperatures and meet the
-design specifications for inlet and outlet temperatures. For transient
-analysis, the $\beta_{\text{eff}}$ and reactivity coefficients are arguably
-more important as they dictate the duration and magnitude of the reactor
-response to a transient initiator. The $\beta_{\text{eff}}$ values at 973 K,
-shown in Table \ref{table:betaeff}, are in excellent agreement with a 2.89
-pcm discrepancy.
+temperature of the reactor. We can raise or lower the average core temperature
+at steady state to meet the design specifications for the inlet and outlet
+temperatures by adjusting the fissile inventory and $k_{\text{eff}}$. In
+transient analysis, the $\beta_{\text{eff}}$ and reactivity coefficients are
+important as they dictate the duration and magnitude of the reactor
+response to a transient initiator. We calculated $\beta_{\text{eff}}$ by
+taking the relative difference between the $k_{\text{eff}}$ values with and
+without \glspl{DNP} in Table \ref{table:keff}. The $\beta_{\text{eff}}$ values
+at 973 K, shown in Table \ref{table:betaeff}, are in excellent agreement with
+a 4.43 pcm discrepancy.
 
@@
 		{Serpent} & 304.08 \pm .81 & {-}\\
-		{Moltres} & 301.19 \pm .20 & 2.89\\
+		{Moltres} & 299.65 \pm .20 & 4.43\\
 		\bottomrule
@@
 
-Moltres also reproduced the six-group neutron spectrum very well from the
-Serpent group constants. Figure \ref{fig:ntspec} compares the energy spectra
-from Serpent and Moltres in the central fuel salt region. The six-group
-neutron spectra overlap exactly over each other. More generally, the plot
-shows the distinctive fast spectrum observed in the \gls{MSFR} with dips in
-the spectrum corresponding to elastic scattering resonances from lithium and
-fluorine. From this plot, we observe that the discrepancies in
+Moltres also closely replicated the six-group neutron spectrum from the
+Serpent group constants. Figure \ref{fig:ntspec} compares the neutron energy
+spectra from Serpent and Moltres in the central fuel salt region. The
+six-group neutron spectra overlap exactly over each other. More generally, the
+plot shows the distinctive fast spectrum observed in the \gls{MSFR} with dips
+in the spectrum corresponding to elastic scattering resonances from lithium
+and fluorine. From this plot, we observe that the discrepancies in
 $k_{\text{eff}}$ arise mainly from discretizing neutron energy into groups
@@
 
-In summary, Moltres can replicate most of the relevant neutronics parameters
-accurately with the group constant data from Serpent. While the
+In summary, Moltres replicated most of the relevant neutronics parameters
+accurately with the group constant data from Serpent. The $\beta_{\text{eff}}$
+and temperature reactivity coefficients, which are important
+parameters for modeling transient reactor behavior, are in excellent
+agreement. While the
 $k_{\text{eff}}$ values have discrepancies on the order of 100 pcm, they are
-relatively small compared to results from other codes. Furthermore, the
-$\beta_{\text{eff}}$ and temperature coefficients, which are important
-parameters for modeling transient reactor behavior, are in excellent
-agreement.
+relatively small compared to the $k_{\text{eff}}$ values from the neutron
+diffusion and SP3 models in OpenFOAM \cite{cervi_development_2019}.

### 2020-06-19 steadystate.tex
Edit neutronics, add steady-state figures, edit transient

@@
 
-For a direct comparison with the steady-state results from the Polimi/TUDelft
-models, the bulk of our steady-state results do not have decay heat modeling.
-Instead, we separately discuss the minor differences borne from decay heat
-modeling in the last subsection.
+For a direct comparison with the steady-state results from the Polimi and
+TUDelft models \cite{aufiero_development_2014}, we will first present our
+steady-state results without modeling decay heat. After this comparison, we
+separately discuss the minor differences borne from decay heat modeling in the
+last subsection.
 
-\section{Thermal-Hydraulics}
+\section{Steady-State Thermal-Hydraulics Results}
 
-Figure \ref{fig:flow-temp} shows the temperature and velocity fields
-of the fuel salt in the core at steady state from Moltres and the
-Polimi/TUDelft models. The results from Moltres show good qualitative
-agreement with the Polimi/TUDelft models; we observe similar flow and hotspot
-features in all three models. There is a large recirculation region near the
-blanket tank walls arising from turbulent flow. Inertial forces dominate over
-viscous forces to form this large eddy. The figure also shows two relatively
-stagnant regions along the central axis at the top and bottom of the core.
-Temperature hotspots form in these regions of recirculation as convection is
-the dominant heat transfer mechanism. The maximum temperature from Moltres,
-1279 K near the bottom of the recirculation zone, is closer to the maximum
-temperature in the Polimi model ($\approx$ 1300 K) than the TUDelft model
-($\approx$ 1200K). The minimum temperature is 922 K at the inlet. Figure
-\ref{fig:flow} provides an alternate view of the flow profile through flow
-streamlines superimposed on the velocity magnitude distribution. The fastest
-salt flow occurs at the inlet, outlet, and at core half-height approximately
-0.40 m away from the central axis.
-
-\begin{figure}[htb!]
+\begin{figure}[t!]
     \centering
-    \begin{subfigure}[t]{.37\textwidth}
+    \begin{subfigure}[t]{.365\textwidth}
         \centering
@@
 
-\begin{figure}[htb!]
+\begin{figure}[t!]
     \centering
@@
 \end{figure}
+
+Figure \ref{fig:flow-temp} shows the temperature and velocity fields of the
+fuel salt in the core at steady state from Moltres and the Polimi/TUDelft
+models. The results from Moltres show good qualitative agreement with the
+Polimi and TUDelft models \cite{aufiero_development_2014}; we observe similar
+flow and hotspot features in all three models. There is a large recirculation
+region near the blanket tank walls arising from turbulent flow. Inertial
+forces dominate over viscous forces to form this large eddy. The figure also
+shows two relatively stagnant regions along the central axis at the top and
+bottom of the core. Temperature hotspots form in these regions of
+recirculation as convection is
+the dominant heat transfer mechanism. The maximum temperature from Moltres,
+1275 K near the bottom of the recirculation zone, is closer to the maximum
+temperature in the Polimi model ($\approx$ 1300 K) than the TUDelft model
+($\approx$ 1200K). The minimum temperature is 922 K at the inlet. Figure
+\ref{fig:flow} provides an alternate view of the flow profile through flow
+streamlines superimposed on the velocity magnitude distribution. The fastest
+salt flow occurs at the inlet, outlet, and at core half-height approximately
+0.40 m away from the central axis.
 
@@
 
-\section{Neutronics}
+\section{Steady-State Neutronics Results}
+
+\subsection{Neutron Flux}
 
 The neutron flux distribution represents the heat source distribution in a
-nuclear reactor. Figure * shows the neutron flux distributions for all six
-neutron energy groups. The distribution forms the symmetrical shape we expect
-to see in a nuclear reactor with the peak at the center of the core. The
-relatively lower temperatures near the center of the core enhances the neutron
-flux peak but it is not a major concern as there are no structural parts
-vulnerable to neutron damage in that region. The total flux at the center is
-*, which is very close to values reported by Fiorina et al. 
-\cite{fiorina_molten_2013} and Aufiero \cite{aufiero_development_2014}, albeit
-their results are from their neutronics models at a fixed uniform temperature.
+nuclear reactor. Figure \ref{fig:neutronflux} shows the neutron flux
+distributions in the core for all six neutron energy groups, and Figure
+\ref{fig:axialradial} shows the axial and radial fluxes along the center of
+the core and at reactor half-height, respectively.
+The distributions are highly symmetric along the
+central and horizontal axes, as expected of a cylindrical reactor design. The
+relatively lower temperatures near the center of the core promotes the neutron
+flux peaking but it is not a major concern as there are no structural parts
+vulnerable to neutron damage in that region. The peak total flux at the center
+is $9.80 \times 10^{15}$ cm$^{-2}$ s$^{-1}$, which is close to values reported
+by Fiorina et al. \cite{fiorina_molten_2013} and Aufiero et al.
+\cite{aufiero_development_2014} as shown in Table \ref{table:peak-flux}. The
+peak flux value from this paper is slightly higher as we used the steady-state
+temperature distribution while Fiorina et al. and Aufiero et al. imposed a
+uniform temperature distribution at 973 K.
 
-\begin{figure}[htb!]
+\begin{figure}[t!]
     \centering
-%    \includegraphics[width=.6\textwidth]{}
+    \begin{subfigure}[t]{.325\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-g1}
+    \end{subfigure}
+    \begin{subfigure}[t]{.325\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-g2}
+    \end{subfigure}
+    \begin{subfigure}[t]{.325\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-g3}
+    \end{subfigure}
+    \begin{subfigure}[t]{.325\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-g4}
+    \end{subfigure}
+    \begin{subfigure}[t]{.325\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-g5}
+    \end{subfigure}
+    \begin{subfigure}[t]{.325\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-g6}
+    \end{subfigure}
     \caption{Neutron flux distributions in the core for neutron energy groups
-    1 to 6.}
+    1 to 6. The y and x axes represent height and radius values of the core
+    relative to the entire reactor geometry.}
     \label{fig:neutronflux}
@@
 
+\begin{figure}[t!]
+    \centering
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{axial-flux}
+    \end{subfigure}
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{radial-flux}
+    \end{subfigure}
+    \caption{Axial (left) and radial (right) neutron flux distributions in the
+    core for neutron energy groups 1 to 6.}
+    \label{fig:axialradial}
+\end{figure}
+
+\begin{table}[t!]
+	\centering
+	\caption{Peak neutron flux values from Moltres (this paper), COMSOL
+	\cite{fiorina_molten_2013}, and OpenFOAM \cite{aufiero_development_2014}
+	models. The COMSOL and OpenFOAM values were calculated with uniform
+	temperature distributions.}
+	\begin{tabular}{l S}
+		\toprule
+		{Model} & {Peak Neutron Flux [$\times 10^{15}$ cm$^{-2}$ s$^{-1}$]}
+		\\
+		\midrule
+		{Moltres (This paper)} & 9.80\\
+		{COMSOL} & 8.6 \\
+		{OpenFOAM} & 9.0 \\
+		\bottomrule
+	\end{tabular}
+	\label{table:peak-flux}
+\end{table}
+
 As mentioned before, the \glspl{DNP} are mobile in \glspl{MSR} and their
 distributions do not directly correspond to the neutron flux distributions.
-The location where delayed neutrons are emitted impacts their neutron
-importance according to their proximity to fissile and parasitic isotopes.
-Figure * shows the \gls{DNP} distributions for all eight \gls{DNP} groups. The
+The location where the \glspl{DNP} decay and emit neutrons impacts their
+neutron
+importance depending on their proximity to fissile and parasitic isotopes.
+Figure \ref{fig:dnp} shows the \gls{DNP} distributions for all eight \gls{DNP}
+groups. The
 precursors from the shortest-lived group (Group 8) predominantly decay within
@@
     \centering
-%    \includegraphics[width=.6\textwidth]{}
+    \begin{subfigure}[t]{.243\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-pre1}
+    \end{subfigure}
+    \begin{subfigure}[t]{.243\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-pre2}
+    \end{subfigure}
+    \begin{subfigure}[t]{.243\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-pre3}
+    \end{subfigure}
+    \begin{subfigure}[t]{.243\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-pre4}
+    \end{subfigure}
+    \begin{subfigure}[t]{.243\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-pre5}
+    \end{subfigure}
+    \begin{subfigure}[t]{.243\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-pre6}
+    \end{subfigure}
+    \begin{subfigure}[t]{.243\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-pre7}
+    \end{subfigure}
+    \begin{subfigure}[t]{.243\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ss-pre8}
+    \end{subfigure}
     \caption{\gls{DNP} distributions in the core for \gls{DNP} groups
-    1 to 8.}
+    1 to 8 (from left to right, top to bottom). Refer to Figure \ref{fig:neutronflux} for the height and radius
+    scales on the y and x axes, respectively. Note the different scales for
+    each distribution.}
     \label{fig:dnp}
@@
 
-An important parameter for \gls{MSR} codes is the overall in-core
-$\beta_{\text{eff}}$ value. This value represents the true effective delayed
-neutron fraction in \glspl{MSR} that accounts for the loss of delayed neutrons
-due to \gls{DNP} decay outside the core where fissions cannot be sustained. 
-** DNP values and comparison **
+\subsection{In-Core \texorpdfstring{$\beta_{\text{eff}}$}{beta_eff}}
+
+An important safety parameter for \glspl{MSR} is the actual in-core
+$\beta_{\text{eff}}$ value. This value represents the actual effective delayed
+neutron fraction in \glspl{MSR} after accounting for the loss of delayed
+neutrons due to out-of-core \gls{DNP} decay. Table * shows
 

### 2020-06-19 transient.tex
Edit neutronics, add steady-state figures, edit transient

@@
 This chapter discusses the transient multiphysics simulation results of the
-\gls{MSFR} from Moltres for four accident scenarios. These scenarios include
+\gls{MSFR} from Moltres for four accident scenarios. These scenarios, adapted
+from Fiorina et al.'s work, include
 unprotected instances of reactivity insertion, loss of flow, pump overspeed,
@@
 heat modeling capabilities. Therefore, they presented results from their
-Polimi and TUDelft model for the four accident scenarios (listed above)
-without decay heat modeling. They only enabled decay heat modeling for the
+Polimi and TUDelft model for the four accident scenarios 
+without decay heat modeling; they enabled decay heat modeling for the
 loss of heat sink transient scenario only. For a fair comparison, we also ran
@@
 loss of heat sink scenario in which we ran the simulations with and without
-the decay heat model.
-
-We detailed the setup for each transient simulation in their respective
-sections.
+the decay heat model. More generally, we imposed various other simplifying
+assumptions in our transient models to match their implementations as closely
+as possible, within Moltres' capabilities. The details of the setup for each
+transient simulation are in their respective sections.
 
@@
 
+The focus of this transient study is the neutronic and thermal-hydraulic
+behavior of the reactor core. Thus, we assume that the heat exchanger and the
+associated power generation equipment (generator turbines, heat sinks, etc.)
+can withstand all variations in the power output during the transients.
+
 \section{Unprotected Loss of Flow}
 
-A loss of forced flow accident can occur in the event of a station
+A loss of forced flow transient can occur in the event of a station
 blackout; the pumps would cease operating due to the loss of AC electrical
@@
 temperature difference between the primary and secondary loop at the location
-of the heat exchanger as shown previously in \ref{eq:hx}.
+of the heat exchanger as shown previously in \ref{eq:hx}. Lastly, we assume
+the turbulent viscosity, $\mu_t$, is directly proportional to the flow rate to
+account for reduced mixing when the flow rate decreases.
 
 \section{Unprotected Pump Overspeed}
+
+In the context of this thesis, pump overspeed refers to a sustained
+increase in pump speed in the primary coolant loop. We followed Fiorina et
+al.'s implementation by ramping up the inlet velocity, $u$, by 50\% from
+the nominal value, $u_0$, according to the following formula:
+
+\begin{align}
+    u(t) &= u_0 [1 + 0.5 (1 - e^{-t / \tau})] \\
+    \intertext{where}
+    \tau &= 5 \text{ s.} \nonumber
+\end{align}
+
+We retain the directly proportional assumption between the flow rate and
+$\mu_t$ that we imposed in the unprotected loss of flow simulation.
+The increased flow rate impacts reactor performance in several ways.
+It affects the neutronics by reducing the in-core $\beta_eff$ as more of the
+shorter-lived precursors will tend to flow out of the core before decaying.
+This net loss of neutrons reduces the reactivity in the core, thereby causing
+core temperatures to fall to counteract this change through temperature
+reactivity feedback. The increased flow rate also enhances the heat transfer
+coefficient on the primary loop side of the heat exchanger and enables the
+reactor to operate at a higher power output. At the same time, the improved
+mixing flattens the temperature distribution in the core.
 

### 2020-06-24 neutronics.tex
Typo in nts chapter, complete dnp section in ss chapter, fix figures

@@
 temperatures by adjusting the fissile inventory and $k_{\text{eff}}$. In
-transient analysis, the $\beta_{\text{eff}}$ and reactivity coefficients are
-important as they dictate the duration and magnitude of the reactor
-response to a transient initiator. We calculated $\beta_{\text{eff}}$ by
+transient analysis, the delayed neutron fraction $\beta$ and reactivity
+coefficients $\alpha$ are
+important as they dictate the duration, shape, and magnitude of the reactor
+response to a transient initiator. The difference between $\beta$ and
+$\beta_{\text{eff}}$ is that $\beta$ is the physical delayed neutron fraction
+while $\beta_{\text{eff}}$ is the delayed neutron fraction weighted on the
+adjoint neutron flux. We will compare the $\beta$ value from Moltres to the
+$\beta_{\text{eff}}$ value from Serpent because Moltres currently lacks a
+adjoint calculation capability. We calculated $\beta$ by
 taking the relative difference between the $k_{\text{eff}}$ values with and
-without \glspl{DNP} in Table \ref{table:keff}. The $\beta_{\text{eff}}$ values
-at 973 K, shown in Table \ref{table:betaeff}, are in excellent agreement with
-a 4.43 pcm discrepancy.
+without \glspl{DNP} in Table \ref{table:keff}. The $\beta$ and
+$\beta_{\text{eff}}$ values at 973 K, shown in Table \ref{table:betaeff}, are
+in good agreement with a 4.43 pcm discrepancy.
 
@@
 	\centering
-	\caption{$\beta_{\text{eff}}$ values from Serpent and Moltres at 973 K.}
+	\caption{$\beta_{\text{eff}}$ and $\beta$ values from Serpent and Moltres,
+	respectively, at 973 K.}
 	\begin{tabular}{l S S}

### 2020-06-24 steadystate.tex
Typo in nts chapter, complete dnp section in ss chapter, fix figures

@@
 Figure \ref{fig:flow-temp} shows the temperature and velocity fields of the
-fuel salt in the core at steady state from Moltres and the Polimi/TUDelft
-models. The results from Moltres show good qualitative agreement with the
-Polimi and TUDelft models \cite{aufiero_development_2014}; we observe similar
-flow and hotspot features in all three models. There is a large recirculation
-region near the blanket tank walls arising from turbulent flow. Inertial
-forces dominate over viscous forces to form this large eddy. The figure also
-shows two relatively stagnant regions along the central axis at the top and
-bottom of the core. Temperature hotspots form in these regions of
-recirculation as convection is
-the dominant heat transfer mechanism. The maximum temperature from Moltres,
-1275 K near the bottom of the recirculation zone, is closer to the maximum
-temperature in the Polimi model ($\approx$ 1300 K) than the TUDelft model
-($\approx$ 1200K). The minimum temperature is 922 K at the inlet. Figure
-\ref{fig:flow} provides an alternate view of the flow profile through flow
-streamlines superimposed on the velocity magnitude distribution. The fastest
-salt flow occurs at the inlet, outlet, and at core half-height approximately
-0.40 m away from the central axis.
+fuel salt in the core at steady state from Moltres and the Polimi and TUDelft
+models. Figure \ref{fig:flow} provides an alternate view of the flow profile
+through flow streamlines superimposed on the velocity magnitude distribution.
+The results from Moltres show good qualitative agreement with the
+Polimi and TUDelft models \cite{fiorina_modelling_2014}; we observe similar
+flow and hotspot features in all three models. Furthermore, the highest salt
+velocities in all three models occur at the inlet, outlet, and at core
+half-height approximately 0.40 m away from the central axis. There is a large
+recirculation region near the blanket tank walls arising from turbulent flow.
+Inertial forces dominate over viscous forces to form this large eddy. The main
+difference between Moltres and Fiorina et al.'s models is the flow profile
+near the central axis at the top and bottom of the core.
+The Polimi and TUDelft models predict relatively stagnant flow in these
+regions without recirculation. Moltres, on the other hand, predicts explicit
+recirculating flow in these regions. This is due to our constant turbulent
+viscosity approximation in Moltres. The k-$\epsilon$ turbulence models in the
+Polimi and TUDelft models predict that the turbulent viscosity in these
+regions is as high as 100 Pa s, much higher than our 40 Pa s approximation.
+
+Nevertheless, similar temperature hotspots form in these regions of
+recirculation and stagnation as convection is the dominant heat transfer
+mechanism. The maximum temperature from Moltres, 1275 K near the bottom of the
+large recirculation zone, is closer to the maximum temperature in the Polimi
+model ($\approx$ 1300 K) than the TUDelft model ($\approx$ 1200K). Similarly,
+we observe cooler temperatures in high-velocity regions. The minimum
+temperature is 924 K at the inlet.
 
@@
 
-\begin{figure}[t!]
+\begin{figure}[b!]
     \centering
@@
     \caption{Neutron flux distributions in the core for neutron energy groups
-    1 to 6. The y and x axes represent height and radius values of the core
+    1 to 6. The y and x axes represent height and radius (in cm) of the core
     relative to the entire reactor geometry.}
@@
 
-\begin{table}[t!]
+\begin{table}[b!]
 	\centering
@@
 	\cite{fiorina_molten_2013}, and OpenFOAM \cite{aufiero_development_2014}
-	models. The COMSOL and OpenFOAM values were calculated with uniform
-	temperature distributions.}
-	\begin{tabular}{l S}
+	models along with the temperature distribution with which the values were
+	obtained.}
+	\begin{tabular}{l l S}
 		\toprule
-		{Model} & {Peak Neutron Flux [$\times 10^{15}$ cm$^{-2}$ s$^{-1}$]}
+		{Model} & {Temperature distribution} & {Peak Neutron Flux [$\times 10^{15}$ cm$^{-2}$ s$^{-1}$]}
 		\\
 		\midrule
-		{Moltres (This paper)} & 9.80\\
-		{COMSOL} & 8.6 \\
-		{OpenFOAM} & 9.0 \\
+		{Moltres (This paper)} & {Steady state} & 9.80\\
+		{COMSOL} & {Uniform, 973 K} & 8.6 \\
+		{OpenFOAM} & {Uniform, 973 K} & 9.0 \\
 		\bottomrule
@@
 
-As mentioned before, the \glspl{DNP} are mobile in \glspl{MSR} and their
-distributions do not directly correspond to the neutron flux distributions.
-The location where the \glspl{DNP} decay and emit neutrons impacts their
-neutron
-importance depending on their proximity to fissile and parasitic isotopes.
-Figure \ref{fig:dnp} shows the \gls{DNP} distributions for all eight \gls{DNP}
-groups. The
-precursors from the shortest-lived group (Group 8) predominantly decay within
-the core as their half-lives are shorter than the time it takes to reach the
-outlet while the precursors from the longest-lived group (Group 1) are
-effectively uniformly distributed due to their long half-lives relative to the
-average circulation time. For the longer-lived groups, the mesh elements with
-relatively high and low \gls{DNP} concentrations near the outlet and inlet,
-respectively, suggest that the mesh is too coarse. Thus, we would recommend
-further refinement for future work involving similar geometries.
-
-\begin{figure}[htb!]
+\begin{figure}[t!]
     \centering
@@
 
-\subsection{In-Core \texorpdfstring{$\beta_{\text{eff}}$}{beta_eff}}
+\subsection{Delayed Neutron Fraction}
 
-An important safety parameter for \glspl{MSR} is the actual in-core
-$\beta_{\text{eff}}$ value. This value represents the actual effective delayed
+As mentioned earlier, the delayed neutron precursors (DNPs) are mobile in
+\glspl{MSR} and their distributions do not directly correspond to the neutron
+flux distributions. The location where the \glspl{DNP} decay and emit neutrons
+impacts their neutron
+importance depending on their proximity to fissile and parasitic isotopes.
+Figure \ref{fig:dnp} shows the \gls{DNP} distributions for all eight \gls{DNP}
+groups. In general, we observe less \glspl{DNP} in the regions with fast salt
+flow, namely
+ The
+precursors from the shortest-lived group (Group 8) predominantly decay within
+the core as their half-lives are shorter than the time it takes to reach the
+outlet while the precursors from the longest-lived group (Group 1) are
+relatively evenly distributed due to their long half-lives. For the
+longer-lived groups, the \gls{DNP} concentrations are not well-resolved on the
+mesh elements adjacent to the outlet and the inlet boundaries, respectively.
+Thus, we recommend careful mesh refinement for future work involving similar
+geometries.
+
+\begin{figure}[b!]
+    \centering
+    \begin{subfigure}[t]{.30\textwidth}
+        \centering
+        \vspace{.9cm}
+        \includegraphics[width=\textwidth]{ss-pre}
+    \end{subfigure}
+    \begin{subfigure}[t]{.69\textwidth}
+        \centering
+        \vspace{0pt}
+        \includegraphics[width=\textwidth]{fiorina-pre}
+    \end{subfigure}
+    \caption{Total \gls{DNP} distribution in the core from Moltres
+    (left), Polimi (center), and TUDelft (right) models.}
+    \label{fig:pre}
+\end{figure}
+
+In Figure \ref{fig:pre}, we compare the total delayed neutron source
+distribution from Moltres with the results from the Polimi and TUDelft models
+\cite{fiorina_modelling_2014}. In contrast to Figure \ref{fig:dnp} which shows
+the precursor distribution, Figure \ref{fig:pre} shows the rate of delayed
+neutron emission, which we calculated by multiplying each \gls{DNP} group
+$C_i$ with its associated decay constant $\lambda_i$.
+We observe that the Polimi and TUDelft models
+feature greater \gls{DNP} retention in the stagnant regions within the core.
+This effect is less pronounced in the Moltres model. There is less build-up
+of \glspl{DNP} at the top of the core in Moltres because most of the
+\glspl{DNP} produced near the center of the core cannot enter the axial
+recirculation zone that only appears in Moltres. We also observe less build-up
+in the large recirculation zone near the blanket tank.
+
+The in-core delayed neutron fraction $\beta_c$ is an important safety
+parameter for \glspl{MSR}. This value represents the actual delayed
 neutron fraction in \glspl{MSR} after accounting for the loss of delayed
-neutrons due to out-of-core \gls{DNP} decay. Table * shows
+neutrons from \glspl{DNP} decaying outside the active core region. In general,
+reactors with smaller $\beta$ values exhibit greater prompt jumps in the
+neutron flux in response to reactivity insertions because they have a greater
+proportion of prompt neutrons under normal operating conditions. This is
+undesirable from a reactor safety perspective because it exposes the reactor
+to relatively more extreme conditions before active safety mechanisms
+activate and scram the reactor. In \glspl{MSR}, this danger is partly
+mitigated by the strong, negative fuel temperature reactivity coefficient. 
+Chapter 6 contains more in-depth discussions for various transient scenarios.
+
+In Table \ref{table:dnf}, we compare the fraction of out-of-core emissions and
+the $\beta_c$ values from Moltres with the Polimi and TUDelft models. We
+calculated the fraction of out-of-core emissions by finding the total amount
+of each \gls{DNP} group in the core and the outer loop, multiplying each total
+by their associated decay time constants $\lambda_i$, and calculating the
+proportion of emissions in the outer loop relative to the grand total. We
+calculated $\beta_c$ by first obtaining the prompt neutron emission rate from
+Moltres and subsequently using in-core delayed neutron emission rate from the
+previous calculation to find the fraction of delayed neutron emission rate
+relative to total emission rate.
+
+\begin{table}[t!]
+	\centering
+	\caption{The fraction of delayed neutrons lost from out-of-core emission
+	and the in-core delayed neutron fraction $\beta_c$ values from Moltres
+	(this paper), and the Polimi and TUDelft models
+	\cite{fiorina_modelling_2014}.}
+	\begin{tabular}{l S S}
+		\toprule
+		{Model} & {Out-of-Core Emission [\%]} & {$\beta_c$ [pcm]}
+		\\
+		\midrule
+		{Moltres (This paper)} & {44.16} & {184.9}\\
+		{Polimi} & {34.80} & {134.3} \\
+		{TUDelft} & {34.85} & {123.8} \\
+		\bottomrule
+	\end{tabular}
+	\label{table:dnf}
+\end{table}
+
+The fraction of out-of-core emissions from our Moltres model
+differs significantly by approximately 10\%, while $\beta_c$ differs by
+60-70 pcm. We attribute the former to the lesser \gls{DNP} retention in the
+stagnant flow regions in the core; the \glspl{DNP} are more evenly distributed
+along the entire primary loop, leading to more delayed neutron emissions in
+the outer loop region. The most likely reason for this is differences in the
+flow pattern in the recirculation zone because convective species transport
+dominates diffusive effects in the \gls{MSFR}. The exact flow pattern in the
+recirculation zones in the Polimi and TUDelft models is likely to differ from
+that in Moltres (Figure \ref{fig:flow}). Figure \ref{fig:flow-temp} also shows
+some minor differences in the magnitude of the flow in the recirculation zones
+between Moltres, and the Polimi and TUDelft models. Although the sizes of the
+arrows representing flow velocity are not normalized to the same scales, a
+quick comparison between the largest arrows and the arrows in the
+recirculation zone indicate that the recirculation zone in Moltres is
+relatively more stagnant. This could explain the concentration of \glspl{DNP}
+along an ``arc'' closer to the center of the core in Moltres as opposed to the
+more even distribution of \glspl{DNP} throughout the whole recirculation zones
+in the Polimi and TUDelft models. The higher peak \glspl{DNP} distribution in
+Moltres also supports this assertion.
+
+In spite of the greater delayed neutron losses, the $\beta_c$ value is higher
+in Moltres than the Polimi and TUDelft models. To account for this
+peculiarity, we note that Fiorina et al. \cite{fiorina_modelling_2014}
+applied adjoint flux weighting for their $\beta_c$ calculation while we report
+our value as the physical fraction without adjoint weighting. The weighting
+results in a significant difference in $\beta_c$ because a large fraction of
+the \glspl{DNP} decay in the recirculation zones, where the neutron importance
+is noticeably diminished.
 

### 2020-06-26 transient.tex
transient chapter figures and edits

@@
 \gls{MSFR} from Moltres for four accident scenarios. These scenarios, adapted
-from Fiorina et al.'s work, include
-unprotected instances of reactivity insertion, loss of flow, pump overspeed,
-and loss of heat sink. The term ``unprotected'' means that there is no
+from Fiorina et al.'s work \cite{fiorina_modelling_2014}, include unprotected
+instances of reactivity insertion, loss of heat sink, loss of flow, and pump
+overspeed accidents. The term ``unprotected'' means that there is no
 external intervention in these scenarios. As such, these simulations
@@
 in the previous chapter as the initial conditions for the transient
-simulations discussed in this chapter. In essence, we imported all
+simulations discussed in this chapter. Specifically, we imported all
 steady-state spatial values for neutron flux, delayed neutron precursor
@@
 As noted by Fiorina et al. \cite{fiorina_modelling_2014}, explicit decay heat
-modeling has a negligible effect on reactivity-, pump-, and enhanced
-cooling-initiated transients. Furthermore, only their Polimi model had decay
+modeling has a negligible effect in reactivity-, and pump-initiated
+transients. Furthermore, only their Polimi model had decay
 heat modeling capabilities. Therefore, they presented results from their
-Polimi and TUDelft model for the four accident scenarios 
-without decay heat modeling; they enabled decay heat modeling for the
-loss of heat sink transient scenario only. For a fair comparison, we also ran
-all transient scenarios without the decay heat model with the exception of the
-loss of heat sink scenario in which we ran the simulations with and without
+Polimi and TUDelft model for the four accident scenarios  without decay heat
+modeling; they enabled decay heat modeling for the loss of heat sink transient
+scenario only. For a direct comparison, we also ran all transient simulations
+without the decay heat model with the exception of the loss of heat sink
+scenario in which we ran two separate simulations with and without
 the decay heat model. More generally, we imposed various other simplifying
@@
 
-We modeled two unprotected reactivity insertion scenarios in Moltres by
-swapping out the original set of group constant data with two new, separate
+We modeled two unprotected step-wise reactivity insertion scenarios in Moltres
+by swapping out the original set of group constant data with two new, separate
 sets of data from Serpent corresponding to 50 pcm and 200 pcm reactivity
@@
 
+Figures \ref{fig:50pcmheat} and \ref{fig:50pcmtemp} show the power output and
+average core temperature increase following the 50 pcm step-wise reactivity
+insertion in the Moltres, Polimi, and TUDelft models. The initial prompt
+response to the reactivity insertion raises power to 4 GW by $t=0.001$ s.
+Figure \ref{fig:50pcmjump} shows the rise in power output specifically during
+the prompt response in a separate plot.
+Power continues to rise at a slower rate up to 4.63 GW at around $t=0.005$ s,
+at which point the negative reactivity from the Doppler effect and salt
+expansion becomes greater
+than the initial +50 pcm insertion. Power continues to fall as the average
+core temperature rises. There is a slight change in slope at $t=0.3$ s. This
+value approximately corresponds to the average half-life of the two
+shortest-lived delayed neutron precursor (DNP) groups ($t_{1/2}=0.195$ s and
+$0.424$ s); the decay of the surplus precursors produced in the initial phase
+negates a fraction of the negative reactivity from the elevated core
+temperature. By $t=3$-$4$ s, most of the heated salt and \glspl{DNP} from the
+initial phase will have circulated around the primary loop and returned to the
+core. The heated salt causes a small, noticeable dip in power before the power
+output stabilises. The core tends to a new equilibrium average temperature
+approximately 7.5 K higher than the initial average temperature.
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.85\textwidth]{50pcm-heat}
+    \caption{Power output following
+    a 50 pcm step-wise reactivity insertion in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:50pcmheat}
+%
+    \centering
+    \includegraphics[width=.85\textwidth]{50pcm-temp}
+    \caption{Average core temperature increase following
+    a 50 pcm step-wise reactivity insertion in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:50pcmtemp}
+\end{figure}
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.7\textwidth]{50pcm-jump}
+    \caption{Power output during the prompt response following
+    a 50 pcm step-wise reactivity insertion in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:50pcmjump}
+\end{figure}
+
+The results from Moltres show good agreement with the results from the Polimi
+and TUDelft models; Moltres reproduced all of the individual features in both
+plots. The most significant difference is in the magnitude of the reactor
+response. Moltres predicts a smaller peak in the power output and a smaller
+overall increase in the average core temperature. This is largely due to the
+more negative temperature reactivity coefficient in Moltres than in the Polimi
+and TUDelft models. The temperature reactivity coefficient $\alpha_T$ in
+Moltres is $-7.184$ pcm K$^{-1}$ (Table \ref{table:alpha}), as opposed to
+approximately $-6.5$ pcm K$^{-1}$ within the relevant temperature range in the
+Polimi and TUDelft models. Therefore, we see a smaller temperature increase in
+the Moltres model for the same reactivity insertion. Multiplying the average
+core temperature increase at $t=10$ s with $\alpha_T$ gives us $-7.184$ pcm
+K$^{-1} \times 7.46$ K $= -53.6$ pcm, which is approximately equal to the 50
+pcm reactivity insertion.
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.85\textwidth]{200pcm-heat}
+    \caption{Power output following
+    a 200 pcm step-wise reactivity insertion in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:200pcmheat}
+%
+    \centering
+    \includegraphics[width=.85\textwidth]{200pcm-temp}
+    \caption{Average core temperature increase following
+    a 200 pcm step-wise reactivity insertion in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:200pcmtemp}
+\end{figure}
+
+The results for the 200 pcm reactivity insertion scenario show similar trends
+to the 50 pcm case. The larger reactivity insertion illicits a stronger
+prompt response in the power output which peaks at 92.1 GW. The average core
+temperature increases much more rapidly and subsequently triggers a sharper drop in power output. This results in the clearer distinction in the rate of
+core temperature increase before and after $t=0.01$ s. In this transient, we
+also observe greater deviation between Moltres and the other models arising
+from the differences in the temperature reactivity coefficients. Overall,
+Moltres' results show good agreement with the Polimi and TUDelft results and
+the differences arise mainly due to the differences in the temperature
+reactivity coefficients.
+
+\clearpage
+
+\section{Unprotected Loss of Heat Sink}
+
+An unprotected loss of heat sink accident can occur when the pumps in the
+intermediate loop fail. The heat exchangers would then lose most of their
+cooling capabilities. We followed Fiorina et al.'s approach in assuming that
+the cooling from the heat exchangers decreases exponentially with a time
+constant of 1 s and all other parameters held constant
+\cite{fiorina_modelling_2014}. As mentioned in the Chapter \ref{chap:ss}, we
+will present two sets of results for this transient: 1) without decay heat modeling, and 2) with decay heat modeling.
+
+\subsection{Without Decay Heat}
+
+Figures \ref{fig:lohsheat} and \ref{fig:lohstemp} show the power output and
+average core temperature increase during the unprotected loss of heat sink
+transient in the Moltres, Polimi, and TUDelft models without decay heat
+modeling. The power output and average core temperature show little change in
+the first two seconds as it takes approximately that amount of time for the
+partially cooled salt to migrate to the center of the core. At $t=2$ s, we
+observe a sharp spike in average core temperature and a corresponding drop
+in power output. The presence of delayed neutron precursors (DNPs) from the
+steady-state operating conditions momentarily halt the increase in temperature
+at around $t=5$ s. The average core temperature continues to rise while the
+power output falls through the rest of the transient.
+
+The results from Moltres show good agreement with the results from the Polimi
+and TUDelft. Moltres reproduced all of the trends that we see in the Polimi
+and TUDelft models. The temporary halt in the temperature increase occurs at
+a lower average core temperature for Moltres than the other two models. This
+is likely due to the difference in the temperature reactivity coefficient
+that we discussed in the reactivity insertion results; a smaller increase in
+the average core temperature produces the same decrease in power output.
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.85\textwidth]{lohs-heat}
+    \caption{Power output during
+    an unprotected loss of heat sink transient in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:lohsheat}
+    \includegraphics[width=.85\textwidth]{lohs-temp}
+    \caption{Average core temperature increase during
+    an unprotected loss of heat sink transient in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:lohstemp}
+\end{figure}
+
+\subsection{With Decay Heat}
+
+\clearpage
+
 \section{Unprotected Loss of Flow}
@@
 changes becomes the dominant driving force for salt flow in the primary loop.
-While Fiorina et al. \cite{fiorina_modelling_2014} and Aufiero et al.
-\cite{aufiero_development_2014} applied the Boussinesq approximation for
-bouyancy-driven flow in their models, it is not directly applicable to the
-\gls{MSFR} model in this thesis because we partitioned the primary loop into
-two separate geometries and used Dirichlet boundary conditions to drive flow.
-Fiorina et al.'s Polimi and TUDelft models featured complete exponential
-coast-downs of the pumps with a time constant of 5 s. The resulting flow rate
-from natural circulation was approximately 18 times smaller than the initial
-flow rate. Thus, for the \gls{MSFR} model in Moltres, we imposed a similar
-exponential decay term with the same time constant on the inflow Dirichlet
-boundary condition to reduce the flow rate to the same final flow rate
-observed by Fiorina et al. Figure * shows the change in flow rate over the
-transient simulation.
+While Fiorina et al. \cite{fiorina_modelling_2014} applied the Boussinesq
+approximation for bouyancy-driven flow in their models, it is not directly
+applicable to the \gls{MSFR} model in this thesis because we partitioned the
+primary loop into two separate geometries and used Dirichlet boundary
+conditions at the inlet to drive flow. Fiorina et al.'s Polimi and TUDelft
+models featured complete exponential coast-downs of the pumps with a time
+constant of 5 s. The resulting flow rate from natural circulation was
+approximately 18 times smaller than the initial flow rate. Figure
+\ref{fig:flowrate} shows that the actual flow rate decreased with a time
+constant of 8 s. Thus, for the \gls{MSFR} model in Moltres, we imposed a
+similar exponential decay term with a time constant of 8 s on the inflow
+Dirichlet boundary condition:
+%
+\begin{align}
+    \text{Flow rate, } v = 0.25862 + (4.5-0.25862) e^{-t/8} 
+    \label{eq:flowrate}
+\end{align}
 
-We also expect the reduced flow rate to decrease the heat transfer rate
-between the primary and secondary loop through the heat exchanger as the heat
-transfer coefficient is dependent on the flow rate. Given the limited
-information available on the heat exchanger specifications, we assumed that
-the heat transfer coefficient is directly proportional to the flow rate, and
-the overall heat transfer rate retains its linear dependence to the
-temperature difference between the primary and secondary loop at the location
-of the heat exchanger as shown previously in \ref{eq:hx}. Lastly, we assume
-the turbulent viscosity, $\mu_t$, is directly proportional to the flow rate to
-account for reduced mixing when the flow rate decreases.
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.7\textwidth]{lof-flow-rate}
+    \caption{The change in flow rate in the Polimi and TUDelft models and the
+    imposed flow rate in Moltres.}
+    \label{fig:flowrate}
+\end{figure}
+
+The reduced flow rate also decreases the heat transfer rate between the
+primary and intermediate loop through the heat exchanger as the heat transfer
+coefficient is dependent on the flow rate. This step was problematic because
+the pointwise heat exchanger implementation in Moltres performs differently
+compared to the heat exchangers that take up ``36\% of the out-of-core part''
+in the Polimi and TUDelft models \cite{fiorina_modelling_2014}. Most of the
+cooling happens in the top half of the heat exchanger where the temperature
+differential between the primary and intermediate loops is the largest. In the
+Polimi and TUDelft models, the overall heat transfer coefficient is the
+``harmonic mean of the heat transfer coefficients on each side of the heat
+exchanger''. For this loss of flow transient, the authors intended to focus on
+the primary loop and assumed that only the pumps in the primary loop failed.
+In addition to this, the authors applied the Dittus-Boelter correlation (cite
+*) for
+the relationship between the primary side heat transfer coefficient and the
+flow rate. The Dittus-Boelter correlation for fluids being cooled is:
+%
+\begin{align}
+    Nu &= 0.023 Re^{0.8} Pr^{0.3} \label{eq:db} \\
+    \intertext{where}
+    Nu &= \text{ Nusselt number,} \nonumber \\
+    Re &= \text{ Reynolds number,} \nonumber \\
+    Pr &= \text{ Prandtl number.} \nonumber
+\end{align}
+
+The only direct relation to flow velocity in the Dittus-Boelter correlation is
+through the $Re$ term, which is directly proportional to flow velocity. This
+gives the following relation between the heat transfer coefficient $h$ and
+flow velocity $v$:
+%
+\begin{align}
+    h \propto v^{0.8} \label{eq:hv}
+\end{align}
+%
+However, this relation provided very different results in the unprotected loss
+of flow and pump overspeed transients compared to the Polimi and TUDelft
+models; it overpredicted the equilibrium power output in the
+unprotected loss of flow transient and underpredicted the same parameter in
+the unprotected pump overspeed transient. Upon further investigation, we found
+that raising the power of $v$ from 0.8 to 1.1 brought the average core
+temperatures closer to the results from the other models in both transients.
+Therefore, we adopted the raised power in this study.
+
+Another issue pertained to the turbulent viscosity $\mu_t$ as a function of
+$v$. When we made a simple approximation of $\mu_t$ being directly
+proportional to $v$, the results
+differed significantly compared to the Polimi and TUDelft models (Appendix *).
+This may have been due to buoyancy-driven flow contributing to turbulence; the
+turbulent energy $k$ equation in the $k$-$\epsilon$ model has an explicit
+source term from buoyancy effects (Cite *). Another point to note is the
+Reynolds number does not change if $\mu$ and $v$ decrease in tandem. This
+preserves the existence of the recirculation zone in the core and it is at
+odds with the results from the Polimi and TUDelft models, which show that the
+recirculation zones disappear during the loss of flow transient. We
+circumvented this issue by letting higher fractions of $\mu_t$ be conserved
+regardless of the final flow velocity. This measure allowed for laminar flow
+to develop in the core and yielded results resembling those from the Polimi
+and TUDelft models.
+
+Figures * show
 
@@
 In the context of this thesis, pump overspeed refers to a sustained
-increase in pump speed in the primary coolant loop. We followed Fiorina et
-al.'s implementation by ramping up the inlet velocity, $u$, by 50\% from
-the nominal value, $u_0$, according to the following formula:
-
-\begin{align}
-    u(t) &= u_0 [1 + 0.5 (1 - e^{-t / \tau})] \\
-    \intertext{where}
-    \tau &= 5 \text{ s.} \nonumber
-\end{align}
-
-We retain the directly proportional assumption between the flow rate and
-$\mu_t$ that we imposed in the unprotected loss of flow simulation.
-The increased flow rate impacts reactor performance in several ways.
+increase in pump speed in the primary coolant loop. The increased flow rate
+impacts reactor performance in several ways.
 It affects the neutronics by reducing the in-core $\beta_eff$ as more of the
@@
 
-\section{Unprotected Loss of Heat Sink}
+We followed Fiorina et al.'s implementation \cite{fiorina_modelling_2014} by
+ramping up the inlet velocity, $u$, by 50\% from the nominal value, $u_0$,
+according to the following formula:
 
-\subsection{Without Decay Heat Model}
+\begin{align}
+    u(t) &= u_0 [1 + 0.5 (1 - e^{-t / \tau})] \\
+    \intertext{where}
+    \tau &= 5 \text{ s.} \nonumber
+\end{align}
+%
+For this transient, we assumed that $\mu_t$ was directly proportional to $v$
+because the buoyancy effects are negligible and the recirculation zones 
+persist throughout the entire duration. Figures \ref{fig:poheat} and
+\ref{fig:potemp} show the power output and average core temperature increase
+during the unprotected pump overspeed transient in the Moltres, Polimi, and
+TUDelft models.
 
-\subsection{With Decay Heat Model}
+Unlike in the loss of flow transient, the $\mu_t \propto v$ approximation
+worked well in pump overspeed transient.
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.85\textwidth]{po-heat}
+    \caption{Power output during
+    an unprotected pump overspeed transient in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:poheat}
+    \includegraphics[width=.85\textwidth]{po-temp}
+    \caption{Average core temperature increase during
+    an unprotected pump overspeed transient in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:potemp}
+\end{figure}

### 2020-06-28 conclusion.tex
Transients and conclusion chapter. Update bib. Add figures. Minor edits.

The unique phenomena in \glspl{MSR}, generally not found in conventional
reactors, necessitate the development of new reactor safety analysis software.
This thesis presents the latest developments in Moltres, namely coupling its
existing neutron diffusion module to the incompressible Navier-Stokes module
in MOOSE, and developing a decay heat model for short-term transients. We
demonstrated and verified some of its current capabilities through a static
neutronics study, and a coupled neutronics/thermal-hydraulics safety analysis
of the \gls{MSFR} concept.

The neutronics study showed good agreement between Moltres and Serpent. With
the relevant group constant data from Serpent,
Moltres could accurately replicate the $k_{eff}$, $\beta$,
$\alpha_T$, and multi-group neutron flux results from Serpent. The
$k_{eff}$ estimates from Moltres were approximately 100 pcm higher for all
measurements between 800 K and 1400 K. This discrepancy is notably smaller
than the discrepancies observed in the neutron diffusion and SP3 models
developed in OpenFOAM \cite{aufiero_extended_2013}. The $\beta$ and $\alpha_T$
values from Moltres had 1.46\% and 0.265\% discrepancies, respectively, to
Serpent's results. Lastly, the normalized six-group neutron flux from Moltres
and Serpent were visually indistinguishable from each other. The journal
article that introduced Moltres \cite{lindsay_introduction_2018} verified its
neutron diffusion model for a two-group thermal-spectrum \gls{MSBR} model;
the results of this study extends code-to-code verification of Moltres'
neutron diffusion model with the six-group, fast-spectrum \gls{MSFR} model.

Although Moltres currently lacks a proper turbulence model, our simplifying
assumption for the turbulent viscosity $\mu_t$ yielded good results for most
of the \gls{MSFR}
steady-state and transient analyses. The steady-state temperature and velocity
distributions showed many similiarities in their shapes and magnitudes to the
Polimi and TUDelft model results \cite{fiorina_modelling_2014}. Our
uniform $\mu_t$ assumption accounted for the minor differences in the flow at
the top of the core and the loss of delayed neutrons to out-of-core emissions.
The results with the decay heat model showed a slight flattening of the
temperature distribution in the core that is in line with our expectations
given the diffusion and advection of the decay heat precursors.

The unprotected reactivity insertion and loss of heat sink results showed the
same trends Fiorina et al. observed in their Polimi and TUDelft models. The
small difference in the temperature reactivity coefficient accounted for the
small difference in the magnitude of the peaks in power output and average
core temperature increase. The differences between the pointwise heat
exchanger in Moltres and the volumetric heat exchanger in the other two
models required minor adjustments in the relationship between flow rate
$\dot{m}$ and the heat transfer coefficient $h$ from the original
Dittus-Boelter correlation for the pump-initiated transients. Assuming a
directly proportional relationship between $\mu_t$ and $\dot{m}$ yielded
results in good agreement with the other two models for the pump overspeed
transient. However, Moltres performed poorly in the loss of flow transient as
we could not incorporate buoyancy-driven flow and its associated effects on
$\mu_t$.

Overall, we have demonstrated that Moltres can handle most of the case
studies that this thesis covered.

\section{Future Work}

Further research and development on Moltres should aim to rectify the issues
mentioned in this thesis. There are three main avenues for improvement.
Firstly, proper 2D/3D heat exchanger implementation
would allow us to move away from the 1D outer loop system and towards a full
2D/3D closed loop. The biggest change from this is users would be able to use
the Boussinesq approximation for buoyancy-driven flow capability in Moltres.
Buoyancy-driven flow is a critical component in loss of forced flow scenarios
and these scenarios in turn are important accident transients in reactor
safety analyses.

Secondly, Moltres would benefit from a proper turbulence model such as the
$k$-$\epsilon$ or $k$-$\omega$ turbulence models. Turbulence effects are
significant in \gls{MSR} designs with fast flow, and they inform optimization
studies for improving flow patterns and eliminating local hotspots in eddies.
Our work in this thesis, particularly for the loss of flow transient, shows
that simplifying assumptions for turbulence lead to erroneous results under
flow conditions that deviate significantly from steady state.

Lastly, a compressible Navier-Stokes model would be essential for modeling
compressible flow effects such as variable temperature-dependent density
changes following a large reactivity insertion and finite wave propagation
speeds in a fluid. The presence of bubbles in the core from the gas sparging
system increases fuel compressibility and enhances compressibility effects
\cite{cervi_development_2019}.

### 2020-06-28 neutronics.tex
Transients and conclusion chapter. Update bib. Add figures. Minor edits.

@@
 \ref{table:keff} shows the $k_{\text{eff}}$ values from Serpent and Moltres at
-973 K and the corresponding salt density, while Table \ref{table:ktemp} shows
+973 K and the corresponding salt density, and Table \ref{table:ktemp} shows
 the $k_{\text{eff}}$ values for other temperatures at 100 K intervals. We
@@
 		\midrule
-		{Serpent} & -4034 \pm 14 & 3.737 \pm 0.013 & 3.424 \pm 0.013 & 7.165
-		\pm 0.013 \\
-		{Moltres} & {-} & {-} & {-} & 7.184\\
+		{Serpent} & -4034 \pm 14 & -3.737 \pm 0.013 & -3.424 \pm 0.013 &
+		-7.165 \pm 0.013 \\
+		{Moltres} & {-} & {-} & {-} & -7.184\\
 		\bottomrule
@@
 parameters for modeling transient reactor behavior, are in excellent
-agreement. While the
-$k_{\text{eff}}$ values have discrepancies on the order of 100 pcm, they are
-relatively small compared to the $k_{\text{eff}}$ values from the neutron
-diffusion and SP3 models in OpenFOAM \cite{cervi_development_2019}.
+agreement. The $k_{\text{eff}}$ values have discrepancies on the order of 100
+pcm and they are relatively small compared to the $k_{\text{eff}}$ values from
+the neutron diffusion and SP3 models in OpenFOAM
+\cite{cervi_development_2019}.

### 2020-06-28 steadystate.tex
Transients and conclusion chapter. Update bib. Add figures. Minor edits.

@@
     (left), Polimi (center), and TUDelft (right) models. The colors represent
-    temperature according to the respective colorbars while the arrows
+    temperature according to the respective colorbars and the arrows
     represent velocity fields.}
@@
 
-While the temperatures at the hotspots are well below the melting point of the
-Ni-alloy structure (1500 K), they may cause undue thermal stress on the
+Although the temperatures at the hotspots are well below the melting point of the Ni-alloy structure (1500 K), they may cause undue thermal stress on the
 blanket tank structure and induce relatively faster salt corrosion rates. A
@@
     \end{subfigure}
-    \caption{Total \gls{DNP} distribution in the core from Moltres
-    (left), Polimi (center), and TUDelft (right) models.}
+    \caption{Total delayed neutron source distribution in the core from
+    Moltres (left), Polimi (center), and TUDelft (right) models.}
     \label{fig:pre}
@@
 The fraction of out-of-core emissions from our Moltres model
-differs significantly by approximately 10\%, while $\beta_c$ differs by
+differs significantly by approximately 10\%, and $\beta_c$ differs by
 60-70 pcm. We attribute the former to the lesser \gls{DNP} retention in the
@@
 peculiarity, we note that Fiorina et al. \cite{fiorina_modelling_2014}
-applied adjoint flux weighting for their $\beta_c$ calculation while we report
-our value as the physical fraction without adjoint weighting. The weighting
-results in a significant difference in $\beta_c$ because a large fraction of
-the \glspl{DNP} decay in the recirculation zones, where the neutron importance
-is noticeably diminished.
+applied adjoint flux weighting for their $\beta_c$ calculation whereas we
+report our value as the physical fraction without adjoint weighting. The
+weighting results in a significant difference in $\beta_c$ because a large
+fraction of the \glspl{DNP} decay in the recirculation zones, where the
+neutron importance is noticeably diminished.
 

### 2020-06-28 transient.tex
Transients and conclusion chapter. Update bib. Add figures. Minor edits.

@@
 prompt response in the power output which peaks at 92.1 GW. The average core
-temperature increases much more rapidly and subsequently triggers a sharper drop in power output. This results in the clearer distinction in the rate of
+temperature increases much more rapidly and subsequently triggers a sharper
+drop in power output. This results in the clearer distinction in the rate of
 core temperature increase before and after $t=0.01$ s. In this transient, we
@@
 \cite{fiorina_modelling_2014}. As mentioned in the Chapter \ref{chap:ss}, we
-will present two sets of results for this transient: 1) without decay heat modeling, and 2) with decay heat modeling.
+will present two sets of results for this transient: 1) without decay heat
+modeling, and 2) with decay heat modeling.
 
@@
     an unprotected loss of heat sink transient in the Moltres, Polimi, and
-    TUDelft models \cite{fiorina_modelling_2014}.}
+    TUDelft models \cite{fiorina_modelling_2014} without decay heat.}
     \label{fig:lohsheat}
@@
     an unprotected loss of heat sink transient in the Moltres, Polimi, and
-    TUDelft models \cite{fiorina_modelling_2014}.}
+    TUDelft models \cite{fiorina_modelling_2014} without decay heat.}
     \label{fig:lohstemp}
 \end{figure}
+
+\clearpage
 
@@
 changes becomes the dominant driving force for salt flow in the primary loop.
-While Fiorina et al. \cite{fiorina_modelling_2014} applied the Boussinesq
-approximation for bouyancy-driven flow in their models, it is not directly
-applicable to the \gls{MSFR} model in this thesis because we partitioned the
+Fiorina et al. \cite{fiorina_modelling_2014} applied the Boussinesq
+approximation for bouyancy-driven flow in their models, but we were unable to
+follow this procedure because we partitioned the
 primary loop into two separate geometries and used Dirichlet boundary
@@
 models featured complete exponential coast-downs of the pumps with a time
-constant of 5 s. The resulting flow rate from natural circulation was
-approximately 18 times smaller than the initial flow rate. Figure
-\ref{fig:flowrate} shows that the actual flow rate decreased with a time
+constant of 5 s. The resulting flow rate $\dot{m}$ from natural circulation was
+approximately 18 times smaller than the initial $\dot{m}$. Figure
+\ref{fig:flowrate} shows that the actual $\dot{m}$ decreased with a time
 constant of 8 s. Thus, for the \gls{MSFR} model in Moltres, we imposed a
@@
 
-The reduced flow rate also decreases the heat transfer rate between the
+The reduced $\dot{m}$ also decreases the heat transfer rate between the
 primary and intermediate loop through the heat exchanger as the heat transfer
-coefficient is dependent on the flow rate. This step was problematic because
+coefficient $h$ is dependent on the $\dot{m}$. This step was problematic
+because
 the pointwise heat exchanger implementation in Moltres performs differently
@@
 differential between the primary and intermediate loops is the largest. In the
-Polimi and TUDelft models, the overall heat transfer coefficient is the
+Polimi and TUDelft models, the overall $h$ is the
 ``harmonic mean of the heat transfer coefficients on each side of the heat
@@
 the primary loop and assumed that only the pumps in the primary loop failed.
-In addition to this, the authors applied the Dittus-Boelter correlation (cite
-*) for
-the relationship between the primary side heat transfer coefficient and the
-flow rate. The Dittus-Boelter correlation for fluids being cooled is:
+In addition to this, the authors applied the Dittus-Boelter correlation
+\cite{dittus_heat_1930} for the relationship between the primary side $h$ and
+$\dot{m}$. The Dittus-Boelter correlation for fluids being cooled is:
 %
@@
 \end{align}
-
-The only direct relation to flow velocity in the Dittus-Boelter correlation is
-through the $Re$ term, which is directly proportional to flow velocity. This
-gives the following relation between the heat transfer coefficient $h$ and
-flow velocity $v$:
+%
+The only direct relation to $\dot{m}$ in the Dittus-Boelter correlation is
+through the $Re$ term, which is directly proportional to flow velocity $v$.
+This gives the following relation between $h$ and$v$:
 %
@@
 $v$. When we made a simple approximation of $\mu_t$ being directly
-proportional to $v$, the results
-differed significantly compared to the Polimi and TUDelft models (Appendix *).
-This may have been due to buoyancy-driven flow contributing to turbulence; the
-turbulent energy $k$ equation in the $k$-$\epsilon$ model has an explicit
-source term from buoyancy effects (Cite *). Another point to note is the
+proportional to $v$, the results differed significantly compared to the Polimi
+and TUDelft models (Appendix *). This is likely due to buoyancy-driven flow
+contributing to turbulence; the turbulent energy $k$ equation in COMSOL's
+$k$-$\epsilon$ model has an explicit source term from buoyancy effects
+\cite{comsol_ab_comsol_2018}. Another point to note is the
 Reynolds number does not change if $\mu$ and $v$ decrease in tandem. This
@@
 recirculation zones disappear during the loss of flow transient. We
-circumvented this issue by letting higher fractions of $\mu_t$ be conserved
-regardless of the final flow velocity. This measure allowed for laminar flow
-to develop in the core and yielded results resembling those from the Polimi
-and TUDelft models.
+circumvented this issue by letting fixed fractions of the initial $\mu_{t,0}$
+be conserved regardless of the final flow velocity, according to the following
+equation:
+%
+\begin{align}
+    \mu_t &= \mu_c + (\mu_{t,0} - \mu_c) e^{-t/8} \label{lofmu} \\
+    \intertext{where}
+    \mu_c &= \text{ conserved fraction of $\mu_{t,0}$ [Pa$\cdot$s].} \nonumber
+\end{align}
+%
+This measure allowed for laminar flow to develop in the core and yielded
+results with closer resemblance to those from the Polimi and TUDelft models.
 
-Figures * show
+Figures \ref{fig:lofheat} and \ref{fig:loftemp} show the power output and
+average core temperature increase during the unprotected loss of flow
+transient in the Moltres, Polimi, and TUDelft models without decay heat
+modeling. The three sets of results from Moltres correspond to $\mu_c =
+\frac{1}{4} \mu_{t,0}, \frac{1}{2} \mu_{t,0}, \text{and } \frac{3}{4}
+\mu_{t,0}$. Moltres
+performs poorer in this transient relative to the two previous transients.
+Although Moltres shows the same decreasing trend in power output, it failed to
+capture the exact individual features in the reactor response. In the Polimi 
+and TUDelft models, Fiorina et al.
+stated that after around $t=15$ s, the ``flow pattern changed in the core and
+the recirculation zones started to disappear''. As a result, there is a sudden
+drop in the average core temperature as the pocket of hot salt leaves the
+core. In Moltres, the wider peak in the average core temperature
+indicates that there was a more gradual change in the flow pattern.
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.85\textwidth]{lof-heat}
+    \caption{Power output during
+    an unprotected loss of flow transient in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:lofheat}
+    \includegraphics[width=.85\textwidth]{lof-temp}
+    \caption{Average core temperature increase during
+    an unprotected loss of flow transient in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:loftemp}
+\end{figure}
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.37\textwidth]{lof-flow-temp}
+    \includegraphics[width=.62\textwidth]{fiorina-lof}
+    \caption{Temperature and velocity fields in the core at $t=300$ s during
+    a loss of flow transient in the Moltres ($\mu_c = \frac{1}{2} \mu_{t,0}$),
+    Polimi, and TUDelft models.}
+    \label{fig:lofflowtemp}
+\end{figure}
+
+Figure \ref{fig:lofflowtemp} shows the flow patterns and temperature
+distribution in the core at $t=300$ s in all three models. Figures
+\ref{fig:lofheat}, \ref{fig:loftemp}, and \ref{fig:lofflowtemp} combined
+highlight the difference between laminar flow in the Moltres model and
+buoyancy-driven flow in the other models, and its impact on the reactor
+response. They show that low-speed laminar flow is a poor substitute for
+buoyancy-driven flow in the context of the MSFR. It is particularly evident
+in the transition from high-speed turbulent flow to low-speed viscous
+flow as Moltres did not capture the intermediate stages accurately. The
+simplifying assumption for the uniform, time-dependent $\mu_t$ is also not
+ideal for a safety analysis code.
+
+The results from this transient inform our goals for Moltres: 1)
+implementing a proper turbulence model, and
+2) developing a new heat exchanger feature that is compatible with the
+buoyancy-driven flow capabilities already present in Moltres.
+
+\clearpage
 
@@
 
-In the context of this thesis, pump overspeed refers to a sustained
+Pump overspeed refers to a sustained
 increase in pump speed in the primary coolant loop. The increased flow rate
-impacts reactor performance in several ways.
-It affects the neutronics by reducing the in-core $\beta_eff$ as more of the
+$\dot{m}$ impacts reactor performance in several ways.
+It affects the neutronics by reducing the in-core $\beta$ as more of the
 shorter-lived precursors will tend to flow out of the core before decaying.
@@
 core temperatures to fall to counteract this change through temperature
-reactivity feedback. The increased flow rate also enhances the heat transfer
+reactivity feedback. The increased $\dot{m}$ also enhances the heat transfer
 coefficient on the primary loop side of the heat exchanger and enables the
@@
 according to the following formula:
-
+%
 \begin{align}
-    u(t) &= u_0 [1 + 0.5 (1 - e^{-t / \tau})] \\
+    u(t) &= u_0 [1 + 0.5 (1 - e^{-t / \tau})] \label{eq:overspeed}
     \intertext{where}
@@
 during the unprotected pump overspeed transient in the Moltres, Polimi, and
-TUDelft models.
-
-Unlike in the loss of flow transient, the $\mu_t \propto v$ approximation
-worked well in pump overspeed transient.
+TUDelft models. Figure \ref{fig:poshort} shows the same results for the first
+20 seconds of the transient. At the start of the transient, the rising flow
+rate cools the core and causes power output to rise sharply. Although the
+average core temperature has a strictly decreasing trend, the temperature at
+the center of the core briefly rises due ot the sharp increase in power
+output. Since this is the region where most of the fissions take place, the
+Doppler effect and salt expansion causes the power output to stall and dip
+briefly before starting to rise again at $t=2.5$ s. The reactor tends to a new
+equilibrium power output and average core temperature. The temperature
+distribution in the core is more evenly distributed because the turbulent
+thermal conductivity $k_t$ is directly proportional to $mu_t$.
 
@@
 \end{figure}
+
+\begin{figure}[htbp!]
+    \centering
+    \begin{subfigure}[t]{.485\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{po-heat-short}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.485\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{po-temp-short}
+    \end{subfigure}
+    \caption{The first 20 s of the power output and average core temperature
+    increase during an unprotected pump overspeed transient.}
+    \label{fig:poshort}
+\end{figure}
+
+The results show excellent agreement with the Polimi and TUDelft models. The
+average core temperature increase in particular reproduces the Polimi and   
+TUDelft results very well and the curve falls between the other two curves.
+The power output is higher because the Moltres \gls{MSFR} model has a
+stronger negative temperature coefficient than the other two models.

### 2020-07-03 introduction.tex
editing

@@
 change, a global phenomenon characterized by shifting weather patterns and
-rising temperatures. Climate change has dire consequences on human health and
-safety due to extreme weather events and the overall impact on food
-production. As a result, urgent measures are necessary to limit greenhouse gas
-emissions, namely CO$_2$ emissions from electricity generation.
+rising temperatures. The concentration of CO$_2$, a greenhouse gas, in the
+atmosphere has been rising steadily throughout modern human history
+\cite{us_department_of_commerce_global_nodate}. Climate change has dire
+consequences on human health and safety due to extreme weather events and the
+overall impact on food production \cite{noauthor_global_nodate}. As a result,
+urgent measures are necessary to limit greenhouse gas emissions, namely CO$_2$
+emissions from electricity generation.
 
@@
 advantage over renewable energy sources is that it provides consistent
-base-load power independent of weather and geographical location (cite mit).
-Employing a diverse mix of these energy sources ensures energy security and
-reliability during and after our transition towards a low-carbon future (cite mit).
+base-load power independent of weather and geographical location
+\cite{petti_future_2018}. Employing a diverse mix of these energy sources
+ensures energy security and reliability in our transition
+towards a low-carbon future \cite{petti_future_2018}.
 
@@
 include perceived safety risks, sustainability concerns, nuclear proliferation
-risks, and the ability to compete economically with other sources of energy.
-A potential solution to the aforementioned issues is
-the \gls{MSR} concept, one of six advanced reactor designs
-selected by the \gls{GIF} \cite{gif_technology_2002} for continued research
-and development.
+risks, and the ability to compete economically with other sources of energy
+\cite{massachusetts_institute_of_technology_future_2003}. A potential solution
+to the aforementioned issues is the \gls{MSR} concept, one of six advanced
+reactor designs selected by the Generation IV International Forum
+\cite{gif_technology_2002} for continued research and development.
 
@@
 challenges in simulating the transient behavior of \glspl{MSR}; the
-neutronics and thermal-hydraulics are tightly coupled due to the strong
-temperature reactivity coefficient. Furthermore, we have to account for the
-movement of \glspl{DNP} as
-they are now generated directly within the primary coolant loop. Therefore,
+interactions between neutronics and thermal-hydraulics are stronger due
+to greater fuel material expansion. Furthermore, \glspl{DNP} in \glspl{MSR}
+can flow freely within the primary coolant
+loop as opposed to being held in place in a solid fuel matrix. Therefore,
 the choice of coupling methods for each set of physics requires careful
-consideration. Existing reactor system-level software and modeling approaches
-for conventional \gls{LWR} analysis contain reactor-specific assumptions in
-multiphysics coupling and other areas that render the techniques less suitable
-for simulating \glspl{MSR}. Thus, making minor modifications to these codes
-without changing the underlying algorithms is not the best approach for
-\gls{MSR} safety analysis.
+consideration.
 
-In the past two decades, researchers have developed several
-new tools for simulating steady-state and transient behavior in \glspl{MSR}.
-Many of the earlier efforts featured simplifications in simulating
-thermal-hydraulics by solving 1-D Navier-Stokes equations or using
-predetermined uniform velocity fields \cite{krepel_dyn3d-msr_2007}
-\cite{kophazi_development_2009}. In more recent years, there has been
-significant progress towards fully coupled, spatial codes that feature
-2-D axisymmetric or full 3-D models. In 2011, Cammi et al.
-\cite{cammi_multi-physics_2011} performed a ``\gls{MPM}'' analysis
-of a simplified 2-D axisymmetric model of a single \gls{MSBR} fuel channel
-using the commercial finite element analysis software COMSOL Multiphysics. The
-physics were implemented through the two-group neutron diffusion
-equations, and the \gls{RANS} standard $k-\epsilon$ turbulence model, for the
-neutronics and thermal-hydraulics respectively. The
-authors emphasized the need for proper full coupling of the multiphysics and
-presented both steady-state and transient results in various
-scenarios such as reactivity insertions, changes in pumping rate, and the
-presence of periodic perturbations. This approach was featured again in a
-later paper by Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D
-axisymmetric model of the
-\gls{MSFR}. The authors presented results from the Politecnico di Milano
-COMSOL-based approach, and another approach by researchers from Delft
-University of Technology, in which they coupled their in-house neutronics and
-thermal-hydraulics codes, DALTON and HEAT respectively. With multigroup
-neutron diffusion and \gls{RANS} formulations on ultra fine meshes, both
-models showed good agreement in the steady-state neutron flux, temperature,
-and \gls{DNP} distributions, and in the power responses following various
-accident transient initiations. Aufiero et al. \cite{aufiero_development_2014}
-concurrently developed
-a full-core 3-D model of the \gls{MSFR} on OpenFOAM, albeit with one-group
-neutron diffusion to reduce computational load. With the 3-D model, the
-authors could simulate the asymmetric reactor response to the failure of a
-single pump in the sixteen-pump \gls{MSFR} configuration. The authors also
-provided some quantitative data supporting the use of implicit coupling over
-explicit coupling to obtain accurate solutions of the transient cases.
-Recognizing the huge computational burden required for full 3-D simulations, 
-later authors came up with innovative ways to alleviate this issue such as
-selective geometrical reduced order modeling for various components of a
-reactor based on the importance of the physical phenomena being simulated
-\cite{zanetti_geometric_2015}, or using a novel, efficient method for
-neutronics calculations \cite{laureau_transient_2017}.
+Most reactor analysis applications are usually reactor-specific by
+design such as PARCS and TRACE for \glspl{LWR}, and SAS4A/SASSYS-1 for
+liquid metal cooled reactors. Thus, these applications generally would not
+account for \gls{MSR}-specific phenomena and are not suitable for \gls{MSR}
+analysis without any modifications to the source code. Some research efforts
+do focus on adapting these applications for \gls{MSR} analysis. Examples
+include the coupling of modified versions of TRACE and PARCS
+\cite{pettersen_coupled_2016}, and the development of VERA-MSR from the
+integrated \gls{LWR} simulation tool VERA \cite{graham_development_2019}.
+Others developed their \gls{MSR} simulation tools from general
+multiphysics or \gls{CFD} applications such as COMSOL
+\cite{fiorina_modelling_2014} and OpenFOAM \cite{aufiero_development_2014}.
 
-The preceding discussion highlights the challenges faced in MSR simulation,
-and the need for a highly efficient simulation tool that incorporates
-implicitly coupled multiphysics with good computational scalability over
-multiple processing units. This paper presents the open source MSR simulation
-tool, Moltres, as a strong contender to overcoming these challenges for
-simulating the MSFR. Moltres is an application code
-built in the Multiphysics Object-Oriented Simulation Environment (MOOSE)
-parallel finite element framework. Similar to COMSOL and OpenFOAM, it solves
+This thesis demonstrates latest capabilities of Moltres
+\cite{lindsay_introduction_2018}, an open source MSR simulation tool. Moltres
+is an application built in the \gls{MOOSE} parallel finite element framework.
+It solves
 the deterministic multigroup neutron diffusion and thermal-hydraulics
 \glspl{PDE} simultaneously on the same mesh. It supports up to 3-D meshes and
-scales well for a large number of processors.
+scales well for a large number of processors. The underlying \gls{MOOSE}
+framework
+provides a range of implicit and explicit methods for the coupling between the
+neutronics and thermal-hydraulics governing equations.
 
-Moltres employs fully implicit coupling between the neutronics and
-thermal-hydraulics governing equations, to fully account for the tightly
-coupled physics expected in MSRs due to the movement of fuel in the salt. We
-implemented flow using the Navier-Stokes equations and zeroth-order
-approximation of eddy viscosity for a more accurate representation of the flow
-profile compared to imposing a predetermined velocity field.
+In the introductory journal article for Moltres, Lindsay et al.
+\cite{lindsay_introduction_2018} demonstrated Moltres' capabilities with
+2D-axisymmetric and 3D models of the \gls{MSRE}. The results showed good
+qualitative agreement with legacy \gls{MSRE} data but there were some minor
+quantitative discrepancies due to a number of differences in the legacy model. 
+Moltres has undergone further development in the past two years. The authors
+from the first paper have since developed various new capabilities in Moltres
+such as support for looping delayed neutron precursors back into the core,
+and a heat exchanger object to set the heat removal rate.
+
+In addition to those new capabilities, this thesis presents two more recent
+developments in Moltres, namely fully integrating \gls{MOOSE}'s incompressible
+Navier-Stokes module into Moltres, and introducing a
+decay heat model. The incompressible Navier-Stokes module provides a more
+accurate representation of the flow profile, precursor movement, and heat
+transfer as opposed to assuming uniform or function-based velocity fields.
 
@@
 
-The main objective of this thesis is to demonstrate Moltres' capabilities in
-modeling multiphysics, steady-state and transient behavior of fast-spectrum
-\glspl{MSR} through the study of the \gls{MSFR} concept. This is achieved by
-first verifying Moltres' neutronics results against Serpent in the context of
-the \gls{MSFR}, and then comparing the coupled neutronics/thermal-hydraulics
-steady-state and transient accident results of the \gls{MSFR} concept. The
-multiphysics results are verified against the results by Fiorina et al.
-\cite{fiorina_modelling_2014} and Aufiero et al.
-\cite{aufiero_development_2014}.
+The main objectives of this thesis are to demonstrate and verify Moltres'
+latest capabilities in modeling multiphysics, steady-state, and transient
+behavior of fast-spectrum \glspl{MSR} through the study of the \gls{MSFR}
+concept. Code-to-code verification is an important exercise in software
+development for ensuring that the application produces accurate and reliable
+results. This thesis covers the \gls{MSFR} concept mainly because 
+there is readily available literature data to verify our results against. The
+steady-state and transient results also feature a significant degree of
+physics interaction between the neutronics and thermal-hydraulics. This
+present work will first present a verification of Moltres' \gls{MSFR}
+neutronics results against Serpent 2, followed by a verification of
+the coupled neutronics/thermal-hydraulics steady-state and transient accident
+results against results from the Polimi and TUDelft models published by
+Fiorina et al. \cite{fiorina_modelling_2014}. The
+secondary objective is to identify areas of improvement in Moltres for future
+research.
 
@@
 
-The outline of this thesis is as follows: Chapter 2 discusses the history and
-features of \glspl{MSR}. The \gls{MSFR} concept is covered in greater detail.
-Chapter 3 details the simulation codes and the general modeling approach for
-generating the results in this thesis. Chapter 4 provides a neutronics
-assessment by comparing key neutronics parameters from Moltres' eigenvalue
-calculations to Serpent's Monte Carlo calculations. Chapter 5 presents
-steady-state results of coupled neutronics/thermal-hydraulics \gls{MSFR}
-simulations in Moltres. Chapters 6 to 9 present transient accident simulation
-results for unprotected reactivity insertions, unprotected loss of flow,
-unprotected pump overspeed, and unprotected loss of heat sink, respectively.
-Lastly, Chapter 10 summarizes the key findings of this thesis and posits some
-potential avenues for future work.
+The outline of this thesis is as follows. Chapter 2 discusses the history and
+features of \glspl{MSR}, and a literature review of existing \gls{MSR}
+simulation tools. The chapter also covers the \gls{MSFR} concept in greater
+detail. Chapter 3 details the simulation codes and the general modeling
+approach for generating the results in this thesis. Chapter 4 provides a
+neutronics assessment by comparing key neutronics parameters from Moltres'
+eigenvalue calculations to Serpent's Monte Carlo calculations. Chapter 5
+presents steady-state results of coupled neutronics/thermal-hydraulics
+\gls{MSFR} simulations in Moltres. Chapter 6 presents transient accident
+simulation results for unprotected reactivity insertions, unprotected loss of
+heat sink, unprotected loss of flow, and unprotected pump overspeed,
+respectively. Lastly, Chapter 7 summarizes the key findings in this thesis
+and posits some potential avenues for future work.

### 2020-07-03 model.tex
editing

@@
-The main simulation tool used in this work for studying the \gls{MSFR} is
-Moltres. While we used Moltres for the steady-state and transient analysis, it
-requires a dedicated neutron transport solver such as Serpent 2 to generate
-neutron energy group constant data for the multigroup neutron diffusion
-calculations in Moltres. This chapter provides brief introductions of the
-Serpent and Moltres codes, and the general modeling approach for the
+This work demonstrates the \gls{MSFR} simulation capabilities of Moltres, a
+multiphysics simulation tool for \glspl{MSR} \cite{lindsay_introduction_2018}.
+To run simulations on Moltres, it requires input
+group constant data from a neutron transport solver for the
+multigroup neutron diffusion calculations, and a mesh file representing the
+geometry of the reactor. This work uses Serpent 2 \cite{leppanen_serpent_2014}
+for the former and Trelis \cite{noauthor_trelis_2018} for the latter.
+This chapter provides brief introductions of
+Serpent 2, MOOSE, and Moltres, and the general modeling approach for the
 multiphysics simulations in Moltres.
@@
 Serpent 2 \cite{leppanen_serpent_2014} is a continuous-energy Monte Carlo
-reactor physics code under
-active development at the VTT Technical Research Centre of Finland. It was
-created in 2004 out of the need for better general-purpose neutron transport
-codes for group constant generation in lattice geometries, and has since grown
-to support more general capabilities, for an active user base of more than 500
-people worldwide. Serpent is highly parallelizable due to its support for
-both MPI and OpenMP parallel programming APIs, and highly validated and
+neutron transport application under
+active development led by the VTT Technical Research Centre of Finland. It was
+created in 2004 for group constant generation in lattice geometries, and has
+since grown to support more general capabilities. Serpent 2 is highly
+parallelizable, supporting
+both MPI and OpenMP parallel programming APIs. It has also been validated and
 verified against experimental data and other well-established codes.
 
-In Serpent, each neutron is tracked through a combination of ray-tracing-based
-surface tracking and rejection sampling-based delta-tracking. Users may define
-the number of neutrons and the number of active and inactive cycles for each
+In Serpent 2, each neutron is tracked through a combination of
+ray-tracing-based surface tracking and rejection sampling-based delta
+tracking. Users may define the number of neutron histories and the number of
+active and inactive cycles for each
 simulation. Inactive cycles are required for fission source distribution
@@
 determined stochastically based on neutron interaction data from established
-nuclear data libraries (e.g. ENDF, JEFF). These nuclear data libraries provide
-continuous-energy cross section data at discrete temperatures. For other
-temperatures, Serpent has a built-in Doppler-broadening preprocessor that
-extrapolates the relevant cross section data from a lower temperature.
+nuclear data libraries (e.g. ENDF \cite{chadwick_endf/b-vii.1_2011}, JEFF
+\cite{oecd/nea_jeff-3.1.2_2014}). These nuclear data libraries provide
+continuous-energy cross section data at discrete temperatures. Beyond the
+discrete library temperatures, Serpent 2 has a built-in Doppler-broadening
+preprocessor that extrapolates the relevant cross section data from a lower
+temperature \cite{leppanen_serpent_2014}.
 
-In the context of this project, we used Serpent with the JEFF-3.1.2 nuclear
+In the context of this work, Serpent 2 uses the JEFF-3.1.2 nuclear
 data library \cite{oecd/nea_jeff-3.1.2_2014} to generate group constants
-for Moltres. The reactor geometry is based on the axisymmetric \gls{MSFR}
-neutronics benchmark geometry published in several previous papers involving
-the \gls{MSFR}, as shown in Fig. \ref{fig:msfrgeom}. The relevant group
-constant data are collapsed into six neutron energy groups, and calculated for
-temperature values of 800 K to 1300 K at 50 K intervals.
+needed by Moltres. The relevant group constant data are collapsed into six
+neutron energy groups, and calculated at discrete temperature values from 800
+K to 1300 K at 100 K intervals. Table \ref{table:bound} shows the upper bounds
+of each neutron energy group. The group constants relevant for neutronics
+calculations in Moltres are the
+macroscopic fission $\Sigma^f_{g}$, removal $\Sigma^r_{g}$, and scattering
+$\Sigma^s_{g' \rightarrow g}$ neutron cross sections,
+neutron diffusion coefficients $D_g$, average fission energies $\epsilon_g$,
+average neutron yields $\nu$, inverse neutron speeds $1/v$, flux spectra
+$\chi$, \gls{DNP} decay constants $\lambda_i$, and effective delayed neutron
+fractions $\beta_{eff}$. These group constants are extracted from
+the Serpent 2 output files using a Python script available from the Github
+repository that holds the Moltres source code \cite{lindsay_moltres_2017}. The
+script rewrites the group constants into a Moltres-compatible format.
 
-The group constants relevant for neutronics calculations in Moltres are the
-macroscopic fission, absorption, and scattering neutron cross sections,
-neutron diffusion coefficients, average fission energies, average neutron
-yields, inverse neutron speeds, flux spectra, \gls{DNP} decay constants, and
-effective delayed neutron fractions. These group constants are extracted from
-the Serpent output files using a Python script available from the Github
-repository that holds the Moltres source code. The script rewrites the group
-constants into a Moltres-compatible format.
+Serpent 2 provides standard geometric surfaces (e.g. planes, cylinders, cones)
+for defining reactor geometries. In this work, the reactor geometry is the
+same axisymmetric \gls{MSFR} geometry from the paper by Fiorina et al.
+\cite{fiorina_modelling_2014}, as shown in Fig. \ref{fig:msfrgeom}.
 
-\section{Moltres}
+\begin{table}[htb!]
+	\centering
+	\caption{Neutron energy group upper bounds used in Serpent 2.}
+	\begin{tabular}{c S}
+		\toprule
+		{Group number} & {Upper bound [MeV]}\\
+		\midrule
+		1 & 20\\
+		2 & 2.2313\\
+		3 & 0.4979\\
+		4 & 0.0247875\\
+		5 & 0.0055308\\
+		6 & 0.0007485\\
+		\bottomrule
+	\end{tabular}
+	\label{table:bound}
+\end{table}
 
-Moltres is a \gls{MOOSE}-based application for coupled
-neutronics/thermal-hydraulics simulations of \glspl{MSR}. \gls{MOOSE}
-\cite{gaston_physics-based_2015} is a highly parallelizable,
+\section{MOOSE and Moltres}
+
+Moltres \cite{lindsay_moltres_2017} is a \gls{MOOSE}-based
+\cite{gaston_physics-based_2015} application for
+coupled neutronics/thermal-hydraulics simulations of \glspl{MSR}. MOOSE
+provides foundational framework for computationally solving \glspl{PDE} and
+pre-built physics modules for common applications such as the Navier-Stokes
+physics module. Moltres is also a MOOSE physics module for simulating
+\glspl{MSR}. This section describes MOOSE, the Navier-Stokes physics module,
+and Moltres.
+
+\subsection{MOOSE}
+
+\gls{MOOSE} \cite{gaston_physics-based_2015} is a highly parallelizable,
 finite element framework developed at \gls{INL} for simplifying the process of
@@
     \caption{2-D axisymmetric model of the MSFR. The red box indicates the
-    central core region relevant to the modeling approach in Moltres.}
+    central core region in the modeling approach in Moltres.}
     \label{fig:core}
@@
 interpolation method of choice.
-
-\begin{table}[htb!]
-	\centering
-	\caption{Neutron energy group upper bounds used in Serpent.}
-	\begin{tabular}{c S}
-		\toprule
-		{Group number} & {Upper bound [MeV]}\\
-		\midrule
-		1 & 20\\
-		2 & 2.2313\\
-		3 & 0.4979\\
-		4 & 0.0247875\\
-		5 & 0.0055308\\
-		6 & 0.0007485\\
-		\bottomrule
-	\end{tabular}
-	\label{table:bound}
-\end{table}
 

### 2020-07-03 msr.tex
editing

--- 2020-03-09 msr.tex
@@
 \glspl{MSR} are one of six advanced reactor designs shortlisted by
-the \gls{GIF} in 2001 for promising significant advances in safety,
-sustainability, efficiency, and cost over existing designs in operation
-today. This has attracted significant attention and resources towards
+the Generation IV Forum in 2001 for promising significant advances in safety,
+sustainability, efficiency, and cost over existing designs. This
+has attracted significant attention and resources towards
 \gls{MSR} research, most noticeable by the number of start-up companies that
@@
 that earned the concept the label of being a Generation IV reactor. Lastly,
-we present the reference specifications of the \gls{MSFR} concept studied in
-this work.
+this chapter presents the reference specifications of the \gls{MSFR} concept
+studied in this work.
 
@@
 
-The first \gls{MSR}, named the \gls{ARE}, dates back to the 1940s,
+The first \gls{MSR}, named the \gls{ARE}, dates back to the 1940s
 as part of the US Aircraft Nuclear Propulsion program
-\cite{rosenthal_molten-salt_1970}; the molten salt concept was considered due
-to the stability of molten salts at high temperatures and neutron radiation.
-The 2.5 MW$_{\text{th}}$ reactor was built at \gls{ORNL}, where it achieved
-criticality on November 1954 and generated 100 MWh over nine days
-\cite{rosenthal_molten-salt_1970}. The fuel
-consisted of enriched uranium in a molten salt mixture of NaF, ZrF$_4$, and
-UF$_4$, and was moderated by blocks of beryllium oxide. The project ultimately
-never came to fruition as the development of intercontinental ballistic
-missiles effectively eliminated the need for long-range nuclear-powered
-bomber aircraft.
+\cite{rosenthal_molten-salt_1970}. Researchers recommended the molten fluoride
+salts in particular for high uranium solubility, chemical stability, low vapor
+pressure even at high temperatures, decent heat transfer properties,
+resistance against radiation damage, and reduced corrosive effects on some
+common structural material \cite{rosenthal_molten-salt_1970}. They
+subsequently built the 2.5 MW$_{\text{th}}$ ARE reactor at \gls{ORNL}, where
+it achieved criticality on November 1954 and generated 100 MWh over nine days.
+The fuel consisted of enriched uranium in a molten salt mixture of NaF,
+ZrF$_4$, and UF$_4$, and the reactor used blocks of BeO for neutron
+moderation. The aircraft program
+ultimately never came to fruition as the development of intercontinental
+ballistic missiles effectively eliminated the need for long-range
+nuclear-powered bomber aircraft.
 
@@
 research into adapting \glspl{MSR} for civilian power generation
-\cite{rosenthal_molten-salt_1970}. One of the key findings from the
-research was that breeding $^{233}$U from $^{232}$Th gave better performance
-than breeding $^{239}$Pu from $^{238}$U in thermal-spectrum reactors.
+\cite{rosenthal_molten-salt_1970}. One key finding from the
+research was that the thorium fuel cycle had a better breeding ratio than the
+$^{238}$U-to-$^{239}$Pu fuel cycle in thermal-spectrum reactors.
 Ultimately, these efforts culminated in the design, construction, and
-successful operation of the \gls{MSRE}, a graphite moderated thermal
-\gls{MSR}. Although the experiment did not include breeding, scientists at
-\gls{ORNL} obtained a wealth of other experimental data and new insights from
-the study of this reactor. The \gls{MSRE} had a
+successful operation of the \gls{MSRE}, a graphite-moderated thermal
+\gls{MSR}. The \gls{MSRE} had a
 graphite-moderated design with a LiF-BeF$_2$-ZrF$_4$-UF$_4$ fuel salt mixture,
@@
 MW$_{\text{th}}$ due to a miscalculation of heat transfer capabilities
-\cite{haubenreich_experience_1970}.
+\cite{haubenreich_experience_1970}. In January 1969, the \gls{MSRE} became the
+first reactor to run on $^{233}$U fuel.
 
-Design of the \gls{MSRE} commenced in the summer of 1960, with construction
-starting in early 1962 \cite{haubenreich_experience_1970}. The reactor
-achieved zero-power criticality in June 1965, and 30 days of continuous
-operation at full power in December 1966. The reactor operated at full power
-for the most of the following 15 months, during which the researchers carried
-out various experiments. Soon after shutdown, the $^{235}$U fuel was replaced
-with $^{233}$U and in January 1969, the \gls{MSRE} became the first reactor to
-run on $^{233}$U fuel. Further experiments were run, including xenon
-stripping, fission product deposition, tritium behavior, and plutonium
-addition studies, before the \gls{MSRE} was permanently shut down to conserve
-remaining funding for other related activities \cite{macpherson_molten_1985}.
-
-Building on the \gls{MSRE}'s hugely successful run, \gls{ORNL} proposed a
+Building on their experience with the \gls{MSRE}, \gls{ORNL} proposed a
 new program for the construction and operation of a demonstration reactor
-based on the \gls{MSBR} concept that they had developed
-\cite{macpherson_molten_1985}. The \gls{MSBR} is a thermal-spectrum, single
-fluid reactor with fertile $^{232}$Th isotopes mixed directly into the FLiBe
-molten salt for $^{233}$U breeding \cite{gehin_liquid_2016}. Like the
+based on the Molten Salt Breeder Reactor (MSBR) concept that they had
+developed \cite{macpherson_molten_1985}. The \gls{MSBR} is a thermal-spectrum,
+single fluid reactor with fertile $^{232}$Th isotopes mixed directly into the
+FLiBe molten salt for $^{233}$U breeding \cite{gehin_liquid_2016}. Like the
 \gls{MSRE}, the \gls{MSBR} relies on continuous online reprocessing to add
-fertile material and remove \gls{FP} neutron poisons. As a breeder
-reactor, the doubling time (the minimum amount of time required to produce
-enough fissile material to start up another \gls{MSBR}) was estimated to be
-approximately 22 years. However, \gls{ORNL} failed to secure funding on two
-separate occasions in 1972 and 1974; they lost out to the competing
-\gls{LMFBR} program which had a head start and wider political and technical
-support. Nevertheless, from a technical perspective, two independent
-technology evaluation and design studies of the \gls{MSR} had ``reported
-favorably on the promise of the system'' \cite{macpherson_molten_1985}.
+fertile material and remove fission product neutron poisons. Researchers
+estimated the doubling time (the minimum amount of time required to produce
+enough fissile material to start up another \gls{MSBR}) to be
+approximately 22 years. However, \gls{ORNL} failed to secure funding for the
+new program in their two attempts in 1972 and 1974. Nevertheless, from a
+technical perspective, two independent
+technology evaluation and design studies of the \gls{MSR} had reported
+favorably on the promise of the system \cite{macpherson_molten_1985}.
 
@@
 1970s. In 1980, \gls{ORNL} published a report describing a new \gls{MSR}
-concept, called the \gls{DMSR} \cite{gehin_liquid_2016}. This design was
-developed in response to the fuel reprocessing restrictions introduced by
-President Ford in 1976; the \gls{DMSR} is designed to operate as a
-once-through converter system without fuel reprocessing. While it is largely
-fueled by 19.75 \% \gls{HALEU}, the initial core loading includes thorium to
+concept, called the \gls{DMSR} \cite{gehin_liquid_2016} with denatured
+$^{235}$U fuel (i.e. low-enriched uranium). The \gls{ORNL} researchers
+developed this design in response to the fuel reprocessing restrictions
+introduced by President Ford in 1976. The \gls{DMSR} would operate as a
+once-through
+converter system without fuel reprocessing. While the fuel largely consists of
+19.75 \% \gls{HALEU}, the initial core loading includes thorium to
 boost its conversion ratio throughout its lifetime. It has a continuous
@@
 policy thresholds. The design also includes a gas sparging system for removing
-gaseous \glspl{FP}, while noble metals naturally plate out onto the walls of
-the coolant loop. A significant drawback in the \gls{MSBR} design is the
-extensive neutron damage in the graphite moderator that necessitated frequent
+gaseous fission products, while noble metals plate out onto the walls of
+the coolant loop. The older \gls{MSBR} design had a significant drawback; the
+extensive neutron damage in the graphite moderator necessitated frequent
 replacement (every four years) throughout its operational lifetime. The
 \gls{DMSR} avoids this issue by having a lower power density while maintaining
-the overall power output of 2250 MW$_{\text{th}}$. As a result, the graphite
-moderator was projected to last for the entirety of the \gls{DMSR}'s design
-lifetime.
+the overall power output of 2250 MW$_{\text{th}}$. As a result, researchers
+projected that the graphite moderator would last for the entirety of the
+\gls{DMSR}'s design lifetime.
 
@@
 with separate loops for the fuel salt and the blanket salt. The blanket is a
-1 m-wide tank surrounding the core. The relatively harder neutron spectrum
-arising from the absence of moderators and the choice of chloride over
-fluoride salt favors $^{239}$Pu breeding over the thorium cycle. Some
-experiments were performed to study molten salt chemistry but no reactor
-prototypes were built. The UK program was eventually shut down just like its
+1 m-wide tank surrounding the core. The absence of moderators and the choice
+of chloride over fluoride salt resulted in a relatively hard neutron spectrum
+which favors $^{239}$Pu breeding over the thorium cycle. The UK researchers
+performed some experiments to study molten salt chemistry but there were no
+reactor prototypes. The UK program eventually shut down just like its
 US counterpart partly due to the successful demonstration of the Prototype
@@
 Following a lull lasting through the late 20th century, \gls{MSR} research
-picked up pace due to renewed interest initiated by the \gls{GIF} in 2001.
-Today, there are numerous \gls{MSR} concepts under active development led by
-various national and commercial bodies. Many differences exist between
-different concepts and the only common denominator is the use of nuclear fuel
-dissolved in molten salt. Table * is a non-exhaustive list of \glspl{MSR}
-under development.
+picked up pace due to renewed interest initiated by the Generation IV Forum in
+2001. Today, there are numerous \gls{MSR} concepts under active development
+led by various national and commercial bodies.
 
@@
 
-As mentioned in the introduction section, the most significant difference
+As mentioned in Chapter \ref{chap:intro}, the most significant difference
 between \glspl{MSR} and other reactor concepts is the liquid fuel in
 \glspl{MSR}; fissile and/or fertile material is dissolved in high temperature,
-commonly eutectic mixtures of molten salts. Most \gls{MSR} designs are
-circulating-fuel reactors. The primary coolant loop containing the fuel salt
-transfers heat through a heat exchanger to the clean, secondary/intermediate
+commonly eutectic mixtures of molten salts. Molten salt-cooled, solid-fuel
+reactors also exist but this thesis will focus on liquid-fuel reactors.
+The primary coolant loop containing the fuel salt
+transfers heat through a heat exchanger to the clean, intermediate
 loop. The liquid fuel form allows for continuous online fuel reprocessing,
-and the removal of gaseous \glspl{FP} via a gas sparging system.
+and the removal of gaseous fission products via a gas sparging system.
 
-The flexibility of \glspl{MSR} is best illustrated by the various designs
-under development today. Graphite-moderated thermal-spectrum \glspl{MSR} are
-typically straightforward \gls{LEU} burners, or $^{232}$Th/$^{233}$U
-iso-breeders/breeders, while epithermal- and fast-spectrum \glspl{MSR} have
-the additional options of operating as \gls{TRU} fuel burners or
+The various \gls{MSR} designs under development today illustrate the
+flexibility of this reactor concept. Graphite-moderated thermal-spectrum
+\glspl{MSR} are typically straightforward \gls{LEU} burners or
+$^{232}$Th/$^{233}$U iso-breeders/breeders, while epithermal- and
+fast-spectrum \glspl{MSR} can operate as \gls{TRU} fuel burners or
 $^{238}$U/$^{239}$Pu breeders. Breeder designs can be further categorized into
-one- or two-fluid designs; two-fluid designs feature separate blanket molten
+one- or two-fluid designs. Two-fluid designs feature separate blanket molten
 salt mixtures that contain higher proportions of fertile material than the
-fuel salt mixture.
+fuel salt mixture. Examples of one-fluid designs include the Integral Molten
+Salt Reactor from Terrestrial Energy \cite{leblanc_integral_2015} and the
+\gls{MSR} design from Transatomic Power
+\cite{transatomic_power_corporation_technical_2016} while two-fluid designs
+include the \gls{MSFR} \cite{serp_molten_2014} and \gls{MOSART}
+\cite{ignatiev_molten_2014}.
 
@@
 
-\glspl{MSR} are generally deemed to be safer than \glspl{LWR} due to their
-reliance on natural physical phenomena for passive safety.
-The most immediate and significant safety characteristic is the strong
-negative temperature reactivity feedback of the fuel salt. This is due to
-greater temperature-induced volumetric expansion in liquid fuel than solid
+\glspl{MSR} rely on natural physical phenomena for passive safety such as the
+strong negative temperature reactivity feedback of the fuel salt due to
+greater temperature-induced expansion in liquid fuel than solid
 fuel. Combined with the Doppler broadening of resonance capture cross sections
-present in both fuel forms, we expect to observe a smaller temperature
+present in both fuel forms, there would be a smaller temperature
 increase following an unprotected reactivity insertion. The overall
-temperature reactivity coefficient varies widely between different \gls{MSR}
+temperature reactivity coefficient varies widely among \gls{MSR}
 designs due to other structures, such as moderators and reflectors, present in
 the core. In particular, graphite moderators tend to have slightly positive
-temperature reactivity coefficients. This was notably observed in the
-\gls{MSBR} concept, but the total temperature reactivity coefficient was still
-relatively large and negative \cite{rykhlevskii_modeling_2019}. This
-phenomenon provides a large degree of control and stability as it is always
+temperature reactivity coefficients. \gls{MSBR} concept has this issue, but
+the total temperature reactivity coefficient is still relatively large and
+negative \cite{rykhlevskii_modeling_2019}. The negative temperature reactivity
+feedback provides a large degree of control and stability as it is always
 present in an \gls{MSR} regardless of the operating conditions.
@@
 addition to the strong negative temperature
-reactivity coefficient, diminishes the likelihood and severity of unprotected
+reactivity coefficient, diminish the likelihood and severity of unprotected
 criticality accidents in \glspl{MSR} \cite{elsheikh_safety_2013}. In the
-unlikely situation where an \gls{MSR} encounters a severe runaway reaction,
-we can rely on another passive safety feature called freeze plugs. A freeze
-plug is a plug of solidified salt at the bottom of the core actively cooled by
-fans or other cooling systems to keep its temperature just below the freezing
-point of the salt \cite{aji_experimental_2020}. When temperatures in the core
-exceed a certain threshold during a dangerous transient, the freeze plug melts
-and the molten salt in the core drains into a containment tank designed to
-keep the salt in a subcritical configuration. This is especially easy to
-achieve for
-thermal-spectrum \glspl{MSR} as the absence of moderators in the containment
-tank would automatically bring the multiplication factor down below unity
-\cite{elsheikh_safety_2013}. \glspl{MSR} also typically have large
-margin-to-boiling under nominal operating conditions so that fuel salt boiling
-does not occur. Furthermore, the reactor vessel is consequently subject to
-much lower stresses as \glspl{MSR} operate at near-atmospheric pressure
-levels. Thus, the probability of pipe ruptures due to high pressure is low.
+unlikely situation in which an \gls{MSR} encounters a severe runaway reaction,
+\glspl{MSR} rely on another passive safety feature: freeze plugs. Various
+freeze plugs designs exist for different \glspl{MSR}. The freeze plug
+in the \gls{MSFR} concept is a plug of solidified salt at the bottom of the
+core actively cooled by fans or other cooling systems to keep its temperature
+just below the freezing point of the salt \cite{aji_experimental_2020}. In
+general, when temperatures in the core exceed a certain threshold during a
+dangerous transient, the freeze plug melts and the molten salt in the core
+drains into a containment tank designed to keep the salt in a subcritical
+configuration. This is especially easy to achieve in thermal-spectrum
+\glspl{MSR} as the absence of moderators in the containment tank would
+automatically drive the multiplication factor down below unity
+\cite{elsheikh_safety_2013}.
 
-\glspl{MSR} may be more resilient to pump failure accidents as natural
-circulation could passively sustain enough heat transfer through the heat
-exchanger to remove any remaining fission and decay heat after the failure. If
-natural circulation proves insufficient, the aforementioned freeze plug can
-drain the salt out of the core. Decay heat in \glspl{MSR} with online
+During pump failure accidents, natural circulation can passively sustain
+enough heat transfer to remove decay heat and prevent catastrophic structural
+failure. If natural circulation proves insufficient, the aforementioned freeze
+plug can drain the salt out of the core. Decay heat in \glspl{MSR} with online
 reprocessing is typically lower than that in \glspl{LWR} due to the
-continuous removal of \glspl{FP}. For example, the decay heat in an \gls{MSFR}
-after reaching equilibrium salt composition is approximately 3.5\% compared
-to 6\% in \glspl{LWR} \cite{brovchenko_design-related_2013}.
+continuous removal of fission products. For example, the decay heat in an
+\gls{MSFR} after reaching equilibrium salt composition is approximately
+3.5\% of full reactor power compared to 6\% in \glspl{LWR}
+\cite{brovchenko_design-related_2013}.
+
+\glspl{MSR} also typically have large margin-to-boiling under nominal
+operating conditions so that fuel salt boiling does not occur
+\cite{elsheikh_safety_2013}. Furthermore, the reactor vessel is consequently
+subject to much lower stresses as \glspl{MSR} operate at near-atmospheric
+pressure levels. Thus, the probability of pipe ruptures due to high pressure
+is low.
 
 However, there are also some safety disadvantages associated with \glspl{MSR}.
-Firstly, \glspl{MSR} have smaller fractions of \glspl{DNP} in the active core
-region as some of them decay in the external loop regions; this complicates
-reactor control and may result in faster transients due to the decrease in
-average neutron lifetime. Secondly, structural materials in the core must be
-resistant to salt corrosion under high temperatures and neutron irradiation
-over the lifetime of the reactor. To a lesser extent, pipes should also be
-corrosion resistant to prevent pipe ruptures. Lastly, fuel salt overcooling is
-a potentially dangerous accident scenario as the strong temperature
-coefficient can effectively cause a large reactivity insertion. Overcooling
-may also cause salt to freeze and restrict flow, thereby causing a complete
-loss of flow accident.
+Firstly, \glspl{MSR} have smaller fractions of delayed neutron precursors in
+the active core region as some of them decay in the external loop regions.
+This complicates reactor control and may result in faster transients due to
+the decrease in average neutron lifetime. Secondly, hot molten salts are
+generally corrosive and the corrosion mechanism is different from conventional
+corrosion induced by water and other common agents \cite{yoshioka_7_2017}. The
+intense gamma and neutron radiation in the core may also accelerate corrosion
+in the structural components. Lastly, overcooling in pipes and heat exchangers
+may pose operational challenges as the salt can freeze and restrict flow,
+thereby causing a loss of flow accident \cite{ho_molten_2013}.
 
-\subsection{Other Factors}
+\subsection{Other Features}
 
-Other factors for assessing a reactor design include sustainability,
-economics, waste management, and non-proliferation.
-
-Breeder \gls{MSR} designs easily score well in the sustainability category.
-\glspl{MSR} generally have good neutron economy due to little structural
-material in the core and continuous online removal of neutron poisons. Noble
-metal neutron poisons also naturally plate out onto the inner walls of the
-loop which are generally regions of lower neutronic importance. Both
+\glspl{MSR} have several positive sustainability features.
+\glspl{MSR} have good neutron economy due to the continuous removal of
+neutron poisons by the online reprocessing and gas sparging systems
+\cite{kamei_recent_2012}. Both
 $^{232}$Th/$^{233}$U and $^{238}$U/$^{239}$Pu fuel cycles are viable
 candidates for breeding in \glspl{MSR}, with the former being more suited for
-thermal reactors and the latter being more suited for fast reactors. Another
-benefit of this is that some \glspl{MSR} can potentially start with an initial
-core loading of $^{239}$Pu with a fast spectrum before transitioning towards
-$^{233}$U fuel with a $^{232}$Th feed; this measure effectively circumvents
-the issue of having insufficient $^{233}$U inventory to start up the reactor.
+thermal reactors and the latter being more suited for fast reactors.
 
@@
 $^{233}$U. This reduces the overall radiotoxicity and long-term decay heat
-associated with long-lived plutonium and \gls{MA} isotopes. However, there are
-other radionuclides such as $^{231}$Pa, $^{229}$Th, and $^{230}$U that may
-pose long-term radiological concerns. Nevertheless, this feature complements
-\gls{TRU} burning in fast spectrum \glspl{MSR} to reduce overall levels
-of \gls{TRU} waste going into long-term storage in nuclear waste repositories.
+associated with long-lived plutonium and \gls{MA} isotopes. The combination
+of \gls{TRU} fuel and $^{232}$Th feed in fast spectrum \glspl{MSR} would help
+reduce overall levels of \gls{TRU} waste going into long-term storage in
+nuclear waste repositories \cite{merle-lucotte_launching_2011}.
 
-There are some nuclear non-proliferation concerns with the thorium fuel cycle
-in \glspl{MSR}.
-The main concern involves the separation of the intermediate $^{233}$Pa
+Nuclear non-proliferation concerns in thorium-fueled \glspl{MSR} involve the
+separation of the intermediate $^{233}$Pa
 isotope from the fuel salt. $^{233}$Pa decays into $^{233}$U with a half-life
-of approximately 27 days and $^{233}$U is at least as potent as $^{235}$U for
-nuclear weapons production. The highly radioactive $^{232}$U provides some
-level of proliferation resistance but it can be sidestepped by
-separating away its $^{232}$Pa precursor, which has a half-life of only 1.31
-days. Strong regulatory framework and close monitoring of \glspl{MSR} would be
-essential to ensure non-proliferation.
+of approximately 27 days and the $^{233}$U produced is equivalent in potency
+to $^{239}$Pu for nuclear weapons production \cite{grape_10_2017}. The highly
+radioactive $^{232}$U
+by-product provides some level of proliferation resistance but nuclear
+proliferators can sidestep this complication by
+separating the $^{232}$U away from the combined $^{232}$Pa/$^{233}$Pa stream
+after most of the $^{232}$Pa has decayed into $^{232}$U at a short half-life
+of 1.31 days \cite{grape_10_2017}. Safeguards by design and close monitoring
+of \glspl{MSR} are therefore essential to avoiding spent nuclear fuel
+diversion.
 
-The economic prospects are uncertain at the current stage of \gls{MSR}
-development but we can qualitatively list the potential impacts. Capital costs
-are expected to be greater due to the need for increased corrosion resistance
-and the supporting reprocessing facilities. This is in spite of the reduced
-pressure constraints on the reactor pressure vessel for \glspl{MSR} resulting
-from the significantly lower operating pressure. There are potential cost
-savings from eliminating fuel pellet and fuel assembly fabrication. Higher
-fuel utilization is achievable as neutron poisons are continuously removed
-from the core via the continuous online reprocessing and gas sparging systems.
-Online fuel reprocessing also minimizes reactor downtime because refuelling
-can be performed at any time. The higher operating temperature results in
-greater thermal efficiencies and presents new opportunities beyond
-conventional electricity generation, namely industrial process heat
-applications and hydrogen production.
+Economic analyses of \glspl{MSR} are preliminary at the current
+stage of \gls{MSR} development. Qualitatively, many technical factors favor
+\glspl{MSR} over \gls{LWR}. Some of these factors include: smaller reactor
+core size due to low operating pressures, higher thermal efficiency and
+cheaper air-cooling systems due to high operating temperatures, reduction in
+fuel fabrication costs, and shorter reactor shutdown periods due to online
+refueling \cite{dolan_1_2017}. Moir \cite{moir_recommendations_2008}
+published data from \gls{ORNL} reports on the estimated electricity costs for
+a 1000 MWe \gls{MSR} with 20\% enriched $^{235}$U fuel and no reprocessing
+throughout the reactor's 30-year lifetime. Adjusted to year 2000 dollars, the
+electricity costs for this type of \gls{MSR} was 7\% and 8\% lower than that
+of pressurized water reactor and coal power plants. These differences are
+within the $\pm$10\% uncertainty range but there is room for some optimism.
+The author also notes that costs could be lower in an \gls{MSR} with
+fertile feed and online reprocessing could reach fuel self sufficiency.
 
-Beyond these factors, \glspl{MSR} still require significant R\&D efforts and
-engineering demonstrations for experimental validation of various components
-before a full commercial model can be commissioned. Work towards creating a
-safety and licensing framework for \glspl{MSR} has picked up pace only in
-recent years due to the growing interest from commercial \gls{MSR} developers.
+In terms of technological readiness, \glspl{MSR} still require significant
+R\&D efforts and engineering demonstrations for experimental validation of
+various components before a full commercial model can be commissioned. Work
+towards creating a safety and licensing framework for \glspl{MSR} has picked
+up pace only in recent years due to the growing interest from commercial
+\gls{MSR} developers.
+
+\section{Literature Review of MSR Simulation Tools}
+
+In the past two decades, researchers have developed several
+new tools for simulating steady-state and transient behavior in \glspl{MSR}.
+Many of the earlier efforts featured simplifications in simulating
+thermal-hydraulics by solving 1-D Navier-Stokes equations or using
+predetermined uniform velocity fields \cite{krepel_dyn3d-msr_2007,
+kophazi_development_2009}. In more recent years, there has been
+significant progress towards fully coupled, spatial codes that feature
+2-D axisymmetric or full 3-D models. In 2011, Cammi et al.
+\cite{cammi_multi-physics_2011} performed a ``\gls{MPM}'' analysis
+of a simplified 2-D axisymmetric model of a single \gls{MSBR} fuel channel
+using the commercial finite element analysis software COMSOL Multiphysics. The
+physics were implemented through the two-group neutron diffusion equations,
+and the \gls{RANS} standard $k$-$\epsilon$ turbulence model, for the
+neutronics and thermal-hydraulics respectively. The
+authors emphasized the need for proper full coupling of the multiphysics and
+presented both steady-state and transient results in various
+scenarios such as reactivity insertions, changes in pumping rate, and the
+presence of periodic perturbations. This approach was featured again in a
+later paper by Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D
+axisymmetric model of the
+\gls{MSFR}. The authors presented results from the Politecnico di Milano
+COMSOL-based approach, and another approach by researchers from Delft
+University of Technology, in which they coupled their in-house neutronics and
+thermal-hydraulics codes, DALTON and HEAT respectively. With multigroup
+neutron diffusion and \gls{RANS} formulations on ultra fine meshes, both
+models showed good agreement in the steady-state neutron flux, temperature,
+and \gls{DNP} distributions, and in the power responses following various
+accident transient initiations. Aufiero et al. \cite{aufiero_development_2014}
+concurrently developed
+a full-core 3-D model of the \gls{MSFR} on OpenFOAM, albeit with one-group
+neutron diffusion to reduce computational load. With the 3-D model, the
+authors could simulate the asymmetric reactor response to the failure of a
+single pump in the sixteen-pump \gls{MSFR} configuration. The authors also
+provided some quantitative data supporting the use of implicit coupling over
+explicit coupling to obtain accurate solutions of the transient cases.
+Recognizing the huge computational burden required for full 3-D simulations, 
+later authors came up with innovative ways to alleviate this issue such as
+selective geometrical reduced order modeling for various components of a
+reactor based on the importance of the physical phenomena being simulated
+\cite{zanetti_geometric_2015}, or using a novel, efficient method for
+neutronics calculations \cite{laureau_transient_2017}.
+
+\section{Molten Salt Fast Reactor}
+
+The \gls{MSFR} is a European reference fast-spectrum \gls{MSR} concept
+\cite{mathieu_thorium_2006, merle_optimized_2007}. Table \ref{table:msfr} and
+Figure \ref{fig:msfr} show the main specifications and schematic view of the
+\gls{MSFR}, respectively. Developed from the \gls{MSBR} design, the
+\gls{MSFR} is intended to run primarily on a closed thorium fuel cycle with
+continuous online fuel reprocessing. Several reasons motivated the omission of
+graphite moderators from the original \gls{MSBR} design. Graphite is
+susceptible to
+long-term radiation damage and replacement is likely to be necessary during
+the operating lifetime of the reactor. Graphite also has a positive
+temperature coefficient of reactivity; eliminating graphite from the design
+ensures a greater safety margin \cite{mathieu_thorium_2006}. While negative
+temperature coefficients are attainable with very thermalized spectra,
+breeding ratios deteriorated significantly due to parasitic absorption in the
+large volume of graphite needed for thermalization
+\cite{mathieu_thorium_2006}.
+%
+\begin{figure}[htb!] 
+	\centering
+	\includegraphics[width=0.8\textwidth]{MSFR}
+	\caption{Schematic view of the MSFR concept. Figure reproduced from
+	Brovchenko et al. \cite{brovchenko_design-related_2013}.}
+	\label{fig:msfr}
+\end{figure}
+%
+\begin{table}[htb!]
+    \small
+	\caption{Main specifications of the \gls{MSFR} concept
+				\cite{serp_molten_2014}.}
+	\centering
+	\begin{tabular}{ l r }
+		\hline
+		Parameter & Value \\
+		\hline
+		Thermal/Electric output [MW$_{\text{th}}$/MW$_{\text{e}}$] & 3000 /
+		1500 
+		\\
+		Salt volume [m$^3$] & 18 \\
+		Salt fraction in core & 0.5 \\
+		Number of circulation loops & 16 \\
+		Nominal flow rate [kg s$^{-1}$] & 18500  \\
+		Nominal circulation time [s] & 4.0 \\
+		Inlet/outlet temperature [K] & 923 / 1023 \\
+		Blanket volume [m$^3$] & 7.3\\
+		\hline
+	\end{tabular}
+	\label{table:msfr}
+\end{table}
+
+In the \gls{MSFR} design, fuel salt flows vertically upward through a 9 m$^3$
+central core region. At the top of the core, the flow separates into sixteen
+smaller external loops, each of which passes through a heat exchanger before
+being pumped back into the bottom of the core. The salt also passes through
+the online salt reprocessing and gas sparging systems located along the
+external loops. A toroidal blanket tank containing fertile salt for breeding
+surrounds the core radially. The top and bottom of the core are enclosed
+by nickel alloy reflectors. A layer of boron carbide behind the blanket tanks
+protects the peripheral equipment from excessive neutron damage. During severe
+accidents when core temperatures rise to dangerous levels, the actively
+fan-cooled freeze plug at the bottom of the core melts and drains the fuel
+salt into a containment vessel designed to keep the salt subcritical.
+
+Although the \gls{MSFR} is primarily designed to operate on the thorium fuel
+cycle, it supports a range of start-up fuel and feed compositions. This
+versatility is particularly important for the first few \glspl{MSFR} to be
+deployed due to the lack of $^{233}$U reserves required for the initial core
+loading. In general, the fuel and blanket salts are approximately composed of
+eutectic mixtures of 77.5\% LiF - 22.5\% AcF$_4$; AcF$_4$ represents
+actinide fluorides such as uranium, thorium, plutonium, and other \gls{TRU}
+fluorides. For an initial composition consisting of $^{232}$Th and $^{233}$U,
+the benchmark value for the amount of $^{233}$U for criticality under
+normal operating conditions is 2.5 mol\%. However, a neutronic benchmark
+study by Brovchenko et al. \cite{brovchenko_neutronic_2019} shows that
+different neutronics software with different nuclear data clearly provide
+different $k_{eff}$ estimates even with the same isotopic compositions and
+temperature distributions. The individual co-authors adjusted the
+ratio of $^{232}$Th to $^{233}$U slightly to achieve exact criticality at a
+uniform temperature of 973 K for their own neutronics software
+\cite{brovchenko_neutronic_2019}. This thesis
+performs the same exercise to adjust the inlet and outlet temperatures to
+match nominal values.
+
+The thermal and electric power output of the \gls{MSFR} are 3000 MW
+$_{\text{th}}$ and 1500 MW$_{\text{e}}$, respectively. The high thermal
+efficiency ($\eta_{th}=0.5$) is due to the high operating
+temperature. The inlet and outlet temperature
+specifications of the fuel salt are 923 K and 1023 K, respectively, for a
+minimum 50 K temperature buffer between the
+operating temperatures and the melting point of the salt (873 K)
+\cite{euratom_final_2015}. The \gls{MSFR} has
+heat exchangers and an intermediate coolant loop to isolate the power
+conversion system from the highly radioactive fuel salt. This also serves as
+a layer of containment between the radioactive material and the outside
+environment. The exact composition of the intermediate coolant is not
+finalized yet but potential candidates include NaF-NaBF$_4$, FLiNaK,
+LiF-ZrF$_4$, and FliBe \cite{merle_concept_2017}.
+
+\subsection{Model Reactor Geometry}
+
+This present work uses the same 2-D square-cylindrical \gls{MSFR} design to
+benchmark our results against results published by Fiorina et al.
+\cite{fiorina_modelling_2014}. The design is a 2-D axisymmetric representation
+of the \gls{MSFR} with the sixteen individual external loops homogenized into
+a single outer loop as shown in Figure \ref{fig:msfrgeom}. For the multigroup
+group constants calculations in Serpent, the 2-D axisymmetric model is
+extended into a 3-D model by a 360-degrees rotation about the
+central axis. The material definitions are the same as those specified in the
+reference \gls{MSFR} model. Accordingly, the pump and heat exchanger regions
+are assumed to be composed of 100\% fuel salt. While this may not be entirely
+accurate, the exact details of the pump and heat exchanger systems are still
+under active study, and this external loop region is presumed to be of little
+neutronic importance due to its position behind the strong boron carbide
+neutron absorber layer.
+%
+\begin{figure}[t!] 
+	\centering
+	\includegraphics[width=0.5\textwidth]{msfr-geom}
+	\caption{2-D axisymmetric model of the \gls{MSFR} core used for the
+	simulations in Serpent. All dimensions are in meters.
+	\cite{brovchenko_neutronic_2019}}
+	\label{fig:msfrgeom}
+\end{figure}
+
+Although this present work uses the same 2-D axisymmetric geometry for
+generating
+group constant data from Serpent, there are two minor differences between the
+\gls{MSFR} geometry used in the Moltres model, and the Polimi and TUDelft
+models \cite{fiorina_modelling_2014}. Firstly, the reactor geometry for
+Moltres excludes the 2-cm-thick structural material around the blanket tank
+that separates the fuel and blanket salts. The thickness is much smaller than
+the rest of the regions in the geometry and caused issues for meshing. The
+neutronics results in Chapter \ref{chap:nts} shows that the overall $k_{eff}$
+and other parameters from Moltres show good agreement with that from Serpent.
+It also has no direct impact on the temperature distribution results because
+this present work solves for the temperature distribution in the primary loop
+with homogeneous Neumann boundary conditions on the fuel salt-to-wall
+interface. This approach is for consistency with the Polimi and TUDelft
+models.
+
+The second difference pertains to the out-of-core section of the primary loop.
+The focus for Moltres development has been on core multiphysics over detailed
+out-of-core multiphysics. The Moltres model simulates the outer loop as a 1-D
+pipe with a pointwise heat sink to represent the heat exchanger. To simulate
+the pumps, a Dirichlet boundary condition for velocity at the inlet boundaries
+drives the flow in the central core region of the primary loop. At every
+timestep, Moltres converges the core and outer loop calculations using Picard
+iterations. This approach shares some similarities with the
+geometric multiscale modeling approach by Zanetti et al.
+\cite{zanetti_geometric_2015}. Future models could create a better
+representation of the primary loop by implementing a whole continuous loop
+with pressure increases and drops corresponding to the pumps and heat
+exchangers.
+
+\subsection{Material Specifications}
+
+This section details the material specifications of the various reactor
+components in the \gls{MSFR}.
+
+\subsubsection{Molten Salt}
+
+The reference start-up salt composition for the fuel and blanket salts is
+77.5\% LiF - 22.5\% AcF$_4$ (actinide fluorides)
+\cite{merle-lucotte_launching_2011}. Typically, researchers working with the
+\gls{MSFR}
+model tweak the exact actinide composition by varying the $^{232}$Th to
+$^{233}$U ratio to obtain a $k_{\text{eff}}$ value of 1 at a uniform
+temperature of 973 K \cite{brovchenko_neutronic_2019}. Thus, the exact
+actinide compositions vary depending on
+the nuclear data library and neutron transport code. This present work uses
+a fuel salt composition of 77.5\% LiF - 19.913\% ThF$_4$ - 2.587\%
+$^{233}$UF$_4$ for all calculations, which is a simplifying assumption but a
+fuel burnup analysis is complex and out of scope for this thesis. Table
+\ref{table:prip} shows relevant physical properties of the fuel and blanket
+salts.
+%
+\begin{table}[htb!]
+\small
+\centering
+\caption{Properties of the fuel and blanket salts LiF-AcF$_4$.}
+\begin{tabular}{l l r r}
+\toprule
+Property & Formula & {Value at 973 K} & Validity Range\\
+\midrule
+Melting temperature [K] & 841 & {N/A} & 1 bar \\
+Boiling temperature [K] & 1874 & {N/A} & 1 bar \\
+Density, $\rho$ [kg m$^{-3}$] & $4094-0.882 \cdot (T-1008)$ & $4125$ & 893-1123 K \\
+Dynamic viscosity, $\mu$ [Pa s] & $\rho \cdot 5.55 \times 10^{-8} \cdot e^{3689/T}$ & $1.015$ & 898-1121 K \\
+Thermal conductivity, $k$ [W m$^{-1}$ K$^{-1}$] & $0.928+8.397 \times 10^{-5} \cdot T$ & $0.01010$ & 891-1020 K \\
+Specific heat, $c_p$ [J kg$^{-1}$ K$^{-1}$] & $-1111+2.78\cdot T$ & $1594$ & 867-907 K \\
+\bottomrule
+\end{tabular}
+\label{table:prip}
+\end{table}
+
+\subsubsection{Structural Materials}
+
+The reflectors on the periphery of the reactor core and the blanket tank are
+made of a NiCrW Hastelloy (metal alloy) \cite{brovchenko_neutronic_2019}.
+Table \ref{table:refl} details the elemental composition
+of the Hastelloy. The alloy has a density
+of 10 g$\cdot$cm$^{-3}$. The \gls{MSFR} also includes a 20 cm
+layer of boron carbide (B$_4$C) to protect the heat exchangers and pumps from
+neutron irradiation. The reference specifications indicate that natural boron
+is used, which is composed of 19.8 \% $^{10}$B and 80.2 \% $^{11}$B, with an
+overall density of 2.52 g$\cdot$cm$^{-3}$. 
+%
+\begin{table}[htb!]
+\footnotesize
+\centering
+\caption{Composition (mol \%) of the NiCrW Hastelloy.}
+\begin{tabular}{l l l l l l l l l l l l l}
+\toprule
+Ni & W & Cr & Mo & Fe & Ti & C & Mn & Si & Al & B & P & S \\
+\midrule
+79.432 & 9.976 & 8.014 & 0.736 & 0.632 & 0.295 & 0.294 & 0.257 & 0.252 & 0.052 & 0.033 & 0.023 & 0.004 \\
+\bottomrule
+\end{tabular}
+\label{table:refl}
+\end{table}

### 2020-07-03 neutronics.tex
editing

@@
-In this chapter, we compare key neutronics results between Serpent and Moltres
+This chapter compares key neutronics results between Serpent and Moltres
 for a static model of the \gls{MSFR}, i.e. no salt flow, and uniform
@@
 exercise also serves as a verification for the group constant data
-generated for Moltres from Serpent. For visibility, we list the important
-details of the static model below.
+generated for Moltres from Serpent. For clarity, here are the important
+details of the static model.
 
@@
 
-We calculated estimates of $k_{\text{eff}}$ values in Serpent and Moltres for
-the static \gls{MSFR} model. Moltres solves the six-group neutron diffusion equations as a
-steady-state eigenvalue problem to find the $k_{\text{eff}}$ as it does not
-currently have the capability to calculate the adjoint flux. Table
-\ref{table:keff} shows the $k_{\text{eff}}$ values from Serpent and Moltres at
-973 K and the corresponding salt density, and Table \ref{table:ktemp} shows
+Moltres solves the six-group neutron diffusion equations as a
+steady-state eigenvalue problem to find the $k_{\text{eff}}$ for the static
+\gls{MSFR} model. Table
+\ref{table:keff} shows the $k_{\text{eff}}$ values from Serpent 2 and Moltres
+at 973 K and the corresponding salt density, and Table \ref{table:ktemp} shows
 the $k_{\text{eff}}$ values for other temperatures at 100 K intervals. We
@@
 which we attribute to two main factors: the accuracy of the neutron diffusion
-model, and the omission of the blanket tank structural material. The neutron
-diffusion model is not as accurate as the other S$_{\text{N}}$ or
+model, and the omission of the blanket tank structural material in Moltres.
+The neutron diffusion model is not as accurate as the other S$_{\text{N}}$ or
 SP$_{\text{N}}$ deterministic methods nor the Monte Carlo approach in Serpent.
@@
 	\centering
-	\caption{$k_{\text{eff}}$ values from Serpent and Moltres at 973 K.}
+	\caption{$k_{\text{eff}}$ values from Serpent 2 and Moltres at 973 K.}
 	\begin{tabular}{l S c}
 		\toprule
-		{Code} & {$k_{\text{eff}}$} & {Difference wrt Serpent [pcm]}\\
+		{Code} & {$k_{\text{eff}}$} \\
 		\midrule
-		{Serpent} & 1.00662 \pm 0.00005 & {-}\\
-		{Moltres with \glspl{DNP}} & 1.0079400 \pm 0.0000010 & 132\\
-		{Moltres without \glspl{DNP}} & 1.0049197 \pm 0.0000010 & {-}\\
+		{Serpent 2} & 1.00662 \pm 0.00005 \\
+		{Moltres with \glspl{DNP}} & 1.0079400 \pm 0.0000010 \\
+		{Moltres without \glspl{DNP}} & 1.0049197 \pm 0.0000010 \\
 		\bottomrule
@@
 	\centering
-	\caption{$k_{\text{eff}}$ values from Serpent and Moltres at various
+	\caption{$k_{\text{eff}}$ values from Serpent 2 and Moltres at various
 	temperatures from 800 K to 1400 K.}
@@
 		\toprule
-		{Temperature [K]} & {$k_{\text{eff}}$ $\pm$ 1-$\sigma$ uncertainty
-		(Serpent)} & {$k_{\text{eff}}$ (Moltres)} & {Difference wrt Serpent
-		[pcm]}
+		{Temperature [K]} & {$k_{\text{eff}}$ $\pm$ $\sigma$
+		(Serpent 2)} & {$k_{\text{eff}}$ (Moltres)} & {Difference wrt Serpent
+		2 [pcm]}
 		\\
@@
 
-The absolute value of $k_{\text{eff}}$ largely impacts the steady-state
+The absolute value of $k_{\text{eff}}$ impacts the final steady-state
 temperature of the reactor. We can raise or lower the average core temperature
@@
 important as they dictate the duration, shape, and magnitude of the reactor
-response to a transient initiator. The difference between $\beta$ and
+transient response. The difference between $\beta$ and
 $\beta_{\text{eff}}$ is that $\beta$ is the physical delayed neutron fraction
@@
 	\centering
-	\caption{$\beta_{\text{eff}}$ and $\beta$ values from Serpent and Moltres,
-	respectively, at 973 K.}
+	\caption{$\beta_{\text{eff}}$ and $\beta$ values from Serpent 2 and
+	Moltres, respectively, at 973 K.}
 	\begin{tabular}{l S S}
@@
 values of the reactivity given the relatively linear trend within the relevant
-temperature range (Figure \ref{fig:reactivity}). Table \ref{table:alpha} shows
+temperature range (Figure \ref{fig:reactivity}). The temperature range
+extends below the melting point of the fuel salt (873 K) to ensure that
+the data covers the relevant range between 873 K and 900 K. Table
+\ref{table:alpha} shows
 the temperature coefficients as described prior. The total temperature
@@
 		\toprule
-		{Code} & {$\alpha_D$ (log) [pcm]} & {$\alpha_D$ (linear) [pcm
+		{Software} & {$\alpha_D$ (log) [pcm]} & {$\alpha_D$ (linear) [pcm
 		K$^{-1}$]} & {$\alpha_\rho$ [pcm K$^{-1}$]} & {$\alpha_T$ [pcm
@@
     \includegraphics[width=.8\textwidth]{nt-spec}
-    \caption{Reactivity values from Serpent and Moltres. The Doppler
-    reactivity values were calculated at a fixed density of 4.1249 g
-    cm$^{-3}$. The density reactivity values were calculated at a fixed
-    temperature of 973 K.}
+    \caption{The fine-group and six-group neutron energy spectra from Serpent
+    2 and Moltres normalized per unit lethargy.}
     \label{fig:ntspec}
@@
 In summary, Moltres replicated most of the relevant neutronics parameters
-accurately with the group constant data from Serpent. The $\beta_{\text{eff}}$
-and temperature reactivity coefficients, which are important
-parameters for modeling transient reactor behavior, are in excellent
-agreement. The $k_{\text{eff}}$ values have discrepancies on the order of 100
-pcm and they are relatively small compared to the $k_{\text{eff}}$ values from
-the neutron diffusion and SP3 models in OpenFOAM
-\cite{cervi_development_2019}.
+accurately with the group constant data from Serpent. Moltres agrees with the
+high fidelity simulation in Serpent 2 for the $\beta_{\text{eff}}$ and
+temperature reactivity coefficients, which are important parameters for
+modeling transient reactor behavior. The $k_{\text{eff}}$ values have
+discrepancies on the order of 100 pcm and they are relatively small compared
+to the $k_{\text{eff}}$ values from the neutron diffusion and SP3 models in
+OpenFOAM \cite{cervi_development_2019}.

### 2020-07-03 abstract.tex
more editing

This thesis presents the Molten Salt Fast Reactor simulation capabilities of
Moltres, a simulation tool for molten salt reactors.

### 2020-07-03 introduction.tex
more editing

@@
 
-This thesis demonstrates latest capabilities of Moltres
-\cite{lindsay_introduction_2018}, an open source MSR simulation tool. Moltres
-is an application built in the \gls{MOOSE} parallel finite element framework.
-It solves
-the deterministic multigroup neutron diffusion and thermal-hydraulics
-\glspl{PDE} simultaneously on the same mesh. It supports up to 3-D meshes and
-scales well for a large number of processors. The underlying \gls{MOOSE}
-framework
-provides a range of implicit and explicit methods for the coupling between the
-neutronics and thermal-hydraulics governing equations.
-
-In the introductory journal article for Moltres, Lindsay et al.
-\cite{lindsay_introduction_2018} demonstrated Moltres' capabilities with
-2D-axisymmetric and 3D models of the \gls{MSRE}. The results showed good
-qualitative agreement with legacy \gls{MSRE} data but there were some minor
-quantitative discrepancies due to a number of differences in the legacy model. 
-Moltres has undergone further development in the past two years. The authors
-from the first paper have since developed various new capabilities in Moltres
-such as support for looping delayed neutron precursors back into the core,
-and a heat exchanger object to set the heat removal rate.
-
-In addition to those new capabilities, this thesis presents two more recent
-developments in Moltres, namely fully integrating \gls{MOOSE}'s incompressible
-Navier-Stokes module into Moltres, and introducing a
-decay heat model. The incompressible Navier-Stokes module provides a more
-accurate representation of the flow profile, precursor movement, and heat
-transfer as opposed to assuming uniform or function-based velocity fields.
+Moltres is an open source MSR simulation tool that falls under the latter
+category \cite{lindsay_introduction_2018}. It is an application built in the
+\gls{MOOSE} \cite{gaston_physics-based_2015} parallel finite element
+framework. Lindsay et al. first presented the tool in 2017 and demonstrated
+its capabilities by simulating 2-D and 3-D models of the \gls{MSRE}
+\cite{lindsay_introduction_2018}. The results showed good qualitative
+agreement with the original design calculations by \gls{MSRE} researchers at
+\gls{ORNL}. There has since been new developments in Moltres allowing for
+more accurate \gls{MSR} simulations.
 
@@
 
-The main objectives of this thesis are to demonstrate and verify Moltres'
+This thesis demonstrates latest capabilities of Moltres
+\cite{lindsay_introduction_2018}, an open source MSR simulation tool.
+In particular, this thesis presents two more recent
+developments in Moltres, namely fully integrating \gls{MOOSE}'s incompressible
+Navier-Stokes module into Moltres, and introducing a
+decay heat model.
+The main objective of this thesis is to verify Moltres'
 latest capabilities in modeling multiphysics, steady-state, and transient
@@
 simulation tools. The chapter also covers the \gls{MSFR} concept in greater
-detail. Chapter 3 details the simulation codes and the general modeling
+detail. Chapter 3 details the software and the general modeling
 approach for generating the results in this thesis. Chapter 4 provides a

### 2020-07-03 model.tex
more editing

@@
 for the former and Trelis \cite{noauthor_trelis_2018} for the latter.
-This chapter provides brief introductions of
-Serpent 2, MOOSE, and Moltres, and the general modeling approach for the
+This chapter provides brief introductions to
+Serpent 2, \gls{MOOSE}, and Moltres, and the modeling approach for the
 multiphysics simulations in Moltres.
@@
 
-\section{MOOSE and Moltres}
-
-Moltres \cite{lindsay_moltres_2017} is a \gls{MOOSE}-based
-\cite{gaston_physics-based_2015} application for
-coupled neutronics/thermal-hydraulics simulations of \glspl{MSR}. MOOSE
-provides foundational framework for computationally solving \glspl{PDE} and
-pre-built physics modules for common applications such as the Navier-Stokes
-physics module. Moltres is also a MOOSE physics module for simulating
-\glspl{MSR}. This section describes MOOSE, the Navier-Stokes physics module,
-and Moltres.
-
-\subsection{MOOSE}
+\section{MOOSE}
 
@@
 can inherit properties from related old objects to simplify implementation and
-reduce code duplication. Overall, this approach
+reduce code duplication. These objects are commonly referred to as kernels in
+\gls{MOOSE}. Overall, this approach
 is helpful for many researchers, a significant fraction of whom do not possess
@@
 
-\gls{MOOSE} itself relies heavily on libMesh \cite{kirk_libmesh:_2006} and
+\gls{MOOSE} itself relies on libMesh \cite{kirk_libmesh:_2006} and
 PETSc \cite{satish_petsc_2019} for its mesh handling and \gls{PDE} solver
@@
 
+\gls{MOOSE} includes a set of built-in physics modules for commonly studied
+phenomena such as the heat conduction, flow dynamics and solid mechanics. This
+work uses \gls{MOOSE}'s Navier-Stokes module for simulating
+incompressible salt flow in the \gls{MSFR}. Peterson et al. verified the
+incompressible flow capabilities in the
+Navier-Stokes module and presented results for common \gls{CFD} problems such
+as the lid-driven cavity, axisymmetric channel, and flow-over-a-sphere
+problems \cite{peterson_overview_2017}.
+
+\section{Moltres}
+
+Moltres is an application built in the \gls{MOOSE} parallel finite element
+framework \cite{lindsay_introduction_2018}. Similar to the physics modules in
+\gls{MOOSE}, Moltres contains the necessary kernels representing various
+physics and boundary conditions for solving for the neutron flux, delayed
+neutron precursor concentration, and temperature. Together with the
+Navier-Stokes module, it solves
+the deterministic multigroup neutron diffusion and thermal-hydraulics
+\glspl{PDE} simultaneously on the same mesh. Moltres supports up to 3-D meshes
+and scales well for a large number of processors. The underlying \gls{MOOSE}
+framework provides a range of implicit and explicit methods for the coupling
+between the neutronics and thermal-hydraulics governing equations.
+
+In the introductory journal article for Moltres, Lindsay et al.
+\cite{lindsay_introduction_2018} demonstrated Moltres' capabilities with
+2D-axisymmetric and 3D models of the \gls{MSRE}. The results showed good
+qualitative agreement with legacy \gls{MSRE} data with some minor
+quantitative discrepancies due to a number of differences in the legacy model. 
+Since then, Moltres has undergone further development in the past three years.
+The authors of the first paper have since developed various new capabilities
+in Moltres, most significantly providing support for looping delayed neutron
+precursors back into the core, and a pointwise heat removal kernel to simulate
+a heat exchanger. The present author demonstrated these capabilities in a
+2-D axisymmetric model of the \gls{MSFR} with uniform salt flow
+\cite{park_safety_2019}.
+
+Building on the prior progress, this thesis presents two more recent
+developments in Moltres, namely the new kernels required to couple
+incompressible flow with the delayed neutron precursor looping capability, and
+introducing a decay heat model to simulate decay heat from fission products.
+The incompressible flow profile from \gls{MOOSE}'s Navier-Stokes module
+provides a more accurate representation of the flow profile, precursor
+movement, and heat transfer as opposed to assuming uniform velocity fields
+featured in the previous papers \cite{lindsay_introduction_2018,
+park_safety_2019}.
+
+\section{Modeling Approach}
+
+This section discusses the neutronics and thermal-hydraulics
+\glspl{PDE} that Moltres solves and other relevant details specific to the
+\gls{MSFR} model in this work.
+
+As mentioned in the previous chapter, the fuel salt loop is divided into two
+regions, the central core region where most of the fissions take place, and
+the outer loop region where the heat exchanger is located. The central core
+region specifically refers to the central region indicated by
+the red box in Figure \ref{fig:core}. The outer loop is simplified
+into a 1-D pipe as it is a subcritical region and its main purposes are: to
+introduce an out-of-core residence time for the \glspl{DNP}, and to contain
+the heat removal kernel to simulate the heat exchanger.
+Accordingly, this section provides separate descriptions for the governing
+equations in the central core region and the outer loop region. There is also
+a third region comprised of the blanket salt, the NiCrW Hastelloy reflectors, 
+and other material. This region is important for capturing a more accurate
+estimate of neutron leakage than would
+imposing fixed albedo neutron flux boundary conditions.
+
 \begin{figure}[htb!]
@@
 
-As mentioned in the previous chapter, we divided the fuel salt loop into two
-regions, the central core region where most of the fissions take place, and
-the outer loop region where the heat exchanger is located. The central core
-region specifically refers to the central region indicated by
-the red box in Figure \ref{fig:core}. We simplified the outer loop
-into a 1-D pipe as it is a subcritical region and its main purposes are: to
-introduce a out-of-core residence time for the \glspl{DNP}, and to site the
-heat exchanger-equivalent object in Moltres.
-Accordingly, this section provides separate descriptions for the governing
-equations in the central core region and the outer loop region. There is also
-a third region comprising of the blanket salt, the Ni-alloy reflectors, 
-and other material . This region is important for capturing an accurate
-estimate of neutron leakage as opposed to simpler approximations such as
-imposing fixed albedo boundary conditions on the neutron fluxes.
-
 \subsection{Central Core Region}
@@
 transient scenarios; the center of the reactor is naturally where most of the
-fissions and heat generation occur. Consequently, the \glspl{PDE} solved for
-in this region are in their most complete form relatively to simplifying
-assumptions found in the other regions.
+fissions and heat generation occur.
 
@@
     \phi_{g'} - \lambda_i C_i - \boldsymbol{u} \cdot \nabla C_i + \nabla \cdot
-    D_t \nabla C_i, \label{eq:dnp} \\
+    K \nabla C_i, \label{eq:dnp} \\
     \intertext{where}
-    v_g &= \text{average speed of neutrons in group $g$ [cm s$^{-1}$],} 
+    v_g &= \text{average speed of neutrons in group $g$ [cm$\cdot$s$^{-1}$],} 
     \nonumber \\
-    \phi_g &= \text{neutron flux in group $g$ [cm$^{-2}$ s$^{-1}$],} \nonumber
-    \\
+    \phi_g &= \text{neutron flux in group $g$ [cm$^{-2}\cdot$s$^{-1}$],}
+    \nonumber \\
     t &= \text{time [s],} \nonumber \\
-    D_g &= \text{diffusion coefficient of neutrons in group $g$ [cm$^2$
-    s$^{-1}$],} \nonumber \\
+    D_g &= \text{diffusion coefficient of neutrons in group $g$
+    [cm$^2\cdot$s$^{-1}$],} \nonumber \\
     \Sigma^r_g &= \text{macroscopic cross section for removal of neutrons from
@@
     group $i$ [cm$^{-3}$],} \nonumber \\
-    D_t &= \text{turbulent diffusion of the delayed neutron precursors [cm$^2$
-    s$^{-1}$].} \nonumber
+    K &= \text{turbulent diffusion of the delayed neutron precursors
+    [cm$^2\cdot$s$^{-1}$].} \nonumber
 \end{align}
@@
 accuracy at lower computational costs. Moreover, the central core region
-contains no material interfaces except at its boundaries. The Neutronics
-results section provides a comparison of the \gls{MSFR} multiplication factor
+contains no material interfaces except at its boundaries. Chapter
+\ref{chap:nts} provides a comparison of the \gls{MSFR} multiplication factor
 values and reactivity coefficients between Moltres and Serpent.
@@
 diffusion terms to account for the movement of \glspl{DNP} in the primary
-coolant loop. The turbulent diffusion value is governed by the following
+coolant loop. The turbulent diffusion $K$ is governed by the following
 equation:
@@
 \begin{align}
-    D_t &= \frac{\mu_t}{\rho Sc_t}
+    K &= \frac{\mu_t}{\rho Sc_t}
     \intertext{where}
@@
 %
-
-We assumed $Sc_t = 0.85$ for a fair comparison with the Polimi/TUDelft
-models which used the same value.
+This work assumes $Sc_t = 0.85$ for a fair comparison with the Polimi and
+TUDelft models \cite{fiorina_modelling_2014} which used the same value. It has
+its roots in the Reynolds Analogy which states that turbulent momentum and
+heat transfer largely depend on the same eddies in turbulent flow
+\cite{bartosiewicz_612_2019}. Therefore,
+$Sc_t$ should be close to unity. $Sc_t = 0.85$ is also the default value for
+most commercial \gls{CFD} software \cite{bartosiewicz_612_2019}.
 
@@
 temperatures using one of the many predefined interpolation methods available
-in \gls{MOOSE}. Once again, users have the freedom to select their
+in \gls{MOOSE}. Once again, Moltres allows users to select their
 interpolation method of choice.
 
-For this work, we have six neutron energy groups according to the energy
-boundaries in table \ref{table:bound}, eight \gls{DNP} groups as defined by
-the JEFF-3.1.2 library, and the spline interpolation method. The neutron flux
+The \gls{MSFR} model in this work six neutron energy groups according to the
+energy boundaries in table \ref{table:bound}, and eight \gls{DNP} groups as
+defined by the JEFF-3.1.2 library. The neutron flux
 and \gls{DNP} concentration values were approximated by first-order Lagrange
@@
 For the decay heat model, a previous study on the MSFR by Aufiero et al.
-\cite{aufiero_extended_2013} had shown that using three decay heat precursor
-groups with different half-lives in the form of exponential equations, can
+\cite{aufiero_extended_2013} showed that using three decay heat precursor
+groups with appropriate half-lives in the form of exponential equations, can
 accurately model decay heat in the MSFR for up to 300 seconds after shutdown
-with a relative error of less than 2\%. Thus, we implemented a decay heat
-model using the following equation:
+with a relative error of less than 2\%. Thus, this thesis implements the new
+decay heat modeling capability with the following equation:
 %
 \begin{align}
-	\frac{\partial \omega_k}{\partial t} &= f_k \sum^G_{g=1} \epsilon_{g}
-	\Sigma^f_{g} \phi_{g} - \lambda^k \omega_k - \boldsymbol{u} \cdot \nabla
-	\omega_k + \nabla \cdot D_t \nabla \omega_k, \label{eq:decayheat} \\
+	\frac{\partial \omega_j}{\partial t} &= f_j \sum^G_{g=1} \epsilon_{g}
+	\Sigma^f_{g} \phi_{g} - \lambda_j \omega_j - \boldsymbol{u} \cdot \nabla
+	\omega_j + \nabla \cdot K \nabla \omega_j, \label{eq:decayheat} \\
 	\intertext{where}
-    \omega_k &= \text{total decay heat power density from decay heat
-    precursors in group $k$ [W cm$^{-3}$]} \nonumber \\
-	f_k &= \text{fraction of decay heat to total power at steady state}
+    \omega_j &= \text{total decay heat power density from decay heat
+    precursors in group $j$ [W$\cdot$cm$^{-3}$],} \nonumber \\
+	f_j &= \text{fraction of decay heat from group $j$ to total power under
+	full-power conditions,}
 	\nonumber \\
-	\epsilon_g &= \text{average fission energy per fission [W]} \nonumber \\
-	\lambda^k &= \text{average decay constant of decay heat precursors in
-	group $k$ [s$^{-1}$].} \nonumber
+	\epsilon_g &= \text{average fission energy per fission [W],} \nonumber \\
+	\lambda_j &= \text{average decay constant of decay heat precursors in
+	group $j$ [s$^{-1}$].} \nonumber
 \end{align}
 
-Like the neutron and \gls{DNP} groups, Moltres can take an arbitrary number of
-decay heat groups. In this work, as with the other parameters, we used the
-same decay heat fractions and decay constants, shown in Table
-\ref{eq:decayheat}, used in the Polimi/TUDelft models for three decay heat
+Like the neutron and \gls{DNP} groups, Moltres can accommodate an arbitrary
+number of decay heat groups. This work uses the same decay heat fractions and decay constants, shown in Table
+\ref{eq:decayheat}, used in the Polimi and TUDelft models for three decay heat
 groups.
@@
 	\centering
-	\caption{Decay heat group parameters \cite{fiorina_modelling_2014}.
-	$\lambda_i$ and $f_i$ are the decay constants and decay heat fractions
-	associated to group $i$.}
+	\caption{Decay heat group parameters \cite{fiorina_modelling_2014}.}
 	\begin{tabular}{S S S}
 		\toprule
-		{Decay heat group} & {$\lambda_i$ [s$^{-1}$]} & {$f_i$} \\
+		{Decay heat group $j$} & {$\lambda_j$ [s$^{-1}$]} & {$f_j$} \\
 		\midrule
@@
 
-Fluid dynamics in Moltres can be modeled using the \gls{INS} equations with
-the Boussinesq hypothesis for eddy viscosity. Most of the Navier-Stokes
-capabilities in Moltres is derived from the MOOSE Navier-Stokes module
-\cite{peterson_overview_2017}. The standard \gls{INS} equations are:
+This work models fluid dynamics using the \gls{INS} capabilities from the
+MOOSE Navier-Stokes module \cite{peterson_overview_2017}. The standard
+\gls{INS} equations are:
 %
@@
 
-In addition to the intrinsic molecular viscosity, we introduced an eddy
-viscosity term to approximate turbulent flow effects. The current
-implementation of Moltres does not have a turbulence model such as the
-\gls{RANS} models used in the Polimi/TUDelft models. Thus, we made a
-zeroth-order approximation of the eddy viscosity based on the results reported
-in the Polimi/TUDelft models. The eddy viscosity is assumed to be 40 Pa s.
-Despite the simplicity of this assumption, we were able to reproduce much of
-the flow profile observed in the Polimi/TUDelft models at steady state.
+In addition to the intrinsic molecular viscosity, this thesis introduces an
+eddy viscosity term to approximate turbulent flow effects. The current
+implementation of the Navier-Stokes module does not have a turbulence model
+such as the \gls{RANS} models used in the Polimi and TUDelft models
+\cite{fiorina_modelling_2014}. Thus, this work uses
+a zeroth-order approximation of the eddy viscosity based on the results
+reported in the Polimi/TUDelft models. The eddy viscosity is assumed to be 40
+Pa s. Despite the simplicity of this assumption, the resulting flow profile is
+similar to flow profile in the Polimi and TUDelft models at steady state.
 
-The energy balance equation for temperature is
-given in Eq. \ref{eq:temp}. The diffusion term includes turbulent heat
-diffusivity based on the eddy viscosity $\mu_t$ and the turbulent Prandtl
-number $Pr_t$, which we assume to be equal to 0.85. 
+The energy balance equation for temperature used in this Moltres model is:
 %
@@
     k_t &= \frac{\mu_t}{\rho Pr_t} \\
-    Q_s &= \Big( 1 - \sum^K_{k=1} f_k \Big) \sum^G_{g=1} \epsilon_g \Sigma_g^f
-    \phi_g + \sum^K_{k=1} \omega_k, \label{eq:source}
+    Q_s &= \Big( 1 - \sum^J_{j=1} f_j \Big) \sum^G_{g=1} \epsilon_g \Sigma_g^f
+    \phi_g + \sum^J_{j=1} \omega_j, \label{eq:source}
     \intertext{where}
-    c_p &= \text{specific heat capacity of molten salt [J kg$^{-1}$ K$^{-1}$],
-    } \nonumber \\
+    c_p &= \text{specific heat capacity of molten salt
+    [J$\cdot$kg$^{-1}\cdot$K$^{-1}$],} \nonumber \\
     T &= \text{temperature of molten salt [K]} \nonumber \\
-    \boldsymbol{u} &= \text{velocity of molten salt [m s$^{-1}$],} \nonumber
-    \\
-    k &= \text{thermal conductivity of molten salt [W m$^{-1}$ K$^{-1}$],}
+    \boldsymbol{u} &= \text{velocity of molten salt [m$\cdot$s$^{-1}$],}
     \nonumber \\
-    K &= \text{total number of decay heat groups.} \nonumber
+    k &= \text{thermal conductivity of molten salt
+    [W$\cdot$m$^{-1}\cdot$K$^{-1}$],} \nonumber \\
+    J &= \text{total number of decay heat groups.} \nonumber
 \end{align}
 %
-The first term in the heat source $Q_s$ represents prompt fission heat, and
-the second term represents decay heat from the $K$ decay heat groups.
+The diffusion term includes turbulent heat
+diffusivity based on the eddy viscosity $\mu_t$ and the turbulent Prandtl
+number $Pr_t$. $Pr_t$ is also 0.85 due to the same reasoning provided for
+$Sc_t$. The first term in the heat source $Q_s$ equation represents prompt
+fission heat, and the second term represents decay heat from the $J$ decay
+heat groups.
 
-We expect good qualitative agreement with the Polimi/TUDelft models,
+We expected good qualitative agreement with the Polimi and TUDelft models,
 including the large recirculation region near the blanket tank walls and the

### 2020-07-04 abstract.tex
final edits for 2nd draft

@@
-This thesis presents the Molten Salt Fast Reactor simulation capabilities of
-Moltres, a simulation tool for molten salt reactors.
+Molten salt reactors are a class of advanced nuclear reactors that promise
+numerous improvements over the current fleet of largely light-water reactors.
+As the world continues its transition towards low-carbon electricity
+generation to combat climate change, molten salt reactors are potential
+options in the near-term future for replacing fossil fuel and ageing nuclear
+power plants. At the current state of development, molten salt reactors still
+require extensive research to become viable. This thesis presents latest
+developments in Moltres, a simulation tool for molten salt reactors. These new
+developments are: support for coupling the incompressible Navier-Stokes and
+the delayed neutron precursor looping systems, and a new decay heat model for
+simulating decay heat from fission products at steady state and during
+transients. This work demonstrates these capabilities in conjunction with
+existing capabilities through multiphysics simulations of the Molten Salt Fast
+Reactor concept. This work first verifies the six-group neutron diffusion
+results from Moltres against continuous-energy Monte Carlo neutron transport
+results from Serpent 2. The multiplication factors $k_{\text{eff}}$, delayed
+neutron fractions $\beta$, temperature reactivity coefficient $\alpha_T$, and
+the six-group neutron energy spectra from Moltres agreed with the high
+fidelity simulation results from Serpent 2. The $k_{\text{eff}}$ values have
+small discrepancies on the order of 100 pcm, which is smaller than some
+reported literature values. The decay heat model shows an expected flattening
+of the temperature distribution due to the movement and dispersion of the
+decay heat precursors throughout the primary coolant loop. This work also
+demonstrates and verifies steady state and
+transient multiphysics simulations of the Molten Salt Fast Reactor. The
+transient scenarios under study are unprotected instances of reactivity
+insertion, loss of heat sink, loss of flow, and pump overspeed. The steady
+state and transient results are verified against results from another paper
+that presented results for the same cases. The steady-state temperature and
+velocity distributions, and the peak neutron flux showed good agreement with
+the literature results. Minor differences in the delayed neutron precursor
+distribution and the in-core delayed neutron fraction were explainable with
+the differences in the handling of turbulence in the models. In three of the
+transient results (reactivity insertion, loss of heat sink, and pump
+overspeed), Moltres reproduced the expected magnitude and pattern of the
+reactor response to these transient initiators.
+The loss of flow results showed greater discrepancies that are attributed
+to differences in the fluid dynamics modeling in Moltres and the other models.
+Through the verification studies, this work has also identified avenues
+for further Moltres software development.

### 2020-07-04 conclusion.tex
final edits for 2nd draft

@@
-The unique phenomena in \glspl{MSR}, generally not found in conventional
-reactors, necessitate the development of new reactor safety analysis software.
+The unique phenomena in \glspl{MSR}
+necessitate the development of new reactor safety analysis software.
 This thesis presents the latest developments in Moltres, namely coupling its
@@
 transient. However, Moltres performed poorly in the loss of flow transient as
-we could not incorporate buoyancy-driven flow and its associated effects on
-$\mu_t$.
+this work could not incorporate buoyancy-driven flow and its associated
+effects on $\mu_t$.
 
@@
 Further research and development on Moltres should aim to rectify the issues
-mentioned in this thesis. There are three main avenues for improvement.
-Firstly, proper 2D/3D heat exchanger implementation
+mentioned in this thesis. This work has highlighted three main avenues for
+improvement. Firstly, proper 2D/3D heat exchanger implementation
 would allow us to move away from the 1D outer loop system and towards a full
-2D/3D closed loop. The biggest change from this is users would be able to use
+2D/3D closed loop. The biggest change is that users would be able to use
 the Boussinesq approximation for buoyancy-driven flow capability in Moltres.

### 2020-07-04 introduction.tex
final edits for 2nd draft

@@
 atmosphere has been rising steadily throughout modern human history
-\cite{us_department_of_commerce_global_nodate}. Climate change has dire
+\cite{us_department_of_commerce_global_2020}. Climate change has dire
 consequences on human health and safety due to extreme weather events and the
-overall impact on food production \cite{noauthor_global_nodate}. As a result,
+overall impact on food production \cite{mcmichael_global_2004}. As a result,
 urgent measures are necessary to limit greenhouse gas emissions, namely CO$_2$
@@
 Electricity generation from burning fossil fuels represents the
-largest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019});
+greatest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019});
 replacing it with low-carbon
-alternatives would curb a large fraction of emissions. Nuclear power is
+alternatives would curb a significant fraction of emissions. Nuclear power is
 a viable low-carbon replacement for burning fossil fuels.
@@
 The world would have to ramp up the current rate of reactor deployments to
-displace some of the
-presently large share of energy production from fossil fuel power plants.
-However, several obstacles stand in the way of mass reactor deployments. These
-include perceived safety risks, sustainability concerns, nuclear proliferation
+displace some of the presently large share of energy production from fossil
+fuel power plants. However, several obstacles stand in the way of mass reactor
+deployments. These obstacles include perceived safety risks, sustainability
+concerns, nuclear proliferation
 risks, and the ability to compete economically with other sources of energy
@@
 design such as PARCS and TRACE for \glspl{LWR}, and SAS4A/SASSYS-1 for
-liquid metal cooled reactors. Thus, these applications generally would not
-account for \gls{MSR}-specific phenomena and are not suitable for \gls{MSR}
+liquid metal cooled reactors. Thus, these applications would disregard
+\gls{MSR}-specific phenomena and are inappropriate for \gls{MSR}
 analysis without any modifications to the source code. Some research efforts
@@
 development for ensuring that the application produces accurate and reliable
-results. This thesis covers the \gls{MSFR} concept mainly because 
-there is readily available literature data to verify our results against. The
+results. This thesis covers the \gls{MSFR} concept mainly because it has been
+studied extensively with readily available literature data to verify our
+results against. The
 steady-state and transient results also feature a significant degree of

### 2020-07-04 model.tex
final edits for 2nd draft

@@
 \glspl{PDE} simultaneously on the same mesh. Moltres supports up to 3-D meshes
-and scales well for a large number of processors. The underlying \gls{MOOSE}
+and scales well over a large number of processors. The underlying \gls{MOOSE}
 framework provides a range of implicit and explicit methods for the coupling
@@
     \frac{\partial C_i}{\partial t} &= \beta_i \sum^G_{g'=1} \nu \Sigma^f_{g'}
-    \phi_{g'} - \lambda_i C_i - \boldsymbol{u} \cdot \nabla C_i + \nabla \cdot
+    \phi_{g'} - \lambda_i C_i - \vec{u} \cdot \nabla C_i + \nabla \cdot
     K \nabla C_i, \label{eq:dnp} \\
@@
     \centering
-    \includegraphics[width=.8\textwidth]{mesh}
+    \includegraphics[width=\textwidth]{mesh2}
     \caption{Mesh adopted in Moltres and a close-up view of the mesh around
@@
 Figure \ref{fig:mesh} shows the mesh adopted for the \gls{MSFR} model.
-We assumed vacuum boundary conditions for all six neutron group fluxes along
+This work assumes vacuum boundary conditions for all six neutron group fluxes
+along
 the external boundaries of the geometry, and homogeneous Neumann boundary
@@
 	\frac{\partial \omega_j}{\partial t} &= f_j \sum^G_{g=1} \epsilon_{g}
-	\Sigma^f_{g} \phi_{g} - \lambda_j \omega_j - \boldsymbol{u} \cdot \nabla
+	\Sigma^f_{g} \phi_{g} - \lambda_j \omega_j - \vec{u} \cdot \nabla
 	\omega_j + \nabla \cdot K \nabla \omega_j, \label{eq:decayheat} \\
@@
 \begin{align}
-    \text{Momentum eq.:} && \rho \frac{\partial \boldsymbol{u}}{\partial t} &=
-    -\rho (\boldsymbol{u}
-    \cdot \nabla) \boldsymbol{u} + \nabla \cdot (-p \boldsymbol{I} + \mu [
-    \nabla \boldsymbol{u} + (\nabla \boldsymbol{u})^T] + \boldsymbol{f} &&
+    \text{Momentum eq.:} && \rho \frac{\partial \vec{u}}{\partial t} &=
+    -\rho (\vec{u}
+    \cdot \nabla) \vec{u} + \nabla \cdot (-p \vec{I} + \mu [
+    \nabla \vec{u} + (\nabla \vec{u})^T] + \vec{f} &&
     \label{eq:momemtum} \\
-    \text{Divergence-free:} && \nabla \cdot \boldsymbol{u} &= 0 &&
+    \text{Divergence-free:} && \nabla \cdot \vec{u} &= 0 &&
     \label{eq:divergence}
@@
     && \mu &= \text{ dynamic viscosity [Pa s],} && \nonumber \\
-    && \boldsymbol{f} &= \text{ body force per unit volume [N m$^{-3}$].} &&
+    && \vec{f} &= \text{ body force per unit volume [N$\cdot$m$^{-3}$].} &&
     \nonumber
@@
 \begin{align}
-    \rho c_{p} \frac{\partial T}{\partial t} &= - \rho c_p \boldsymbol{u}
+    \rho c_{p} \frac{\partial T}{\partial t} &= - \rho c_p \vec{u}
     \cdot \nabla T + \nabla \cdot [(k + k_t) \nabla T] + Q_s
@@
     T &= \text{temperature of molten salt [K]} \nonumber \\
-    \boldsymbol{u} &= \text{velocity of molten salt [m$\cdot$s$^{-1}$],}
+    \vec{u} &= \text{velocity of molten salt [m$\cdot$s$^{-1}$],}
     \nonumber \\
@@
 
-We expected good qualitative agreement with the Polimi and TUDelft models,
-including the large recirculation region near the blanket tank walls and the
-resulting high temperatures in that region. There would be some minor
+With this model, the results were expected to show good qualitative agreement
+with the Polimi and TUDelft models, including the large recirculation region
+near the blanket tank walls and the resulting high temperatures in that
+region. The results in Chapter \ref{chap:ss} would be some minor
 discrepancies where the viscosity values are under- or over-predicted, leading
-to slightly inaccurate temperature and precursor concentration values from
-turbulent diffusion.
+to minor inaccuracies in temperature and precursor concentration distributions
+from turbulent diffusion.
+
+\subsubsection{Boundary Conditions}
+
+Table \ref{table:corebc} summarizes the boundary conditions for all variables
+on all of the relevant boundaries. Figure \ref{fig:msfrbc} shows the locations
+of the various boundaries listed in the table.
+
+\begin{table}[htbp!]
+    \small
+	\caption{Boundary conditions in the main reactor geometry (Figure
+	\ref{fig:msfrbc}.}
+	\centering
+	\begin{tabular}{ l l c}
+		\toprule
+		Variable & Boundary & Boundary Condition \\
+		\midrule
+		\multirow{4}{*}{Neutron flux $\phi_g$} & Top & $\frac{d \phi_g}{dx}
+		\big|_{\text{inflow}} = 0$ \\
+        & Outer & $\frac{d \phi_g}{dx} \big|_{\text{inflow}} = 0$ \\
+        & Bottom & $\frac{d \phi_g}{dx} \big|_{\text{inflow}} = 0$ \\
+        & Axial & $\frac{d \phi_g}{dx} = 0$ \\
+        \midrule
+        \multirow{6}{*}{Delayed neutron precursor concentration $C_i$} &
+        Top (Core) & $\frac{d C_i}{dx} = 0$ \\
+        & Bottom (Core) & $\frac{d C_i}{dx} = 0$ \\
+        & Outer (Core) & $\frac{d C_i}{dx} = 0$ \\
+        & Axial (Core) & $\frac{d C_i}{dx} = 0$ \\
+        & Inlet (Core) & $C_i = c$ \\
+        & Outlet (Core) & $u_x \cdot C_i = 0$ \\
+        \midrule
+        \multirow{6}{*}{Decay heat power density $\omega_j$} &
+        Top (Core) & $\frac{d \omega_j}{dx} = 0$ \\
+        & Bottom (Core) & $\frac{d \omega_j}{dx} = 0$ \\
+        & Outer (Core) & $\frac{d \omega_j}{dx} = 0$ \\
+        & Axial (Core) & $\frac{d \omega_j}{dx} = 0$ \\
+        & Inlet (Core) & $\omega_j = c$ \\
+        & Outlet (Core) & $u_x \cdot \omega_j = 0$ \\
+        \midrule
+        \multirow{6}{*}{Radial velocity $u_x$} & Top (Core) & $u_x = 0$ \\
+        & Bottom (Core) & $u_x = 0$ \\
+        & Outer (Core) & $u_x = 0$ \\
+        & Axial (Core) & $u_x = 0$ \\
+        & Inlet (Core) & $u_x = c$ \\
+        & Outlet (Core) & $\frac{d u_x}{dx} = 0$ \\
+        \midrule
+        \multirow{6}{*}{Axial velocity $u_y$} & Top (Core) & $u_y = 0$ \\
+        & Bottom (Core) & $u_y = 0$ \\
+        & Outer (Core) & $u_y = 0$ \\
+        & Axial (Core) & $\frac{d u_y}{dx} = 0$ \\
+        & Inlet (Core) & $u_y = 0$ \\
+        & Outlet (Core) & $\frac{d u_y}{dx} = 0$ \\
+        \midrule
+        \multirow{6}{*}{Temperature $T$} & Top (Core) &
+        $\frac{d T}{dx} = 0$ \\
+        & Bottom (Core) & $\frac{d T}{dx} = 0$ \\
+        & Outer (Core) & $\frac{d T}{dx} = 0$ \\
+        & Axial (Core) & $\frac{d T}{dx} = 0$ \\
+        & Inlet (Core) & $T = c$ \\
+        & Outlet (Core) & $\frac{d T}{dx} = 0$ \\
+		\bottomrule
+	\end{tabular}
+	\label{table:corebc}
+\end{table}
+
+\clearpage
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\textwidth]{msfrbc}
+    \caption{The boundaries in the \gls{MSFR} geometry that are relevant for
+    the boundary conditions mentioned in Table \ref{table:corebc}.}
+    \label{fig:msfrbc}
+\end{figure}
 
@@
 
-Since this region is largely subcritical, the only significant
-neutronics-related phenomena are the drift, and decay of \glspl{DNP}.
+Since this outer loop region is largely subcritical, the only significant
+neutronics-related phenomena are the drift and decay of \glspl{DNP}.
 The governing equation for the \glspl{DNP} is:
@@
     \frac{\partial C_i}{\partial t} &= - \lambda_i C_i - u
-    \frac{\partial C_i}{\partial x} + D_t \frac{\partial^2 C_i}{\partial x^2}.
+    \frac{\partial C_i}{\partial x}.
     \label{eq:dnploop}
@@
 
-Instead of the Navier-Stokes equations, we have a predetermined velocity of
-1.1275 m s$^{-1}$ in the outer loop region. The governing equation for
+A constant velocity of 1.1275 m s$^{-1}$ is applied in the outer loop region.
+The governing equation for
 temperature, derived from equation \ref{eq:temp}, is:
@@
     \intertext{where}
-    \alpha &= \text{heat transfer coefficient [W K$^{-1}$],} \nonumber \\
+    Q_{hx} &= \text{heat removal rate through the heat exchanger [W],} 
+    \nonumber \\
+    \alpha &= \text{heat transfer coefficient [W$\cdot$K$^{-1}$],} \nonumber
+    \\
+    T_i &= \text{temperature of the intermediate salt [K],} \nonumber \\
     x_0 &= \text{position of the point heat exchanger [m].} \nonumber
@@
 The fission heat source term is replaced with a heat exchanger sink term
+$Q_{hx}$
 which depends on the temperature difference between the fuel salt $T$ and the
-secondary loop salt $T_s$. For simplicity, we assumed a constant temperature
-of 823 K in the secondary loop. The heat transfer coefficient was determined
-by assuming that the fuel outlet temperature is 1023 K and calculating the
-heat removal rate to induce a 100 K drop at the given volumetric flow rate
-and heat capacity of the fuel salt. We opted to ignore the diffusion term due
-to the discontinuity of the temperature distribution across the point heat
-exchanger.
+intermediate loop salt $T_i$. For simplicity, this work assumes a constant
+temperature of 823 K in the intermediate loop. The heat transfer coefficient
+was determined by assuming that the fuel outlet temperature is 1023 K and
+calculating the heat removal rate to induce a 100 K drop at the given
+volumetric flow rate and heat capacity of the fuel salt. The resulting value
+for $\alpha$ is 370.668 W$\cdot$K$^{-1}$. This work opted to
+ignore the diffusion term due to the discontinuity of the temperature
+distribution across the point heat exchanger.
 
-\subsection{Boundary Conditions and Flow Transfers}
+\subsubsection{Boundary Conditions}
 
-This subsection details the various boundary conditions for all of the
-tracked variables, and the \gls{DNP} and temperature flow transfers between
-the central core and outer loop regions.
+Table \ref{table:loopbc} summarizes the boundary conditions for all variables
+on the inlet and outlet of the 1-D outer loop region.
 
-Starting with the central core region, for the neutron group fluxes, we
-imposed vacuum boundary conditions on the
-outermost boundaries of the geometry in Figure \ref{fig:core} excluding the
-axial boundary. The \gls{DNP} variables have homogeneous Neumann boundary
-conditions along the axis and the walls in the central core region, and inflow
-and outflow boundary conditions on the inlet and outlet boundaries,
-respectively. The temperature variable shares the same type of boundary
-conditions as the \gls{DNP} variables.
+\begin{table}[htbp!]
+    \small
+	\caption{Boundary conditions in the 1-D outer loop geometry. $u$
+	represents the 1-D velocity in this region.}
+	\centering
+	\begin{tabular}{ l l c}
+		\toprule
+		Variable & Boundary & Boundary Condition \\
+        \midrule
+        \multirow{2}{*}{Delayed neutron precursor concentration $C_i$} &
+        Inlet (Core) & $C_i = c$ \\
+        & Outlet (Core) & $u \cdot C_i = 0$ \\
+        \midrule
+        \multirow{2}{*}{Decay heat power density $\omega_j$} &
+        Inlet (Core) & $\omega_j = c$ \\
+        & Outlet (Core) & $u \cdot \omega_j = 0$ \\
+        \midrule
+        \multirow{2}{*}{Temperature $T$} &
+        Inlet (Core) & $T = c$ \\
+        & Outlet (Core) & $u \cdot T = 0$ \\
+		\bottomrule
+	\end{tabular}
+	\label{table:loopbc}
+\end{table}
 
-The flow rate is dictated by an inflow boundary condition at the core inlet
-for a volumetric flow rate of 4.5 m$^3$ s$^{-1}$. We imposed no-slip boundary
-conditions on the walls of the central core, and homogeneous Neumann boundary
-conditions on the outlet. Along the axial boundary, we have homogeneous
-Dirichlet boundary condition for the radial velocity component, and
-homogeneous Neumann boundary condition for the axial velocity component.
+\subsection{Flow Transfers}
+
+This subsection details the delayed neutron and decay heat precursors, and
+temperature flow transfers between the central core and outer loop regions.
+
+%Starting with the central core region, for the neutron group fluxes, we
+%imposed vacuum boundary conditions on the
+%outermost boundaries of the geometry in Figure \ref{fig:core} excluding the
+%axial boundary. The \gls{DNP} variables have homogeneous Neumann boundary
+%conditions along the axis and the walls in the central core region, and inflow
+%and outflow boundary conditions on the inlet and outlet boundaries,
+%respectively. The temperature variable shares the same type of boundary
+%conditions as the \gls{DNP} variables.
+
+A parabolic flow profile was imposed on the inlet Dirichlet boundary
+condition. The equation for $u_x$ at the inlet is:
+%
+\begin{align}
+    u_x &= -2034.01 \Big[\frac{y}{18.75} - \Big(\frac{y}{18.75}\Big)^2 
+    \Big] \\
+    \intertext{where}
+    y = \text{height along the inlet [cm].}
+\end{align}
+%
+The inlet is 18.75 cm high, and the coefficient at the front is from
+normalizing the equation for a total volumetric flow rate of
+4.5 m$^3$ s$^{-1}$.
 
 At every timestep, Moltres also calculates weighted averages of the
-temperature and \gls{DNP} at the outlet. These values are weighted by the
+temperature and the precursors at the outlet. These values are weighted by the
 outflow velocity values at the outlet according to the following equation:
@@
 \begin{align}
-    \overline{\phi} &= \frac{\int_\mathcal{C} \phi(y) u(y) dy}{
+    \overline{\psi} &= \frac{\int_\mathcal{C} \psi(y) u(y) dy}{
     \int_\mathcal{C} u(y) dy} \\
     \intertext{where}
-    \phi &= \text{variable to be weighted} \nonumber \\
+    \psi &= \text{variable to be weighted} \nonumber \\
     \mathcal{C} &= \text{outlet boundary curve} \nonumber \\

### 2020-07-04 msr.tex
final edits for 2nd draft

@@
 The fuel consisted of enriched uranium in a molten salt mixture of NaF,
-ZrF$_4$, and UF$_4$, and the reactor used blocks of BeO for neutron
+ZrF$_4$, and UF$_4$. Additionally, the reactor used blocks of BeO for neutron
 moderation. The aircraft program
@@
 new program for the construction and operation of a demonstration reactor
-based on the Molten Salt Breeder Reactor (MSBR) concept that they had
+based on the \gls{MSBR} concept that they had
 developed \cite{macpherson_molten_1985}. The \gls{MSBR} is a thermal-spectrum,
@@
 once-through
-converter system without fuel reprocessing. While the fuel largely consists of
-19.75 \% \gls{HALEU}, the initial core loading includes thorium to
-boost its conversion ratio throughout its lifetime. It has a continuous
+converter system without fuel reprocessing. While the fuel consists of 19.75
+\% high-assay low-enriched uranium, the initial core loading includes thorium
+to boost its conversion ratio throughout its lifetime. It has a continuous
 online feed consisting of \gls{HALEU} to maintain criticality, and denatured
@@
 There was a concurrent program at the UK Atomic Energy Authority for the
-development of a 2500 MW$_{\text{e}}$ lead-cooled
-\gls{MCFR} concept \cite{smith_assessment_1974}. It is a dual fluid system,
+development of a 2500 MW$_{\text{e}}$ lead-cooled Molten Chloride Fast Reactor
+concept \cite{smith_assessment_1974}. It is a dual fluid system,
 with separate loops for the fuel salt and the blanket salt. The blanket is a
@@
 picked up pace due to renewed interest initiated by the Generation IV Forum in
-2001. Today, there are numerous \gls{MSR} concepts under active development
-led by various national and commercial bodies.
+2001. Today, various national and commercial bodies are developing numerous
+\gls{MSR} concepts.
 
@@
 the core. In particular, graphite moderators tend to have slightly positive
-temperature reactivity coefficients. \gls{MSBR} concept has this issue, but
-the total temperature reactivity coefficient is still relatively large and
+temperature reactivity coefficients. The \gls{MSBR} concept has this issue,
+but the total temperature reactivity coefficient is still relatively large and
 negative \cite{rykhlevskii_modeling_2019}. The negative temperature reactivity
-feedback provides a large degree of control and stability as it is always
+feedback provides a great degree of control and stability as it is always
 present in an \gls{MSR} regardless of the operating conditions.
@@
 core actively cooled by fans or other cooling systems to keep its temperature
-just below the freezing point of the salt \cite{aji_experimental_2020}. In
-general, when temperatures in the core exceed a certain threshold during a
-dangerous transient, the freeze plug melts and the molten salt in the core
-drains into a containment tank designed to keep the salt in a subcritical
+just below the freezing point of the salt \cite{aji_experimental_2020}. When
+temperatures in the core exceed a certain threshold during a
+dangerous transient, the freeze plug melts and drains the molten salt
+into a containment tank designed to keep the salt in a subcritical
 configuration. This is especially easy to achieve in thermal-spectrum
@@
 
-\glspl{MSR} also typically have large margin-to-boiling under nominal
-operating conditions so that fuel salt boiling does not occur
+\glspl{MSR} also typically have a high margin to boiling under nominal
+operating conditions so that fuel salt boiling is absent
 \cite{elsheikh_safety_2013}. Furthermore, the reactor vessel is consequently
@@
 
-However, there are also some safety disadvantages associated with \glspl{MSR}.
+However, \glspl{MSR} do have some associated safety risks.
 Firstly, \glspl{MSR} have smaller fractions of delayed neutron precursors in
@@
 the decrease in average neutron lifetime. Secondly, hot molten salts are
-generally corrosive and the corrosion mechanism is different from conventional
+corrosive and the corrosion mechanism is different from conventional
 corrosion induced by water and other common agents \cite{yoshioka_7_2017}. The
@@
 may pose operational challenges as the salt can freeze and restrict flow,
-thereby causing a loss of flow accident \cite{ho_molten_2013}.
+causing a loss of flow accident \cite{ho_molten_2013}.
 
@@
 \glspl{MSR} have several positive sustainability features.
-\glspl{MSR} have good neutron economy due to the continuous removal of
-neutron poisons by the online reprocessing and gas sparging systems
+The continuous removal of neutron poisons by the online reprocessing and gas
+sparging systems creates good neutron economy in \glspl{MSR}
 \cite{kamei_recent_2012}. Both
@@
 associated with long-lived plutonium and \gls{MA} isotopes. The combination
-of \gls{TRU} fuel and $^{232}$Th feed in fast spectrum \glspl{MSR} would help
-reduce overall levels of \gls{TRU} waste going into long-term storage in
+of \gls{TRU} fuel and $^{232}$Th feed in fast spectrum \glspl{MSR} contributes
+to lower levels of \gls{TRU} waste going into long-term storage in
 nuclear waste repositories \cite{merle-lucotte_launching_2011}.
@@
 by-product provides some level of proliferation resistance but nuclear
-proliferators can sidestep this complication by
-separating the $^{232}$U away from the combined $^{232}$Pa/$^{233}$Pa stream
+proliferators can sidestep this complication; they can
+separate the $^{232}$U away from the combined $^{232}$Pa/$^{233}$Pa stream
 after most of the $^{232}$Pa has decayed into $^{232}$U at a short half-life
@@
 stage of \gls{MSR} development. Qualitatively, many technical factors favor
-\glspl{MSR} over \gls{LWR}. Some of these factors include: smaller reactor
+\glspl{MSR} over \glspl{LWR}. Some of these factors include: smaller reactor
 core size due to low operating pressures, higher thermal efficiency and
@@
 electricity costs for this type of \gls{MSR} was 7\% and 8\% lower than that
-of pressurized water reactor and coal power plants. These differences are
-within the $\pm$10\% uncertainty range but there is room for some optimism.
+of pressurized water reactors and coal power plants. These differences are
+within the $\pm$10\% uncertainty range but Moir remains optimistic.
 The author also notes that costs could be lower in an \gls{MSR} with
@@
 In terms of technological readiness, \glspl{MSR} still require significant
-R\&D efforts and engineering demonstrations for experimental validation of
+research efforts and engineering demonstrations for experimental validation of
 various components before a full commercial model can be commissioned. Work
@@
 neutron diffusion and \gls{RANS} formulations on ultra fine meshes, both
-models showed good agreement in the steady-state neutron flux, temperature,
-and \gls{DNP} distributions, and in the power responses following various
+models agreed on the steady-state neutron flux, temperature,
+and \gls{DNP} distributions, as well as the power responses following various
 accident transient initiations. Aufiero et al. \cite{aufiero_development_2014}
@@
 single pump in the sixteen-pump \gls{MSFR} configuration. The authors also
-provided some quantitative data supporting the use of implicit coupling over
+provided quantitative data supporting the use of implicit coupling over
 explicit coupling to obtain accurate solutions of the transient cases.
 Recognizing the huge computational burden required for full 3-D simulations, 
-later authors came up with innovative ways to alleviate this issue such as
+later authors came up with innovative ways to alleviate this issue. Methods
+include
 selective geometrical reduced order modeling for various components of a
@@
 breeding ratios deteriorated significantly due to parasitic absorption in the
-large volume of graphite needed for thermalization
-\cite{mathieu_thorium_2006}.
+graphite \cite{mathieu_thorium_2006}.
 %
@@
 	\begin{tabular}{ l r }
-		\hline
+		\toprule
 		Parameter & Value \\
-		\hline
+		\midrule
 		Thermal/Electric output [MW$_{\text{th}}$/MW$_{\text{e}}$] & 3000 /
@@
 		Blanket volume [m$^3$] & 7.3\\
-		\hline
+		\bottomrule
 	\end{tabular}

### 2020-07-04 neutronics.tex
final edits for 2nd draft

@@
 six-group neutron diffusion model in Moltres on a fast-spectrum reactor. This
-exercise also serves as a verification for the group constant data
-generated for Moltres from Serpent. For clarity, here are the important
-details of the static model.
+verification exercise builds on a previous study by Lindsay et al.
+\cite{lindsay_introduction_2018} that had verified Moltres' neutronics
+capabilities with a two-group neutron diffusion model of the \gls{MSRE}.
 

### 2020-07-04 steadystate.tex
final edits for 2nd draft

@@
 With the verification of Moltres' neutronics modeling capabilities in the
-context of the \gls{MSFR}, we now move on to the multiphysics simulation
-results. This chapter covers the steady-state multiphysics simulation results
-from Moltres.
+context of the \gls{MSFR}, this chapter will cover the steady-state
+multiphysics simulation results from Moltres.
 
 The procedure for obtaining the steady-state results involved several steps
-due to the highly coupled \glspl{PDE}. First, we ran a preliminary transient
-simulation of fluid flow in the \gls{MSFR} core, starting from zero inlet
-velocity and gradually ramping it up to match the nominal flow rate (4.5 m$^3$
-s$^{-1}$); otherwise Moltres had difficulty converging to the desired fully
-developed flow profile. We imposed a parabolic flow profile at the inlet.
-Next, we imported these fully developed flow values as initial values for
-velocity in the actual transient simulation modeling the full coupled
-neutronics and thermal-hydraulics. The initial values for the temperature and
-neutron group flux distributions are 953 K and $1 \times 10^{14}$ cm$^{-2}$
-s$^{-1}$ uniformly throughout the geometry. Finally, we assume that steady
+due to the tightly coupled \glspl{PDE}. First, a preliminary transient
+simulation of fluid flow in the \gls{MSFR} core was run, starting from zero
+inlet velocity and gradually ramping up to match the nominal flow rate (4.5
+m$^3$ s$^{-1}$); otherwise Moltres had difficulty converging to the desired
+fully developed flow profile.
+Next, these fully developed flow values were imported as initial values for
+velocity in the full time-dependent simulation modeling the full coupled
+neutronics and thermal-hydraulics multiphysics model. The initial values for
+the temperature and neutron group flux distributions are 953 K and
+$1 \times 10^{14}$ cm$^{-2}$ s$^{-1}$ uniformly throughout the geometry.
+Finally, this work assumes that steady
 state is reached when the volume integral values of every variable remain
@@
 For a direct comparison with the steady-state results from the Polimi and
-TUDelft models \cite{aufiero_development_2014}, we will first present our
-steady-state results without modeling decay heat. After this comparison, we
-separately discuss the minor differences borne from decay heat modeling in the
-last subsection.
+TUDelft models \cite{aufiero_development_2014}, this chapter will first
+present the steady-state results without modeling decay heat. After this
+comparison, the chapter separately discusses the minor differences in the
+results from decay heat modeling in the last subsection.
 
@@
 through flow streamlines superimposed on the velocity magnitude distribution.
-The results from Moltres show good qualitative agreement with the
-Polimi and TUDelft models \cite{fiorina_modelling_2014}; we observe similar
+The results from Moltres show good qualitative agreement with the Polimi and
+TUDelft models \cite{fiorina_modelling_2014}; the plots show similar
 flow and hotspot features in all three models. Furthermore, the highest salt
 velocities in all three models occur at the inlet, outlet, and at core
-half-height approximately 0.40 m away from the central axis. There is a large
-recirculation region near the blanket tank walls arising from turbulent flow.
+half-height approximately 0.40 m away from the central axis. A large
+recirculation region forms near the blanket tank walls from turbulent flow.
 Inertial forces dominate over viscous forces to form this large eddy. The main
@@
 model ($\approx$ 1300 K) than the TUDelft model ($\approx$ 1200K). Similarly,
-we observe cooler temperatures in high-velocity regions. The minimum
+the plots show cooler temperatures in high-velocity regions. The minimum
 temperature is 924 K at the inlet.
@@
 relatively lower temperatures near the center of the core promotes the neutron
-flux peaking but it is not a major concern as there are no structural parts
-vulnerable to neutron damage in that region. The peak total flux at the center
-is $9.80 \times 10^{15}$ cm$^{-2}$ s$^{-1}$, which is close to values reported
-by Fiorina et al. \cite{fiorina_molten_2013} and Aufiero et al.
+flux peaking but it is of little concern as no structural parts vulnerable to
+neutron damage exist in that region. The peak total flux at the center is
+$9.80 \times 10^{15}$ cm$^{-2}\cdot$s$^{-1}$, which is close to values
+reported by Fiorina et al. \cite{fiorina_molten_2013} and Aufiero et al.
 \cite{aufiero_development_2014} as shown in Table \ref{table:peak-flux}. The
-peak flux value from this paper is slightly higher as we used the steady-state
-temperature distribution while Fiorina et al. and Aufiero et al. imposed a
-uniform temperature distribution at 973 K.
+peak flux value from this paper is slightly higher as this work used the
+steady-state temperature distribution while Fiorina et al. and Aufiero et al.
+imposed a uniform temperature distribution at 973 K.
 
@@
 Figure \ref{fig:dnp} shows the \gls{DNP} distributions for all eight \gls{DNP}
-groups. In general, we observe less \glspl{DNP} in the regions with fast salt
-flow, namely
- The
+groups. In general, the figures show less \glspl{DNP} in the regions with fast
+salt flow. The
 precursors from the shortest-lived group (Group 8) predominantly decay within
@@
 relatively evenly distributed due to their long half-lives. For the
-longer-lived groups, the \gls{DNP} concentrations are not well-resolved on the
+longer-lived groups, the \gls{DNP} concentrations are ill-resolved on the
 mesh elements adjacent to the outlet and the inlet boundaries, respectively.
-Thus, we recommend careful mesh refinement for future work involving similar
-geometries.
+Thus, the present author recommends careful mesh refinement for future work
+involving similar geometries.
 
@@
 
-In Figure \ref{fig:pre}, we compare the total delayed neutron source
+Figure \ref{fig:pre} compares the total delayed neutron source
 distribution from Moltres with the results from the Polimi and TUDelft models
@@
 the precursor distribution, Figure \ref{fig:pre} shows the rate of delayed
-neutron emission, which we calculated by multiplying each \gls{DNP} group
+neutron emission, which was calculated by multiplying each \gls{DNP} group
 $C_i$ with its associated decay constant $\lambda_i$.
-We observe that the Polimi and TUDelft models
+The Polimi and TUDelft models
 feature greater \gls{DNP} retention in the stagnant regions within the core.
-This effect is less pronounced in the Moltres model. There is less build-up
-of \glspl{DNP} at the top of the core in Moltres because most of the
+This effect is less pronounced in the Moltres model. Less build-up
+of \glspl{DNP} at the top of the core occurs in Moltres because most of the
 \glspl{DNP} produced near the center of the core cannot enter the axial
-recirculation zone that only appears in Moltres. We also observe less build-up
-in the large recirculation zone near the blanket tank.
+recirculation zone that only appears in Moltres. The figure also shows less
+build-up in the large recirculation zone near the blanket tank.
 
@@
 neutron fraction in \glspl{MSR} after accounting for the loss of delayed
-neutrons from \glspl{DNP} decaying outside the active core region. In general,
-reactors with smaller $\beta$ values exhibit greater prompt jumps in the
+neutrons from \glspl{DNP} decaying outside the active core region.
+Reactors with smaller $\beta$ values exhibit greater prompt jumps in the
 neutron flux in response to reactivity insertions because they have a greater
@@
 
-In Table \ref{table:dnf}, we compare the fraction of out-of-core emissions and
+Table \ref{table:dnf} compares the fraction of out-of-core emissions and
 the $\beta_c$ values from Moltres with the Polimi and TUDelft models. We
@@
 differs significantly by approximately 10\%, and $\beta_c$ differs by
-60-70 pcm. We attribute the former to the lesser \gls{DNP} retention in the
+60-70 pcm. The former is attributed to the lesser \gls{DNP} retention in the
 stagnant flow regions in the core; the \glspl{DNP} are more evenly distributed
 along the entire primary loop, leading to more delayed neutron emissions in
-the outer loop region. The most likely reason for this is differences in the
-flow pattern in the recirculation zone because convective species transport
+the outer loop region. This is likely caused by differences in the
+flow pattern in the recirculation zone because convective transport
 dominates diffusive effects in the \gls{MSFR}. The exact flow pattern in the
@@
 between Moltres, and the Polimi and TUDelft models. Although the sizes of the
-arrows representing flow velocity are not normalized to the same scales, a
+arrows representing flow velocity are on different scales, a
 quick comparison between the largest arrows and the arrows in the
-recirculation zone indicate that the recirculation zone in Moltres is
+recirculation zone indicates that the recirculation zone in Moltres is
 relatively more stagnant. This could explain the concentration of \glspl{DNP}
@@
 in Moltres than the Polimi and TUDelft models. To account for this
-peculiarity, we note that Fiorina et al. \cite{fiorina_modelling_2014}
-applied adjoint flux weighting for their $\beta_c$ calculation whereas we
-report our value as the physical fraction without adjoint weighting. The
-weighting results in a significant difference in $\beta_c$ because a large
+peculiarity, Fiorina et al. \cite{fiorina_modelling_2014} applied adjoint flux
+weighting for their $\beta_c$ calculation whereas this work
+reports the value as the physical fraction without adjoint weighting. The
+weighting results in a significant difference in $\beta_c$ because a greater
 fraction of the \glspl{DNP} decay in the recirculation zones, where the
@@
 the volumetric heat source from the center of the core to the entire loop.
-Thus, we expect to observe a slight flattening of the temperature distribution
+Thus, there should be a slight flattening of the temperature distribution
 across the entire primary loop.

### 2020-07-04 transient.tex
final edits for 2nd draft

@@
 instances of reactivity insertion, loss of heat sink, loss of flow, and pump
-overspeed accidents. The term ``unprotected'' means that there is no
-external intervention in these scenarios. As such, these simulations
+overspeed accidents. The term ``unprotected'' means no external interventions
+occur in these scenarios. As such, these simulations
 give an insight on the \gls{MSFR}'s passive safety capabilities in the absence
-of any active safety system. We used the steady-state configuration presented
-in the previous chapter as the initial conditions for the transient
-simulations discussed in this chapter. Specifically, we imported all
-steady-state spatial values for neutron flux, delayed neutron precursor
-concentration, temperature, velocity, and pressure as the initial state of the
-transient scenarios.
+of any active safety system. This work used the steady-state configuration
+presented in the previous chapter as the initial conditions for the transient
+simulations discussed in this chapter. Specifically, all steady-state spatial
+values for neutron flux, delayed neutron precursor concentration, temperature,
+velocity, and pressure were imported as the initial state of the transient
+scenarios.
 
@@
 Polimi and TUDelft model for the four accident scenarios without decay heat
-modeling; they enabled decay heat modeling for the loss of heat sink transient
-scenario only. For a direct comparison, we also ran all transient simulations
-without the decay heat model with the exception of the loss of heat sink
-scenario in which we ran two separate simulations with and without
-the decay heat model. More generally, we imposed various other simplifying
+modeling and enabled decay heat modeling for the loss of heat sink transient
+scenario only. This work also ran all transient simulations without the decay
+heat model for a fair comparison. The only exception is loss of heat sink
+scenario in which two separate simulations with and without the decay heat
+model were run. More generally, this work imposed various other simplifying
 assumptions in our transient models to match their implementations as closely
@@
 
-\glspl{RIA} are a type of nuclear accident caused by unintended positive
+Reactivity insertion accidents are a type of nuclear accident caused by
+unintended positive
 reactivity insertions. The excess reactivity causes the power output and
@@
 
-We modeled two unprotected step-wise reactivity insertion scenarios in Moltres
-by swapping out the original set of group constant data with two new, separate
-sets of data from Serpent corresponding to 50 pcm and 200 pcm reactivity
-insertions, respectively. We increased the reactivity of the Serpent
-\gls{MSFR} models by increasing the $^{233}$U-to-$^{232}$Th ratio in the fuel
-salt.
+This work modeled two unprotected step-wise reactivity insertion scenarios in
+Moltres by swapping out the original set of group constant data with two new,
+separate sets of data from Serpent corresponding to 50 pcm and 200 pcm
+reactivity insertions, respectively. The reactivity of the Serpent \gls{MSFR}
+models was increased by increasing the $^{233}$U-to-$^{232}$Th ratio in
+the fuel salt.
 
 The focus of this transient study is the neutronic and thermal-hydraulic
-behavior of the reactor core. Thus, we assume that the heat exchanger and the
-associated power generation equipment (generator turbines, heat sinks, etc.)
-can withstand all variations in the power output during the transients.
+behavior of the reactor core. Thus, this work assumes that the heat exchanger
+and the associated power generation equipment (generator turbines, heat sinks,
+etc.) can withstand all variations in the power output during the transients.
 
@@
 than the initial +50 pcm insertion. Power continues to fall as the average
-core temperature rises. There is a slight change in slope at $t=0.3$ s. This
-value approximately corresponds to the average half-life of the two
+core temperature rises. The slight change in slope occurs at $t=0.3$ s. The
+elapsed time approximately corresponds to the average half-life of the two
 shortest-lived delayed neutron precursor (DNP) groups ($t_{1/2}=0.195$ s and
@@
 and TUDelft models; Moltres reproduced all of the individual features in both
-plots. The most significant difference is in the magnitude of the reactor
-response. Moltres predicts a smaller peak in the power output and a smaller
-overall increase in the average core temperature. This is largely due to the
+plots. The magnitude of the reactor response is the most significant
+difference. Moltres predicts a smaller peak in the power output and a smaller
+overall increase in the average core temperature mainly due to the
 more negative temperature reactivity coefficient in Moltres than in the Polimi
@@
 approximately $-6.5$ pcm K$^{-1}$ within the relevant temperature range in the
-Polimi and TUDelft models. Therefore, we see a smaller temperature increase in
-the Moltres model for the same reactivity insertion. Multiplying the average
+Polimi and TUDelft models. Therefore, the results show a smaller temperature
+increase in the Moltres model for the same reactivity insertion. Multiplying
+the average
 core temperature increase at $t=10$ s with $\alpha_T$ gives us $-7.184$ pcm
@@
 The results for the 200 pcm reactivity insertion scenario show similar trends
-to the 50 pcm case. The larger reactivity insertion illicits a stronger
+to the 50 pcm case. The greater reactivity insertion elicits a stronger
 prompt response in the power output which peaks at 92.1 GW. The average core
@@
 from the differences in the temperature reactivity coefficients. Overall,
-Moltres' results show good agreement with the Polimi and TUDelft results and
-the differences arise mainly due to the differences in the temperature
+Moltres' results show good agreement with the Polimi and TUDelft results.
+The differences arise mainly due to the differences in the temperature
 reactivity coefficients.
@@
 intermediate loop fail. The heat exchangers would then lose most of their
-cooling capabilities. We followed Fiorina et al.'s approach in assuming that
-the cooling from the heat exchangers decreases exponentially with a time
+cooling capabilities. This work followed Fiorina et al.'s approach in assuming
+that the cooling from the heat exchangers decreases exponentially with a time
 constant of 1 s and all other parameters held constant
@@
 The results from Moltres show good agreement with the results from the Polimi
-and TUDelft. Moltres reproduced all of the trends that we see in the Polimi
+and TUDelft. Moltres reproduced all of the trends in the Polimi
 and TUDelft models. The temporary halt in the temperature increase occurs at
@@
 is likely due to the difference in the temperature reactivity coefficient
-that we discussed in the reactivity insertion results; a smaller increase in
+discussed in the reactivity insertion results; a smaller increase in
 the average core temperature produces the same decrease in power output.
@@
 Fiorina et al. \cite{fiorina_modelling_2014} applied the Boussinesq
-approximation for bouyancy-driven flow in their models, but we were unable to
-follow this procedure because we partitioned the
-primary loop into two separate geometries and used Dirichlet boundary
+approximation for buoyancy-driven flow in their models, but this approach was
+not possible in Moltres because the primary loop is partitioned into two
+separate geometries and used Dirichlet boundary
 conditions at the inlet to drive flow. Fiorina et al.'s Polimi and TUDelft
 models featured complete exponential coast-downs of the pumps with a time
-constant of 5 s. The resulting flow rate $\dot{m}$ from natural circulation was
-approximately 18 times smaller than the initial $\dot{m}$. Figure
+constant of 5 s. The resulting flow rate $\dot{m}$ from natural circulation
+was approximately 18 times smaller than the initial $\dot{m}$. Figure
 \ref{fig:flowrate} shows that the actual $\dot{m}$ decreased with a time
-constant of 8 s. Thus, for the \gls{MSFR} model in Moltres, we imposed a
-similar exponential decay term with a time constant of 8 s on the inflow
+constant of 8 s. Thus, for the \gls{MSFR} model in Moltres, this work imposed
+a similar exponential decay term with a time constant of 8 s on the inflow
 Dirichlet boundary condition:
@@
 the pointwise heat exchanger implementation in Moltres performs differently
-compared to the heat exchangers that take up ``36\% of the out-of-core part''
-in the Polimi and TUDelft models \cite{fiorina_modelling_2014}. Most of the
-cooling happens in the top half of the heat exchanger where the temperature
-differential between the primary and intermediate loops is the largest. In the
-Polimi and TUDelft models, the overall $h$ is the
+compared with the heat exchangers that take up ``36\% of the out-of-core
+part'' in the Polimi and TUDelft models \cite{fiorina_modelling_2014}. Most of
+the cooling happens in the top half of the heat exchanger where the
+temperature differential between the primary and intermediate loops is the
+largest. In the Polimi and TUDelft models, the overall $h$ is the
 ``harmonic mean of the heat transfer coefficients on each side of the heat
@@
 However, this relation provided very different results in the unprotected loss
-of flow and pump overspeed transients compared to the Polimi and TUDelft
+of flow and pump overspeed transients compared with the Polimi and TUDelft
 models; it overpredicted the equilibrium power output in the
 unprotected loss of flow transient and underpredicted the same parameter in
-the unprotected pump overspeed transient. Upon further investigation, we found
-that raising the power of $v$ from 0.8 to 1.1 brought the average core
-temperatures closer to the results from the other models in both transients.
-Therefore, we adopted the raised power in this study.
+the unprotected pump overspeed transient. Upon further investigation, the
+present author found that raising the power of $v$ from 0.8 to 1.1 brought the
+average core temperatures closer to the results from the other models in both
+transients. Therefore, this work adopted the raised power in this study.
 
 Another issue pertained to the turbulent viscosity $\mu_t$ as a function of
-$v$. When we made a simple approximation of $\mu_t$ being directly
-proportional to $v$, the results differed significantly compared to the Polimi
-and TUDelft models (Appendix *). This is likely due to buoyancy-driven flow
+$v$. Using a simple approximation of $\mu_t$ being directly
+proportional to $v$, the results differed significantly compared with the
+Polimi and TUDelft models. This is likely due to buoyancy-driven flow
 contributing to turbulence; the turbulent energy $k$ equation in COMSOL's
@@
 \cite{comsol_ab_comsol_2018}. Another point to note is the
-Reynolds number does not change if $\mu$ and $v$ decrease in tandem. This
+Reynolds number remains constant if $\mu$ and $v$ decrease in tandem. This
 preserves the existence of the recirculation zone in the core and it is at
@@
 stated that after around $t=15$ s, the ``flow pattern changed in the core and
-the recirculation zones started to disappear''. As a result, there is a sudden
-drop in the average core temperature as the pocket of hot salt leaves the
+the recirculation zones started to disappear''. A sudden drop in the average
+core temperature results as the pocket of hot salt leaves the
 core. In Moltres, the wider peak in the average core temperature
@@
 in the transition from high-speed turbulent flow to low-speed viscous
-flow as Moltres did not capture the intermediate stages accurately. The
-simplifying assumption for the uniform, time-dependent $\mu_t$ is also not
-ideal for a safety analysis code.
+flow as Moltres mispredicted the intermediate stages. The
+simplifying assumption for the uniform, time-dependent $\mu_t$ is also flawed
+in a safety analysis code.
 
@@
 
-We followed Fiorina et al.'s implementation \cite{fiorina_modelling_2014} by
+This workfollowed Fiorina et al.'s implementation
+\cite{fiorina_modelling_2014} by
 ramping up the inlet velocity, $u$, by 50\% from the nominal value, $u_0$,
@@
 %
-For this transient, we assumed that $\mu_t$ was directly proportional to $v$
-because the buoyancy effects are negligible and the recirculation zones 
+For this transient, this work assumed that $\mu_t$ was directly proportional
+to $v$ because the buoyancy effects are negligible and the recirculation zones 
 persist throughout the entire duration. Figures \ref{fig:poheat} and
@@
 Doppler effect and salt expansion causes the power output to stall and dip
-briefly before starting to rise again at $t=2.5$ s. The reactor tends to a new
+briefly before rising again at $t=2.5$ s. The reactor tends to a new
 equilibrium power output and average core temperature. The temperature

### 2020-07-09 steadystate.tex
decay heat sections

@@
 
-\section{Decay Heat}
+\clearpage
 
-The inclusion of a decay heat model effectively redistributes a fraction of
-the volumetric heat source from the center of the core to the entire loop.
-Thus, there should be a slight flattening of the temperature distribution
-across the entire primary loop.
+\section{Steady-State Decay Heat Results}
+
+The movement of decay heat precursors effectively redistributes a fraction of
+the volumetric heat source concentrated at the center of the core to the
+entire primary loop.
+Thus, the temperature distribution in the core should
+show less extreme temperatures compared with the simulation without decay
+heat.
+
+Figure \ref{fig:decayheat} shows the decay heat source distribution in the
+core for groups 1 to 3. The decay constants are $\lambda_j = 0.197$ s$^{-1},
+0.0168$ s$^{-1}, 0.000358$ s$^{-1}$ for groups 1, 2, and 3, respectively.
+Similar to the delayed neutron precursor distributions, the shorter-lived
+decay heat precursors show greater build-up in the recirculation zones
+compared with the longer-lived decay heat precursors that are more evenly
+distributed throughout the core.
+
+\begin{figure}[htbp!]
+    \centering
+    \begin{subfigure}[t]{.325\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{heat1}
+    \end{subfigure}
+    \begin{subfigure}[t]{.325\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{heat2}
+    \end{subfigure}
+    \begin{subfigure}[t]{.325\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{heat3}
+    \end{subfigure}
+    \caption{Decay heat source distribution in the core for decay heat groups
+    1 to 3 (from left to right). Refer to Figure \ref{fig:decayheattemp} for
+    the height and radius scales on the y and x axes, respectively. Note the
+    different scales for each distribution.}
+    \label{fig:decayheat}
+\end{figure}
+
+Figure \ref{fig:decayheattemp} shows the
+difference in core temperatures at steady state with decay heat modeling
+relative to the core temperatures without decay heat in Figure
+\ref{fig:flow-temp}. This figure uses a different colormap scheme to clearly
+distinguish between temperature increases and decreases in the core. The
+hotspots in the recirculation zones are approximately
+2 to 4 K cooler, while the cooled salt flowing in from the inlet is
+approximately 2 K hotter. These regions correspond to the hottest and coldest
+areas in the core and the results agree with the expectation that the extreme
+temperatures would be affected the most by the introduction of decay heat
+precursors. The difference between the outlet and inlet temperatures falls
+from 101.1 K to 98.9 K as some of the decay heat precursors deposit heat in
+the outer loop region. The figure also shows some unphysical oscillations in
+the temperature near the outlet because the mesh near the boundary is too
+coarse. Users can avoid these oscillations near the outlet but using a mesh
+that is progressively finer the closer the element is to the boundary. The
+present author did not resolve this issue due to time constraints.
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.6\textwidth]{decay-heat-temp}
+    \caption{Difference in core temperatures at steady state with decay heat
+    relative to the result without decay heat (Figure \ref{fig:flow-temp}).}
+    \label{fig:decayheattemp}
+\end{figure}

### 2020-07-09 transient.tex
decay heat sections

@@
 
-\subsection{Without Decay Heat}
+\subsection{Without Decay Heat} \label{sec:wodecayheat}
 
@@
 \subsection{With Decay Heat}
+
+Decay heat from fission products poses a great safety risk in a loss of
+heat sink accident. Section \ref{sec:wodecayheat} showed that prompt fission
+power output quickly falls as core temperatures rise. However, decay power
+output is independent of the instantaneous neutron flux. Figure
+\ref{fig:moltresdecaypower} shows that the decay power output remains
+relatively high during a short-term transient. Decay heat becomes the dominant
+heat source from $t=34$ s and falls at a much slower rate than prompt heat.
+Figure \ref{fig:moltresdecaytemp} highlights the
+greater core temperature increase attributed to decay heat as compared with
+the results without decay heat. By $t=120$ s, the model with decay heat
+records an average core temperature increase that is 45 K higher than the
+model without decay heat. The average core temperature reaches approximately
+1220 K and would continue to rise further. This places undue thermal stress
+and accelerate corrosion rates in the Hastelloy structural material. In the
+absence of an auxiliary heat removal system in the primary loop, reactor
+operators would have to rely on the freeze plug to drain the core into a drain
+tank with emergency cooling systems to keep the salt cool.
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.85\textwidth]{moltres-decay-power}
+    \caption{Power output during
+    an unprotected loss of heat sink transient in the Moltres model with and
+    without decay heat.}
+    \label{fig:moltresdecaypower}
+    \includegraphics[width=.85\textwidth]{moltres-decay-temp}
+    \caption{Average core temperature increase during
+    an unprotected loss of heat sink transient in the Moltres model with and
+    without decay heat.}
+    \label{fig:moltresdecaytemp}
+\end{figure}
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.85\textwidth]{decay-temp}
+    \caption{Loop-averaged temperature increase during
+    an unprotected loss of heat sink transient in the Moltres and Polimi
+    models \cite{fiorina_modelling_2014} with decay heat.}
+    \label{fig:polimidecaytemp}
+\end{figure}
+
+Figure \ref{fig:polimidecaytemp} shows the loop-averaged temperature
+increase in the Moltres and Polimi models \cite{fiorina_modelling_2014}. The
+TUDelft model does not have a decay heat modeling capability. The Moltres
+model predicts the same increasing trend in the temperature. The loop-averaged
+temperature rises significantly at the start of the transient and continues to
+rise at a decreasing rate. The loop-averaged temperature increase in the
+Moltres model at $t=120$ s is approximately 17 K lower than that in the Polimi
+model. It is difficult to ascertain the exact cause for this difference
+without the power profile from the Polimi model with decay heat to compare
+with. However, if the decay power output are similar, the stronger negative
+temperature reactivity coefficient would cause the prompt power output in the
+Moltres model to fall faster than the Polimi model. Consequently, the
+loop-averaged temperature would be lower as shown in the figure. Overall, the
+results for the loss of heat sink transient agree with the Polimi and TUDelft
+model results in both cases, with and without decay heat modeling.
 

### 2020-07-09 abstract.tex
Katy's second review edits

@@
-Molten salt reactors are a class of advanced nuclear reactors that promise
+Molten salt reactors is a class of advanced nuclear reactors that promise
 numerous improvements over the current fleet of largely light-water reactors.
 As the world continues its transition towards low-carbon electricity
-generation to combat climate change, molten salt reactors are potential
+generation to combat climate change, molten salt reactors is potential
 options in the near-term future for replacing fossil fuel and ageing nuclear
 power plants. At the current state of development, molten salt reactors still
-require extensive research to become viable. This thesis presents latest
+require extensive research to become viable. This thesis presents the latest
 developments in Moltres, a simulation tool for molten salt reactors. These new
 developments are: support for coupling the incompressible Navier-Stokes and
-the delayed neutron precursor looping systems, and a new decay heat model for
+the delayed neutron precursor looping systems, and a model for
 simulating decay heat from fission products at steady state and during
-transients. This work demonstrates these capabilities in conjunction with
-existing capabilities through multiphysics simulations of the Molten Salt Fast
+transients. This work demonstrates these capabilities through multiphysics
+simulations of the Molten Salt Fast
 Reactor concept. This work first verifies the six-group neutron diffusion
@@
 fidelity simulation results from Serpent 2. The $k_{\text{eff}}$ values have
-small discrepancies on the order of 100 pcm, which is smaller than some
-reported literature values. The decay heat model shows an expected flattening
+small discrepancies on the order of 100 pcm, which is smaller than the
+$-256.7$ pcm discrepancy reported in literature with the same six-group
+neutron diffusion approach. The decay heat model shows an expected flattening
 of the temperature distribution due to the movement and dispersion of the
@@
 transient multiphysics simulations of the Molten Salt Fast Reactor. The
-transient scenarios under study are unprotected instances of reactivity
+transient scenarios under study were unprotected instances of reactivity
 insertion, loss of heat sink, loss of flow, and pump overspeed. The steady
-state and transient results are verified against results from another paper
+state and transient results were verified against results from another paper
 that presented results for the same cases. The steady-state temperature and

### 2020-07-09 introduction.tex
Katy's second review edits

@@
 Greenhouse gas emission from human activities is the main cause of climate
-change, a global phenomenon characterized by shifting weather patterns and
-rising temperatures. The concentration of CO$_2$, a greenhouse gas, in the
-atmosphere has been rising steadily throughout modern human history
-\cite{us_department_of_commerce_global_2020}. Climate change has dire
+change which has dire
 consequences on human health and safety due to extreme weather events and the
-overall impact on food production \cite{mcmichael_global_2004}. As a result,
-urgent measures are necessary to limit greenhouse gas emissions, namely CO$_2$
-emissions from electricity generation.
-
+overall impact on food production \cite{mcmichael_global_2004}.
 Electricity generation from burning fossil fuels represents the
@@
 alternatives would curb a significant fraction of emissions. Nuclear power is
-a viable low-carbon replacement for burning fossil fuels.
-Other low-carbon energy sources include solar-, wind-, and hydro-power. Each
-energy source has its own set of advantages and disadvantages; nuclear power's
-advantage over renewable energy sources is that it provides consistent
-base-load power independent of weather and geographical location
-\cite{petti_future_2018}. Employing a diverse mix of these energy sources
-ensures energy security and reliability in our transition
-towards a low-carbon future \cite{petti_future_2018}.
+a viable low-carbon replacement for burning fossil fuels and it provides
+consistent base-load power independent of weather and geographical location
+\cite{petti_future_2018}. Furthermore, employing a diverse mix of nuclear
+power and renewable sources ensures energy security and reliability in our
+transition towards a low-carbon future \cite{petti_future_2018}.
 
@@
 interactions between neutronics and thermal-hydraulics are stronger due
-to greater fuel material expansion. Furthermore, \glspl{DNP} in \glspl{MSR}
-can flow freely within the primary coolant
+to greater fuel material expansion. Furthermore, fissile material and
+\glspl{DNP} in \glspl{MSR} can flow freely within the primary coolant
 loop as opposed to being held in place in a solid fuel matrix. Therefore,
@@
 Most reactor analysis applications are usually reactor-specific by
-design such as PARCS and TRACE for \glspl{LWR}, and SAS4A/SASSYS-1 for
+design such as TRACE for \glspl{LWR}, and SAS4A/SASSYS-1 for
 liquid metal cooled reactors. Thus, these applications would disregard
@@
 agreement with the original design calculations by \gls{MSRE} researchers at
-\gls{ORNL}. There has since been new developments in Moltres allowing for
-more accurate \gls{MSR} simulations.
+\gls{ORNL}. This thesis presents some of the new developments in Moltres
+allowing for more complex and accurate \gls{MSR} simulations.
 
@@
 This thesis demonstrates latest capabilities of Moltres
-\cite{lindsay_introduction_2018}, an open source MSR simulation tool.
+\cite{lindsay_introduction_2018}.
 In particular, this thesis presents two more recent
@@
 results. This thesis covers the \gls{MSFR} concept mainly because it has been
-studied extensively with readily available literature data to verify our
-results against. The
-steady-state and transient results also feature a significant degree of
-physics interaction between the neutronics and thermal-hydraulics. This
+studied extensively with readily available data in the literature to verify
+our results against. The \gls{MSFR} design also features interesting flow
+patterns that greatly affect the steady-state and transient behavior. This
 present work will first present a verification of Moltres' \gls{MSFR}
-neutronics results against Serpent 2, followed by a verification of
-the coupled neutronics/thermal-hydraulics steady-state and transient accident
+diffusion neutronics against the Monte Carlo neutron tranport software
+Serpent 2, followed by a verification of
+the coupled neutronics/thermal-hydraulics steady-state and accident transient
 results against results from the Polimi and TUDelft models published by
@@
 presents steady-state results of coupled neutronics/thermal-hydraulics
-\gls{MSFR} simulations in Moltres. Chapter 6 presents transient accident
+\gls{MSFR} simulations in Moltres. Chapter 6 presents accident transient
 simulation results for unprotected reactivity insertions, unprotected loss of
-heat sink, unprotected loss of flow, and unprotected pump overspeed,
-respectively. Lastly, Chapter 7 summarizes the key findings in this thesis
+heat sink, unprotected loss of flow, and unprotected pump overspeed. Lastly,
+Chapter 7 summarizes the key findings in this thesis
 and posits some potential avenues for future work.

### 2020-07-09 model.tex
Katy's second review edits

@@
 multiphysics simulation tool for \glspl{MSR} \cite{lindsay_introduction_2018}.
-To run simulations on Moltres, it requires input
+In particular, this work introduces two new capabilities: full support
+for coupling incompressible flow with the existing delayed neutron precursor
+looping capability, and a decay heat model. The former allows users to
+simulate non-trivial flow patterns in the core and simultaneously loop the
+precursors through an external region, and the latter to simulate delayed
+heating from fission products.
+To run simulations with Moltres, the user must provide input
 group constant data from a neutron transport solver for the
@@
 geometry of the reactor. This work uses Serpent 2 \cite{leppanen_serpent_2014}
-for the former and Trelis \cite{noauthor_trelis_2018} for the latter.
+for the former and Trelis/Cubit \cite{noauthor_trelis_2018} for the latter.
 This chapter provides brief introductions to
 Serpent 2, \gls{MOOSE}, and Moltres, and the modeling approach for the
-multiphysics simulations in Moltres.
+\gls{MSFR} multiphysics simulations in this thesis.
 
@@
 since grown to support more general capabilities. Serpent 2 is highly
-parallelizable, supporting
-both MPI and OpenMP parallel programming APIs. It has also been validated and
-verified against experimental data and other well-established codes.
+parallelizable, supporting both MPI and OpenMP parallel programming APIs. It
+has also been validated and verified against experimental data and other well
+established codes \cite{leppanen_calculation_2014}.
 
@@
 
-In the context of this work, Serpent 2 uses the JEFF-3.1.2 nuclear
-data library \cite{oecd/nea_jeff-3.1.2_2014} to generate group constants
-needed by Moltres. The relevant group constant data are collapsed into six
-neutron energy groups, and calculated at discrete temperature values from 800
-K to 1300 K at 100 K intervals. Table \ref{table:bound} shows the upper bounds
-of each neutron energy group. The group constants relevant for neutronics
-calculations in Moltres are the
-macroscopic fission $\Sigma^f_{g}$, removal $\Sigma^r_{g}$, and scattering
-$\Sigma^s_{g' \rightarrow g}$ neutron cross sections,
-neutron diffusion coefficients $D_g$, average fission energies $\epsilon_g$,
-average neutron yields $\nu$, inverse neutron speeds $1/v$, flux spectra
-$\chi$, \gls{DNP} decay constants $\lambda_i$, and effective delayed neutron
-fractions $\beta_{eff}$. These group constants are extracted from
-the Serpent 2 output files using a Python script available from the Github
-repository that holds the Moltres source code \cite{lindsay_moltres_2017}. The
-script rewrites the group constants into a Moltres-compatible format.
-
 Serpent 2 provides standard geometric surfaces (e.g. planes, cylinders, cones)
@@
 \cite{fiorina_modelling_2014}, as shown in Fig. \ref{fig:msfrgeom}.
-
-\begin{table}[htb!]
-	\centering
-	\caption{Neutron energy group upper bounds used in Serpent 2.}
-	\begin{tabular}{c S}
-		\toprule
-		{Group number} & {Upper bound [MeV]}\\
-		\midrule
-		1 & 20\\
-		2 & 2.2313\\
-		3 & 0.4979\\
-		4 & 0.0247875\\
-		5 & 0.0055308\\
-		6 & 0.0007485\\
-		\bottomrule
-	\end{tabular}
-	\label{table:bound}
-\end{table}
 
@@
 terms in the \glspl{PDE}, the initial and boundary conditions, the material
-properties, etc., are represented in \gls{MOOSE} as C++ objects. New objects
-can inherit properties from related old objects to simplify implementation and
-reduce code duplication. These objects are commonly referred to as kernels in
-\gls{MOOSE}. Overall, this approach
-is helpful for many researchers, a significant fraction of whom do not possess
-high-level programming backgrounds, as they are unencumbered by the
-technical details and complexities involved in programing mesh handling
-and \gls{PDE} solving in finite element analysis.
+properties, etc., are represented in \gls{MOOSE} as C++ objects. Child objects
+can inherit properties from parent objects to simplify implementation and
+reduce code duplication. Overall, this approach
+is helpful for many researchers as they are unencumbered by the
+technical details and complexities involved in programing mesh-handling
+and \gls{PDE}-solving in finite element analysis.
 
 \gls{MOOSE} itself relies on libMesh \cite{kirk_libmesh:_2006} and
-PETSc \cite{satish_petsc_2019} for its mesh handling and \gls{PDE} solver
+PETSc \cite{satish_petsc_2019} for mesh handling and \gls{PDE} solver
 functionalities. As a result, \gls{MOOSE} supports adaptive meshing schemes
-and automatic variable scaling amongst other advanced features for improved
-accuracy and performance times. Full
+and automatic variable scaling amongst other advanced featuresin finite
+element analysis. Full
 coupling is maintained by the execution of Newton-based solves on the
@@
 Fully-coupled solves are essential for accurately resolving systems with
-strongly interacting physics. The \gls{MSR} concept is one such example, where
-the neutronics and thermal-hydraulics are tightly coupled through the Doppler
-effect and the temperature dependence of liquid fuel salt density.
+strongly interacting physics. The \gls{MSR} concept is one such example, in
+which the neutronics and thermal-hydraulics are tightly coupled through the
+Doppler effect and the temperature dependence of liquid fuel salt density.
 
 \gls{MOOSE}, and Moltres by extension, are capable of up to 3-D geometry
-modelling. They support a wide range of input mesh file formats listed in a
-\gls{MOOSE} webpage, including the commonly used Exodus II file format.
-Specifically for the 2-D case, axial symmetry is easily and automatically
-imposed by changing one line of code in the input file, without any changes in
+modelling. They support a wide range of input mesh file formats, including the
+commonly used Exodus II file format. Specifically for 2-D geometries, users
+can easily switch between Cartesian and polar coordinates by changing one line
+of code in the input file, without any changes in
 the Cartesian representations of the \glspl{PDE} and boundary conditions in
@@
 Another important feature for reducing computational time is the use of MPI
-for parallel computing. All \gls{MOOSE}-based codes can be compiled and
+for parallel computing. All \gls{MOOSE}-based applications can be compiled and
 run on high performance computing clusters.
 
-\gls{MOOSE} includes a set of built-in physics modules for commonly studied
-phenomena such as the heat conduction, flow dynamics and solid mechanics. This
+\gls{MOOSE} includes a set of built-in physics modules such as the Heat
+Conduction, Navier-Stokes, and Solid Mechanics modules for commonly studied
+physical phenomena. This
 work uses \gls{MOOSE}'s Navier-Stokes module for simulating
@@
 2-D axisymmetric model of the \gls{MSFR} with uniform salt flow
-\cite{park_safety_2019}.
+\cite{park_safety_2019}. The present work also includes these capabilities
+which are discussed in the following section on the modeling approach.
 
 Building on the prior progress, this thesis presents two more recent
-developments in Moltres, namely the new kernels required to couple
+developments in Moltres, namely the new features required to couple
 incompressible flow with the delayed neutron precursor looping capability, and
@@
 featured in the previous papers \cite{lindsay_introduction_2018,
-park_safety_2019}.
+park_safety_2019}. The next section describes these new developments in
+detail.
 
@@
 
-This section discusses the neutronics and thermal-hydraulics
-\glspl{PDE} that Moltres solves and other relevant details specific to the
-\gls{MSFR} model in this work.
+This section discusses the group constant generation in Serpent 2, the
+neutronics and thermal-hydraulics \glspl{PDE} that Moltres solves, and the
+relevant procedures specific to the \gls{MSFR} model in this work.
 
-As mentioned in the previous chapter, the fuel salt loop is divided into two
-regions, the central core region where most of the fissions take place, and
-the outer loop region where the heat exchanger is located. The central core
-region specifically refers to the central region indicated by
+\subsection{Group Constant Generation}
+
+The current work uses the JEFF-3.1.2 nuclear data library
+\cite{oecd/nea_jeff-3.1.2_2014} with Serpent 2 to generate group constants
+needed by Moltres. The relevant group constant data are collapsed into six
+neutron energy groups, and calculated at discrete temperature values from 800
+K to 1300 K at 100 K intervals. Table \ref{table:bound} shows the upper bounds
+of each neutron energy group. The group constants relevant for neutronics
+calculations in Moltres are:
+%
+\begin{align*}
+    &\Sigma^f_{g} \text{: macroscopic fission cross section in group $g$,} \\
+    &\Sigma^r_{g} \text{: macroscopic removal cross section in group $g$,} \\
+    &\Sigma^s_{g' \rightarrow g} \text{: macroscopic scattering cross section
+    from group $g$' to $g$,} \\
+    &D_g \text{: diffusion coefficient of neutrons in group $g$,} \\
+    &\epsilon_g \text{: average fission energy per fission by a neutron from
+    group $g$,} \\
+    &\nu \text{: average neutron yield per fission by a neutron from group
+    $g$,} \\
+    &\frac{1}{v} \text{: inverse neutron speed in group $g$,} \\
+    &\lambda_i \text{: decay constant of \gls{DNP} group $i$,} \\
+    &\beta_{eff} \text{: effective delayed neutron fraction.} \\
+\end{align*}
+%
+These group constants are extracted from
+the Serpent 2 output files using a Python script available from the Github
+repository that holds the Moltres source code \cite{lindsay_moltres_2017}. The
+script rewrites the group constants into a Moltres-compatible format.
+
+\begin{table}[htb!]
+	\centering
+	\caption{Neutron energy group upper bounds used in Serpent 2.}
+	\begin{tabular}{c S}
+		\toprule
+		{Group number} & {Upper bound [MeV]}\\
+		\midrule
+		1 & 20\\
+		2 & 2.2313\\
+		3 & 0.4979\\
+		4 & 0.0247875\\
+		5 & 0.0055308\\
+		6 & 0.0007485\\
+		\bottomrule
+	\end{tabular}
+	\label{table:bound}
+\end{table}
+
+\subsection{Central Core Region}
+
+As mentioned in the Chapter \ref{chap:msr}, the fuel salt loop is divided into
+two regions, the central core region where most of the fissions take place,
+and the outer loop region where the heat exchanger is located. The central
+core region specifically refers to the central region indicated by
 the red box in Figure \ref{fig:core}. The outer loop is simplified
@@
 \end{figure}
-
-\subsection{Central Core Region}
 
@@
 
-While the limitations of the multigroup neutron diffusion compared to other
-deterministic and Monte Carlo methods, particularly for flux values near
+While the limitations of the multigroup neutron diffusion method compared to
+other deterministic and Monte Carlo methods, particularly for flux values near
 boundaries, are well-documented, the diffusion model provides acceptable
@@
 This work assumes vacuum boundary conditions for all six neutron group fluxes
-along
-the external boundaries of the geometry, and homogeneous Neumann boundary
-conditions along the axial symmetry boundary. For the \gls{DNP}
-concentrations, we imposed homogeneous Neumann boundary conditions on the
-walls, and inflow and outflow boundary conditions on the inlet and outlet
-boundaries respectively. The inlet \gls{DNP} concentration values were
+along the external boundaries of the geometry, and homogeneous Neumann
+boundary conditions along the axial symmetry boundary. For the \gls{DNP}
+concentrations, this work imposed homogeneous Neumann boundary conditions on
+the walls, and inflow and outflow boundary conditions on the inlet and outlet
+boundaries, respectively. The inlet \gls{DNP} concentration values were
 imported from the outlet values of the 1-D outer loop pipe at the same
-timestep.
+timestep. Table \ref{table:corebc} describes these boundary conditions
+mathematically.
 
@@
     precursors in group $j$ [W$\cdot$cm$^{-3}$],} \nonumber \\
-	f_j &= \text{fraction of decay heat from group $j$ to total power under
-	full-power conditions,}
-	\nonumber \\
-	\epsilon_g &= \text{average fission energy per fission [W],} \nonumber \\
+	f_j &= \text{fraction of total power attributable to decay heat group
+	$j$,} \nonumber \\
+	\epsilon_g &= \text{average fission energy per fission initiated by a
+	neutron in group $g$ [W],} \nonumber \\
 	\lambda_j &= \text{average decay constant of decay heat precursors in
@@
 
-Like the neutron and \gls{DNP} groups, Moltres can accommodate an arbitrary
-number of decay heat groups. This work uses the same decay heat fractions and decay constants, shown in Table
-\ref{eq:decayheat}, used in the Polimi and TUDelft models for three decay heat
-groups.
+Like the neutron energy and \gls{DNP} groups, Moltres can accommodate an
+arbitrary number of decay heat groups. The current work uses the same decay
+heat fractions and decay constants, shown in Table \ref{eq:decayheat}, used in
+the Polimi and TUDelft models for three decay heat groups.
 
@@
     -\rho (\vec{u}
-    \cdot \nabla) \vec{u} + \nabla \cdot (-p \vec{I} + \mu [
-    \nabla \vec{u} + (\nabla \vec{u})^T] + \vec{f} &&
+    \cdot \nabla) \vec{u} + \nabla \cdot [-p \vec{I} + \mu [
+    \nabla \vec{u} + (\nabla \vec{u})^T]] + \vec{f} &&
     \label{eq:momemtum} \\
@@
     && p &= \text{ pressure [Pa],} && \nonumber \\
-    && \mu &= \text{ dynamic viscosity [Pa s],} && \nonumber \\
+    && \mu &= \text{ dynamic viscosity [Pa$\cdot$s],} && \nonumber \\
     && \vec{f} &= \text{ body force per unit volume [N$\cdot$m$^{-3}$].} &&
@@
 
-In addition to the intrinsic molecular viscosity, this thesis introduces an
+In addition to the intrinsic molecular viscosity in the \gls{INS} equations,
+this thesis introduces an
 eddy viscosity term to approximate turbulent flow effects. The current
@@
 reported in the Polimi/TUDelft models. The eddy viscosity is assumed to be 40
-Pa s. Despite the simplicity of this assumption, the resulting flow profile is
+Pa$\cdot$s. Despite the simplicity of this assumption, the resulting flow profile is
 similar to flow profile in the Polimi and TUDelft models at steady state.
@@
 on all of the relevant boundaries. Figure \ref{fig:msfrbc} shows the locations
-of the various boundaries listed in the table.
+of the various boundaries listed in the table. This thesis introduces the
+CoupledOutflow boundary condition for $C_i$ and $\omega_j$ that allows users
+to couple these variables to the outlet velocity components (e.g. $u_x$,
+$u_y$). Without this new feature, users could only use uniform or fixed
+function-based velocity profiles in conjunction with the precursor looping
+capability.
 
@@
 	\caption{Boundary conditions in the main reactor geometry (Figure
-	\ref{fig:msfrbc}.}
+	\ref{fig:msfrbc}).}
 	\centering
@@
 		\multirow{4}{*}{Neutron flux $\phi_g$} & Top & $\frac{d \phi_g}{dx}
-		\big|_{\text{inflow}} = 0$ \\
-        & Outer & $\frac{d \phi_g}{dx} \big|_{\text{inflow}} = 0$ \\
-        & Bottom & $\frac{d \phi_g}{dx} \big|_{\text{inflow}} = 0$ \\
-        & Axial & $\frac{d \phi_g}{dx} = 0$ \\
+		\big|_{\text{inflow}} = 0$ \\[.5ex]
+        & Outer & $\frac{d \phi_g}{dx} \big|_{\text{inflow}} = 0$ \\[.5ex]
+        & Bottom & $\frac{d \phi_g}{dx} \big|_{\text{inflow}} = 0$ \\[.5ex]
+        & Axial & $\frac{d \phi_g}{dx} = 0$ \\[.5ex]
         \midrule
         \multirow{6}{*}{Delayed neutron precursor concentration $C_i$} &
-        Top (Core) & $\frac{d C_i}{dx} = 0$ \\
-        & Bottom (Core) & $\frac{d C_i}{dx} = 0$ \\
-        & Outer (Core) & $\frac{d C_i}{dx} = 0$ \\
-        & Axial (Core) & $\frac{d C_i}{dx} = 0$ \\
+        Top (Core) & $\frac{d C_i}{dx} = 0$ \\[.5ex]
+        & Bottom (Core) & $\frac{d C_i}{dx} = 0$ \\[.5ex]
+        & Outer (Core) & $\frac{d C_i}{dx} = 0$ \\[.5ex]
+        & Axial (Core) & $\frac{d C_i}{dx} = 0$ \\[.5ex]
         & Inlet (Core) & $C_i = c$ \\
@@
         \multirow{6}{*}{Decay heat power density $\omega_j$} &
-        Top (Core) & $\frac{d \omega_j}{dx} = 0$ \\
-        & Bottom (Core) & $\frac{d \omega_j}{dx} = 0$ \\
-        & Outer (Core) & $\frac{d \omega_j}{dx} = 0$ \\
-        & Axial (Core) & $\frac{d \omega_j}{dx} = 0$ \\
+        Top (Core) & $\frac{d \omega_j}{dx} = 0$ \\[.5ex]
+        & Bottom (Core) & $\frac{d \omega_j}{dx} = 0$ \\[.5ex]
+        & Outer (Core) & $\frac{d \omega_j}{dx} = 0$ \\[.5ex]
+        & Axial (Core) & $\frac{d \omega_j}{dx} = 0$ \\[.5ex]
         & Inlet (Core) & $\omega_j = c$ \\
@@
         & Inlet (Core) & $u_x = c$ \\
-        & Outlet (Core) & $\frac{d u_x}{dx} = 0$ \\
+        & Outlet (Core) & $\frac{d u_x}{dx} = 0$ \\[.5ex]
         \midrule
@@
         & Outer (Core) & $u_y = 0$ \\
-        & Axial (Core) & $\frac{d u_y}{dx} = 0$ \\
+        & Axial (Core) & $\frac{d u_y}{dx} = 0$ \\[.5ex]
         & Inlet (Core) & $u_y = 0$ \\
-        & Outlet (Core) & $\frac{d u_y}{dx} = 0$ \\
+        & Outlet (Core) & $\frac{d u_y}{dx} = 0$ \\[.5ex]
         \midrule
         \multirow{6}{*}{Temperature $T$} & Top (Core) &
-        $\frac{d T}{dx} = 0$ \\
-        & Bottom (Core) & $\frac{d T}{dx} = 0$ \\
-        & Outer (Core) & $\frac{d T}{dx} = 0$ \\
-        & Axial (Core) & $\frac{d T}{dx} = 0$ \\
+        $\frac{d T}{dx} = 0$ \\[.5ex]
+        & Bottom (Core) & $\frac{d T}{dx} = 0$ \\[.5ex]
+        & Outer (Core) & $\frac{d T}{dx} = 0$ \\[.5ex]
+        & Axial (Core) & $\frac{d T}{dx} = 0$ \\[.5ex]
         & Inlet (Core) & $T = c$ \\
-        & Outlet (Core) & $\frac{d T}{dx} = 0$ \\
+        & Outlet (Core) & $\frac{d T}{dx} = 0$ \\[.5ex]
 		\bottomrule
@@
 \glspl{DNP} outside the central core region by simulating its flow in a
-separate 1-D pipe geometry. This outer loop pipe simulation is implicitly
+separate 1-D pipe geometry. This outer loop pipe calculation is implicitly
 coupled to the active core simulation through Picard iterations in MOOSE's
 MultiApp functionality and inlet/outlet boundary values. For this work with
-the \gls{MSFR} model, we assumed a pipe
-length of 2.255 m with salt flowing at 1.1275 m s$^{-1}$ for an average
-out-of-core residence time of 2 seconds to follow the design specifications.
+the \gls{MSFR} model, the pipe length is 2.255 m with salt flowing at 1.1275 m
+s$^{-1}$ for an out-of-core residence time of 2 s. The present author derived
+these parameters from the reference specifications of 4 s cycle
+time and 50\% out-of-core salt fraction (Table \ref{table:msfr}).
 
@@
 
-Since this outer loop region is largely subcritical, the only significant
-neutronics-related phenomena are the drift and decay of \glspl{DNP}.
-The governing equation for the \glspl{DNP} is:
+The outer loop region is largely subcritical because most of it is adjacent to
+the boron carbide absorber as shown in Figure \ref{fig:core}. Therefore, the
+only significant neutronics-related phenomena are the drift and decay of
+\glspl{DNP}. The governing equation for the \glspl{DNP} is:
 %
@@
 
-A constant velocity of 1.1275 m s$^{-1}$ is applied in the outer loop region.
-The governing equation for
-temperature, derived from equation \ref{eq:temp}, is:
+A constant velocity of 1.1275 m s$^{-1}$ is applied in the outer loop region
+to maintain the nominal 2 s out-of-core residence time. The governing equation
+for temperature, derived from equation \ref{eq:temp}, is:
 %
@@
 
-The fission heat source term is replaced with a heat exchanger sink term
-$Q_{hx}$
+In the outer loop region, the fission heat source term is replaced with a heat
+exchanger sink term $Q_{hx}$
 which depends on the temperature difference between the fuel salt $T$ and the
@@
 \begin{align}
-    u_x &= -2034.01 \Big[\frac{y}{18.75} - \Big(\frac{y}{18.75}\Big)^2 
+    u_x &= -\xi \Big[\frac{y}{h} - \Big(\frac{y}{h}\Big)^2 
     \Big] \\
     \intertext{where}
-    y = \text{height along the inlet [cm].}
+    \xi &= \text{normalizing constant [m$\cdot$s$^{-1}$],} \nonumber \\
+    y &= \text{height along the inlet [m],} \nonumber \\
+    h &= \text{total height of the inlet $= 0.1875$ m.} \nonumber
 \end{align}
 %
-The inlet is 18.75 cm high, and the coefficient at the front is from
-normalizing the equation for a total volumetric flow rate of
-4.5 m$^3$ s$^{-1}$.
+$\xi$ is a normalizing constant that depends on the total volumetric flow
+rate, $\dot{V}$. Solving the following set of equations in $v$ and $\dot{V}$:
+%
+\begin{align}
+    v &= \xi [\frac{y}{h} - \frac{y^2}{h^2}] \\
+    \dot{V} &= \int^h_0 \int^{2 \pi}_0 v r d\theta dy \\
+    \intertext{where}
+    r &= \text{radius [m],} \nonumber \\
+    \theta &= \text{azimuthal angle [rad],} \nonumber
+\end{align}
+%
+gives $\xi = 20.3401$ m$\cdot$s$^{-1}$ for $\dot{V} = 4.5$ m$^3\cdot$s$^{-1}$.
 
@@
     \mathcal{C} &= \text{outlet boundary curve} \nonumber \\
-    u &= \text{outflow velocity perpendicular to the outlet boundary [m
-    s$^{-1}$].} \nonumber
+    u &= \text{outflow velocity perpendicular to the outlet boundary
+    [m$\cdot$s$^{-1}$].} \nonumber
 \end{align}

### 2020-07-09 msr.tex
Katy's second review edits

@@
 this chapter presents the reference specifications of the \gls{MSFR} concept
-studied in this work.
+studied in this work and summarizes the state of the art in the international
+literature regarding modeling and simulation of the \gls{MSFR}.
 
-\section{History}
+\section{\gls{MSR} History}
 
@@
 as part of the US Aircraft Nuclear Propulsion program
-\cite{rosenthal_molten-salt_1970}. Researchers recommended the molten fluoride
+\cite{rosenthal_molten-salt_1970}. Researchers recommended molten fluoride
 salts in particular for high uranium solubility, chemical stability, low vapor
-pressure even at high temperatures, decent heat transfer properties,
+pressure even at high temperatures, good heat transfer properties,
 resistance against radiation damage, and reduced corrosive effects on some
@@
 
-\section{Features}
+\section{\gls{MSR} Features}
 
@@
 new tools for simulating steady-state and transient behavior in \glspl{MSR}.
-Many of the earlier efforts featured simplifications in simulating
+Many earlier efforts featured simplifications in simulating
 thermal-hydraulics by solving 1-D Navier-Stokes equations or using
@@
 concurrently developed
-a full-core 3-D model of the \gls{MSFR} on OpenFOAM, albeit with one-group
+a full-core 3-D model of the \gls{MSFR} in OpenFOAM, albeit with one-group
 neutron diffusion to reduce computational load. With the 3-D model, the
@@
 continuous online fuel reprocessing. Several reasons motivated the omission of
-graphite moderators from the original \gls{MSBR} design. Graphite is
+graphite moderators from the original \gls{MSBR} design. Firstly, graphite is
 susceptible to
 long-term radiation damage and replacement is likely to be necessary during
-the operating lifetime of the reactor. Graphite also has a positive
+the operating lifetime of the reactor. Secondly, graphite also has a positive
 temperature coefficient of reactivity; eliminating graphite from the design
-ensures a greater safety margin \cite{mathieu_thorium_2006}. While negative
-temperature coefficients are attainable with very thermalized spectra,
-breeding ratios deteriorated significantly due to parasitic absorption in the
-graphite \cite{mathieu_thorium_2006}.
+ensures a greater safety margin \cite{mathieu_thorium_2006}. Lastly, while
+negative temperature coefficients are attainable with very thermalized
+spectra, breeding ratios deteriorated significantly due to parasitic
+absorption in the graphite \cite{mathieu_thorium_2006}.
 %
@@
 		Thermal/Electric output [MW$_{\text{th}}$/MW$_{\text{e}}$] & 3000 /
-		1500 
+		1300 
 		\\
@@
 surrounds the core radially. The top and bottom of the core are enclosed
-by nickel alloy reflectors. A layer of boron carbide behind the blanket tanks
+by nickel alloy reflectors. A layer of boron carbide behind the blanket tank
 protects the peripheral equipment from excessive neutron damage. During severe
@@
 study by Brovchenko et al. \cite{brovchenko_neutronic_2019} shows that
-different neutronics software with different nuclear data clearly provide
+different neutronics software with different nuclear data provide
 different $k_{eff}$ estimates even with the same isotopic compositions and
@@
 The thermal and electric power output of the \gls{MSFR} are 3000 MW
-$_{\text{th}}$ and 1500 MW$_{\text{e}}$, respectively. The high thermal
-efficiency ($\eta_{th}=0.5$) is due to the high operating
+$_{\text{th}}$ and 1300 MW$_{\text{e}}$, respectively. The high thermal
+efficiency ($\eta_{th}=0.43$) is due to the high operating
 temperature. The inlet and outlet temperature
@@
 operating temperatures and the melting point of the salt (873 K)
-\cite{euratom_final_2015}. The \gls{MSFR} has
-heat exchangers and an intermediate coolant loop to isolate the power
+\cite{euratom_final_2015}. The \gls{MSFR} has heat exchangers and an
+intermediate molten salt coolant loop to separate the power
 conversion system from the highly radioactive fuel salt. This also serves as
@@
 
-This present work uses the same 2-D square-cylindrical \gls{MSFR} design to
-benchmark our results against results published by Fiorina et al.
+The present work uses the same 2-D square-cylindrical \gls{MSFR} design to
+compare our results to results published by Fiorina et al.
 \cite{fiorina_modelling_2014}. The design is a 2-D axisymmetric representation
@@
 group constants calculations in Serpent, the 2-D axisymmetric model is
-extended into a 3-D model by a 360-degrees rotation about the
+extended into a 3-D model by a 360-degree rotation about the
 central axis. The material definitions are the same as those specified in the
-reference \gls{MSFR} model. Accordingly, the pump and heat exchanger regions
-are assumed to be composed of 100\% fuel salt. While this may not be entirely
-accurate, the exact details of the pump and heat exchanger systems are still
-under active study, and this external loop region is presumed to be of little
-neutronic importance due to its position behind the strong boron carbide
-neutron absorber layer.
+reference \gls{MSFR} model \cite{brovchenko_neutronic_2019}. Accordingly, the
+pump and heat exchanger regions
+are assumed to be composed of 100\% fuel salt. While this is an approximation,
+the outer loop is shielded from the core by a layer of boron carbide and sees
+sufficiently low neutron fluxes that we may neglect its exact neutronics.
 %
@@
 
-Although this present work uses the same 2-D axisymmetric geometry for
-generating
-group constant data from Serpent, there are two minor differences between the
-\gls{MSFR} geometry used in the Moltres model, and the Polimi and TUDelft
-models \cite{fiorina_modelling_2014}. Firstly, the reactor geometry for
+Although the present work uses the same 2-D axisymmetric geometry for
+generating group constant data from Serpent, there are two minor differences
+between the \gls{MSFR} geometry modeled herein, and the geometry used in the
+Polimi and TUDelft models \cite{fiorina_modelling_2014}. First, the reactor
+geometry for
 Moltres excludes the 2-cm-thick structural material around the blanket tank
 that separates the fuel and blanket salts. The thickness is much smaller than
-the rest of the regions in the geometry and caused issues for meshing. The
-neutronics results in Chapter \ref{chap:nts} shows that the overall $k_{eff}$
+the rest of the regions in the geometry and it complicated mesh generation in
+the adjacent regions near the boundaries. The
+neutronics results in Chapter \ref{chap:nts} show that the overall $k_{eff}$
 and other parameters from Moltres show good agreement with that from Serpent.
 It also has no direct impact on the temperature distribution results because
-this present work solves for the temperature distribution in the primary loop
+the present work solves for the temperature distribution in the primary loop
 with homogeneous Neumann boundary conditions on the fuel salt-to-wall
@@
 77.5\% LiF - 22.5\% AcF$_4$ (actinide fluorides)
-\cite{merle-lucotte_launching_2011}. Typically, researchers working with the
-\gls{MSFR}
-model tweak the exact actinide composition by varying the $^{232}$Th to
-$^{233}$U ratio to obtain a $k_{\text{eff}}$ value of 1 at a uniform
-temperature of 973 K \cite{brovchenko_neutronic_2019}. Thus, the exact
-actinide compositions vary depending on
-the nuclear data library and neutron transport code. This present work uses
+\cite{merle-lucotte_launching_2011}. Some researchers working with the
+\gls{MSFR} model tweak the exact actinide composition by varying the
+$^{232}$Th to $^{233}$U ratio to obtain a $k_{\text{eff}}$ value of 1 at a
+uniform temperature of 973 K \cite{brovchenko_neutronic_2019} or for other
+unmentioned reasons \cite{fiorina_molten_2013}. Thus, the exact
+actinide compositions vary depending on the nuclear data library, neutron
+transport code, and the purpose of the study. The present work uses
 a fuel salt composition of 77.5\% LiF - 19.913\% ThF$_4$ - 2.587\%
-$^{233}$UF$_4$ for all calculations, which is a simplifying assumption but a
-fuel burnup analysis is complex and out of scope for this thesis. Table
-\ref{table:prip} shows relevant physical properties of the fuel and blanket
-salts.
+$^{233}$UF$_4$ for all calculations, which gives steady-state inlet and outlet
+temperatures approximately equal to the nominal values in Table
+\ref{table:msfr}. Table \ref{table:prip} shows relevant physical properties of
+the fuel and blanket salts.
 %
@@
 \centering
-\caption{Properties of the fuel and blanket salts LiF-AcF$_4$.}
+\caption{Properties of the fuel and blanket salts LiF-AcF$_4$
+\cite{brovchenko_neutronic_2019}.}
 \begin{tabular}{l l r r}
@@
 Melting temperature [K] & 841 & {N/A} & 1 bar \\
-Boiling temperature [K] & 1874 & {N/A} & 1 bar \\
 Density, $\rho$ [kg m$^{-3}$] & $4094-0.882 \cdot (T-1008)$ & $4125$ & 893-1123 K \\
@@
 The reflectors on the periphery of the reactor core and the blanket tank are
-made of a NiCrW Hastelloy (metal alloy) \cite{brovchenko_neutronic_2019}.
-Table \ref{table:refl} details the elemental composition
-of the Hastelloy. The alloy has a density
-of 10 g$\cdot$cm$^{-3}$. The \gls{MSFR} also includes a 20 cm
+made of a NiCrW Hastelloy (metal alloy) \cite{brovchenko_neutronic_2019}, and
+Table \ref{table:refl} details its elemental composition. This alloy has a
+density of 10 g$\cdot$cm$^{-3}$. The \gls{MSFR} also includes a 20 cm
 layer of boron carbide (B$_4$C) to protect the heat exchangers and pumps from

### 2020-07-09 neutronics.tex
Katy's second review edits

--- 2020-07-03 neutronics.tex
@@
-This chapter compares key neutronics results between Serpent and Moltres
-for a static model of the \gls{MSFR}, i.e. no salt flow, and uniform
-temperature distribution to assess the accuracy of the
+This chapter verifies Moltres' ability to reproduce key neutronics parameters
+using group constant data from Serpent 2, which is essential for accurate
+neutronics calculations in the subsequent multiphysics simulations. The model
+under study is a static model of the \gls{MSFR}, i.e. no salt flow, and
+uniform temperature distribution to assess the accuracy of the
 six-group neutron diffusion model in Moltres on a fast-spectrum reactor. This
-exercise also serves as a verification for the group constant data
-generated for Moltres from Serpent. For clarity, here are the important
-details of the static model.
+verification exercise builds on the previous study by Lindsay et al.
+\cite{lindsay_introduction_2018} that verified Moltres' neutronics
+capabilities with a two-group neutron diffusion model of the \gls{MSRE}.
 
-\textbf{Details of the static \gls{MSFR} model on Moltres}
+\textbf{Details of the static \gls{MSFR} model in Moltres}
 \begin{itemize}
     \item No salt flow (static salt)
+    \begin{center}
+        $v_{salt} = 0$ m$\cdot$s$^{-1}$
+    \end{center}
     \item Uniform temperature of 973 K throughout the 2D core model
+    \begin{center}
+        $T = 973$ K
+    \end{center}
     \item Six neutron energy groups
+    \begin{center}
+        $G = 6$
+    \end{center}
     \item Eight delayed neutron precursor groups
+    \begin{center}
+        $I = 8$
+    \end{center}
     \item Vacuum boundary conditions on the outer boundaries of the 2D core
     mesh
+    \begin{center}
+        $\frac{d \phi}{dx}\big|_{\text{inflow}} = 0$ m$^{-2}\cdot$s$^{-1}$
+    \end{center}
 \end{itemize}
@@
 
-Moltres solves the six-group neutron diffusion equations as a
+Moltres solves the six-group neutron diffusion equations (Equation
+\ref{eq:neut}) as a
 steady-state eigenvalue problem to find the $k_{\text{eff}}$ for the static
@@
 at 973 K and the corresponding salt density, and Table \ref{table:ktemp} shows
-the $k_{\text{eff}}$ values for other temperatures at 100 K intervals. We
-observe small discrepancies on the order of 100 pcm between the two codes
-which we attribute to two main factors: the accuracy of the neutron diffusion
+the $k_{\text{eff}}$ values for other temperatures at 100 K intervals. Two
+main factors contribute to the small discrepancies on the order of 100 pcm
+between the two applications: the accuracy of the neutron diffusion
 model, and the omission of the blanket tank structural material in Moltres.
@@
 SP$_{\text{N}}$ deterministic methods nor the Monte Carlo approach in Serpent.
-Regarding the omission of the blanket tank material, we replaced the
-2 cm-thick structural material with blanket salt. This replacement may be
-partly responsible for the higher $k_{\text{eff}}$ value calculated by Moltres
-as the macroscopic fission cross sections of the blanket salt are non-zero for
-the higher neutron energy groups. Nevertheless, the discrepancy is smaller
+Regarding the omission of the blanket tank material, the present author
+replaced the 2 cm-thick structural material with blanket salt. This
+replacement is partly responsible for the higher $k_{\text{eff}}$ value
+calculated by Moltres as fissions occur in the blanket salt. Nevertheless, the
+discrepancy is smaller
 than the 228.5 pcm and 256.7 pcm discrepancies reported by Cervi et al.
 \cite{cervi_development_2019} for their six-group $SP_3$ and neutron
-diffusion methods, respectively, in OpenFOAM. Aufiero et al.
-\cite{aufiero_development_2014} used the same neutron diffusion model in
-OpenFOAM for their transient analysis of the \gls{MSFR}, albeit with one
-neutron energy group to reduce computational load.
+diffusion methods, respectively.
 
@@
 at steady state to meet the design specifications for the inlet and outlet
-temperatures by adjusting the fissile inventory and $k_{\text{eff}}$. In
-transient analysis, the delayed neutron fraction $\beta$ and reactivity
-coefficients $\alpha$ are
-important as they dictate the duration, shape, and magnitude of the reactor
-transient response. The difference between $\beta$ and
-$\beta_{\text{eff}}$ is that $\beta$ is the physical delayed neutron fraction
-while $\beta_{\text{eff}}$ is the delayed neutron fraction weighted on the
-adjoint neutron flux. We will compare the $\beta$ value from Moltres to the
-$\beta_{\text{eff}}$ value from Serpent because Moltres currently lacks a
-adjoint calculation capability. We calculated $\beta$ by
+temperatures by adjusting the fissile inventory. On the other hand, the
+delayed neutron fraction, $\beta$, and reactivity coefficients, $\alpha$, are
+clearer indicators of transient reactor behavior in an accident transient.
+$\beta$ primarily affects the magnitude of the initial change in power and the
+time delay towards the new equilibrium power, while $\alpha$ affects the
+magnitude of the change in reactor power and temperature.
+
+This work compares the $\beta$ value from Moltres to the
+$\beta_{\text{eff}}$ value from Serpent because Moltres currently lacks an
+adjoint calculation capability.
+The difference between $\beta$ and $\beta_{\text{eff}}$ is
+that $\beta$ is the unweighted delayed neutron fraction
+while $\beta_{\text{eff}}$ is the delayed neutron fraction weighted by the
+adjoint neutron flux. The present author calculated $\beta$ by
 taking the relative difference between the $k_{\text{eff}}$ values with and
@@
 
-The temperature reactivity feedback arises mainly from Doppler broadening of
-resonance absorption peaks and temperature-induced density changes. Although
-Doppler coefficients typically show logarithmic dependence to temperature, we
-report them, along with the other coefficients, as linear gradient
-values of the reactivity given the relatively linear trend within the relevant
-temperature range (Figure \ref{fig:reactivity}). The temperature range
-extends below the melting point of the fuel salt (873 K) to ensure that
-the data covers the relevant range between 873 K and 900 K. Table
-\ref{table:alpha} shows
-the temperature coefficients as described prior. The total temperature
+Temperature reactivity feedback arises mainly from Doppler broadening of
+resonance absorption peaks and thermal expansion. The current work reports
+the reactivity, $\rho$, values for temperatures from 800 K to 1400 K at 100 K
+intervals (Figure \ref{fig:reactivity}). The slopes represent the total,
+Doppler, density $\alpha$ values. The temperature range
+extends below the melting point of the fuel salt (841 K) to ensure that
+the data covers the relevant range between 841 K and 900 K. Table
+\ref{table:alpha} shows the various $\alpha$ values calculated using the
+linear least squares approach. The total temperature
 coefficients from Serpent and Moltres show excellent agreement with a
@@
     reactivity values were calculated at a fixed density of 4.1249 g
-    cm$^{-3}$. The density reactivity values were calculated at a fixed
-    temperature of 973 K.}
+    cm$^{-3}$. The thermal expansion reactivity values were calculated at a
+    fixed temperature of 973 K.}
     \label{fig:reactivity}
@@
 	for the temperature range of 800 K to 1400 K.}
-	\begin{tabular}{l S S S S}
+	\begin{tabular}{l S S S}
 		\toprule
-		{Software} & {$\alpha_D$ (log) [pcm]} & {$\alpha_D$ (linear) [pcm
+		{Software} & {$\alpha_D$ [pcm
 		K$^{-1}$]} & {$\alpha_\rho$ [pcm K$^{-1}$]} & {$\alpha_T$ [pcm
@@
 		\midrule
-		{Serpent} & -4034 \pm 14 & -3.737 \pm 0.013 & -3.424 \pm 0.013 &
+		{Serpent} & -3.737 \pm 0.013 & -3.424 \pm 0.013 &
 		-7.165 \pm 0.013 \\
-		{Moltres} & {-} & {-} & {-} & -7.184\\
+		{Moltres} & {-} & {-} & -7.184\\
 		\bottomrule
@@
 Serpent group constants. Figure \ref{fig:ntspec} compares the neutron energy
-spectra from Serpent and Moltres in the central fuel salt region. The
-six-group neutron spectra overlap exactly over each other. More generally, the
+spectra from Serpent and Moltres in the central fuel salt region. More
+generally, the
 plot shows the distinctive fast spectrum observed in the \gls{MSFR} with dips
@@
 
-In summary, Moltres replicated most of the relevant neutronics parameters
-accurately with the group constant data from Serpent. Moltres agrees with the
-high fidelity simulation in Serpent 2 for the $\beta_{\text{eff}}$ and
+In summary, Moltres replicated the relevant neutronics parameters
+accurately using the group constant data from Serpent 2. Moltres agrees with
+the high fidelity simulation in Serpent 2 for the $\beta_{\text{eff}}$ and
 temperature reactivity coefficients, which are important parameters for
 modeling transient reactor behavior. The $k_{\text{eff}}$ values have
-discrepancies on the order of 100 pcm and they are relatively small compared
-to the $k_{\text{eff}}$ values from the neutron diffusion and SP3 models in
-OpenFOAM \cite{cervi_development_2019}.
+discrepancies on the order of 100 pcm which are relatively small compared
+to other \gls{MSR} multiphysics simulation tools (e.g. the neutron diffusion
+and SP3 models in OpenFOAM \cite{cervi_development_2019}).

### 2020-07-10 abstract.tex
Katy comment edits

@@
 As the world continues its transition towards low-carbon electricity
-generation to combat climate change, molten salt reactors is potential
-options in the near-term future for replacing fossil fuel and ageing nuclear
+generation to combat climate change, molten salt reactors is a potential
+option in the near-term future for replacing fossil fuel and ageing nuclear
 power plants. At the current state of development, molten salt reactors still
 require extensive research to become viable. This thesis presents the latest
-developments in Moltres, a simulation tool for molten salt reactors. These new
-developments are: support for coupling the incompressible Navier-Stokes and
-the delayed neutron precursor looping systems, and a model for
+developments in Moltres, a simulation tool for molten salt reactors. These
+new developments are: support for coupling the incompressible Navier-Stokes
+and the delayed neutron precursor looping systems, and a model for
 simulating decay heat from fission products at steady state and during

### 2020-07-10 model.tex
Katy comment edits

@@
 has also been validated and verified against experimental data and other well
-established codes \cite{leppanen_calculation_2014}.
+established neutron transport applications \cite{leppanen_calculation_2014}.
 
@@
 two regions, the central core region where most of the fissions take place,
-and the outer loop region where the heat exchanger is located. The central
-core region specifically refers to the central region indicated by
-the red box in Figure \ref{fig:core}. The outer loop is simplified
-into a 1-D pipe as it is a subcritical region and its main purposes are: to
-introduce an out-of-core residence time for the \glspl{DNP}, and to contain
-the heat removal kernel to simulate the heat exchanger.
+and the outer loop region where the heat exchanger is located. The red box in
+Figure \ref{fig:core} indicates the central core region. The outer loop is
+simplified into a 1-D pipe as it is a subcritical region. Its main purposes
+are to introduce an out-of-core residence time for the \glspl{DNP} and to
+contain the heat removal kernel to simulate the heat exchanger.
 Accordingly, this section provides separate descriptions for the governing
@@
 a third region comprised of the blanket salt, the NiCrW Hastelloy reflectors, 
-and other material. This region is important for capturing a more accurate
-estimate of neutron leakage than would
-imposing fixed albedo neutron flux boundary conditions.
+and other material.
 
@@
 In addition to the intrinsic molecular viscosity in the \gls{INS} equations,
-this thesis introduces an
-eddy viscosity term to approximate turbulent flow effects. The current
-implementation of the Navier-Stokes module does not have a turbulence model
-such as the \gls{RANS} models used in the Polimi and TUDelft models
-\cite{fiorina_modelling_2014}. Thus, this work uses
-a zeroth-order approximation of the eddy viscosity based on the results
-reported in the Polimi/TUDelft models. The eddy viscosity is assumed to be 40
-Pa$\cdot$s. Despite the simplicity of this assumption, the resulting flow profile is
-similar to flow profile in the Polimi and TUDelft models at steady state.
+this thesis introduces an eddy viscosity term, $\mu_t$, to approximate
+turbulent flow effects. The current implementation of the Navier-Stokes module
+does not have a turbulence model. The options for turbulence modeling in
+\gls{CFD} include direct numerical simulations (DNS) and large eddy
+simulations (LES) for higher fidelity flow simulations, \gls{RANS} methods for
+balanced compromises between accuracy and computational speed, and lumped
+parameter and sub-channel methods for even faster performance with greater
+accuracy costs \cite{moorthi_review_2018}. The Polimi and TUDelft models used
+\gls{RANS} methods to model salt flow \cite{fiorina_modelling_2014}. This work
+uses a zeroth-order approximation of $\mu_t$ based on the calculated $\mu_t$
+values reported in the Polimi and TUDelft models. The models
+predicted spatial $\mu_t$ values ranging from 0-110 Pa$\cdot$s, with most
+of the core lying in the 30-50 Pa$\cdot$s range. Thus, the present work uses
+the approximated value $\mu_t=$ 40 Pa$\cdot$s. Despite the simplicity of this
+approximation, the resulting flow profile is similar to flow profile in the
+Polimi and TUDelft models at steady state.
 
@@
 Table \ref{table:loopbc} summarizes the boundary conditions for all variables
-on the inlet and outlet of the 1-D outer loop region.
+on the inlet and outlet of the 1-D outer loop region. The inlet boundary
+conditions are all Dirichlet boundary conditions. The inlet boundary values
+are set by the outflow from the central core region that this inlet is
+connected to in the actual reactor geometry. The outlet boundary conditions
+are all outflow boundary conditions as shown in the table.
 

### 2020-07-10 msr.tex
Katy comment edits

@@
 
-In spite of this setback, research into \glspl{MSR} continued through the late
-1970s. In 1980, \gls{ORNL} published a report describing a new \gls{MSR}
-concept, called the \gls{DMSR} \cite{gehin_liquid_2016} with denatured
-$^{235}$U fuel (i.e. low-enriched uranium). The \gls{ORNL} researchers
-developed this design in response to the fuel reprocessing restrictions
-introduced by President Ford in 1976. The \gls{DMSR} would operate as a
-once-through
-converter system without fuel reprocessing. While the fuel consists of 19.75
-\% high-assay low-enriched uranium, the initial core loading includes thorium
-to boost its conversion ratio throughout its lifetime. It has a continuous
-online feed consisting of \gls{HALEU} to maintain criticality, and denatured
-$^{235}$U to keep uranium enrichment levels below nuclear non-proliferation
-policy thresholds. The design also includes a gas sparging system for removing
-gaseous fission products, while noble metals plate out onto the walls of
-the coolant loop. The older \gls{MSBR} design had a significant drawback; the
-extensive neutron damage in the graphite moderator necessitated frequent
-replacement (every four years) throughout its operational lifetime. The
-\gls{DMSR} avoids this issue by having a lower power density while maintaining
-the overall power output of 2250 MW$_{\text{th}}$. As a result, researchers
-projected that the graphite moderator would last for the entirety of the
-\gls{DMSR}'s design lifetime.
+Two other \gls{MSR} designs emerged in the 1970s: the \gls{DMSR} from
+\gls{ORNL} \cite{gehin_liquid_2016} and the \gls{MCFR} from the UK Atomic
+Energy Authority \cite{smith_assessment_1974}. \gls{ORNL}
+researchers developed the \gls{DMSR}, named after its planned use of denatured
+$^{235}$U fuel (i.e. low-enriched uranium), as a once-through converter system
+without fuel reprocessing in response to fuel reprocessing restrictions
+introduced by President Ford in 1976. On the other hand, UK researchers
+developed the lead-cooled \gls{MCFR} to enhance $^{239}$Pu breeding with the
+relatively fast neutron spectrum. Both designs failed to gain traction and did
+not inspire any working prototypes.
 
-There was a concurrent program at the UK Atomic Energy Authority for the
-development of a 2500 MW$_{\text{e}}$ lead-cooled Molten Chloride Fast Reactor
-concept \cite{smith_assessment_1974}. It is a dual fluid system,
-with separate loops for the fuel salt and the blanket salt. The blanket is a
-1 m-wide tank surrounding the core. The absence of moderators and the choice
-of chloride over fluoride salt resulted in a relatively hard neutron spectrum
-which favors $^{239}$Pu breeding over the thorium cycle. The UK researchers
-performed some experiments to study molten salt chemistry but there were no
-reactor prototypes. The UK program eventually shut down just like its
-US counterpart partly due to the successful demonstration of the Prototype
-Fast Reactor which had achieved criticality in 1974.
-
-Following a lull lasting through the late 20th century, \gls{MSR} research
-picked up pace due to renewed interest initiated by the Generation IV Forum in
-2001. Today, various national and commercial bodies are developing numerous
-\gls{MSR} concepts.
+Following a lull lasting through the late 20th century, researchers at
+\gls{CNRS} began research into \glspl{MSR} in 1997
+\cite{heuer_simulation_2010}. Starting from the \gls{MSBR} design, they
+performed parametric studies on the design based on safety, breeding, and
+other performance metrics \cite{mathieu_thorium_2006}. Their efforts
+culminated in the \gls{MSFR} concept, a fast-spectrum breeder \gls{MSR}
+designed to run on the thorium fuel cycle \cite{merle_optimized_2007}. In
+2008, the Generation IV International Forum highlighted the \gls{MSFR} among
+other \gls{MSR} designs for further development \cite{gif_generation_2008}.
+The \gls{MSFR} has also benefited from collaborative research through two
+European-funded projects, the EVOL and SAMOFAR projects
+\cite{euratom_final_2015}. Today, numerous other national and commercial
+bodies are also developing their own \gls{MSR} designs.
 
@@
 \cite{transatomic_power_corporation_technical_2016} while two-fluid designs
-include the \gls{MSFR} \cite{serp_molten_2014} and \gls{MOSART}
-\cite{ignatiev_molten_2014}.
+include the \gls{MSFR} \cite{merle_optimized_2007} and \gls{MOSART}
+\cite{ignatiev_molten_2014}. The current work involves simulations of
+the \gls{MSFR} in Moltres and Section \ref{sec:msfr} describes this reactor
+design in further detail.
 
@@
 
-Economic analyses of \glspl{MSR} are preliminary at the current
-stage of \gls{MSR} development. Qualitatively, many technical factors favor
-\glspl{MSR} over \glspl{LWR}. Some of these factors include: smaller reactor
-core size due to low operating pressures, higher thermal efficiency and
-cheaper air-cooling systems due to high operating temperatures, reduction in
-fuel fabrication costs, and shorter reactor shutdown periods due to online
-refueling \cite{dolan_1_2017}. Moir \cite{moir_recommendations_2008}
-published data from \gls{ORNL} reports on the estimated electricity costs for
-a 1000 MWe \gls{MSR} with 20\% enriched $^{235}$U fuel and no reprocessing
-throughout the reactor's 30-year lifetime. Adjusted to year 2000 dollars, the
-electricity costs for this type of \gls{MSR} was 7\% and 8\% lower than that
-of pressurized water reactors and coal power plants. These differences are
-within the $\pm$10\% uncertainty range but Moir remains optimistic.
-The author also notes that costs could be lower in an \gls{MSR} with
-fertile feed and online reprocessing could reach fuel self sufficiency.
-
 In terms of technological readiness, \glspl{MSR} still require significant
-research efforts and engineering demonstrations for experimental validation of
+research efforts and experiments for the validation of
 various components before a full commercial model can be commissioned. Work
@@
 new tools for simulating steady-state and transient behavior in \glspl{MSR}.
-Many earlier efforts featured simplifications in simulating
-thermal-hydraulics by solving 1-D Navier-Stokes equations or using
-predetermined uniform velocity fields \cite{krepel_dyn3d-msr_2007,
-kophazi_development_2009}. In more recent years, there has been
+Earlier efforts featured simplifications in simulating
+thermal-hydraulics by using predetermined 1-D velocity fields. One such effort
+\cite{krepel_dyn3d-msr_2007} used DYN3D-MSR, an \gls{MSR} simulation tool
+adapted from the \gls{LWR} simulation tool DYN3D while another
+\cite{kophazi_development_2009} adapted DALTON, originally a simulation
+tool for high-temperature reactors. Both works studied the \gls{MSRE} and
+simulated flow through advection terms in the temperature and precursor
+equations in conjunction with full 3-D calculations for neutron flux and heat
+conduction.
+
+In more recent years, there has been
 significant progress towards fully coupled, spatial codes that feature
@@
 scenarios such as reactivity insertions, changes in pumping rate, and the
-presence of periodic perturbations. This approach was featured again in a
-later paper by Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D
-axisymmetric model of the
-\gls{MSFR}. The authors presented results from the Politecnico di Milano
-COMSOL-based approach, and another approach by researchers from Delft
-University of Technology, in which they coupled their in-house neutronics and
-thermal-hydraulics codes, DALTON and HEAT respectively. With multigroup
+presence of periodic perturbations. Nagy et al. \cite{nagy_steady-state_2014}
+introduced an improved version of DALTON which coupled with a \gls{CFD}
+application and incorporated incompressible
+Navier-Stokes flow to simulate salt flow more accurately. The authors
+validated their model against \gls{MSRE} experimental data before presenting
+multiphysics simulation results of a different graphite-moderated breeder
+\gls{MSR}.
+
+Both DALTON and the COMSOL-based tool featured again in a later paper by
+Fiorina et al.
+\cite{fiorina_modelling_2014} in 2014 for a 2-D axisymmetric model of the
+\gls{MSFR}. The authors compared \gls{MSFR} simulations results from both
+tools, in which they dubbed the COMSOL and DALTON/HEAT tools the Polimi and
+TUDelft models, respectively, after the universities where each tool was
+developed. With multigroup
 neutron diffusion and \gls{RANS} formulations on ultra fine meshes, both
@@
 authors could simulate the asymmetric reactor response to the failure of a
-single pump in the sixteen-pump \gls{MSFR} configuration. The authors also
+single pump in the sixteen-pump \gls{MSFR} configuration. Aufiero et al. also
 provided quantitative data supporting the use of implicit coupling over
 explicit coupling to obtain accurate solutions of the transient cases.
+
+Separately, Zhang et al. \cite{zhang_couple_2014} introduced another
+self-contained \gls{MSR} simulation tool named COUPLE. Similar to the COMSOL
+and OpenFOAM-based models, COUPLE solves the multigroup neutron diffusion
+equations and the incompressible Navier-Stokes equations with the
+$k$-$\epsilon$ turbulence model. The authors verified COUPLE's neutronics and
+thermal-hydraulics against MCNP and a FLUENT-based model, respectively, before
+presenting their multiphysics steady-state results with a 2-D axisymmetric
+model of the \gls{MSFR}.
+
 Recognizing the huge computational burden required for full 3-D simulations, 
-later authors came up with innovative ways to alleviate this issue. Methods
-include
-selective geometrical reduced order modeling for various components of a
-reactor based on the importance of the physical phenomena being simulated
-\cite{zanetti_geometric_2015}, or using a novel, efficient method for
-neutronics calculations \cite{laureau_transient_2017}.
+some authors came up with innovative ways to alleviate this issue. Zanetti et
+al. \cite{zanetti_geometric_2015} introduced a geometric multiscale modeling
+approach for various components of a reactor based on the importance of
+the physical phenomena being simulated. The authors employed a 3-D 
+single-channel model for the \gls{MSRE} core and modeled the other
+regions in the reactor such as the upper and lower plenums and the heat
+exchanger using a lumped parameter approach. They also applied flux
+factorization to decouple the temporal and spatial components in the neutron
+diffusion equation. Laureau et al. \cite{laureau_transient_2017} devised a
+different method, focusing solely making neutronics calculations more
+efficient while retaining OpenFOAM's \gls{CFD} capabilities. The authors
+adapted fission matrices, commonly used in Monte Carlo neutronics
+applications, for efficient transient calculations in a modified version of
+Serpent 2 \cite{leppanen_serpent_2014} and coupled it with OpenFOAM for
+multiphysics transient simulations of the \gls{MSFR}.
 
-\section{Molten Salt Fast Reactor}
+\section{Molten Salt Fast Reactor} \label{sec:msfr}
 
@@
 	\caption{Main specifications of the \gls{MSFR} concept
-				\cite{serp_molten_2014}.}
+				\cite{merle_concept_2017}.}
 	\centering
@@
 actinide fluorides such as uranium, thorium, plutonium, and other \gls{TRU}
-fluorides.  For an initial composition consisting of $^{232}$Th and $^{233}$U,
+fluorides. Using fluoride salts instead of chloride salts results in a
+relatively thermal neutron spectrum. Therefore, the \gls{MSFR} has a more
+thermal neutron spectrum than a typical sodium fast reactor. For an initial
+composition consisting of $^{232}$Th and $^{233}$U,
 the benchmark value for the amount of $^{233}$U for criticality under
@@
 the outer loop is shielded from the core by a layer of boron carbide and sees
-sufficiently low neutron fluxes that we may neglect its exact neutronics.
+sufficiently low neutron fluxes that we may neglect its exact neutronic
+properties.
 %
@@
 uniform temperature of 973 K \cite{brovchenko_neutronic_2019} or for other
-unmentioned reasons \cite{fiorina_molten_2013}. Thus, the exact
-actinide compositions vary depending on the nuclear data library, neutron
-transport code, and the purpose of the study. The present work uses
+motivations \cite{fiorina_molten_2013} \cite{pettersen_coupled_2016}. Thus,
+the exact actinide compositions vary depending on the nuclear data library,
+neutron transport code, and the purpose of the study. The present work uses
 a fuel salt composition of 77.5\% LiF - 19.913\% ThF$_4$ - 2.587\%
@@
 Density, $\rho$ [kg m$^{-3}$] & $4094-0.882 \cdot (T-1008)$ & $4125$ & 893-1123 K \\
-Dynamic viscosity, $\mu$ [Pa s] & $\rho \cdot 5.55 \times 10^{-8} \cdot e^{3689/T}$ & $1.015$ & 898-1121 K \\
-Thermal conductivity, $k$ [W m$^{-1}$ K$^{-1}$] & $0.928+8.397 \times 10^{-5} \cdot T$ & $0.01010$ & 891-1020 K \\
+Dynamic viscosity, $\mu$ [Pa s] & $\rho \cdot 5.55 \times 10^{-8} \cdot e^{3689/T}$ & $0.01015$ & 898-1121 K \\
+Thermal conductivity, $k$ [W m$^{-1}$ K$^{-1}$] & $0.928+8.397 \times 10^{-5} \cdot T$ & $1.010$ & 891-1020 K \\
 Specific heat, $c_p$ [J kg$^{-1}$ K$^{-1}$] & $-1111+2.78\cdot T$ & $1594$ & 867-907 K \\

### 2020-07-10 steadystate.tex
Katy comment edits

@@
-With the verification of Moltres' neutronics modeling capabilities in the
-context of the \gls{MSFR}, this chapter will cover the steady-state
-multiphysics simulation results from Moltres.
+Building on Chapter \ref{chap:nts} which verified Moltres' neutronics modeling
+capabilities in the context of the \gls{MSFR}, this chapter will cover the
+steady-state multiphysics simulation results of the \gls{MSFR} using Moltres.
+The steady-state results depend heavily on both neutronics and
+thermal-hydraulics solving capabilities in Moltres. Thus, this exercise is the
+first verification step for Moltres as a multiphysics \gls{MSR} simulation
+tool. This chapter will specifically discuss the temperature, velocity,
+neutron flux, and precursor distributions and compare them with data from the
+literature. The steady-state operating conditions will also serve as the
+initial conditions for the subsequent accident transient simulations.
 
-The procedure for obtaining the steady-state results involved several steps
+This chapter will first
+present a summary of the procedure for obtaining the steady-state results.
+Next, the chapter discusses the steady-state results without modeling decay
+heat in direct comparison with the steady-state results from the Polimi and
+TUDelft models \cite{aufiero_development_2014}. After this
+comparison, the chapter separately discusses the minor differences in the
+results from decay heat modeling in the last subsection.
+
+\section{Simulation Procedure}
+
+The procedure for obtaining the results for the steady-state operating
+conditions involved several steps
 due to the tightly coupled \glspl{PDE}. First, a preliminary transient
@@
 $1 \times 10^{14}$ cm$^{-2}$ s$^{-1}$ uniformly throughout the geometry.
-Finally, this work assumes that steady
-state is reached when the volume integral values of every variable remain
-constant (up to 6 sig. fig.) for at least four seconds in the simulation; this
+Finally, this work assumes that steady state is reached when the volume
+integral values of every variable remain constant (up to 6 significant
+figures) for at least four seconds in the simulation; this
 time period corresponds to the nominal circulation time of the \gls{MSFR}.
 
-For a direct comparison with the steady-state results from the Polimi and
-TUDelft models \cite{aufiero_development_2014}, this chapter will first
-present the steady-state results without modeling decay heat. After this
-comparison, the chapter separately discusses the minor differences in the
-results from decay heat modeling in the last subsection.
-
 \section{Steady-State Thermal-Hydraulics Results}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{.365\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{flow-temp}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.625\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{flow-temp-fiorina}
-    \end{subfigure}
-    \caption{Temperature and velocity fields in the core from Moltres
-    (left), Polimi (center), and TUDelft (right) models. The colors represent
-    temperature according to the respective colorbars and the arrows
-    represent velocity fields.}
-    \label{fig:flow-temp}
-\end{figure}
-
-\begin{figure}[t!]
-    \centering
-    \includegraphics[width=.6\textwidth]{flow}
-    \caption{Fuel salt flow streamlines and velocity magnitude in the core.
-    The colors represent velocity magnitude according to the colorbar on the
-    right.}
-    \label{fig:flow}
-\end{figure}
 
@@
 Polimi and TUDelft models predict that the turbulent viscosity in these
-regions is as high as 100 Pa s, much higher than our 40 Pa s approximation.
+regions is as high as 100 Pa$\cdot$s, much higher than our 40 Pa$\cdot$s
+approximation.
+
+\begin{figure}[htbp!]
+    \centering
+    \begin{subfigure}[t]{.365\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{flow-temp}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.625\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{flow-temp-fiorina}
+    \end{subfigure}
+    \caption{Temperature and velocity fields in the core from Moltres
+    (left), Polimi (center), and TUDelft (right) models. The colors represent
+    temperature according to the respective color bars and the arrows
+    represent velocity fields.}
+    \label{fig:flow-temp}
+\end{figure}
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.6\textwidth]{flow}
+    \caption{Fuel salt flow streamlines and velocity magnitude in the core.
+    The colors represent velocity magnitude according to the color bar on the
+    right.}
+    \label{fig:flow}
+\end{figure}
 
@@
 Although the temperatures at the hotspots are well below the melting point of the Ni-alloy structure (1500 K), they may cause undue thermal stress on the
-blanket tank structure and induce relatively faster salt corrosion rates. A
+blanket tank structure and accelerate salt corrosion rates. A
 sudden, large reactivity insertion could push fuel salt temperatures above the
@@
 the reservoir of hot fuel salt may cause unpredictable behavior during
-transient scenarios when the flow profile undergoes a drastic change.
+transient scenarios when the flow profile fluctuates.
 Thus, Rouch et al. \cite{rouch_preliminary_2014} developed an improved
-hourglass-shaped design to optimize flow distribution and prevent these
+hourglass-shaped design for the central core region to optimize flow
+distribution and prevent these
 recirculation zones and hotspots from forming. A study of this new design
@@
 central and horizontal axes, as expected of a cylindrical reactor design. The
-relatively lower temperatures near the center of the core promotes the neutron
-flux peaking but it is of little concern as no structural parts vulnerable to
-neutron damage exist in that region. The peak total flux at the center is
+lower temperatures near the center of the core promote the neutron
+flux peaking but it is of little concern as no neutronically vulnerable
+structures exist in that region. The peak total flux at the center is
 $9.80 \times 10^{15}$ cm$^{-2}\cdot$s$^{-1}$, which is close to values
@@
 peak flux value from this paper is slightly higher as this work used the
-steady-state temperature distribution while Fiorina et al. and Aufiero et al.
-imposed a uniform temperature distribution at 973 K.
+heterogeneous temperature distribution in Figure \ref{fig:flow-temp} while
+Fiorina et al. and Aufiero et al. imposed uniform temperature distributions
+at 973 K.
 
@@
 
-\clearpage
-
 \section{Steady-State Decay Heat Results}

### 2020-07-14 acks.tex
final comment edits

I wish to express my deepest gratitude to my advisor, Dr. Kathryn Huff, for
her guidance, patience, and encouragement in the past few years. My academic
journey thus far has been filled with many ups and downs and it would have
been much more difficult without her support. I would also like to thank Dr.
Tomasz Kozlowski for his help and invaluable comments as the second reader of
this thesis. I wish to acknowledge the funding support from the Singapore
Nuclear Safety \& Research Initiative since the start of my graduate studies.

I wish to thank my groupmates Dr. Andrei Rykhlevskii, Anshuman Chaube, and
Gwendolyn Chee for their input and ideas contributing to the completion of
this work. I also greatly appreciate Gavin Davis' and Matthew Kozak's help
with proofreading my writing.

Finally, no words can adequately express how thankful I am for my family's
unwavering support. I owe a lot to my parents and their efforts for all the
opportunities I've benefitted from.

### 2020-07-14 conclusion.tex
final comment edits

@@
-The unique phenomena in \glspl{MSR}
-necessitate the development of new reactor safety analysis software.
-This thesis presents the latest developments in Moltres, namely coupling its
-existing neutron diffusion module to the incompressible Navier-Stokes module
-in MOOSE, and developing a decay heat model for short-term transients. We
-demonstrated and verified some of its current capabilities through a static
-neutronics study, and a coupled neutronics/thermal-hydraulics safety analysis
+Unique phenomena in \glspl{MSR} necessitate the development of new reactor
+safety analysis software. This thesis presents the latest developments in
+Moltres, namely coupling its existing neutron diffusion module to the
+incompressible Navier-Stokes module in MOOSE and developing a decay heat model
+for short-term transients. This work demonstrated and verified some of
+Moltres' current capabilities through a static-model
+neutronics study and a coupled neutronics/thermal-hydraulics safety analysis
 of the \gls{MSFR} concept.
 
-The neutronics study showed good agreement between Moltres and Serpent. With
-the relevant group constant data from Serpent,
+The neutronics study showed good agreement between Moltres and Serpent 2. With
+the relevant group constant data from Serpent 2,
 Moltres could accurately replicate the $k_{eff}$, $\beta$,
-$\alpha_T$, and multi-group neutron flux results from Serpent. The
+$\alpha_T$, and multi-group neutron flux results from Serpent 2. The
 $k_{eff}$ estimates from Moltres were approximately 100 pcm higher for all
@@
 values from Moltres had 1.46\% and 0.265\% discrepancies, respectively, to
-Serpent's results. Lastly, the normalized six-group neutron flux from Moltres
-and Serpent were visually indistinguishable from each other. The journal
-article that introduced Moltres \cite{lindsay_introduction_2018} verified its
-neutron diffusion model for a two-group thermal-spectrum \gls{MSBR} model;
-the results of this study extends code-to-code verification of Moltres'
+Serpent 2's results. Lastly, the normalized six-group neutron fluxes from
+Moltres all had less than 1\% discrepancy compared with Serpent 2.
+The results of this study extend
+the code-to-code verification of Moltres'
 neutron diffusion model with the six-group, fast-spectrum \gls{MSFR} model.
 
-Although Moltres currently lacks a proper turbulence model, our simplifying
-assumption for the turbulent viscosity $\mu_t$ yielded good results for most
-of the \gls{MSFR}
+Although Moltres currently lacks a proper turbulence model, the uniform
+turbulent viscosity $\mu_t$ approximation was accurate enough to reproduce
+most of the expected results for the \gls{MSFR}
 steady-state and transient analyses. The steady-state temperature and velocity
 distributions showed many similiarities in their shapes and magnitudes to the
-Polimi and TUDelft model results \cite{fiorina_modelling_2014}. Our
-uniform $\mu_t$ assumption accounted for the minor differences in the flow at
+Polimi and TUDelft model results \cite{fiorina_modelling_2014}. The uniform
+$\mu_t$ approximation was responsible for the minor differences in the flow at
 the top of the core and the loss of delayed neutrons to out-of-core emissions.
@@
 The unprotected reactivity insertion and loss of heat sink results showed the
-same trends Fiorina et al. observed in their Polimi and TUDelft models. The
+same trends Fiorina et al. \cite{fiorina_modelling_2014} observed in their
+Polimi and TUDelft models. The
 small difference in the temperature reactivity coefficient accounted for the
@@
 models required minor adjustments in the relationship between flow rate
-$\dot{m}$ and the heat transfer coefficient $h$ from the original
+$\dot{V}$ and the heat transfer coefficient $h$ from the original
 Dittus-Boelter correlation for the pump-initiated transients. Assuming a
-directly proportional relationship between $\mu_t$ and $\dot{m}$ yielded
+directly proportional relationship between $\mu_t$ and $\dot{V}$ yielded
 results in good agreement with the other two models for the pump overspeed
@@
 effects on $\mu_t$.
-
-Overall, we have demonstrated that Moltres can handle most of the case
-studies that this thesis covered.
 
@@
 would allow us to move away from the 1D outer loop system and towards a full
-2D/3D closed loop. The biggest change is that users would be able to use
+2D/3D closed loop. If this were implemented, users would be able to use
 the Boussinesq approximation for buoyancy-driven flow capability in Moltres.
 Buoyancy-driven flow is a critical component in loss of forced flow scenarios
-and these scenarios in turn are important accident transients in reactor
-safety analyses.
+which are important to reactor safety analyses.
 
-Secondly, Moltres would benefit from a proper turbulence model such as the
-$k$-$\epsilon$ or $k$-$\omega$ turbulence models. Turbulence effects are
-significant in \gls{MSR} designs with fast flow, and they inform optimization
+Secondly, Moltres would benefit from a proper turbulence model such as
+$k$-$\epsilon$ \cite{jones_prediction_1972} or $k$-$\omega$
+\cite{wilcox_turbulence_2006}. Turbulence effects are
+significant in \gls{MSR} designs with fast flow and they inform optimization
 studies for improving flow patterns and eliminating local hotspots in eddies.
-Our work in this thesis, particularly for the loss of flow transient, shows
+The current work, particularly the loss of flow accident scenario, shows
 that simplifying assumptions for turbulence lead to erroneous results under
@@
 Lastly, a compressible Navier-Stokes model would be essential for modeling
-compressible flow effects such as variable temperature-dependent density
+effects such as variable temperature-dependent density
 changes following a large reactivity insertion and finite wave propagation
-speeds in a fluid. The presence of bubbles in the core from the gas sparging
+speeds in the salt. The presence of bubbles in the core from the gas sparging
 system increases fuel compressibility and enhances compressibility effects

### 2020-07-14 introduction.tex
final comment edits

@@
 the coupled neutronics/thermal-hydraulics steady-state and accident transient
-results against results from the Polimi and TUDelft models published by
-Fiorina et al. \cite{fiorina_modelling_2014}. The
+results against two sets of results published by
+Fiorina et al. \cite{fiorina_modelling_2014}. The two sets of results arose
+from a collaborative benchmarking exercise by researchers at Politecnico di
+Milano and Technical University of Delft with two separate \gls{MSR}
+simulation tools. Section \ref{sec:litrev} discusses these tools
+in greater detail. The
 secondary objective is to identify areas of improvement in Moltres for future
-research.
+development.
 

### 2020-07-14 model.tex
final comment edits

@@
 Serpent 2 provides standard geometric surfaces (e.g. planes, cylinders, cones)
-for defining reactor geometries. In this work, the reactor geometry is the
-same axisymmetric \gls{MSFR} geometry from the paper by Fiorina et al.
+for defining reactor geometries. In this work, the reactor geometry uses the
+same axisymmetric \gls{MSFR} geometry as the Polimi and TUDelft models
 \cite{fiorina_modelling_2014}, as shown in Fig. \ref{fig:msfrgeom}.
@@
     group $i$ [cm$^{-3}$],} \nonumber \\
-    K &= \text{turbulent diffusion of the delayed neutron precursors
-    [cm$^2\cdot$s$^{-1}$].} \nonumber
+    K &= \text{turbulent diffusion coefficient of the delayed neutron
+    precursors [cm$^2\cdot$s$^{-1}$].} \nonumber
 \end{align}
@@
 \begin{align}
-    u_x &= -\xi \Big[\frac{y}{h} - \Big(\frac{y}{h}\Big)^2 
+    u_x &= -\xi \Big[\frac{y}{H} - \Big(\frac{y}{H}\Big)^2 
     \Big] \\
@@
     y &= \text{height along the inlet [m],} \nonumber \\
-    h &= \text{total height of the inlet $= 0.1875$ m.} \nonumber
+    H &= \text{total height of the inlet $= 0.1875$ m.} \nonumber
 \end{align}
@@
 \begin{align}
-    v &= \xi [\frac{y}{h} - \frac{y^2}{h^2}] \\
+    v &= \xi [\frac{y}{H} - \frac{y^2}{H^2}] \\
     \dot{V} &= \int^h_0 \int^{2 \pi}_0 v r d\theta dy \\

### 2020-07-14 msr.tex
final comment edits

@@
 
-\section{Literature Review of MSR Simulation Tools}
+\section{Literature Review of MSR Simulation Tools} \label{sec:litrev}
 
@@
 Both DALTON and the COMSOL-based tool featured again in a later paper by
-Fiorina et al.
-\cite{fiorina_modelling_2014} in 2014 for a 2-D axisymmetric model of the
-\gls{MSFR}. The authors compared \gls{MSFR} simulations results from both
-tools, in which they dubbed the COMSOL and DALTON/HEAT tools the Polimi and
-TUDelft models, respectively, after the universities where each tool was
-developed. With multigroup
-neutron diffusion and \gls{RANS} formulations on ultra fine meshes, both
-models agreed on the steady-state neutron flux, temperature,
+Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D axisymmetric
+model of the \gls{MSFR}. The authors, from Politecnico di Milano and
+Technical University of Delft, compared \gls{MSFR} transient safety
+analysis results from both tools. They referred to the COMSOL and DALTON/HEAT
+\gls{MSFR} models as the Polimi and TUDelft models, respectively, after the
+universities where each application was developed. This thesis also follows
+this naming convention. With
+multigroup neutron diffusion and \gls{RANS} formulations on ultra fine meshes,
+both models agreed on the steady-state neutron flux, temperature,
 and \gls{DNP} distributions, as well as the power responses following various
-accident transient initiations. Aufiero et al. \cite{aufiero_development_2014}
-concurrently developed
+accident transient initiations.
+Aufiero et al. \cite{aufiero_development_2014} concurrently developed
 a full-core 3-D model of the \gls{MSFR} in OpenFOAM, albeit with one-group
@@
 $k$-$\epsilon$ turbulence model. The authors verified COUPLE's neutronics and
-thermal-hydraulics against MCNP and a FLUENT-based model, respectively, before
-presenting their multiphysics steady-state results with a 2-D axisymmetric
+thermal-hydraulics against MCNP and a FLUENT-based model, respectively, and
+presented their multiphysics steady-state results for a 2-D axisymmetric
 model of the \gls{MSFR}.
@@
 The present work uses the same 2-D square-cylindrical \gls{MSFR} design to
-compare our results to results published by Fiorina et al.
+compare the present results to results from the Polimi and TUDelft models
 \cite{fiorina_modelling_2014}. The design is a 2-D axisymmetric representation

### 2020-07-14 steadystate.tex
final comment edits

@@
 
-\section{Steady-State Thermal-Hydraulics Results}
+\section{Steady-State Thermal-Hydraulics Results} \label{sec:ss-th}
 
@@
 Inertial forces dominate over viscous forces to form this large eddy. The main
-difference between Moltres and Fiorina et al.'s models is the flow profile
-near the central axis at the top and bottom of the core.
+difference between Moltres, and the Polimi and TUDelft models is the flow
+profile near the central axis at the top and bottom of the core.
 The Polimi and TUDelft models predict relatively stagnant flow in these
@@
 The Polimi and TUDelft models
-feature greater \gls{DNP} retention in the stagnant regions within the core.
-This effect is less pronounced in the Moltres model. Less build-up
-of \glspl{DNP} at the top of the core occurs in Moltres because most of the
-\glspl{DNP} produced near the center of the core cannot enter the axial
-recirculation zone that only appears in Moltres. The figure also shows less
-build-up in the large recirculation zone near the blanket tank.
+feature greater \gls{DNP} retention in the stagnant regions in the core. This
+effect is less pronounced in the Moltres model, most notably at the top of the
+core near the central axis and in the recirculation zone adjacent to the
+blanket tank. As mentioned in Section \ref{sec:ss-th}, only the Moltres model
+features a small recirculation zone at the top of the core. This local flow
+pattern is responsible for diverting the shorter-lived \glspl{DNP} away from
+the top of the core and reducing the build-up in that region.
 
-The in-core delayed neutron fraction $\beta_c$ is an important safety
-parameter for \glspl{MSR}. This value represents the actual delayed
-neutron fraction in \glspl{MSR} after accounting for the loss of delayed
-neutrons from \glspl{DNP} decaying outside the active core region.
+The in-core delayed neutron fraction, which we denote as $\beta_c$, is an
+important safety parameter for \glspl{MSR}. This value represents the actual
+delayed neutron fraction in \glspl{MSR} after accounting for the loss of
+delayed neutrons from \glspl{DNP} decaying outside the active core region.
 Reactors with smaller $\beta$ values exhibit greater prompt jumps in the
@@
 proportion of prompt neutrons under normal operating conditions. This is
-undesirable from a reactor safety perspective because it exposes the reactor
-to relatively more extreme conditions before active safety mechanisms
+undesirable from a reactor safety perspective because it reduces the time
+available for active safety mechanisms to
 activate and scram the reactor. In \glspl{MSR}, this danger is partly
@@
 Table \ref{table:dnf} compares the fraction of out-of-core emissions and
-the $\beta_c$ values from Moltres with the Polimi and TUDelft models. We
-calculated the fraction of out-of-core emissions by finding the total amount
-of each \gls{DNP} group in the core and the outer loop, multiplying each total
-by their associated decay time constants $\lambda_i$, and calculating the
-proportion of emissions in the outer loop relative to the grand total. We
-calculated $\beta_c$ by first obtaining the prompt neutron emission rate from
-Moltres and subsequently using in-core delayed neutron emission rate from the
-previous calculation to find the fraction of delayed neutron emission rate
-relative to total emission rate.
+the $\beta_c$ values from Moltres with the Polimi and TUDelft models. The
+present author calculated the fraction of out-of-core emissions by finding the
+average concentration of each \gls{DNP} group in the core and the outer loop,
+multiplying each average by their associated decay time constants $\lambda_i$
+for the average emission rate, and calculating the
+proportion of emissions in the outer loop relative to the total sum. The
+present author calculated $\beta_c$ by first obtaining the prompt neutron
+emission rate from Moltres and subsequently using the in-core delayed neutron
+emission rate from the previous calculation to find the fraction of delayed
+neutron emission rate relative to total emission rate.
 
@@
 stagnant flow regions in the core; the \glspl{DNP} are more evenly distributed
-along the entire primary loop, leading to more delayed neutron emissions in
-the outer loop region. This is likely caused by differences in the
-flow pattern in the recirculation zone because convective transport
-dominates diffusive effects in the \gls{MSFR}. The exact flow pattern in the
-recirculation zones in the Polimi and TUDelft models is likely to differ from
+along the entire primary loop and a greater fraction of delayed neutron
+emissions occur in the outer loop region. The flow patterns are responsible
+for the distribution of shorter-lived \gls{DNP} since convection is the
+dominant mode of species transport in the \gls{MSFR}. The exact flow pattern
+in the recirculation zones in the Polimi and TUDelft models differs from
 that in Moltres (Figure \ref{fig:flow}). Figure \ref{fig:flow-temp} also shows
@@
 recirculation zone indicates that the recirculation zone in Moltres is
-relatively more stagnant. This could explain the concentration of \glspl{DNP}
+more stagnant. This could explain the concentration of \glspl{DNP}
 along an ``arc'' closer to the center of the core in Moltres as opposed to the
 more even distribution of \glspl{DNP} throughout the whole recirculation zones
-in the Polimi and TUDelft models. The higher peak \glspl{DNP} distribution in
+in the Polimi and TUDelft models. The higher peak \gls{DNP} distribution in
 Moltres also supports this assertion.
@@
 In spite of the greater delayed neutron losses, the $\beta_c$ value is higher
-in Moltres than the Polimi and TUDelft models. To account for this
-peculiarity, Fiorina et al. \cite{fiorina_modelling_2014} applied adjoint flux
+in Moltres than the Polimi and TUDelft models. Fiorina et al.
+\cite{fiorina_modelling_2014} applied adjoint flux
 weighting for their $\beta_c$ calculation whereas this work
@@
 The movement of decay heat precursors effectively redistributes a fraction of
-the volumetric heat source concentrated at the center of the core to the
-entire primary loop.
-Thus, the temperature distribution in the core should
+the fission heat source concentrated at the center of the core to other parts
+of the primary loop. Thus, the temperature distribution in the core should
 show less extreme temperatures compared with the simulation without decay
@@
 core for groups 1 to 3. The decay constants are $\lambda_j = 0.197$ s$^{-1},
-0.0168$ s$^{-1}, 0.000358$ s$^{-1}$ for groups 1, 2, and 3, respectively.
+0.0168$ s$^{-1}, 0.000358$ s$^{-1}$ for decay heat groups $j=$ 1, 2, and 3,
+respectively.
 Similar to the delayed neutron precursor distributions, the shorter-lived
@@
 relative to the core temperatures without decay heat in Figure
-\ref{fig:flow-temp}. This figure uses a different colormap scheme to clearly
-distinguish between temperature increases and decreases in the core. The
-hotspots in the recirculation zones are approximately
+\ref{fig:flow-temp}. This figure uses a different color map to clearly
+distinguish between positive and negative temperature differences in the core.
+The hotspots in the recirculation zones are approximately
 2 to 4 K cooler, while the cooled salt flowing in from the inlet is
@@
 the temperature near the outlet because the mesh near the boundary is too
-coarse. Users can avoid these oscillations near the outlet but using a mesh
-that is progressively finer the closer the element is to the boundary. The
-present author did not resolve this issue due to time constraints.
+coarse. Users can avoid these oscillations near the outlet by using a mesh
+that is progressively finer towards the boundary. The
+present author could not resolve this issue due to time constraints.
 

### 2020-07-14 transient.tex
Katy comment edits
final comment edits

@@
-This chapter discusses the transient multiphysics simulation results of the
-\gls{MSFR} from Moltres for four accident scenarios. These scenarios, adapted
-from Fiorina et al.'s work \cite{fiorina_modelling_2014}, include unprotected
-instances of reactivity insertion, loss of heat sink, loss of flow, and pump
-overspeed accidents. The term ``unprotected'' means no external interventions
-occur in these scenarios. As such, these simulations
-give an insight on the \gls{MSFR}'s passive safety capabilities in the absence
-of any active safety system. This work used the steady-state configuration
+This chapter discusses Moltres transient multiphysics simulations of the
+\gls{MSFR} in four accident scenarios. These scenarios, adapted from the
+\gls{MSFR} transient analyses with the Polimi and TUDelft models
+\cite{fiorina_modelling_2014}, include unprotected reactivity
+insertion, loss of heat sink, loss of flow, and pump overspeed accidents. The
+term ``unprotected'' signify accident scenarios without treactor SCRAM. As
+such, these simulations provide an insight on the
+\gls{MSFR}'s passive safety capabilities in the absence
+of active safety systems. This work used the steady-state configuration
 presented in the previous chapter as the initial conditions for the transient
@@
 modeling has a negligible effect in reactivity-, and pump-initiated
-transients. Furthermore, only their Polimi model had decay
-heat modeling capabilities. Therefore, they presented results from their
-Polimi and TUDelft model for the four accident scenarios without decay heat
-modeling and enabled decay heat modeling for the loss of heat sink transient
-scenario only. This work also ran all transient simulations without the decay
-heat model for a fair comparison. The only exception is loss of heat sink
-scenario in which two separate simulations with and without the decay heat
-model were run. More generally, this work imposed various other simplifying
-assumptions in our transient models to match their implementations as closely
-as possible, within Moltres' capabilities. The details of the setup for each
-transient simulation are in their respective sections.
+transients. Furthermore, only their Polimi model had this capability.
+Therefore, they featured decay heat modeling for the loss of heat sink
+accident scenario only. The current work also ran all transient simulations
+without the decay heat model for a fair comparison. The only exception is the
+loss of heat sink scenario in which two separate simulations with and without
+the decay heat model were run. This work imposed simplifying assumptions in
+our transient models to match the implementations in
+\cite{fiorina_modelling_2014} as closely as possible, within Moltres'
+capabilities. The details of the setup for each
+transient simulation appear in the following sections.
 
 \section{Unprotected Reactivity Insertion}
+In reactivity insertion accidents, excess reactivity would drive an increase
+in flux, power, and temperature. In
+\glspl{MSR}, a positive reactivity insertion could occur if the online
+refueling system injected excess fissile material into the core. Excessively
+high neutron fluences and temperatures could negatively impact reactor
+structural integrity and increase the risk of containment breach.
 
-Reactivity insertion accidents are a type of nuclear accident caused by
-unintended positive
-reactivity insertions. The excess reactivity causes the power output and
-temperatures in nuclear reactors to rise to potentially dangerous levels. In
-\glspl{MSR}, a positive reactivity insertion could occur when the online
-refueling system injects excess fissile material into the core. Excessively
-high neutron fluences and temperatures negatively impact reactor structural
-integrity and increase the risk of containment breach.
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.7\textwidth]{insertion}
+    \caption{Step-wise 50 pcm and 200 pcm reactivity insertions used to
+    initiate the accident transients.}
+    \label{fig:insertion}
+\end{figure}
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.7\textwidth]{50pcm-jump}
+    \caption{Power output during the prompt response following a 50 pcm
+    step-wise unprotected reactivity insertion in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:50pcmjump}
+\end{figure}
 
@@
 models was increased by increasing the $^{233}$U-to-$^{232}$Th ratio in
-the fuel salt.
+the fuel salt. Figure \ref{fig:insertion} shows the step-wise reactivity
+insertions used to model the accident scenarios.
 
-The focus of this transient study is the neutronic and thermal-hydraulic
-behavior of the reactor core. Thus, this work assumes that the heat exchanger
+The neutronic and thermal-hydraulic behaviors of the reactor core are the
+focus of this transient study. Thus, this work assumes that the heat exchanger
 and the associated power generation equipment (generator turbines, heat sinks,
@@
 
+Figure \ref{fig:50pcmjump} shows the rise in power output during
+the initial prompt response. The prompt
+response raises the power output to 4 GW by $t=0.001$ s.
 Figures \ref{fig:50pcmheat} and \ref{fig:50pcmtemp} show the power output and
-average core temperature increase following the 50 pcm step-wise reactivity
-insertion in the Moltres, Polimi, and TUDelft models. The initial prompt
-response to the reactivity insertion raises power to 4 GW by $t=0.001$ s.
-Figure \ref{fig:50pcmjump} shows the rise in power output specifically during
-the prompt response in a separate plot.
+average core temperature increase beyond $t=0.001$ s following the 50 pcm
+step-wise reactivity insertion in the Moltres, Polimi, and TUDelft models.
 Power continues to rise at a slower rate up to 4.63 GW at around $t=0.005$ s,
@@
 than the initial +50 pcm insertion. Power continues to fall as the average
-core temperature rises. The slight change in slope occurs at $t=0.3$ s. The
+core temperature rises. A slight change in slope occurs at $t=0.3$ s. The
 elapsed time approximately corresponds to the average half-life of the two
@@
 negates a fraction of the negative reactivity from the elevated core
-temperature. By $t=3$-$4$ s, most of the heated salt and \glspl{DNP} from the
+temperature. By $t=3$ s, most of the heated salt and \glspl{DNP} from the
 initial phase will have circulated around the primary loop and returned to the
 core. The heated salt causes a small, noticeable dip in power before the power
-output stabilises. The core tends to a new equilibrium average temperature
+output stabilizes.
+
+The average core temperature rises steadily from the start of the transient
+until $t=3$ s when the heated salt from the initial phase returns to the core.
+This event is characterized by the small peak in the average core temperature
+at $t=3$ s. The subsequent drop in power output halts the temperature
+increase and the core tends to a new equilibrium average temperature
 approximately 7.5 K higher than the initial average temperature.
@@
     \includegraphics[width=.85\textwidth]{50pcm-heat}
-    \caption{Power output following
-    a 50 pcm step-wise reactivity insertion in the Moltres, Polimi, and
+    \caption{Power output following a 50 pcm step-wise unprotected reactivity
+    insertion in the Moltres, Polimi, and
     TUDelft models \cite{fiorina_modelling_2014}.}
@@
     \includegraphics[width=.85\textwidth]{50pcm-temp}
-    \caption{Average core temperature increase following
-    a 50 pcm step-wise reactivity insertion in the Moltres, Polimi, and
+    \caption{Average core temperature increase following a 50 pcm step-wise
+    unprotected reactivity insertion in the Moltres, Polimi, and
     TUDelft models \cite{fiorina_modelling_2014}.}
     \label{fig:50pcmtemp}
-\end{figure}
-
-\begin{figure}[htbp!]
-    \centering
-    \includegraphics[width=.7\textwidth]{50pcm-jump}
-    \caption{Power output during the prompt response following
-    a 50 pcm step-wise reactivity insertion in the Moltres, Polimi, and
-    TUDelft models \cite{fiorina_modelling_2014}.}
-    \label{fig:50pcmjump}
 \end{figure}
@@
     \includegraphics[width=.85\textwidth]{200pcm-heat}
-    \caption{Power output following
-    a 200 pcm step-wise reactivity insertion in the Moltres, Polimi, and
+    \caption{Power output following a 200 pcm step-wise unprotected reactivity
+    insertion in the Moltres, Polimi, and
     TUDelft models \cite{fiorina_modelling_2014}.}
@@
     \includegraphics[width=.85\textwidth]{200pcm-temp}
-    \caption{Average core temperature increase following
-    a 200 pcm step-wise reactivity insertion in the Moltres, Polimi, and
+    \caption{Average core temperature increase following a 200 pcm step-wise
+    unprotected reactivity insertion in the Moltres, Polimi, and
     TUDelft models \cite{fiorina_modelling_2014}.}
@@
 temperature increases much more rapidly and subsequently triggers a sharper
-drop in power output. This results in the clearer distinction in the rate of
-core temperature increase before and after $t=0.01$ s. In this transient, we
-also observe greater deviation between Moltres and the other models arising
-from the differences in the temperature reactivity coefficients. Overall,
+drop in power output. This more clearly distinguishes the rate of
+core temperature increase before and after $t=0.01$ s. The difference in
+$\alpha_T$ causes greater deviations in the results between Moltres and the
+other models. Overall,
 Moltres' results show good agreement with the Polimi and TUDelft results.
@@
 An unprotected loss of heat sink accident can occur when the pumps in the
-intermediate loop fail. The heat exchangers would then lose most of their
+intermediate loop fail and reactor is not SCRAMed. The heat exchangers would
+then lose most of their
 cooling capabilities. This work followed Fiorina et al.'s approach in assuming
@@
 
-Decay heat from fission products poses a great safety risk in a loss of
+Decay heat from fission products poses a great safety risk in an unprotected
+loss of
 heat sink accident. Section \ref{sec:wodecayheat} showed that prompt fission
@@
 heat source from $t=34$ s and falls at a much slower rate than prompt heat.
-Figure \ref{fig:moltresdecaytemp} highlights the
-greater core temperature increase attributed to decay heat as compared with
+Figure \ref{fig:moltresdecaytemp} highlights the greater core temperature
+increase arising from decay heat as compared with
 the results without decay heat. By $t=120$ s, the model with decay heat
 records an average core temperature increase that is 45 K higher than the
-model without decay heat. The average core temperature reaches approximately
+model without decay heat. The absolute core temperature reaches approximately
 1220 K and would continue to rise further. This places undue thermal stress
-and accelerate corrosion rates in the Hastelloy structural material. In the
+and accelerates salt-induced corrosion in the Hastelloy structural material.
+In the
 absence of an auxiliary heat removal system in the primary loop, reactor
@@
 tank with emergency cooling systems to keep the salt cool.
-
-\begin{figure}[htbp!]
-    \centering
-    \includegraphics[width=.85\textwidth]{moltres-decay-power}
-    \caption{Power output during
-    an unprotected loss of heat sink transient in the Moltres model with and
-    without decay heat.}
-    \label{fig:moltresdecaypower}
-    \includegraphics[width=.85\textwidth]{moltres-decay-temp}
-    \caption{Average core temperature increase during
-    an unprotected loss of heat sink transient in the Moltres model with and
-    without decay heat.}
-    \label{fig:moltresdecaytemp}
-\end{figure}
-
-\begin{figure}[htbp!]
-    \centering
-    \includegraphics[width=.85\textwidth]{decay-temp}
-    \caption{Loop-averaged temperature increase during
-    an unprotected loss of heat sink transient in the Moltres and Polimi
-    models \cite{fiorina_modelling_2014} with decay heat.}
-    \label{fig:polimidecaytemp}
-\end{figure}
 
@@
 
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.85\textwidth]{moltres-decay-power}
+    \caption{Power output during
+    an unprotected loss of heat sink transient in the Moltres model with and
+    without decay heat.}
+    \label{fig:moltresdecaypower}
+    \includegraphics[width=.85\textwidth]{moltres-decay-temp}
+    \caption{Average core temperature increase during
+    an unprotected loss of heat sink transient in the Moltres model with and
+    without decay heat.}
+    \label{fig:moltresdecaytemp}
+\end{figure}
+
 \clearpage
-
-\section{Unprotected Loss of Flow}
-
-A loss of forced flow transient can occur in the event of a station
-blackout; the pumps would cease operating due to the loss of AC electrical
-power. Natural circulation resulting from temperature-dependent density
-changes becomes the dominant driving force for salt flow in the primary loop.
-Fiorina et al. \cite{fiorina_modelling_2014} applied the Boussinesq
-approximation for buoyancy-driven flow in their models, but this approach was
-not possible in Moltres because the primary loop is partitioned into two
-separate geometries and used Dirichlet boundary
-conditions at the inlet to drive flow. Fiorina et al.'s Polimi and TUDelft
-models featured complete exponential coast-downs of the pumps with a time
-constant of 5 s. The resulting flow rate $\dot{m}$ from natural circulation
-was approximately 18 times smaller than the initial $\dot{m}$. Figure
-\ref{fig:flowrate} shows that the actual $\dot{m}$ decreased with a time
-constant of 8 s. Thus, for the \gls{MSFR} model in Moltres, this work imposed
-a similar exponential decay term with a time constant of 8 s on the inflow
-Dirichlet boundary condition:
-%
-\begin{align}
-    \text{Flow rate, } v = 0.25862 + (4.5-0.25862) e^{-t/8} 
-    \label{eq:flowrate}
-\end{align}
 
@@
     \centering
-    \includegraphics[width=.7\textwidth]{lof-flow-rate}
-    \caption{The change in flow rate in the Polimi and TUDelft models and the
-    imposed flow rate in Moltres.}
-    \label{fig:flowrate}
+    \includegraphics[width=.85\textwidth]{decay-temp}
+    \caption{Loop-averaged temperature increase during
+    an unprotected loss of heat sink transient in the Moltres and Polimi
+    models \cite{fiorina_modelling_2014} with decay heat.}
+    \label{fig:polimidecaytemp}
 \end{figure}
-
-The reduced $\dot{m}$ also decreases the heat transfer rate between the
-primary and intermediate loop through the heat exchanger as the heat transfer
-coefficient $h$ is dependent on the $\dot{m}$. This step was problematic
-because
-the pointwise heat exchanger implementation in Moltres performs differently
-compared with the heat exchangers that take up ``36\% of the out-of-core
-part'' in the Polimi and TUDelft models \cite{fiorina_modelling_2014}. Most of
-the cooling happens in the top half of the heat exchanger where the
-temperature differential between the primary and intermediate loops is the
-largest. In the Polimi and TUDelft models, the overall $h$ is the
-``harmonic mean of the heat transfer coefficients on each side of the heat
-exchanger''. For this loss of flow transient, the authors intended to focus on
-the primary loop and assumed that only the pumps in the primary loop failed.
-In addition to this, the authors applied the Dittus-Boelter correlation
-\cite{dittus_heat_1930} for the relationship between the primary side $h$ and
-$\dot{m}$. The Dittus-Boelter correlation for fluids being cooled is:
-%
-\begin{align}
-    Nu &= 0.023 Re^{0.8} Pr^{0.3} \label{eq:db} \\
-    \intertext{where}
-    Nu &= \text{ Nusselt number,} \nonumber \\
-    Re &= \text{ Reynolds number,} \nonumber \\
-    Pr &= \text{ Prandtl number.} \nonumber
-\end{align}
-%
-The only direct relation to $\dot{m}$ in the Dittus-Boelter correlation is
-through the $Re$ term, which is directly proportional to flow velocity $v$.
-This gives the following relation between $h$ and$v$:
-%
-\begin{align}
-    h \propto v^{0.8} \label{eq:hv}
-\end{align}
-%
-However, this relation provided very different results in the unprotected loss
-of flow and pump overspeed transients compared with the Polimi and TUDelft
-models; it overpredicted the equilibrium power output in the
-unprotected loss of flow transient and underpredicted the same parameter in
-the unprotected pump overspeed transient. Upon further investigation, the
-present author found that raising the power of $v$ from 0.8 to 1.1 brought the
-average core temperatures closer to the results from the other models in both
-transients. Therefore, this work adopted the raised power in this study.
-
-Another issue pertained to the turbulent viscosity $\mu_t$ as a function of
-$v$. Using a simple approximation of $\mu_t$ being directly
-proportional to $v$, the results differed significantly compared with the
-Polimi and TUDelft models. This is likely due to buoyancy-driven flow
-contributing to turbulence; the turbulent energy $k$ equation in COMSOL's
-$k$-$\epsilon$ model has an explicit source term from buoyancy effects
-\cite{comsol_ab_comsol_2018}. Another point to note is the
-Reynolds number remains constant if $\mu$ and $v$ decrease in tandem. This
-preserves the existence of the recirculation zone in the core and it is at
-odds with the results from the Polimi and TUDelft models, which show that the
-recirculation zones disappear during the loss of flow transient. We
-circumvented this issue by letting fixed fractions of the initial $\mu_{t,0}$
-be conserved regardless of the final flow velocity, according to the following
-equation:
-%
-\begin{align}
-    \mu_t &= \mu_c + (\mu_{t,0} - \mu_c) e^{-t/8} \label{lofmu} \\
-    \intertext{where}
-    \mu_c &= \text{ conserved fraction of $\mu_{t,0}$ [Pa$\cdot$s].} \nonumber
-\end{align}
-%
-This measure allowed for laminar flow to develop in the core and yielded
-results with closer resemblance to those from the Polimi and TUDelft models.
-
-Figures \ref{fig:lofheat} and \ref{fig:loftemp} show the power output and
-average core temperature increase during the unprotected loss of flow
-transient in the Moltres, Polimi, and TUDelft models without decay heat
-modeling. The three sets of results from Moltres correspond to $\mu_c =
-\frac{1}{4} \mu_{t,0}, \frac{1}{2} \mu_{t,0}, \text{and } \frac{3}{4}
-\mu_{t,0}$. Moltres
-performs poorer in this transient relative to the two previous transients.
-Although Moltres shows the same decreasing trend in power output, it failed to
-capture the exact individual features in the reactor response. In the Polimi 
-and TUDelft models, Fiorina et al.
-stated that after around $t=15$ s, the ``flow pattern changed in the core and
-the recirculation zones started to disappear''. A sudden drop in the average
-core temperature results as the pocket of hot salt leaves the
-core. In Moltres, the wider peak in the average core temperature
-indicates that there was a more gradual change in the flow pattern.
-
-\begin{figure}[htbp!]
-    \centering
-    \includegraphics[width=.85\textwidth]{lof-heat}
-    \caption{Power output during
-    an unprotected loss of flow transient in the Moltres, Polimi, and
-    TUDelft models \cite{fiorina_modelling_2014}.}
-    \label{fig:lofheat}
-    \includegraphics[width=.85\textwidth]{lof-temp}
-    \caption{Average core temperature increase during
-    an unprotected loss of flow transient in the Moltres, Polimi, and
-    TUDelft models \cite{fiorina_modelling_2014}.}
-    \label{fig:loftemp}
-\end{figure}
-
-\begin{figure}[htbp!]
-    \centering
-    \includegraphics[width=.37\textwidth]{lof-flow-temp}
-    \includegraphics[width=.62\textwidth]{fiorina-lof}
-    \caption{Temperature and velocity fields in the core at $t=300$ s during
-    a loss of flow transient in the Moltres ($\mu_c = \frac{1}{2} \mu_{t,0}$),
-    Polimi, and TUDelft models.}
-    \label{fig:lofflowtemp}
-\end{figure}
-
-Figure \ref{fig:lofflowtemp} shows the flow patterns and temperature
-distribution in the core at $t=300$ s in all three models. Figures
-\ref{fig:lofheat}, \ref{fig:loftemp}, and \ref{fig:lofflowtemp} combined
-highlight the difference between laminar flow in the Moltres model and
-buoyancy-driven flow in the other models, and its impact on the reactor
-response. They show that low-speed laminar flow is a poor substitute for
-buoyancy-driven flow in the context of the MSFR. It is particularly evident
-in the transition from high-speed turbulent flow to low-speed viscous
-flow as Moltres mispredicted the intermediate stages. The
-simplifying assumption for the uniform, time-dependent $\mu_t$ is also flawed
-in a safety analysis code.
-
-The results from this transient inform our goals for Moltres: 1)
-implementing a proper turbulence model, and
-2) developing a new heat exchanger feature that is compatible with the
-buoyancy-driven flow capabilities already present in Moltres.
 
@@
 core temperatures to fall to counteract this change through temperature
-reactivity feedback. The increased $\dot{m}$ also enhances the heat transfer
+reactivity feedback. The increased $\dot{V}$ also enhances the heat transfer
 coefficient on the primary loop side of the heat exchanger and enables the
@@
 
-This workfollowed Fiorina et al.'s implementation
+This work followed Fiorina et al.'s implementation
 \cite{fiorina_modelling_2014} by
@@
 For this transient, this work assumed that $\mu_t$ was directly proportional
-to $v$ because the buoyancy effects are negligible and the recirculation zones 
-persist throughout the entire duration. Figures \ref{fig:poheat} and
+to $v$ because the buoyancy effects are assumed to be negligible with forced
+flow and the recirculation zones 
+persist throughout the entire duration.
+
+Setting the exact dependence between the heat transfer coefficient $h$ and
+$\dot{V}$ was problematic because
+the pointwise heat exchanger implementation in Moltres performs differently
+compared with the heat exchangers of finite volume in the Polimi and TUDelft
+models \cite{fiorina_modelling_2014}. In a heat exchanger of finite volume,
+most of the cooling occurs in the first half of the heat exchanger where the
+temperature differential between the primary and intermediate loops is the
+largest. In the Polimi and TUDelft models, the overall $h$ is a
+harmonic mean of the $h_i$ on each side of the heat exchanger, given as:
+%
+\begin{align}
+    h &= \frac{2}{\frac{1}{h_1} + \frac{1}{h_2}}, \\
+    \intertext{where}
+    h &= \text{overall heat transfer coefficient
+    [W$\cdot$K$^{-1}$],} \nonumber \\
+    h_1 &= \text{heat transfer coefficient on the primary loop side
+    [W$\cdot$K$^{-1}$],} \nonumber \\
+    h_2 &= \text{heat transfer coefficient on the intermediate loop side
+    [W$\cdot$K$^{-1}$].} \nonumber
+\end{align}
+%
+In addition to this, Fiorina et al. applied the Dittus-Boelter correlation
+\cite{dittus_heat_1930} for the relationship between $h_1$ and
+$\dot{V}$. The Dittus-Boelter correlation for fluids being cooled is:
+%
+\begin{align}
+    Nu &= 0.023 Re^{0.8} Pr^{0.3}, \label{eq:db} \\
+    \intertext{where}
+    Nu &= \text{Nusselt number,} \nonumber \\
+    &\text{\ \ \ \ \ the ratio of convective to conductive heat
+    transfer at a boundary in a fluid,} \nonumber \\
+    Re &= \text{Reynolds number,} \nonumber \\
+    &\text{\ \ \ \ \ the ratio of inertial forces to viscous
+    forces within a fluid,} \nonumber \\
+    Pr &= \text{Prandtl number,} \nonumber \\
+    &\text{\ \ \ \ \ the ratio of momentum diffusivity to thermal
+    diffusivity.} \nonumber
+\end{align}
+%
+The only direct relation to $\dot{V}$ in the Dittus-Boelter correlation is
+through the Reynolds number, $Re$, which is directly proportional to
+flow velocity $v$. This gives the following relation between $h$ and $v$:
+%
+\begin{align}
+    h &\propto v^{0.8} \label{eq:hv} \\
+    h &= \text{heat transfer coefficient [W$\cdot$K$^{-1}$],} \nonumber \\
+    v &= \text{flow velocity [m$\cdot$s$^{-1}$].} \nonumber
+\end{align}
+%
+However, this relation provided very different results in the unprotected pump
+overspeed and loss of flow transients compared with the Polimi and TUDelft
+models. This approach underpredicted the equilibrium power output in the
+unprotected pump overspeed transient and overpredicted the same parameter in
+the unprotected loss of flow transient. Upon further investigation, I
+found that raising the power of $v$ from 0.8 to 1.1 brought the
+average core temperatures closer to the results from the other models in both
+transients. Therefore, this work adopted the raised power in this thesis.
+
+Figures \ref{fig:poheat} and
 \ref{fig:potemp} show the power output and average core temperature increase
@@
 average core temperature has a strictly decreasing trend, the temperature at
-the center of the core briefly rises due ot the sharp increase in power
+the center of the core briefly rises due to the sharp increase in power
 output. Since this is the region where most of the fissions take place, the
@@
 
-The results show excellent agreement with the Polimi and TUDelft models. The
-average core temperature increase in particular reproduces the Polimi and   
-TUDelft results very well and the curve falls between the other two curves.
-The power output is higher because the Moltres \gls{MSFR} model has a
-stronger negative temperature coefficient than the other two models.
+In both sets of results, Moltres reproduced all of the transient features
+found in the Polimi and TUDelft models. The average core temperature profile
+falls between the Polimi and TUDelft results, while the power output is
+approximately 0.1 GW higher because the Moltres \gls{MSFR} model has a
+greater $\alpha_T$ than the other two models.
+
+\section{Unprotected Loss of Flow}
+
+A loss of forced flow transient can occur in the event of a station blackout
+without reactor SCRAM; the pumps would cease operating due to the loss of AC
+electrical
+power. Natural circulation resulting from temperature-dependent density
+changes becomes the dominant driving force for salt flow in the primary loop.
+Fiorina et al. \cite{fiorina_modelling_2014} applied the Boussinesq
+approximation for buoyancy-driven flow in the Polimi and TUDelft models, but
+this approach was not possible in Moltres because the primary loop is
+partitioned into two separate geometries and used Dirichlet boundary
+conditions at the inlet to drive flow. The Polimi and TUDelft
+models featured complete exponential coast-downs of the pumps with a time
+constant of 5 s. The final flow rate $\dot{V}_f$ from natural circulation
+was approximately 18 times smaller than the initial $\dot{V}_0$. Figure
+\ref{fig:flowrate} shows that the actual $\dot{V}$ decreased with a time
+constant of 8 s. Thus, for the \gls{MSFR} model in Moltres, this work imposed
+a similar exponential decay term with a time constant of 8 s on the inflow
+Dirichlet boundary condition:
+%
+\begin{align}
+    \dot{V} &= \dot{V}_f + (\dot{V}_0-\dot{V}_f) e^{-t/8},
+    \label{eq:flowrate} \\
+    \intertext{where}
+    \dot{V}_f &= \text{final volumetric flow rate} = 0.25862
+    \text{ m$^3\cdot$s$^{-1}$,} \nonumber \\
+    \dot{V}_0 &= \text{initial volumetric flow rate} = 4.5
+    \text{ m$^3\cdot$s$^{-1}$.} \nonumber
+\end{align}
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.7\textwidth]{lof-flow-rate}
+    \caption{The change in flow rate in the Polimi and TUDelft models and the
+    imposed flow rate in Moltres.}
+    \label{fig:flowrate}
+\end{figure}
+%
+The reduced $\dot{V}$ also decreases the heat transfer rate between the
+primary and intermediate loop through the heat exchanger as the heat transfer
+coefficient $h$ is dependent on the $\dot{V}$. For this loss of flow
+transient, Fiorina et al. \cite{fiorina_modelling_2014} intended to focus
+on the primary loop and assumed that only the pumps in the primary loop
+failed.
+
+An issue arose pertaining to the turbulent viscosity $\mu_t$ as a function of
+$v$. When using a simple approximation of $\mu_t$ being directly proportional
+to $v$, the results from Moltres differed significantly compared with the
+Polimi and TUDelft models. The difference is due to buoyancy-driven flow
+contributing to turbulence; the turbulent energy $k$ equation in COMSOL's
+$k$-$\epsilon$ model has an explicit source term from buoyancy effects
+\cite{comsol_ab_comsol_2018}. Another point to note is the
+Reynolds number remains constant if $\mu$ and $v$ decrease in tandem. This
+preserves the existence of the recirculation zone in the core and it is at
+odds with the results from the Polimi and TUDelft models, which show that the
+recirculation zones disappear during the loss of flow transient. The current
+work circumvented this issue by letting fixed fractions of the initial
+$\mu_{t,0}$ be conserved regardless of the final flow velocity, according to
+the following equation:
+%
+\begin{align}
+    \mu_t &= \mu_c + (\mu_{t,0} - \mu_c) e^{-t/8} \label{lofmu} \\
+    \intertext{where}
+    \mu_c &= \text{ conserved fraction of $\mu_{t,0}$ [Pa$\cdot$s].} \nonumber
+\end{align}
+%
+This measure allowed for laminar flow to develop in the core and yielded
+results showing better qualitative agreement with those from the Polimi and
+TUDelft models. The subsequent paragraphs discuss these results.
+
+Figures \ref{fig:lofheat} and \ref{fig:loftemp} show the power output and
+average core temperature increase during the unprotected loss of flow
+transient in the Moltres, Polimi, and TUDelft models without decay heat
+modeling. The three sets of results from Moltres correspond to $\mu_c =
+\frac{1}{4} \mu_{t,0}, \frac{1}{2} \mu_{t,0}, \text{and } \frac{3}{4}
+\mu_{t,0}$.
+Although Moltres shows the same decreasing trend in power output, it failed to
+capture the exact individual features in the reactor response. In the Polimi 
+and TUDelft models, Fiorina et al. \cite{fiorina_modelling_2014}
+stated that after around $t=15$ s, the ``flow pattern changed in the core and
+the recirculation zones started to disappear''. The pocket of hot salt leaves
+the core and causes a sudden drop in the average core temperature. In Moltres,
+the wider peak in the average core temperature
+indicates that there was a more gradual change in the flow pattern.
+
+Figure \ref{fig:lofflowtemp} shows the flow patterns and temperature
+distribution in the core at $t=300$ s in all three models. Figures
+\ref{fig:lofheat}, \ref{fig:loftemp}, and \ref{fig:lofflowtemp} combined
+highlight the difference between laminar flow in the Moltres model and
+buoyancy-driven flow in the other models, and its impact on the reactor
+response. They show that low-speed laminar flow is a poor substitute for
+buoyancy-driven flow in the context of the MSFR. It is particularly evident
+in the transition from high-speed turbulent flow to low-speed viscous
+flow as Moltres mispredicted the intermediate stages. As for $\mu_t$, the need
+to fine-tune this parameter confirms that the uniform, function-based $\mu_t$
+approach is inappropriate for safety analysis in a loss of flow accident.
+
+The results from this transient inform our goals for Moltres: 1)
+implementing a proper turbulence model, and
+2) developing a new heat exchanger feature that is compatible with the
+buoyancy-driven flow capabilities already present in Moltres.
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.85\textwidth]{lof-heat}
+    \caption{Power output during
+    an unprotected loss of flow transient in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:lofheat}
+    \includegraphics[width=.85\textwidth]{lof-temp}
+    \caption{Average core temperature increase during
+    an unprotected loss of flow transient in the Moltres, Polimi, and
+    TUDelft models \cite{fiorina_modelling_2014}.}
+    \label{fig:loftemp}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.37\textwidth]{lof-flow-temp}
+    \includegraphics[width=.62\textwidth]{fiorina-lof}
+    \caption{Temperature and velocity fields in the core at $t=300$ s during
+    a loss of flow transient in the Moltres ($\mu_c = \frac{1}{2} \mu_{t,0}$),
+    Polimi, and TUDelft models.}
+    \label{fig:lofflowtemp}
+\end{figure}

### 2020-07-16 msr.tex
minor edit

@@
 uniform temperature of 973 K \cite{brovchenko_neutronic_2019} or for other
-motivations \cite{fiorina_molten_2013} \cite{pettersen_coupled_2016}. Thus,
+purposes \cite{fiorina_molten_2013, pettersen_coupled_2016}. Thus,
 the exact actinide compositions vary depending on the nuclear data library,

### 2020-07-20 abstract.tex
final comments and edits from katy

@@
-Molten salt reactors is a class of advanced nuclear reactors that promise
+Molten salt reactors, a class of advanced nuclear reactors, promise
 numerous improvements over the current fleet of largely light-water reactors.
@@
 transient multiphysics simulations of the Molten Salt Fast Reactor. The
-transient scenarios under study were unprotected instances of reactivity
-insertion, loss of heat sink, loss of flow, and pump overspeed. The steady
-state and transient results were verified against results from another paper
-that presented results for the same cases. The steady-state temperature and
+transient scenarios under study include unprotected instances of reactivity
+insertion, loss of heat sink, loss of flow, and pump overspeed. This thesis
+verified the steady state and transient results against data from the
+literature for the same case studies. The steady-state temperature and
 velocity distributions, and the peak neutron flux showed good agreement with

### 2020-07-20 model.tex
final comments and edits from katy

@@
 geometry of the reactor. This work uses Serpent 2 \cite{leppanen_serpent_2014}
-for the former and Trelis/Cubit \cite{noauthor_trelis_2018} for the latter.
+for the former and Trelis/CUBIT \cite{noauthor_trelis_2018} for the latter.
 This chapter provides brief introductions to
@@
 functionalities. As a result, \gls{MOOSE} supports adaptive meshing schemes
-and automatic variable scaling amongst other advanced featuresin finite
+and automatic variable scaling amongst other advanced features in finite
 element analysis. Full
@@
 on all of the relevant boundaries. Figure \ref{fig:msfrbc} shows the locations
-of the various boundaries listed in the table. This thesis introduces the
-CoupledOutflow boundary condition for $C_i$ and $\omega_j$ that allows users
-to couple these variables to the outlet velocity components (e.g. $u_x$,
-$u_y$). Without this new feature, users could only use uniform or fixed
+of the various boundaries listed in the table. The CoupledOutflow boundary
+condition for $C_i$ and $\omega_j$ is a new feature in Moltres that allows
+users to couple these variables to the outlet velocity components (e.g. $u_x$,
+$u_y$). Without this boundary condition, users could only use uniform or fixed
 function-based velocity profiles in conjunction with the precursor looping
@@
 s$^{-1}$ for an out-of-core residence time of 2 s. The present author derived
-these parameters from the reference specifications of 4 s cycle
+these parameters from the reference specifications of 4s cycle
 time and 50\% out-of-core salt fraction (Table \ref{table:msfr}).
@@
 A constant velocity of 1.1275 m s$^{-1}$ is applied in the outer loop region
-to maintain the nominal 2 s out-of-core residence time. The governing equation
+to maintain the nominal 2s out-of-core residence time. The governing equation
 for temperature, derived from equation \ref{eq:temp}, is:
@@
 connected to in the actual reactor geometry. The outlet boundary conditions
-are all outflow boundary conditions as shown in the table.
+are all outflow boundary conditions as shown in Table \ref{table:loopbc}.
 

### 2020-07-20 msr.tex
final comments and edits from katy

@@
 
-Two other \gls{MSR} designs emerged in the 1970s: the \gls{DMSR} from
-\gls{ORNL} \cite{gehin_liquid_2016} and the \gls{MCFR} from the UK Atomic
-Energy Authority \cite{smith_assessment_1974}. \gls{ORNL}
-researchers developed the \gls{DMSR}, named after its planned use of denatured
-$^{235}$U fuel (i.e. low-enriched uranium), as a once-through converter system
-without fuel reprocessing in response to fuel reprocessing restrictions
-introduced by President Ford in 1976. On the other hand, UK researchers
-developed the lead-cooled \gls{MCFR} to enhance $^{239}$Pu breeding with the
-relatively fast neutron spectrum. Both designs failed to gain traction and did
-not inspire any working prototypes.
+In spite of this setback, research into \glspl{MSR} continued through the late
+1970s. In 1980, \gls{ORNL} published a report describing a new \gls{MSR}
+concept, called the \gls{DMSR} \cite{gehin_liquid_2016} with denatured
+$^{235}$U fuel (i.e. low-enriched uranium). The \gls{ORNL} researchers
+developed this design in response to the fuel reprocessing restrictions
+introduced by President Ford in 1976. The \gls{DMSR} would operate as a
+once-through
+converter system without fuel reprocessing. While the fuel consists of 19.75
+\% high-assay low-enriched uranium, the initial core loading includes thorium
+to boost its conversion ratio throughout its lifetime. It has a continuous
+online feed consisting of \gls{HALEU} to maintain criticality, and denatured
+$^{235}$U to keep uranium enrichment levels below nuclear non-proliferation
+policy thresholds. The design also includes a gas sparging system for removing
+gaseous fission products, while noble metals plate out onto the walls of
+the coolant loop. The older \gls{MSBR} design had a significant drawback; the
+extensive neutron damage in the graphite moderator necessitated frequent
+replacement (every four years) throughout its operational lifetime. The
+\gls{DMSR} avoids this issue by having a lower power density while maintaining
+the overall power output of 2250 MW$_{\text{th}}$. As a result, researchers
+projected that the graphite moderator would last for the entirety of the
+\gls{DMSR}'s design lifetime.
+
+A concurrent program at the UK Atomic Energy Authority led the
+development of a 2500 MW$_{\text{e}}$ lead-cooled Molten Chloride Fast Reactor
+concept \cite{smith_assessment_1974}. It is a dual fluid system,
+with separate loops for the fuel salt and the blanket salt. The blanket is a
+1 m-wide tank surrounding the core. The absence of moderators and the choice
+of chloride over fluoride salt resulted in a relatively hard neutron spectrum
+which favors $^{239}$Pu breeding over the thorium cycle. The UK researchers
+performed some experiments to study molten salt chemistry but there were no
+reactor prototypes. The UK program eventually shut down just like its
+US counterpart partly due to the successful demonstration of the Prototype
+Fast Reactor which had achieved criticality in 1974.
 
@@
 \cite{heuer_simulation_2010}. Starting from the \gls{MSBR} design, they
-performed parametric studies on the design based on safety, breeding, and
+performed parametric studies based on safety, breeding, and
 other performance metrics \cite{mathieu_thorium_2006}. Their efforts
@@
 The \gls{MSFR} has also benefited from collaborative research through two
-European-funded projects, the EVOL and SAMOFAR projects
-\cite{euratom_final_2015}. Today, numerous other national and commercial
-bodies are also developing their own \gls{MSR} designs.
+European-funded projects, the \gls{EVOL} \cite{euratom_final_2015} and
+\gls{SAMOFAR} \cite{kloosterman_20_2017} projects. Under the \gls{EVOL}
+project, researchers further optimized the \gls{MSFR} design based on
+neutronic and thermal-hydraulic safety analyses. They also diversified
+start-up fuel options beyond the original $^{233}$U-based fuel composition to
+include $^{235}$U, $^{239}$Pu, and \gls{TRU}, and established fuel
+reprocessing schemes for the \gls{MSFR}. The \gls{SAMOFAR} project, which
+stated approximately 2 years after the end of the \gls{EVOL} project,
+supported more comprehensive safety assessments of the reactor and the
+reprocessing plant, and funded a number of experiments for validation of the
+\gls{MSFR}'s safety features. This thesis presents steady-state and
+transient results of the \gls{MSFR} in Moltres. Therefore, Section
+\ref{sec:msfr} describes the \gls{MSFR} design in further detail.
+
+Today, numerous other national and commercial
+bodies are also developing their own \gls{MSR} designs. These include the
+Integral Molten Salt Reactor from Terrestrial Energy
+\cite{leblanc_integral_2015}, the Transatomic Power \gls{MSR}
+\cite{transatomic_power_corporation_technical_2016}, the Russian \gls{MOSART}
+\cite{ignatiev_molten_2014}, and the Chinese Thorium Molten Salt Reactor
+project \cite{zhang_review_2018}.
 
@@
 fuel salt mixture. Examples of one-fluid designs include the Integral Molten
-Salt Reactor from Terrestrial Energy \cite{leblanc_integral_2015} and the
-\gls{MSR} design from Transatomic Power
-\cite{transatomic_power_corporation_technical_2016} while two-fluid designs
-include the \gls{MSFR} \cite{merle_optimized_2007} and \gls{MOSART}
-\cite{ignatiev_molten_2014}. The current work involves simulations of
-the \gls{MSFR} in Moltres and Section \ref{sec:msfr} describes this reactor
-design in further detail.
+Salt Reactor \cite{leblanc_integral_2015} and the Transatomic Power
+\gls{MSR} design \cite{transatomic_power_corporation_technical_2016} while
+two-fluid designs include the \gls{MSFR} \cite{merle_optimized_2007} and
+\gls{MOSART} \cite{ignatiev_molten_2014}.
 
@@
 
+This subsection briefly covers other \gls{MSR} features that are not the focus
+of this thesis for completeness.
+
 \glspl{MSR} have several positive sustainability features.
@@
 diversion.
+
+Economic analyses of \glspl{MSR} are preliminary at the current
+stage of \gls{MSR} development. Qualitatively, many technical factors favor
+\glspl{MSR} over \glspl{LWR}. Some of these factors include: smaller reactor
+core size due to low operating pressures, higher thermal efficiency and
+cheaper air-cooling systems due to high operating temperatures, reduction in
+fuel fabrication costs, and shorter reactor shutdown periods due to online
+refueling \cite{dolan_1_2017}.
 
@@
 
-Both DALTON and the COMSOL-based tool featured again in a later paper by
-Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D axisymmetric
+Both DALTON and the COMSOL-based tool featured again in a 2014 paper by
+Fiorina et al. \cite{fiorina_modelling_2014} for a 2-D axisymmetric
 model of the \gls{MSFR}. The authors, from Politecnico di Milano and
@@
 relatively thermal neutron spectrum. Therefore, the \gls{MSFR} has a more
-thermal neutron spectrum than a typical sodium fast reactor. For an initial
+thermal neutron spectrum than a typical fast-spectrum reactors. For an initial
 composition consisting of $^{232}$Th and $^{233}$U,

### 2020-07-20 neutronics.tex
final comments and edits from katy

@@
 uniform temperature distribution to assess the accuracy of the
-six-group neutron diffusion model in Moltres on a fast-spectrum reactor. This
+six-group neutron diffusion model in Moltres on a fast-spectrum reactor. Table
+\ref{table:static} lists relevant details of the static \gls{MSFR} model
+in Moltres. This
 verification exercise builds on the previous study by Lindsay et al.
@@
 
-\textbf{Details of the static \gls{MSFR} model in Moltres}
-\begin{itemize}
-    \item No salt flow (static salt)
-    \begin{center}
-        $v_{salt} = 0$ m$\cdot$s$^{-1}$
-    \end{center}
-    \item Uniform temperature of 973 K throughout the 2D core model
-    \begin{center}
-        $T = 973$ K
-    \end{center}
-    \item Six neutron energy groups
-    \begin{center}
-        $G = 6$
-    \end{center}
-    \item Eight delayed neutron precursor groups
-    \begin{center}
-        $I = 8$
-    \end{center}
-    \item Vacuum boundary conditions on the outer boundaries of the 2D core
-    mesh
-    \begin{center}
-        $\frac{d \phi}{dx}\big|_{\text{inflow}} = 0$ m$^{-2}\cdot$s$^{-1}$
-    \end{center}
-\end{itemize}
+\begin{table}[htbp!]
+    \small
+	\caption{Details of the static \gls{MSFR} model in Moltres.}
+	\centering
+	\begin{tabular}{ l c}
+		\toprule
+		Detail & Mathematical description \\
+        \midrule
+        No salt flow (static salt) & $v_{salt} = 0$ m$\cdot$s$^{-1}$ \\
+        Uniform temperature of 973 K throughout the 2D core model & $T = 973$
+        K \\
+        Six neutron energy groups & $G = 6$ \\
+        Eight delayed neutron precursor groups & $I = 8$ \\
+        Vacuum boundary conditions for neutron flux &
+        $\frac{d \phi}{dx}\big|_{\text{inflow}} = 0$ m$^{-2}\cdot$s$^{-1}$ \\
+		\bottomrule
+	\end{tabular}
+	\label{table:static}
+\end{table}
 
@@
 SP$_{\text{N}}$ deterministic methods nor the Monte Carlo approach in Serpent.
-Regarding the omission of the blanket tank material, the present author
+Regarding the omission of the blanket tank material, I
 replaced the 2 cm-thick structural material with blanket salt. This
@@
 while $\beta_{\text{eff}}$ is the delayed neutron fraction weighted by the
-adjoint neutron flux. The present author calculated $\beta$ by
+adjoint neutron flux. I calculated $\beta$ by
 taking the relative difference between the $k_{\text{eff}}$ values with and

### 2020-07-20 steadystate.tex
change to plasma color map
final comments and edits from katy

@@
     \centering
-    \begin{subfigure}[t]{.365\textwidth}
+    \begin{subfigure}[t]{.35\textwidth}
         \centering
-        \includegraphics[width=\textwidth]{flow-temp}
+        \includegraphics[width=\textwidth]{flow-temp-plasma}
     \end{subfigure}
@@
     \centering
-    \includegraphics[width=.6\textwidth]{flow}
+    \includegraphics[width=.58\textwidth]{flow-plasma}
     \caption{Fuel salt flow streamlines and velocity magnitude in the core.
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-g1}
+        \includegraphics[width=\textwidth]{g1}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-g2}
+        \includegraphics[width=\textwidth]{g2}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-g3}
+        \includegraphics[width=\textwidth]{g3}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-g4}
+        \includegraphics[width=\textwidth]{g4}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-g5}
+        \includegraphics[width=\textwidth]{g5}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-g6}
+        \includegraphics[width=\textwidth]{g6}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-pre1}
+        \includegraphics[width=\textwidth]{pre1}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-pre2}
+        \includegraphics[width=\textwidth]{pre2}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-pre3}
+        \includegraphics[width=\textwidth]{pre3}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-pre4}
+        \includegraphics[width=\textwidth]{pre4}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-pre5}
+        \includegraphics[width=\textwidth]{pre5}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-pre6}
+        \includegraphics[width=\textwidth]{pre6}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-pre7}
+        \includegraphics[width=\textwidth]{pre7}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{ss-pre8}
+        \includegraphics[width=\textwidth]{pre8}
     \end{subfigure}
@@
         \vspace{.9cm}
-        \includegraphics[width=\textwidth]{ss-pre}
+        \includegraphics[width=\textwidth]{pre}
     \end{subfigure}
@@
 
-Figure \ref{fig:decayheat} shows the decay heat source distribution in the
-core for groups 1 to 3. The decay constants are $\lambda_j = 0.197$ s$^{-1},
-0.0168$ s$^{-1}, 0.000358$ s$^{-1}$ for decay heat groups $j=$ 1, 2, and 3,
-respectively.
-Similar to the delayed neutron precursor distributions, the shorter-lived
-decay heat precursors show greater build-up in the recirculation zones
-compared with the longer-lived decay heat precursors that are more evenly
-distributed throughout the core.
-
 \begin{figure}[htbp!]
@@
         \centering
-        \includegraphics[width=\textwidth]{heat1}
+        \includegraphics[width=\textwidth]{heat1-plasma}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{heat2}
+        \includegraphics[width=\textwidth]{heat2-plasma}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{heat3}
+        \includegraphics[width=\textwidth]{heat3-plasma}
     \end{subfigure}
@@
 \end{figure}
+
+Figure \ref{fig:decayheat} shows the decay heat source distribution in the
+core for groups 1 to 3. Readers may refer to Table
+\ref{table:decayheat} for the associated decay time constants.
+Similar to the delayed neutron precursor distributions, the shorter-lived
+decay heat precursors show greater build-up in the recirculation zones
+compared with the longer-lived decay heat precursors that are more evenly
+distributed throughout the core.
 
@@
 from 101.1 K to 98.9 K as some of the decay heat precursors deposit heat in
-the outer loop region. The figure also shows some unphysical oscillations in
+the outer loop region. The figure also shows some nonphysical oscillations in
 the temperature near the outlet because the mesh near the boundary is too
-coarse. Users can avoid these oscillations near the outlet by using a mesh
+coarse for the given outlet velocities. Users can avoid these oscillations
+near the outlet by using a mesh
 that is progressively finer towards the boundary. The
@@
     \centering
-    \includegraphics[width=.6\textwidth]{decay-heat-temp}
+    \includegraphics[width=.55\textwidth]{decay-heat-temp}
     \caption{Difference in core temperatures at steady state with decay heat

### 2020-07-20 transient.tex
final comments and edits from katy

@@
 insertion, loss of heat sink, loss of flow, and pump overspeed accidents. The
-term ``unprotected'' signify accident scenarios without treactor SCRAM. As
+term ``unprotected'' signifies accident scenarios without treactor SCRAM. As
 such, these simulations provide an insight on the
@@
 without the power profile from the Polimi model with decay heat to compare
-with. However, if the decay power output are similar, the stronger negative
+with. However, if the decay power output were similar, the stronger negative
 temperature reactivity coefficient would cause the prompt power output in the
@@
 
-\clearpage
-
 \section{Unprotected Pump Overspeed}
@@
 
-This work followed Fiorina et al.'s implementation
+This work followed the Polimi and TUDelft models
 \cite{fiorina_modelling_2014} by
@@
 %
-In addition to this, Fiorina et al. applied the Dittus-Boelter correlation
-\cite{dittus_heat_1930} for the relationship between $h_1$ and
+In addition to this, the Polimi and TUDelft models used the Dittus-Boelter
+correlation \cite{dittus_heat_1930} for the relationship between $h_1$ and
 $\dot{V}$. The Dittus-Boelter correlation for fluids being cooled is:
@@
     \intertext{where}
-    Nu &= \text{Nusselt number,} \nonumber \\
-    &\text{\ \ \ \ \ the ratio of convective to conductive heat
-    transfer at a boundary in a fluid,} \nonumber \\
-    Re &= \text{Reynolds number,} \nonumber \\
-    &\text{\ \ \ \ \ the ratio of inertial forces to viscous
-    forces within a fluid,} \nonumber \\
-    Pr &= \text{Prandtl number,} \nonumber \\
-    &\text{\ \ \ \ \ the ratio of momentum diffusivity to thermal
-    diffusivity.} \nonumber
+    Nu &= \text{Nusselt number} =
+    \frac{\text{convective heat transfer}}{\text{conductive heat transfer}}
+    = \frac{hL}{k}, \nonumber \\
+    Re &= \text{Reynolds number} =
+    \frac{\text{inertial forces}}{\text{viscous forces}} =
+    \frac{\rho u L}{\mu}, \nonumber \\
+    Pr &= \text{Prandtl number} =
+    \frac{\text{momentum diffusivity}}{\text{thermal diffusivity}} =
+    \frac{c_p \mu}{k}. \nonumber
 \end{align}
@@
 \end{align}
-
+%
 \begin{figure}[htbp!]
     \centering
-    \includegraphics[width=.7\textwidth]{lof-flow-rate}
+    \includegraphics[width=.6\textwidth]{lof-flow-rate}
     \caption{The change in flow rate in the Polimi and TUDelft models and the
@@
 An issue arose pertaining to the turbulent viscosity $\mu_t$ as a function of
-$v$. When using a simple approximation of $\mu_t$ being directly proportional
+$v$. When using this simple approximation, in which $\mu_t$ being directly
+proportional
 to $v$, the results from Moltres differed significantly compared with the
@@
     \centering
-    \includegraphics[width=.37\textwidth]{lof-flow-temp}
+    \includegraphics[width=.37\textwidth]{lof-plasma}
     \includegraphics[width=.62\textwidth]{fiorina-lof}

### 2020-07-22 abstract.tex
writing checklist and Kozlowski review comments

@@
 As the world continues its transition towards low-carbon electricity
-generation to combat climate change, molten salt reactors is a potential
-option in the near-term future for replacing fossil fuel and ageing nuclear
+generation to combat climate change, building molten salt reactors is a
+potential
+option in the near-term future for replacing fossil fuel and aging nuclear
 power plants. At the current state of development, molten salt reactors still
@@
 developments in Moltres, a simulation tool for molten salt reactors. These
-new developments are: support for coupling the incompressible Navier-Stokes
+new developments are: the support for coupling the incompressible
+Navier-Stokes
 and the delayed neutron precursor looping systems, and a model for
-simulating decay heat from fission products at steady state and during
+simulating decay heat from fission products at steady-state and during
 transients. This work demonstrates these capabilities through multiphysics
@@
 small discrepancies on the order of 100 pcm, which is smaller than the
-$-256.7$ pcm discrepancy reported in literature with the same six-group
-neutron diffusion approach. The decay heat model shows an expected flattening
-of the temperature distribution due to the movement and dispersion of the
+$-256.7$ pcm discrepancy reported in the literature with the same six-group
+neutron diffusion approach. The decay heat model showed an expected flattening
+of the temperature distribution due to the movement of the
 decay heat precursors throughout the primary coolant loop. This work also
-demonstrates and verifies steady state and
+demonstrates and verifies steady-state and
 transient multiphysics simulations of the Molten Salt Fast Reactor. The
@@
 insertion, loss of heat sink, loss of flow, and pump overspeed. This thesis
-verified the steady state and transient results against data from the
+verifies the steady-state and transient results against data from the
 literature for the same case studies. The steady-state temperature and
@@
 reactor response to these transient initiators.
-The loss of flow results showed greater discrepancies that are attributed
-to differences in the fluid dynamics modeling in Moltres and the other models.
+The loss of flow results showed greater discrepancies that resulted from
+differences in the fluid dynamics modeling in Moltres and the other models.
 Through the verification studies, this work has also identified avenues

### 2020-07-22 msr.tex
writing checklist and Kozlowski review comments

@@
-\glspl{MSR} are one of six advanced reactor designs shortlisted by
+\glspl{MSR} are one of six advanced reactor classes shortlisted by
 the Generation IV Forum in 2001 for promising significant advances in safety,
 sustainability, efficiency, and cost over existing designs. This
-has attracted significant attention and resources towards
+attracted significant attention and resources towards
 \gls{MSR} research, most noticeable by the number of start-up companies that
@@
 replacement (every four years) throughout its operational lifetime. The
-\gls{DMSR} avoids this issue by having a lower power density while maintaining
+\gls{DMSR} avoids this issue by running on a lower power density while
+maintaining
 the overall power output of 2250 MW$_{\text{th}}$. As a result, researchers
@@
 which favors $^{239}$Pu breeding over the thorium cycle. The UK researchers
-performed some experiments to study molten salt chemistry but there were no
-reactor prototypes. The UK program eventually shut down just like its
+performed some experiments to study molten salt chemistry but they did not
+build reactor prototypes. The UK program eventually shut down just like its
 US counterpart partly due to the successful demonstration of the Prototype
@@
 reprocessing schemes for the \gls{MSFR}. The \gls{SAMOFAR} project, which
-stated approximately 2 years after the end of the \gls{EVOL} project,
+started approximately two years after the end of the \gls{EVOL} project,
 supported more comprehensive safety assessments of the reactor and the
@@
 \glspl{MSR} are typically straightforward \gls{LEU} burners or
-$^{232}$Th/$^{233}$U iso-breeders/breeders, while epithermal- and
+$^{232}$Th/$^{233}$U breeders, while epithermal- and
 fast-spectrum \glspl{MSR} can operate as \gls{TRU} fuel burners or
@@
 fuel. Combined with the Doppler broadening of resonance capture cross sections
-present in both fuel forms, there would be a smaller temperature
+present in both fuel forms, an \gls{MSR} would see a smaller temperature
 increase following an unprotected reactivity insertion. The overall
@@
 Firstly, \glspl{MSR} have smaller fractions of delayed neutron precursors in
-the active core region as some of them decay in the external loop regions.
+the active core region as some of the decay in the outer loop regions.
 This complicates reactor control and may result in faster transients due to
@@
 
-In more recent years, there has been
-significant progress towards fully coupled, spatial codes that feature
+In more recent years, researchers made significant progress towards fully
+coupled, spatially resolved computational tools that feature
 2-D axisymmetric or full 3-D models. In 2011, Cammi et al.
-\cite{cammi_multi-physics_2011} performed a ``\gls{MPM}'' analysis
+\cite{cammi_multi-physics_2011} performed a ``Multi-Physics Modelling''
+analysis
 of a simplified 2-D axisymmetric model of a single \gls{MSBR} fuel channel
@@
 Navier-Stokes flow to simulate salt flow more accurately. The authors
-validated their model against \gls{MSRE} experimental data before presenting
+validated their model against \gls{MSRE} experimental data and presented
 multiphysics simulation results of a different graphite-moderated breeder
@@
 
-Both DALTON and the COMSOL-based tool featured again in a 2014 paper by
-Fiorina et al. \cite{fiorina_modelling_2014} for a 2-D axisymmetric
-model of the \gls{MSFR}. The authors, from Politecnico di Milano and
-Technical University of Delft, compared \gls{MSFR} transient safety
-analysis results from both tools. They referred to the COMSOL and DALTON/HEAT
+Fiorina et al. \cite{fiorina_modelling_2014} adopted both DALTON and the
+COMSOL-based application in their study of a 2-D axisymmetric model of the
+\gls{MSFR}. The authors, from Politecnico di Milano and Technical University
+of Delft, compared \gls{MSFR} transient safety analysis results from both
+tools. They referred to the COMSOL and DALTON/HEAT
 \gls{MSFR} models as the Polimi and TUDelft models, respectively, after the
@@
 ensures a greater safety margin \cite{mathieu_thorium_2006}. Lastly, while
-negative temperature coefficients are attainable with very thermalized
-spectra, breeding ratios deteriorated significantly due to parasitic
-absorption in the graphite \cite{mathieu_thorium_2006}.
+negative temperature coefficients are attainable with extremely thermalized
+spectra, these configurations require large volumes of graphite which
+significantly deteriorate breeding ratios due to parasitic absorption
+\cite{mathieu_thorium_2006}.
 %
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{MSFR}
+	\includegraphics[width=0.7\textwidth]{MSFR}
 	\caption{Schematic view of the MSFR concept. Figure reproduced from
@@
 central core region. At the top of the core, the flow separates into sixteen
-smaller external loops, each of which passes through a heat exchanger before
+smaller outer loops, each of which passes through a heat exchanger before
 being pumped back into the bottom of the core. The salt also passes through
 the online salt reprocessing and gas sparging systems located along the
-external loops. A toroidal blanket tank containing fertile salt for breeding
+outer loops. A toroidal blanket tank containing fertile salt for breeding
 surrounds the core radially. The top and bottom of the core are enclosed
@@
 ratio of $^{232}$Th to $^{233}$U slightly to achieve exact criticality at a
-uniform temperature of 973 K for their own neutronics software
+uniform temperature of 973 K for their neutronics software
 \cite{brovchenko_neutronic_2019}. This thesis
@@
 
-The thermal and electric power output of the \gls{MSFR} are 3000 MW
+The thermal and electric power outputs of the \gls{MSFR} are 3000 MW
 $_{\text{th}}$ and 1300 MW$_{\text{e}}$, respectively. The high thermal
@@
 \cite{fiorina_modelling_2014}. The design is a 2-D axisymmetric representation
-of the \gls{MSFR} with the sixteen individual external loops homogenized into
+of the \gls{MSFR} with the sixteen individual outer loops homogenized into
 a single outer loop as shown in Figure \ref{fig:msfrgeom}. For the multigroup
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{msfr-geom}
+	\includegraphics[width=0.65\textwidth]{msfr-geom-legend}
 	\caption{2-D axisymmetric model of the \gls{MSFR} core used for the
@@
 Although the present work uses the same 2-D axisymmetric geometry for
-generating group constant data from Serpent, there are two minor differences
+generating group constant data from Serpent, two minor differences exist
 between the \gls{MSFR} geometry modeled herein, and the geometry used in the
@@
 pipe with a pointwise heat sink to represent the heat exchanger. To simulate
-the pumps, a Dirichlet boundary condition for velocity at the inlet boundaries
+the pumps, a Dirichlet boundary condition for inlet velocity
 drives the flow in the central core region of the primary loop. At every

### 2020-08-27 conclusion.tex

@@
 steady-state and transient analyses. The steady-state temperature and velocity
-distributions showed many similiarities in their shapes and magnitudes to the
+distributions showed many similarities in their shapes and magnitudes to the
 Polimi and TUDelft model results \cite{fiorina_modelling_2014}. The uniform
-$\mu_t$ approximation was responsible for the minor differences in the flow at
-the top of the core and the loss of delayed neutrons to out-of-core emissions.
+$\mu_t$ approximation was responsible for some differences in the flow in the
+recirculation zones. In turn, these differences in the recirculation zones
+resulted in
+differences in the loss of delayed neutrons to out-of-core emissions.
 The results with the decay heat model showed a slight flattening of the
@@
 exchanger in Moltres and the volumetric heat exchanger in the other two
-models required minor adjustments in the relationship between flow rate
+models required adjustments in the relationship between flow rate
 $\dot{V}$ and the heat transfer coefficient $h$ from the original
@@
 
-Further research and development on Moltres should aim to rectify the issues
+Further research and development of Moltres should aim to rectify the issues
 mentioned in this thesis. This work has highlighted three main avenues for
-improvement. Firstly, proper 2D/3D heat exchanger implementation
-would allow us to move away from the 1D outer loop system and towards a full
-2D/3D closed loop. If this were implemented, users would be able to use
+improvement. Firstly, a proper 2-D/3-D heat exchanger implementation would
+provide users with the option to simulate a fully closed
+2-D/3-D coolant loop. If this were implemented, users would be able to use
 the Boussinesq approximation for buoyancy-driven flow capability in Moltres.
@@
 \cite{wilcox_turbulence_2006}. Turbulence effects are
-significant in \gls{MSR} designs with fast flow and they inform optimization
+significant in \gls{MSR} designs and they inform optimization
 studies for improving flow patterns and eliminating local hotspots in eddies.
@@
 that simplifying assumptions for turbulence lead to erroneous results under
-flow conditions that deviate significantly from steady state.
+flow conditions that deviate significantly from steady-state.
 

### 2020-08-27 model.tex

@@
 heating from fission products.
-To run simulations with Moltres, the user must provide input
+To run simulations with Moltres, the user must provide
 group constant data from a neutron transport solver for the
-multigroup neutron diffusion calculations, and a mesh file representing the
+multigroup neutron diffusion calculations and a mesh file representing the
 geometry of the reactor. This work uses Serpent 2 \cite{leppanen_serpent_2014}
@@
 active development led by the VTT Technical Research Centre of Finland. It was
-created in 2004 for group constant generation in lattice geometries, and has
+created in 2004 for generating group constants in lattice geometries and has
 since grown to support more general capabilities. Serpent 2 is highly
@@
 is helpful for many researchers as they are unencumbered by the
-technical details and complexities involved in programing mesh-handling
+technical details and complexities involved in programming mesh-handling
 and \gls{PDE}-solving in finite element analysis.
@@
 functionalities. As a result, \gls{MOOSE} supports adaptive meshing schemes
-and automatic variable scaling amongst other advanced features in finite
+and automatic variable scaling, amongst other advanced features in finite
 element analysis. Full
@@
 \gls{MOOSE}, and Moltres by extension, are capable of up to 3-D geometry
-modelling. They support a wide range of input mesh file formats, including the
+modeling. They support a wide range of mesh file formats, including the
 commonly used Exodus II file format. Specifically for 2-D geometries, users
@@
 precursors back into the core, and a pointwise heat removal kernel to simulate
-a heat exchanger. The present author demonstrated these capabilities in a
-2-D axisymmetric model of the \gls{MSFR} with uniform salt flow
-\cite{park_safety_2019}. The present work also includes these capabilities
-which are discussed in the following section on the modeling approach.
+a heat exchanger. The present author demonstrated these capabilities in an
+earlier work \cite{park_safety_2019} with a 2-D axisymmetric model of the
+\gls{MSFR} with uniform salt flow. The present work also includes these
+capabilities which are discussed in the following section on the modeling
+approach.
 
@@
 Accordingly, this section provides separate descriptions for the governing
-equations in the central core region and the outer loop region. There is also
-a third region comprised of the blanket salt, the NiCrW Hastelloy reflectors, 
-and other material.
+equations in the central core region and the outer loop region.
 
@@
     \centering
-    \includegraphics[width=.5\textwidth]{central-core}
+    \includegraphics[width=.75\textwidth]{central-core-legend}
     \caption{2-D axisymmetric model of the MSFR. The red box indicates the
@@
 
-The central core region is of greatest interest to us during steady state and
+The central core region is of greatest interest to us during steady-state and
 transient scenarios; the center of the reactor is naturally where most of the
@@
 
-The neutron flux calculations in the central core region are
-performed using the standard formulations for the time-dependent multigroup
+Moltres performs neutron flux calculations in the central core region
+using the standard formulations for the time-dependent multigroup
 neutron diffusion equations and \gls{DNP} concentration equations as shown in
@@
     scattering from $g'$ to $g$ [cm$^{-1}$],} \nonumber \\
-    \chi^p_g &= \text{prompt fission spectrum for neutrons in group $g$,}
+    \chi^p_g &= \text{prompt fission spectrum for neutrons in group $g$ [ - ],
+    } \nonumber \\
+    G &= \text{total number of discrete neutron groups [ - ],} \nonumber \\
+    \nu &= \text{average number of neutrons produced per fission [ - ],}
     \nonumber \\
-    G &= \text{total number of discrete neutron groups,} \nonumber \\
-    \nu &= \text{average number of neutrons produced per fission,} \nonumber
-    \\
     \Sigma^f_{g} &= \text{macroscopic fission cross section for neutron in
     group $g$ [cm$^{-1}$],} \nonumber \\
-    \chi^d_g &= \text{delayed fission spectrum for neutrons in group $g$,}
+    \chi^d_g &= \text{delayed fission spectrum for neutrons in group $g$
+    [ - ],} \nonumber \\
+    I &= \text{total number of delayed neutron precursor groups [ - ],}
     \nonumber \\
-    I &= \text{total number of delayed neutron precursor groups,} \nonumber \\
-    \beta &= \text{total delayed neutron fraction,} \nonumber \\
-    \beta_i &= \text{delayed neutron fraction of precursor group $i$,}
+    \beta &= \text{total delayed neutron fraction [ - ],} \nonumber \\
+    \beta_i &= \text{delayed neutron fraction of precursor group $i$ [ - ],}
     \nonumber \\
@@
     \rho &= \text{ density of the fuel salt [kg m$^{-3}$],} \nonumber \\
-    Sc_t &= \text{ turbulent Schmidt number.} \nonumber
+    Sc_t &= \text{ turbulent Schmidt number [ - ].} \nonumber
 \end{align}
@@
 TUDelft models \cite{fiorina_modelling_2014} which used the same value. It has
-its roots in the Reynolds Analogy which states that turbulent momentum and
+its roots in the Reynolds Analogy, which states that turbulent momentum and
 heat transfer largely depend on the same eddies in turbulent flow
@@
 
-Moltres users may vary the total number of neutron energy groups as
+Moltres users can use an arbitrary number of neutron energy groups as
 long as they provide Moltres with the appropriate group constant data. The
@@
 
-The \gls{MSFR} model in this work six neutron energy groups according to the
+This work uses six neutron energy groups according to the
 energy boundaries in table \ref{table:bound}, and eight \gls{DNP} groups as
@@
 	f_j &= \text{fraction of total power attributable to decay heat group
-	$j$,} \nonumber \\
+	$j$ [ - ],} \nonumber \\
 	\epsilon_g &= \text{average fission energy per fission initiated by a
@@
 In addition to the intrinsic molecular viscosity in the \gls{INS} equations,
-this thesis introduces an eddy viscosity term, $\mu_t$, to approximate
+this thesis includes an eddy viscosity term, $\mu_t$, to approximate
 turbulent flow effects. The current implementation of the Navier-Stokes module
@@
 uses a zeroth-order approximation of $\mu_t$ based on the calculated $\mu_t$
-values reported in the Polimi and TUDelft models. The models
-predicted spatial $\mu_t$ values ranging from 0-110 Pa$\cdot$s, with most
-of the core lying in the 30-50 Pa$\cdot$s range. Thus, the present work uses
+values reported in the Polimi and TUDelft models. The models predicted spatial
+$\mu_t$ values ranging from 0 to 110 Pa$\cdot$s, with most values
+falling within the 30 to 50 Pa$\cdot$s range. Thus, the present work uses
 the approximated value $\mu_t=$ 40 Pa$\cdot$s. Despite the simplicity of this
-approximation, the resulting flow profile is similar to flow profile in the
-Polimi and TUDelft models at steady state.
+approximation, the resulting flow profile is similar to the flow profile in
+the Polimi and TUDelft models at steady-state.
 
@@
     [W$\cdot$m$^{-1}\cdot$K$^{-1}$],} \nonumber \\
-    J &= \text{total number of decay heat groups.} \nonumber
+    J &= \text{total number of decay heat groups [ - ].} \nonumber
 \end{align}
@@
 near the blanket tank walls and the resulting high temperatures in that
-region. The results in Chapter \ref{chap:ss} would be some minor
-discrepancies where the viscosity values are under- or over-predicted, leading
-to minor inaccuracies in temperature and precursor concentration distributions
-from turbulent diffusion.
+region. The results in Chapter \ref{chap:ss} show minor discrepancies in
+regions where the viscosity values were under- or over-predicted.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{msfrbc}
+    \includegraphics[width=.8\textwidth]{msfrbc}
     \caption{The boundaries in the \gls{MSFR} geometry that are relevant for
@@
 
-\subsection{Flow Transfers}
+\subsection{Central Core and Outer Loop Coupling}
 
 This subsection details the delayed neutron and decay heat precursors, and
-temperature flow transfers between the central core and outer loop regions.
+temperature coupling between the central core and outer loop regions.
 
@@
 
-A parabolic flow profile was imposed on the inlet Dirichlet boundary
+This work uses a parabolic flow profile on the inlet Dirichlet boundary
 condition. The equation for $u_x$ at the inlet is:
@@
     \intertext{where}
-    \psi &= \text{variable to be weighted} \nonumber \\
-    \mathcal{C} &= \text{outlet boundary curve} \nonumber \\
+    \psi &= \text{variable to be weighted [ - ]} \nonumber \\
+    \mathcal{C} &= \text{outlet boundary area [ - ]} \nonumber \\
     u &= \text{outflow velocity perpendicular to the outlet boundary
@@
 
-This outflow value from the central core region is transferred to the 1-D
-outer loop region as input for the inhomogeneous Dirichlet boundary
-condition at the inlet boundary. Likewise, the outflow value from the outer
+Moltres transfers this outflow value from the central core region to the 1-D
+outer loop region, to be used as the boundary value for the inhomogeneous
+Dirichlet boundary
+condition at the inlet. Likewise, the outflow value from the outer
 loop region is used for the inflow value in the central core region. No

### 2020-08-27 neutronics.tex

@@
 SP$_{\text{N}}$ deterministic methods nor the Monte Carlo approach in Serpent.
-Regarding the omission of the blanket tank material, I
-replaced the 2 cm-thick structural material with blanket salt. This
+Regarding the omission of the blanket tank material, this model
+replaces the 2 cm-thick structural material with blanket salt. This
 replacement is partly responsible for the higher $k_{\text{eff}}$ value
@@
 while $\beta_{\text{eff}}$ is the delayed neutron fraction weighted by the
-adjoint neutron flux. I calculated $\beta$ by
+adjoint neutron flux. This work calculates $\beta$ by
 taking the relative difference between the $k_{\text{eff}}$ values with and
@@
 extends below the melting point of the fuel salt (841 K) to ensure that
-the data covers the relevant range between 841 K and 900 K. Table
+the data cover the relevant range between 841 K and 900 K. Table
 \ref{table:alpha} shows the various $\alpha$ values calculated using the
@@
 in the spectrum corresponding to elastic scattering resonances from lithium
-and fluorine. From this plot, we observe that the discrepancies in
-$k_{\text{eff}}$ arise mainly from discretizing neutron energy into groups
-rather than the neutron diffusion model itself. We could obtain a more
-accurate representation of the neutronics in the \gls{MSFR} by using more
-neutron energy groups but this would adversely impact simulation times in the
+and fluorine. We could obtain a more accurate representation of the neutronics
+in the \gls{MSFR} by using more neutron energy groups with appropriate energy
+boundaries but this would adversely impact simulation times in the
 subsequent multiphysics finite element analyses.

### 2020-08-27 steadystate.tex

@@
 $1 \times 10^{14}$ cm$^{-2}$ s$^{-1}$ uniformly throughout the geometry.
-Finally, this work assumes that steady state is reached when the volume
+Finally, this work assumes that model has reached steady-state when the volume
 integral values of every variable remain constant (up to 6 significant
 figures) for at least four seconds in the simulation; this
-time period corresponds to the nominal circulation time of the \gls{MSFR}.
+period of time corresponds to the nominal circulation time of the \gls{MSFR}.
 
@@
 Figure \ref{fig:flow-temp} shows the temperature and velocity fields of the
-fuel salt in the core at steady state from Moltres and the Polimi and TUDelft
+fuel salt in the core at steady-state from Moltres and the Polimi and TUDelft
 models. Figure \ref{fig:flow} provides an alternate view of the flow profile
-through flow streamlines superimposed on the velocity magnitude distribution.
+with flow streamlines superimposed on the velocity magnitude distribution.
 The results from Moltres show good qualitative agreement with the Polimi and
@@
 half-height approximately 0.40 m away from the central axis. A large
-recirculation region forms near the blanket tank walls from turbulent flow.
-Inertial forces dominate over viscous forces to form this large eddy. The main
+recirculation region forms near the blanket tank walls. The main
 difference between Moltres, and the Polimi and TUDelft models is the flow
@@
     \includegraphics[width=.58\textwidth]{flow-plasma}
-    \caption{Fuel salt flow streamlines and velocity magnitude in the core.
+    \caption{Fuel salt flow streamlines and velocity magnitude in the core
+    from Moltres.
     The colors represent velocity magnitude according to the color bar on the
@@
 
-Although the temperatures at the hotspots are well below the melting point of the Ni-alloy structure (1500 K), they may cause undue thermal stress on the
+Although the temperatures at the hotspots are well below the melting point of
+the Ni-alloy structure (1500 K), they may cause undue thermal stress on the
 blanket tank structure and accelerate salt corrosion rates. A
@@
 transient scenarios when the flow profile fluctuates.
-Thus, Rouch et al. \cite{rouch_preliminary_2014} developed an improved
-hourglass-shaped design for the central core region to optimize flow
-distribution and prevent these
-recirculation zones and hotspots from forming. A study of this new design
-using Moltres is a potential subject for future work when a proper turbulence
-model is in place.
 
@@
     1 to 6. The y and x axes represent height and radius (in cm) of the core
-    relative to the entire reactor geometry.}
+    relative to the entire reactor geometry. Note the different scales for
+    each flux distribution.}
     \label{fig:neutronflux}
@@
 		\midrule
-		{Moltres (This paper)} & {Steady state} & 9.80\\
+		{Moltres (This paper)} & {Steady-state} & 9.80\\
 		{COMSOL} & {Uniform, 973 K} & 8.6 \\
@@
 longer-lived groups, the \gls{DNP} concentrations are ill-resolved on the
-mesh elements adjacent to the outlet and the inlet boundaries, respectively.
+mesh elements adjacent to the outlet and the inlet boundaries.
 Thus, the present author recommends careful mesh refinement for future work
@@
 neutron flux in response to reactivity insertions because they have a greater
-proportion of prompt neutrons under normal operating conditions. This is
-undesirable from a reactor safety perspective because it reduces the time
+proportion of prompt neutrons under normal operating conditions. This effect
+is undesirable from a reactor safety perspective because it reduces the time
 available for active safety mechanisms to
-activate and scram the reactor. In \glspl{MSR}, this danger is partly
-mitigated by the strong, negative fuel temperature reactivity coefficient. 
+activate and scram the reactor. In \glspl{MSR}, the strong, negative fuel
+temperature reactivity coefficient partly mitigates the sharp increase in
+neutron flux. 
 Chapter 6 contains more in-depth discussions for various transient scenarios.
@@
 
-The fraction of out-of-core emissions from our Moltres model
+The fraction of out-of-core emissions from the Moltres model
 differs significantly by approximately 10\%, and $\beta_c$ differs by
@@
 reports the value as the physical fraction without adjoint weighting. The
-weighting results in a significant difference in $\beta_c$ because a greater
+adjoint weighting results in a smaller $\beta_c$ because a greater
 fraction of the \glspl{DNP} decay in the recirculation zones, where the
-neutron importance is noticeably diminished.
+neutron importance is noticeably diminished. Fast reactors also typically have
+smaller $\beta_{\text{eff}}$ because the fission cross sections of fertile
+isotopes (e.g. $^{238}$U and $^{232}Th$) monotonically decrease with
+decreasing neutron energy \cite{ott_introductory_1985}. On average, delayed
+neutrons are less energetic than prompt neutrons.
 
@@
 Figure \ref{fig:decayheattemp} shows the
-difference in core temperatures at steady state with decay heat modeling
+difference in core temperatures at steady-state with decay heat modeling
 relative to the core temperatures without decay heat in Figure
@@
 coarse for the given outlet velocities. Users can avoid these oscillations
-near the outlet by using a mesh
-that is progressively finer towards the boundary. The
-present author could not resolve this issue due to time constraints.
+near the outlet by using a mesh that is progressively finer towards the
+boundary. This exercise is left as an objective for future work.
 
@@
     \includegraphics[width=.55\textwidth]{decay-heat-temp}
-    \caption{Difference in core temperatures at steady state with decay heat
+    \caption{Difference in core temperatures at steady-state with decay heat
     relative to the result without decay heat (Figure \ref{fig:flow-temp}).}

### 2020-08-27 transient.tex

@@
 insertion, loss of heat sink, loss of flow, and pump overspeed accidents. The
-term ``unprotected'' signifies accident scenarios without treactor SCRAM. As
+term ``unprotected'' signifies accident scenarios without reactor SCRAM. As
 such, these simulations provide an insight on the
@@
 transients. Furthermore, only their Polimi model had this capability.
-Therefore, they featured decay heat modeling for the loss of heat sink
+Therefore, they included decay heat modeling for the loss of heat sink
 accident scenario only. The current work also ran all transient simulations
@@
     \caption{Power output during the prompt response following a 50 pcm
-    step-wise unprotected reactivity insertion in the Moltres, Polimi, and
-    TUDelft models \cite{fiorina_modelling_2014}.}
+    step-wise unprotected reactivity insertion in the Moltres model.}
     \label{fig:50pcmjump}
@@
 temperature. By $t=3$ s, most of the heated salt and \glspl{DNP} from the
-initial phase will have circulated around the primary loop and returned to the
+initial phase will have circulated the primary loop and returned to the
 core. The heated salt causes a small, noticeable dip in power before the power
@@
 Moltres model at $t=120$ s is approximately 17 K lower than that in the Polimi
-model. It is difficult to ascertain the exact cause for this difference
-without the power profile from the Polimi model with decay heat to compare
-with. However, if the decay power output were similar, the stronger negative
+model. Unfortunately, the power profile from the Polimi model is not available
+for direct comparison.
+However, if the decay power output were similar, the stronger negative
 temperature reactivity coefficient would cause the prompt power output in the
@@
     \caption{Power output during
-    an unprotected loss of heat sink transient in the Moltres model with and
-    without decay heat.}
+    an unprotected loss of heat sink transient in the Moltres model with
+    decay heat.}
     \label{fig:moltresdecaypower}
@@
 increase in pump speed in the primary coolant loop. The increased flow rate
-$\dot{m}$ impacts reactor performance in several ways.
-It affects the neutronics by reducing the in-core $\beta$ as more of the
-shorter-lived precursors will tend to flow out of the core before decaying.
-This net loss of neutrons reduces the reactivity in the core, thereby causing
-core temperatures to fall to counteract this change through temperature
-reactivity feedback. The increased $\dot{V}$ also enhances the heat transfer
-coefficient on the primary loop side of the heat exchanger and enables the
-reactor to operate at a higher power output. At the same time, the improved
+$\dot{V}$ impacts reactor performance in several ways. It reduces the in-core
+$\beta$ as more of the shorter-lived precursors will tend to flow out of the
+core and decay in the subcritical outer loop region.
+The increased $\dot{V}$ also enhances the heat transfer
+coefficient on the primary loop side of the heat exchanger and introduces
+cooler fuel salt to the core. The reactivity increase from the temperature
+reactivity feedback outweighs the reactivity decrease from the loss of delayed
+neutrons. Thus, the overall neutron flux and power output is higher at the
+new equilibrium state. At the same time, the improved
 mixing flattens the temperature distribution in the core.
@@
 unprotected pump overspeed transient and overpredicted the same parameter in
-the unprotected loss of flow transient. Upon further investigation, I
+the unprotected loss of flow transient. Upon further investigation, it was
 found that raising the power of $v$ from 0.8 to 1.1 brought the
@@
 20 seconds of the transient. At the start of the transient, the rising flow
-rate cools the core and causes power output to rise sharply. Although the
+rate cools the core by convecting heat away from the center and causes power
+output to rise sharply. Although the
 average core temperature has a strictly decreasing trend, the temperature at
@@
 distribution in the core is more evenly distributed because the turbulent
-thermal conductivity $k_t$ is directly proportional to $mu_t$.
+thermal conductivity $k_t$ is directly proportional to $\mu_t$.
 
@@
 this approach was not possible in Moltres because the primary loop is
-partitioned into two separate geometries and used Dirichlet boundary
+partitioned into two separate geometries and uses Dirichlet boundary
 conditions at the inlet to drive flow. The Polimi and TUDelft
@@
 An issue arose pertaining to the turbulent viscosity $\mu_t$ as a function of
-$v$. When using this simple approximation, in which $\mu_t$ being directly
-proportional
-to $v$, the results from Moltres differed significantly compared with the
-Polimi and TUDelft models. The difference is due to buoyancy-driven flow
+$v$. This simple approximation, in which $\mu_t$ being directly proportional
+to $v$, causes the results from Moltres to differ significantly compared with
+the Polimi and TUDelft models. The difference is due to buoyancy-driven flow
 contributing to turbulence; the turbulent energy $k$ equation in COMSOL's
@@
 recirculation zones disappear during the loss of flow transient. The current
-work circumvented this issue by letting fixed fractions of the initial
-$\mu_{t,0}$ be conserved regardless of the final flow velocity, according to
+work circumvented this issue by retaining fixed fractions of the initial
+$\mu_{t,0}$ regardless of the final flow velocity, according to
 the following equation:
@@
     \intertext{where}
-    \mu_c &= \text{ conserved fraction of $\mu_{t,0}$ [Pa$\cdot$s].} \nonumber
+    \mu_c &= \text{ fixed fraction of $\mu_{t,0}$ independent with time
+    [Pa$\cdot$s].} \nonumber
 \end{align}
@@
 results showing better qualitative agreement with those from the Polimi and
-TUDelft models. The subsequent paragraphs discuss these results.
+TUDelft models.
 
@@
 the wider peak in the average core temperature
-indicates that there was a more gradual change in the flow pattern.
+indicates a more gradual change in the flow pattern.
 
